### 2025-01-15 src/main.tex
sync with overleaf

%%
%% This is file `sample-nime-paper.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-nime-paper.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% History of NIME Template Updaters:
%% Modified by Florent Berthaut, Charles Martin, Yichen Wang, 21 November 2024
%% Modified by Adnan Marquez-Borbon 30 November 2022
%% Modified by Courtney Reed 28 November 2022
%% Modified by Joe Wright 14 December 2019
%% Modified by Niccolò Granieri 10 October 2018
%% Modified by Angelo Fraietta 23 December 2018
%% Modified by Angelo Fraietta 22 November 2018
%% Modified by Rodrigo Schramm on 22 September 2018
%% Modified by Luke Dahl on 17 October 2-17
%% Modified by Cumhur Erkut on <2016-10-11 Tue>
%% Modified by Edgar Berdahl on 5 November 2014
%% Modified by Baptiste Caramiaux on 25 November 2013
%% Modified by Kyogu Lee on 7 October 2012
%% Modified by Georg Essl on 7 November 2011
%%
%% And here's the template:
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{nimeart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{nimeart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{nimeart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{cc}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{}
%% These commands are for a PROCEEDINGS abstract or paper.
%% Make sure this is up to date with the correct edition of NIME.
\acmConference[NIME '25]{International Conference on New Interfaces for Musical Expression}{June 24--27,
  2025}{Canberra, Australia}
%% This suppresses the ACM Reference Format printing.
\settopmatter{printacmref=false}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The NIME of the Title Is Hello}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \city{San Antonio}
  \state{Texas}
  \country{USA}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. This uses ``nimeart'', a fork of the ``acmart''
  documentclass with additions for the NIME conference. This
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{images/sampleteaser.pdf}
  \caption{Two synthesisers on a table, 2019.}
  \Description{Enjoying looking at two synthesisers.
  A Korg Monotron and a Korg Volca FM.}
  \label{fig:teaser}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|nimeart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|nimeart|'' document class can
be used to prepare many different kinds of documentation --- a
double-anonymous initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|nimeart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{nimeart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\texttt{acmsmall}}: The default journal template style.
\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
\item {\texttt{acmtog}}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\texttt{sigconf}}: The default proceedings template style.
\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \texttt{\string\acmSubmissionID} command to print the
  submission's unique ID on each page of the work.
\item{\texttt{authorversion}}: Produces a version of the work suitable
  for posting by the author.
\item{\texttt{screen}}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{nimeart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|nimeart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification.  As an exception, multiple authors may share one
affiliation. Authors' names should not be abbreviated; use full first
names wherever possible. Include authors' e-mail addresses whenever
possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Anonymisation and Blind Review}
Anonymisation is a \textbf{crucial} component for NIME paper submissions. We kindly ask authors to ensure that all information that could introduce bias in the review process is anonymised.

To anonymise author names, please specify \verb|anonymous,review| in square brackets when submitting your paper.
\begin{verbatim}
  \documentclass[sigchi,anonymous,review]{nimepaper}
\end{verbatim}

You can also hide text in the paper like this:

\begin{verbatim}
% Both examples from acmart manual.
\begin{anonsuppress}
  This is the continuation of the previous work by the author
  \cite{prev1, prev2}.
\end{anonsuppress}
\end{verbatim}

or
\begin{verbatim}
This work was performed at \anon{NSA}.
This work was performed at \anon[No Such Agency]{NSA}.
\end{verbatim}

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``nimeart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
\verb|\paragraph|, and \verb|\subparagraph|. The sectioning levels up to
\verb|\subsusection| should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

Below are examples of sectioning commands.

\subsection{Subsection}
\label{sec:subsection}

This is a subsection.

\subsubsection{Subsubsection}
\label{sec:subsubsection}

This is a subsubsection.

\paragraph{Paragraph}

This is a paragraph.

\subparagraph{Subparagraph}

This is a subparagraph.

\section{Tables}

The ``\verb|nimeart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{images/sample-bela.jpg}
  \caption{A Bela board in action at a workshop (2017).}
  \Description{Somebody manipulating sensors on a Bela board.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
  Artifacts: \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

%% The Ethical Standards section is mandatory for all NIME submissions.
\section{Ethical Standards}

To ensure objectivity and transparency in research and to ensure that accepted principles of ethical and professional conduct have been followed, authors must include a section “Ethical Standards” before the References.
This section should include (if relevant): information regarding sources of funding, potential conflicts of interest (financial or non-financial), informed consent if the research involved human participants, statement on welfare of animals if the research involved animals or any other information or context that helps ethically situate your research.
For help with the ethics section, feel free to ask on the NIME forum: \url{https://forum.nime.org}.

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-references}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-nime-paper.tex'.

### 2025-01-15 src/article.tex
Create article.tex

\section{Abstract}\label{abstract}

With the advent of digital musical instruments in the 20th Century there
has been a decoupling between interface and sound generation when
playing. Consequently, aspects of the sensory experience of playing have
been lost, namely the sensation of vibration and mechanical impedance
that a traditional musical instrument provides.

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order \cite{Tagliavini2009}
(https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
items in the collection are not in playing condition or are too fragile
to continue to be available for the public to play. One of those fragile
instruments in the Tagliavini collection is a 1547 Harpsichord by
Alessandro Trasuntino \cite{Treccani2019}, which was used as the basis
for an interactive exhibition commisioned by San Colombano.

Project was also designed to allow others to easily recreate and iterate
on the designs.

First a small cross-sectional model was used as a proof of concept. Once
a sensing system was decided upon, a larger 49-key scale model was
commissioned with and then built by luthier Roberto Livi. The models are
designed to provide only a tactile and haptic response for the player. A
sensor system was installed into the model and messages were sent to
computer that triggered sounds from a sample library which the player
listens to through headphones. The paper looks at the design decisions,
processes, and considerations to be made when undertaking a project of
this kind.

Overview of fabrication methods, techniques that can be employed.
Finally a discussion of the limitations and areas in which the core
project can be improved and developed further.

\section{Introduction}\label{introduction}

With the advent of digital musical instruments in the 20th Century there
has been a decoupling between interface and sound generation when
playing. Consequently, aspects of the sensory experience of playing have
been lost, namely the sensation of vibration and mechanical impedance
that a traditional musical instrument provides.

In this context, a `traditional' instrument is one whose interface for
generating sound and the sound generating mechanism are fundamentally
coupled. A piano requires a key press which triggers a hammer, which
strikes a string which vibrates a body and air cavity. This system is
tied together and cannot be separated. A synthesizer, be it a theremin,
a control voltage synthesizer, a MIDI controller, these are all
instruments whose interface are decoupled from their sound generation.
This decoupling presents a problem that then needs to be addressed. How
do we re-introduce this dimension to the experience of playing a musical
instrument? How do we, and is it possible to, parameterise and preserve
this experience? This is central problem that is trying to be solved in
{[}1, 2{]}.

This is particularly pertinent as it relates to conservation of musical
instruments, as they represent both historic objects, but also tools for
creating music. Historical musical instruments present an interesting
problem for cultural heritage as they are both a physical object but
also a tool for making music. This begs the question, can we use haptics
to preserve experiences of the past? The proposed project intends to
develop further the methodologies implemented in musical haptics
projects {[}3{]} such as {[}2{]}. The final outcome will be to create a
framework for digital conservation of historic musical instruments such
as the 1638 Ioannes Ruckers and the 1646 Ioannes Couchet double-manual
harpsichords as referenced in the NEMUS project {[}4{]}.

Systems exist already for the conversion of piano keyboard. The Yamaha
disklavier, the Don Buchla designed MOOG Piano Bar, Bosendorfer, PNOscan
by QRS focus on reproduction. These same systems could be applied to a
harpsichord, but the limitation is a single data stream per key.

\begin{itemize}
\tightlist
\item
  existing sysetms

  \begin{itemize}
  \tightlist
  \item
    Yamaha Disklaviare
  \item
    Bosendorfer

    \begin{itemize}
    \tightlist
    \item
      290SE
    \item
      CEUS
    \end{itemize}
  \item
    Don Buchla / Moog Piano Bar
  \item
    PNOscan MIDI 9 QRS Music

    \begin{itemize}
    \tightlist
    \item
      PNOmation
    \end{itemize}
  \item
    Piano Disc
  \end{itemize}
\end{itemize}

\section{Background}\label{background}

Prvious projects looking at how to approach the problems of increasing
engagement and retaining the interction with historical musical
instruments in a museum scenario.

The paper by McAlpine \cite{McAlpine2014} also has very similar
motivations to the project presented here. Both were started from a
musical instrument museum, housing a collection of historic instruments
gifted by a collector (Benton Fletcher / Taglivaini) who wished for
collection to remain playable, which is restricted growing fragility of
some items in the respective collections. Bothe projects center around
the creation of a custioom MIDI controller for interfacing with sound
synthesis engine. The keyboard consisted of two modified commercial MIDI
keyboard controllers in a two manual harpsichord arrangement. Weherease
the McAlpine keyboard used weighted keys, this interface extends the
idea by using the mechanism and strings under tension to provide the
haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
but before taking such a leap we look at what affect a more literal
haptic keyboard may provide to the performer experience.

The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
looked at reconstructing a full tromba marina to be used as a digital
interface. The one of the limitations flagged by McAlpine was the
oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
Moderna ``attempted to avoid this problem by constructing a fully
functioning replica of the tromba marina'' \cite{Baldwin}. This project
applies the same approach to the harpsichord.m where the projects
deviate is a questions of scale.

Where the tromba moderna considers a single data stream this project has
had to address 98 and the problems that come with increasing scale.
Material costs mitigated by creating a scaled version of the sound
board, provinding only the length of string necessary under comparable
tension and thus resistance to the jack quill.

Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
sourcing of hardware and software for the project.

\subsection{Why\ldots{}}\label{why}

\subsubsection{\ldots a replica?}\label{a-replica}

\subsubsection{\ldots an interface?}\label{an-interface}

\begin{itemize}
\tightlist
\item
  maintain problems
\item
  probe for research

  \begin{itemize}
  \tightlist
  \item
    on haptics how it affects performance
  \end{itemize}
\end{itemize}

\subsection{Motivations}\label{motivations}

The project was carried out in collaboratoiun with San Colombano. The
results of the project would be presented as an interactive exhibition
for the public.

\subsubsection{the problem}\label{the-problem}

Central philosophy of the Tagliavini Collection is that the instrument
should remain playable to the public. The condition of some instruments
is such that restoration would required a replacement of vast majority
of the instrument, would be too costly, would risk the instrument given
its fragility, or all of those in combination.

This is not an uncommon problem faced by musical instrument museums
worldwide. The central question the project tries to answer is ``how do
we conserve musical instruments in order to continue interacting with
them''

\subsubsection{the solution}\label{the-solution}

The approach taken was to recreate the interface of these instruments
not for acoustics purposes be purely for the mechanical resoponse. As
exhibition for the general public in a musical instruments there were
aesthetic and functional constraints placed on the design of the
interface.

\subsubsection{Criteria}\label{criteria}

A keyboard interface that provides a comparable physical response to a
harpsichord. Notes should be assigned per jack and not per key.

\subsection{Keyboard}\label{keyboard}

\subsubsection{Anatomy of Harpsichord
mechanics}\label{anatomy-of-harpsichord-mechanics}

\begin{itemize}
\item
  simple
\item
  aesthetics
\item
  split workload
\item
  parts

  \begin{itemize}
  \tightlist
  \item
    jack

    \begin{itemize}
    \tightlist
    \item
      quill
    \item
      jack body
    \item
      pivot
    \item
      tongue
    \end{itemize}
  \item
    key
  \item
    body
  \end{itemize}
\end{itemize}

\subsubsection{Design}\label{design}

The desicion was made to design the elctronic sna dthe model keyboard in
tandem. Allow for quicker development time and iteration on the idea.
Likely the first implementation would need re-designed and redesiginng
the electronics to fit the form of the keyboard was more cost effective
than building or carrying out major modifications to the model.

The model was designed to have two internal chambers a front and rear.
The front and back that was xxx by yyy by 370mm between the faceplate
and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber though
a small gap between the under side of the soundboard and the key guides.
The key guide needs to be deep enough to allow for the full movement of
the key and stop the key becoming unseated. The chamber then extends to
a support wall which divides the front and rear.

Discussion of spatial constraints and affects on electyronics design and
installation is covered in

\section{setup}

The rear chamber was desigbned to house comonents for audio processing,
which is discussed further in \textbackslash section

https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf

For the initial prototyping stage, modifying an existing harpsichord was
considered though ultimately discarded. An existing instrument would
have provided a test for scaling the electronics, but internal
measurments and layout would have been too different. A harpsichord that
was economic enough to be deconstructed would also have been of a more
modern mechanical design and not of the design that was being recreated
with the model. Prototyping stage would likely be invasive and damage or
heavy modification likely. As such it was not a good use of funds as a
modern instrument fit for destruction did not offer benefits that the
small model with easy internal access did not already provide.

\subsection{Building Process}\label{building-process}

PARAGRAPH FROM ROBERTO

\subsubsection{3-Key Prototype}\label{key-prototype}

\begin{itemize}
\tightlist
\item
  proof of concept
\item
  available on loan from museum
\item
  easy access to internal structure
\end{itemize}

\subsubsection{Fig: Construction Design
Diagram}\label{fig-construction-design-diagram}

\subsubsection{49-Key Prototype}\label{key-prototype-1}

For the final interface a 49-key (4-octave) was constructed by luthier
Roberto Livi. The keyboard is in the same style† as the 3-key model with
to jacks per key.

The keyboard is fully enclosed and internally has two chambers, front
and rear, in which the electronics could be installed \figure(underside
49-key).

Keyboard to be used for exhibition at san colombano where it would be
available to play by general public.

Timescale: from initial discussion to delivery of interface to the NEMUS
lab was 8-months. Active assembly of the instrument was 3 months

Constructed From:

\begin{itemize}
\tightlist
\item
  for exhibition
\item
  problems of scale
\item
  problems of timescale
\end{itemize}

\subsubsection{Fig: Construction Design
Diagram}\label{fig-construction-design-diagram-1}

\subsubsection{Design Constraints}\label{design-constraints}

\begin{itemize}
\tightlist
\item
  no electronics should be visible.
\item
  it should be robust and reliable, easily maintained by museum staff
  without complex technical processes
\item
  it should not augment or fix the mechanics, but present them with
  their original limitation
\end{itemize}

Another constraint was that the design of the sensor system and the
model keyboard would have to carried out at a distance in separate
workshops. Time constraints meant the keyboard was being fabricated for
a sensor that did not yet exist.

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\FIGURE, but there was still a limit given the structural and aesthatic
requirements.

\subsubsection{Sensor Criteria}\label{sensor-criteria}

A set of criteria was drawn to which the final sensor system would have
to satisfy. This criteria was used as a means of evaluating the
feasability throughout the prototyping phase.

\begin{itemize}
\tightlist
\item
  Non-invasive: The sensor system should not require major modifications
  to the interface on which it is installed.
\item
  Low Latency: Reading sensors and then triggering an output should have
  a low latency. Using {[}paper on latency{]} a target was set as X ms
  with an upper limit set to 25ms beyond which the sensors would be
  deemed unusable
\item
  Reliable: Data obtained should be reliable and accurate enough not to
  result in false positives or false negatives without extensive
  filtering
\item
  Scalable: The system should be scalable both economically and
  temporally. Given the open soucre nature of the porject, it could not
  be prohibitvely expensive to carry out. Also any system that would
  work on a single key would have to scale financially to potentially 50
  or more. The time required to install and calibrate the sensors would
  also have to scale.
\item
  Expandable: Any sensor system should be functionally expandable to
  allow for the usage of the data for control over other parameters.
\end{itemize}

\subsubsection{Triggering a note}\label{triggering-a-note}

For the first iteration the data from the sensor would simply be used to
test when a threshold had been passed. On passing the threshold a
message would be sent to trigger the playback of a corresponding note.

The message format covered in \textbackslash section

\section{Related Work and Motivation}\label{related-work-and-motivation}

\begin{itemize}
\tightlist
\item
  \cite{HapticKey}
\item
  \cite{mcPherson1} \cite{mcPherson2}
\item
  \cite{Mudd2013}
\item
  \cite{FritzClaudia2017Leon}
\end{itemize}

\section{Hardware Design}\label{hardware-design}

The hardware was designed with the xx in mind.

Using a similalr apporach to @HapticKey the first step was to fabricate
or procure a model of a single working key. San Colombano provided a
model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
3-key model was used a test-bed for potential sensors while designs were
drawn for a full-scale model. The cross-sectional nature of the model
meant it was easy to fit and remove sensors. Small size meant that it
would also be easy to transfer between workshops.

3-key model also had a sliding jack rail which allowed for easy
disengagement of the physical mechanical response, which allowed for
some preliminary tests on the presence and absence of mechanism.

\subsection{Failed sensors}\label{failed-sensors}

A number of sensors were tried before the final system was decided upon.
An inertial measurment unit (IMU) was mounted to a single key which
contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
a consistent set of conditions to trigger sound, but this was not easily
differntiable when other keys were played. The fitting time and cost of
a single IMU could b reduced when purchased in bulk but was far greater
than other systems considered.

A reed sensor and a hall sensor were mounted and magnets embedded within
the jacks. Adjusting the hieight of the sensor provided a reliable means
of detecting a threshold, inparticular the hall sensors which could be
partially calibrarted physically and also in softwrae. The time taken to
calibarte a single key was time consuming. Ideally each sensor would be
placed on a thread for adjustment, but given the constraints on the
internal space it was decided any adjustment of heights would likely
mean an unnacceptable length of time to install and calibrate. Embedding
magnets into the jacks was also did not satisfy the criteria of being
scalable and non-invasive.

Light dependant resistor mounted to the jack rail. as jacks were pressed
less light would be let through. Usable data but very dependant on
lighting conditions. Modifying the jack also did not satisfy the
non-intrusive criteria and the placement outside the internal chamber of
the model meant hiding electronics would be more difficult.

Finally, force sensitive resistors placed under the jack of each key was
considered but was never implemented. The sensors were to be placed
under each jack either above or below the padding on which the bottom of
the jack rests. Further discussion of this apporach is in
\textbackslash section.

\section{Electronics}\label{electronics}

Design of electronics carried out in tandem with construction of full
model. Since the Electronics were designed at a distanced to the final
full model some flexbility was required in the design. Design decisions
that were not as optimal but would allow for change in wiring and
fitting should there be any issues during installation.

Electronics designed so that they would not require any equipment
outside what can commonly be found in a maker space at a univeristy.

The equipment required including soldering suitable for throughole
technology (THT) components and 3D printing.

Laser cutting and CNC routing fascilities were used during the research
for the project but not required for the final implementation.

The idea behind this was to allow for the project to be reasonably
recreated without procurement of any additional specialist equipment.

NEMUS lab under construction during the first section of the project.
Facilities available at the maker spaces of Bologna (Alma Labor) and
Edinburgh (uCreate) used for what was deemed reasonable assumption in
terms of available equipment.

\begin{itemize}
\tightlist
\item
  Over arching themes

  \begin{itemize}
  \tightlist
  \item
    simplicity of design

    \begin{itemize}
    \tightlist
    \item
      Limited resources
    \item
      easily reproducible
    \end{itemize}
  \item
    availability and lead times
  \item
    iteration and design from distance
  \end{itemize}
\end{itemize}

\subsubsection{QRE1113}\label{qre1113}

Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
combination infrared LED and phototransistor sensitive to IR light. The
phototransistor of QRE1113 provides data on how much light is present
and in particular how much light is being relfected from a surface.
Since refelected light will be proportional to the distance of a surfcae
it is a good, close proximity, distance sensor. Distance is the way in
which the instruments in \citeMcpherson1\&2 function as well the Moog
Piano Bar from which it too inspiration.

The limitation for taking a distance approach is that there is not one
but 2 jacks per key. For each jack to be measured indepeedntly the data
would need tp com fropm the jacks themselves.

They \emph{could} be used and individual trigger points set in
calibration, but this creates new problems.

\begin{itemize}
\tightlist
\item
  The triggering would need to be bespoke for every key as fluxuations
  in the dimensions of every piece mean that the threshold would be
  different.

  \begin{itemize}
  \tightlist
  \item
    Moving a problem into additional time in the calibtraryoin setup
    phase.
  \end{itemize}
\item
  Remebering the constraint that system isshould noyt fix problem, there
  would be no easy way to tell if a jack had nbot reseated.

  \begin{itemize}
  \tightlist
  \item
    In playing a harpihchord when the jack does not go to its rest
    position, not note would play. This fault should be carried over
    intp the final instrument.
  \end{itemize}
\end{itemize}

Another common application of the QRE1113 is within line following
robots, where an array of QRE1113s ar used in conjunction with black
tape along which some motorised system can following

Applied to the jacks was a greyscale gradient printed the length full
travel range of the jack.

For the first iteration this was performed on an inkjet printer and
double-sided tape. For the final iteration the stickers were printed on
(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
reduced the manual labour for cutting stickers meaning the process would
scale from the 3-key to the 49-key model.

An initial sticker design was used that contained only the gradient. It
was found that the jacks wer sticking in their slots andthe edges on the
stickers easily peeled around the printing The sticker was redesigned to
be the full length of the jack which avoided the problem of peeling. The
larger are also mean the stickers adhered betterto the jack Some
trimming was still required of each sticker to avoid them adding extraa
resistence to the jack's movement.

During testing it was found that their was a considerable amount
cross-talk bwteen adjacent sensors. This meant the reading from key
would be affected by the movement of the key next to it. The cross-talk
was assumed to be as a result of light reflecting from other jacks.
Given that IR is used this was difficult to confirm by sight. A set of
baffles was designed for the pcb that would wrap round the jack and stop
light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
printers using PLA plastic. A dark pigmented PLA was used to avoid IR
light simply being diffused. After the baffles were fitted the
cross-talk was entirely eliminated.

\begin{itemize}
\tightlist
\item
  previous use
\item
  problem of what to sense
\item
  satisfied both criteria and design restrictions
\item
  usage

  \begin{itemize}
  \tightlist
  \item
    Sweet spot is around 6mm distance
  \end{itemize}
\end{itemize}

\subsubsection{sensor surface}\label{sensor-surface}

\begin{itemize}
\tightlist
\item
  linear gradient
\item
  on Jack
\item
  printed and cut on vinyl {[}find name{]}
\item
  cut to travel size
\item
  then to jack size
\item
  jacks varnished on side
\item
  alternatives

  \begin{itemize}
  \tightlist
  \item
    CNC plywood and mdf

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \item
    laser cut plywood and mdf

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \item
    3D printed jack body

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \end{itemize}
\end{itemize}

\subsection{Arduino}\label{arduino}

\subsubsection{Nano BLE}\label{nano-ble}

Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
appealing as it would allow for easy changes between chipsets without
requiring any rewiring and thus no new PCBs would need to be created if
a better alternative was found later in the project. Another benefit to
the Arduino Nano boards was that typically had native-USB meaning the
could be programmed as a hardware USB MIDI device directly.

The Nano BLE specifically was used in the initial stages for the
integrated IMU.

When it was finalised that the prokect would proceed using the QRE1113 a
number of microcontroller units (MCU) were condsidered.

The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
Arduino Nano ESP32 and STM32 NUCLEO-L031K6.

A rough performance test was later carried out using an STM32
NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
more performative. To execute a single sensor read cycle of the firmware
the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
on average. The Nano ESP32 had high variability so it was decided to
continue with the Nano BLE. Further discussion of using ESP32 continues
in

\section{Going forward}

BLE functionality also meant if the cable connection between the nano
and computer running audio synthesis was impractical, BLE MIDI was
available as a fallback. The Nano BLE also has a 12-bit DAC, which
provided a contingency should the default 10-bit not be sensitive enough
for the signal from the sensors.

Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
functionality, but it was found that 2 ADC channels were in fact not
useable and would have required additional multiplexers.

\begin{itemize}
\tightlist
\item
  Nano form factor small
\item
  allows for easy changes in chipset
\item
  ``Native USB''
  https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
\item
  BLE Model

  \begin{itemize}
  \tightlist
  \item
    Performative
  \item
    BLE capabilities for MIDI BLE transmission
  \item
    All analog pins available
  \item
    available at the time
  \item
    12-bit DAC
  \end{itemize}
\end{itemize}

\subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}

There would need to be the ability to store values on non-volatile
memory so that the electroincs could be powered-down without losing
data.

Adjusting the values through firmware would not have been flexible or
practical. Instead an Ferroelectric RAM unit was chosen and connected
using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
was primarily chosen for it's ready availability.

Most microcontrollers have some variety of non-volatile program memory
that can be read and writ accessed but in most cases this memory is
cleared during compilation and re-flashing. The small cost was far
outweighed by the potential for data loss if using onboard RAM.

\emph{It is worth noting that the FRAM model chosen stopped being
available during the project, meaning any future iteration would need to
adjust for another board in firmware}

The firmware was structure in a way that the interface could be changed
so long as the code required for any other board would comply with
function arguments and outputs. Worth noting the fragility and
volatility when designing this kind of project and how easily it is for
the project to become broken.

SD card would provide an easier way to interact with calibration data.
The posiiton of the controller board did not permit easy access.
reliability and speed was preffered over storage capacity.

\begin{itemize}
\tightlist
\item
  On board available, but volatile
\item
  retained across firmware changes
\item
  small,
\item
  could have been sd card or anything else
\item
  available
\item
  interfaced via rotary encoder
\end{itemize}

\subsection{RGB LEDs}\label{rgb-leds}

for problems of scale:

Addressing the problem of scale of identifying which key's threshold is
currently being edited.

The LEDs also allowed for easy identification for keys that were outside
of expected values and when a key was triggered. This allowed for visual
identification of malfunctioning or uncalibrated keys.

The LEDs are on the same PCB as the sensors and are hidden when the
nameboard is in place.

Would sugggest that this is a necessity of this kind to provide a means
to identify and diagnose LEDs with state.

TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
to assembling, but were avoid in favour of Addressable LEDs with
intergrated driver. The intergrated LEDs share the same data line which
has the limitation where a fualty LED will stop all following LEDs from
working. An individual PWM driver per PCB would allow avoid situations
where failutre cascades, but would add complexity in assembly as the
parts are typically SMD. Also, given the LEDs are fixed to the board and
could not be repaired in situ, the solution would be to replace the PCB
regardless of which approach was taken. Given the LEDs were only used in
debugging and calibration and otehrwise would not be visible, a simpler
design approach was taken.

\begin{itemize}
\tightlist
\item
  locating keys
\item
  displaying key state
\end{itemize}

\subsection{Multiplexing}\label{multiplexing}

CD4051BE Texas Instruments

problem of scale

For the 3-key model, individual jacks could be connected to one of the 8
available ADC pins.

The BLE Nano has 8 ADC pins connected to one ADC channel via a
multiplexer

Multiplexer added to expand to the

Transistor / Op-amp circuit could have been used similar to the one in
\cite{mcpherson}.

Benefits were keeping with THT based designed for easy assembly.

An 8-channel multiplexer was chosen as the model has 49-keys and a
division of 7 PCBs each containing 7 sensors

The multiplexer reduced 7 signals to a single pin. Three pins are
required pins for addressing the muliplexewrs and all PCBs are addressed
simulatenously.

\begin{itemize}
\tightlist
\item
  limited ADC channels
\item
  ADC channels already multiplexed
\item
  simple
\item
  8 channel
\item
  THT
\item
  7 into 49

  \begin{itemize}
  \tightlist
  \item
    7 PCBS
  \item
    each with 7 sensors
  \end{itemize}
\end{itemize}

\subsection{PCB design}\label{pcb-design}

\begin{itemize}
\tightlist
\item
  EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
  deprecated EAGLE and the programme will no longer be supported from
  2026 Project files will need to be transitioned to another standard,
  likely the KiCAD programme.
\end{itemize}

trade off between limiting number of boards required and allowing
compensation for variation in pitch.

Smaller PCBs means that spacing could be adjusted to accomodate the
changes in pitch between jacks. It also allowed for a modular approach
where any PCB could swapped for minimal cost.

\subsubsection{Sensor board}\label{sensor-board}

The 49 QRE1113s are divided across 7 PCBs each containing

\begin{itemize}
\item
  x7 QRE1113
\item
  x7 100R
\item
  x7 10kR
\item
  x1 CD4051BE Multipexer
\item
  jack / sensor pitch by board
\item
  7 boards adjust for pitch/spacing error
\item
  modular, replaceable
\item
  orientation problems

  \begin{itemize}
  \tightlist
  \item
    legs break easily
  \item
    end mount like A.P.Mc. time consuming
  \item
    problem

    \begin{itemize}
    \tightlist
    \item
      milling out keys
    \end{itemize}
  \end{itemize}
\item
  designed from distance
\item
  striped lines across pcbs
\item
  problems

  \begin{itemize}
  \tightlist
  \item
    additional terminals uneeded resulted in lots of additional work
  \item
    long cabling increase risk of breaking
  \end{itemize}
\item
  solutions

  \begin{itemize}
  \tightlist
  \item
    striping all adc lines with FFC
  \item
    solder pads to select channel
  \end{itemize}
\end{itemize}

\subsubsection{controller board}\label{controller-board}

Controller board designed around Arduino nano form factor. The Nano is
connected through headers to allow easy switching between chips. Should
another chip in a Nano for factor be found it could supplant the current
Nano without any need for a redesign.

The first version of the board esclusively used 2.54mm pitch headers.
The reasons for this were flexibility when creating cable looms and to
leave open the possiility of rewuring the PCB.

This was the case for the LEDs as they initially wer to be powered on
5V. It was discovered that the 3v regulator of the nano was sufficient
to power all LEDs at a lower brightness while reducing current draw.
Since brightness was not importnat.

The 2.54mm pitch headers had two major drawbacks. First, the orientation
of the cables could be reversed, which in the best case scenario would
simply mean the system did not work correctly and in the worst case it
could potentially damaged some components.

Secondly, the headers do not connect securely and wait of the cabling
was enough for components to become disconnected as the harpsichord was
handled.

To address this, once the wiring for the controller board was finalised
a new design was drafted this time using JST-PH connectors. No
commercial cables were immediately available in the configuration and
length required, as such a cable loom needed to be made which meant
crimping cables by hand. A crimping tool from JST is prohibitively
expensive compared to the cost of other components in the project*. The
crimping tool also appeared to be uncommon and was not readily available
in the workshops at the projects disposal. Eventually a tool was found,
but an alternative approach was also researched in order to maintain the
criteria of accessibility.

A locator was adapted from a creative commons stl model
https://www.thingiverse.com/thing:1646016 that allowed a generic
crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
locator was printed using a Bambu X1 3D printer.

\emph{** A second alterntaive was assembling cables from other sets that
were at hand.} \emph{* from RS
https://uk.rs-online.com/web/p/crimp-tools/6880877 £508 inc VAT} \emph{†
model available in the git repository}

\begin{itemize}
\tightlist
\item
  mount for nano to allow switching
\item
  initially push fit 2.54mm header

  \begin{itemize}
  \tightlist
  \item
    not a secure fit
  \end{itemize}
\item
  JST-PH used

  \begin{itemize}
  \tightlist
  \item
    crimping required custom cables

    \begin{itemize}
    \tightlist
    \item
      could be assembled from other cables
    \item
      length limited
    \end{itemize}
  \item
    prohibitevly expensive to crimpers
  \item
    3D print helper to allow for crimping with generic crimpers
  \end{itemize}
\end{itemize}

\subsection{3D printing}\label{d-printing}

Access to a 3D printer is was vital to the success of the project. Given
the proliferation of 3D printing, particular in maker spaces in
universities it was deemed an unreasonable requirement. 3D printed parts
include the support brackets for PCBs, the baffles for the sensors,
locator for crimping JST-PH crimps and a mount for the MacMini used for
audio synthesis.

The printer used throughout the project was a Bambu X1-Carbon using
recycled PLA filament.

Support brackets and baffles were made for the project while the MacMini
bracket and JST locator were sourced from creative commons models.

Any colour of filament can be used, though it is advised to use a dark
coloured filament for the baffles.

See each corresponding section for more detail on 3D printed components.

\begin{itemize}
\item
  https://www.thingiverse.com/thing:1646016
\item
  https://www.thingiverse.com/thing:2886460
\item
  Supports

  \begin{itemize}
  \tightlist
  \item
    Push fit
  \item
    adjustable for best average placement
  \end{itemize}
\item
  Blinkers

  \begin{itemize}
  \tightlist
  \item
    reduce noise from other senors
  \end{itemize}
\item
  mounts

  \begin{itemize}
  \tightlist
  \item
    MacMini
  \end{itemize}
\end{itemize}

\subsection{Power}\label{power}

All electronics draws around 1.1amps of current at 5V with some
fluctuation during boot.

Arduino BLE has some functionality disabled to limit current draw.

Ideally the unit could be bus powered from USB.

The sensors are in a state of constantly drawing power. It is possible
that powering only when data is required could be reduce power enough.

Current designs have a separate MIDI device for each jack row. Combining
the devices and also reducing current draw to below the common 0.5 amps
for a USB port would certainly be a challenge.

\section{Firmware Design}\label{firmware-design}

Arduino platform chosen for wide adoption, relative ease of use and
library support for components.

Firmare checks components are connected then reads the FRAM to see if
threshold data is already present. The rotary encoder and it's momentary
switch are polled for any change in state. A key is selected with the
rotary encoder while the momentary switch is used for changing between
states for key selection and threshold adjustment A double click of the
rotary encoder writes the current threshold values to FRAM. The sensors
for the current multiplexer channels are polled. If any sensor reads
above the threshold when it was previously below a MIDI On message is
sent. For the reverse case where the value was below the threshold when
it was previously above, a MIDI Note Off message ois sent.

\begin{itemize}
\tightlist
\item
  Arduino platform

  \begin{itemize}
  \tightlist
  \item
    library support and portability
  \item
    dependancy management
  \item
    wide spread use
  \end{itemize}
\item
  core algorithm

  \begin{itemize}
  \tightlist
  \item
    check components connected
  \item
    READ data from FRAM
  \item
    select key with rotary

    \begin{itemize}
    \tightlist
    \item
      click to edit, double to save
    \end{itemize}
  \item
    current threshold and reading sent over serial and plotted
  \item
    when over threshold send MIDI note
  \end{itemize}
\end{itemize}

\subsection{Project structure}\label{project-structure}

The project is separated into three repositories encapsulating

\begin{itemize}
\tightlist
\item
  Firmware
\item
  Keyboard CAD Data
\item
  PCB CAD Data
\end{itemize}

\subsubsection{Firmware}\label{firmware}

Firmware repository contains all firmware for the full and smaller
models and components.

\subsubsection{Keyboard}\label{keyboard-1}

The Keyboard repository contains all plans for fabricating the keyboard,
measurements and 3D models of the jacks.

\subsubsection{PCB}\label{pcb}

PCB repository contains the EAGLE files for creating the sensor and
controller boards.

Given the regular spacing of components, scripts for automatic placement
of components based on parameter of jack pitch spacing and number of
sensors per jack.

\section{Implementation}\label{implementation}

\subsection{Installation}\label{installation}

The electronic components were partly pre-assembled at the university of
Edinburgh. Though all measurments were known ahead, enough flexibiliyty
was needed to allow for proboems to solved in situ.

The final assembly took place at the NEMUS lab at the univerity of
bologna over the final week in October 2024 for delivery at San
Colombano for exhibition in the first week of November in 2024.

A misreading in the original clearance between the keys and the PCB
meant that the keys wer block entirely from moving. A 10mm by 15mm
section was milled out of each lke \FIGURE to create extra clearance.
The milling required the removal, shortening and refitting of the
leather pads for the jacks \FIGURE. The extra work added an extra day to
the full install.

A redesign of the PCBs meant that removing the material would no longer
be necessary.

The PCBs shared power and data lines for components with the exception
of the output from the multiplexer. Ribbon cable was cut to size to
allow for some room to readjust the distance of the sensors from the
jacks. The data signals from each PCB were connected using a separate
cable loom. Soldering cables and inspecting each join required another
day work. This did allow for readjustments to the positon of the
controller board Measurements of the rear chamber did not match with the
MacMini that was to be used a different layout was decided for fitting
the controller board/ Cutting cables to size allowed for a great deal of
felxibility but at the cost of time. After the first install it was
clearer what constraints were in place for fittig and an alternative
wiring system using flat flex cable was implement (discussed in
\textbackslash section)

Fitting PCBs required the removal of all keys. The support brackets were
assembled and screwed into the the underside. With the harpsichord on
it's back the sensors were tested to ensure an expected range.
Self-tapping round head screws were used to to attach the brackets. The
screws were not over tightened to allow for brackets to move. The
brackets were not easily adjusted once the key were replaced. Adjusted
to a best average with subsequent adjustment carried out on the
microcontroller.

As the PCBs were fitted the jacks had the gradient stickers applied to
them. The jacks had been coated with a varnish on the short side to
provide a better surface for the stikcers to adhere to. After the
stickers were applied excess material was trimmed from each jack. The
jack were then check that they moved freely in their slot.

After fitting the PCBs preliminary calibration was carried out. Before
the keys were reffitted the jacks wer put back in place. The data from
each key was confirmed to be in a similar range. Any key showing
anomalous dat had it's jack removed and re-inspected. The PCB fitting
process and refitting required another day as faults were found in
wiring or misalignement of the stickers. After the sensors were
confirmed to be functioning the keys were put back.

The calibration workflow was with Dr.~Craig Webb who was unfamiliar with
the system. Adjustments were made to the firmware to to make it more
intuitive for the user. Addressable LEDs became vital as it allowed for
easy identification of the key being calibrated. They were also used to
identify any keys whose current reading was beyond what was expected.

The rotary encoder was used as the interface to select a key, adjust
it's threshold and update the Ferroelectric RAM (FRAM) with new values.

A process was implemented where the key was selected by playing, but
this caused problems in situations where the data was unusual due to
misalignement of the sensors

This would likely be a better system for when the keyboard was fully
operational

The thresholds were calibrated visually utilising the serial plotter of
the Arduino IDE and MIDI was confirmed using the open source MIDI
Monitor (https://github.com/krevis/MIDIApps).

\begin{itemize}
\tightlist
\item
  partial pre-assembly
\item
  handsolder components

  \begin{itemize}
  \tightlist
  \item
    analog lines
  \end{itemize}
\item
  connecting supports
\end{itemize}

\subsection{Context}\label{context}

Designed as part of an exhibition for the museum San Colombano in
Bologna. Instruments is to be played by general public as a means of
interacting with items in the collection that are no longer in playing
condition. Audience of one who listens via headphones.

The strings of the interface are damped with felt strips. The damping
did change the tension and thereforew the feel of the strings so
adjustments needed to be made to the tuning to accomodate for this.

\begin{itemize}
\tightlist
\item
  in museum context
\item
  strings damped
\item
  headphones
\item
  interfacing with kontakt spitfire library
\item
  display

  \begin{itemize}
  \tightlist
  \item
    secure storage and access
  \item
    interface to the public
  \end{itemize}
\end{itemize}

\subsection{Calibration}\label{calibration}

Finer calibration carried out with the guidance of Catalina Vincens as
an expert harpsichord performer

\begin{itemize}
\tightlist
\item
  sensor surface

  \begin{itemize}
  \tightlist
  \item
    printed gradient
  \end{itemize}
\end{itemize}

\section{Conclusion}\label{conclusion}

\begin{itemize}
\tightlist
\item
  scale
\item
  time scale
\item
  working at distance
\item
  limitations

  \begin{itemize}
  \tightlist
  \item
    latency
  \item
    deprecation

    \begin{itemize}
    \tightlist
    \item
      software

      \begin{itemize}
      \tightlist
      \item
        EAGLE
      \end{itemize}
    \item
      hardware

      \begin{itemize}
      \tightlist
      \item
        FRAM
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{itemize}

\section{Going forward}\label{going-forward}

\begin{itemize}
\tightlist
\item
  Velocity, aftertouch
\item
  second jack row
\item
  reducing latency
\item
  alternative sensor surfaces
\end{itemize}

A limitation in the design is the variablility in positioning of tage,
especially when they need to be reapplied. A jig was designed for
applying tags, but could not be printed in time for installation. Even
with a jig, some tags had to be radjusted by hand regardless. Long term
this hasd impliccation on how to maintain the interface.

Taken a different approach to what surface is used. Preliminary test
were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
cutting results in a loss of material that has too much ariablity. The
edge o f the jack body will be charrred. The material also needs to be
reliable. on the advice of two luthiers, roberto livi and Jonathan Santa
Maria Bouquet at Edinburgh, Plywood and MDF have problems with
variations in humidity, tear-out and delamination.

Some tests were carried out using a notched jack body that was 3D
printed. The notch means that as the jack raises, the forward material
acts like a shutter. No reasons that this could not be CNC though a
variety of materials would be needed. Traditional material used is xx
though need to test of the material is naturally reflective enough or if
a coating would beed to be applied.

Outputs of the NEMUS project will also include numerically simulated
non-linear plucked string models with which the device can interact. A
bespoke sound synthesis engine to leverage data from both jack rows
which will also be open-sourced. Design for instrument to be compatible
with standard MIDI instruments as well software designed for the full
data output.

A second interface designed more closely to the Trasuntino style
instrument commisioned by the NEMUS project with Roberto Livi. Using
lessons learned for the first iteration the second can make better
acommodations for internal electronics. Interface to be used for a
performer study and used for performance with a numerically simulated
sound syntheiss engine.

Testing current sensors in conjunction with force sensitive resistors to
research if a impulse response for the pluck can be extrapolated.

Future novel designs for an interface that would utilise velocity and
aftertouch messages in software. This would be applicable for numericla
simulations of strings. A bespoke piec e of sowftare would also allow
for direct mapping of jack rows to registers. There is enough data to
implemenet functionality for velocity and aftertouch MIDI messages and
simply needs to be activated in the firmware.

Evaluation

\subsection{improvements}\label{improvements}

\begin{itemize}
\tightlist
\item
  ESP-IDF

  \begin{itemize}
  \tightlist
  \item
    latency half
  \item
    workload high
  \end{itemize}
\end{itemize}

A current limitation is the latency in triggering MIDI output. Currently
the latency is around 20 ms, but should ideally be closer to 5ms.

An obvious approach would be merely changing the microcontroller. A test
was carried out using an Arduino Nano ESP32 which is a NORA-W106, a
module with a ESP32-S3. The benefit of this microcontroller is that it
would allow simply the board to be swapped and the rest of the PCB
designs would be unaffected.

Using the Espressif development framework ESP-IDF and follwong the same
logic as the original firmware source, the latency was reduced to around
10ms. ESP-IDF allows access to functionality of the ESP32 that is not
exposed in the Arduino libraries. It is possible that the 10ms latency
could be improved upon but there would be a cost in the form of
develoment time. Consideration would need to be made on how to port
existing code or whether to completely re-write the firmware.

\begin{itemize}
\tightlist
\item
  avoid soldering cabling
\item
  power indicators on pcbs

  \begin{itemize}
  \tightlist
  \item
    debug if problem is power related Add power small power LEDs to the
    each PCB. It was not common for a board to not be functioning and
    thye result to be that the power weas not correctly connected.
    Especially when the ower is connected via somethgin such as an FFC
    cable where it is easy to plug-in the wrong way esepcially when
    fitting is to be carried out by someone unfmailiar wiyth the system.
    Power LEDs would quickly confirm whether the problem was power
    related. They would have to be small enough and low enough current
    draw to avoid glare or being mistaken with other indcator LEDs To
    continue using THT components, the choice would have to be moade
    carefully.
  \end{itemize}
\end{itemize}

\section{Acknowledgments}\label{acknowledgments}

A project of this breadth developed across 3 institutes and 2 countries
cannot be carried without the support and technical knowledge of many
skilled and passionate people.

We acknowledge the invaluable contributions of

\begin{itemize}
\tightlist
\item
  Unibo

  \begin{itemize}
  \tightlist
  \item
    Alma Labor

    \begin{itemize}
    \tightlist
    \item
      Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
      of bologna for their 3D printing expertise and generosity with
      electronics equipment
    \end{itemize}
  \item
    Fisica Tecnica

    \begin{itemize}
    \tightlist
    \item
      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
      Costantini
    \end{itemize}
  \end{itemize}
\item
  San Colombano

  \begin{itemize}
  \tightlist
  \item
    Catalina Vincens
  \item
    Roberto Livi
  \end{itemize}
\item
  Edinburgh

  \begin{itemize}
  \tightlist
  \item
    uCreate

    \begin{itemize}
    \tightlist
    \item
      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
      Edinburgh for their support with soldering facilities.
    \end{itemize}
  \item
    ECA

    \begin{itemize}
    \tightlist
    \item
      Richard Collins and the Digital Making Staff at the Edinburgh
      College of Art for 3D printing instruction and CNC support.
    \item
      Joe Hathaway of Digital Development at the College of Art for
      providing electronics components
    \item
      St Cecilia's Jenny Nex
    \end{itemize}
  \end{itemize}
\end{itemize}

\section{Reference}\label{reference}

\subsection{Example Trasuntino}\label{example-trasuntino}

\begin{itemize}
\tightlist
\item
  {[}1538{]}
  https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
\item
  {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
\end{itemize}

### 2025-01-15 src/article.tex
Updates from Overleaf

@@
 space was added into the design of the internal chamber of the model
-\FIGURE, but there was still a limit given the structural and aesthatic
+\\FIGURE, but there was still a limit given the structural and aesthatic
 requirements.
 
-\subsubsection{Sensor Criteria}\label{sensor-criteria}
-
-A set of criteria was drawn to which the final sensor system would have
-to satisfy. This criteria was used as a means of evaluating the
-feasability throughout the prototyping phase.
-
-\begin{itemize}
-\tightlist
-\item
-  Non-invasive: The sensor system should not require major modifications
-  to the interface on which it is installed.
-\item
-  Low Latency: Reading sensors and then triggering an output should have
-  a low latency. Using {[}paper on latency{]} a target was set as X ms
-  with an upper limit set to 25ms beyond which the sensors would be
-  deemed unusable
-\item
-  Reliable: Data obtained should be reliable and accurate enough not to
-  result in false positives or false negatives without extensive
-  filtering
-\item
-  Scalable: The system should be scalable both economically and
-  temporally. Given the open soucre nature of the porject, it could not
-  be prohibitvely expensive to carry out. Also any system that would
-  work on a single key would have to scale financially to potentially 50
-  or more. The time required to install and calibrate the sensors would
-  also have to scale.
-\item
-  Expandable: Any sensor system should be functionally expandable to
-  allow for the usage of the data for control over other parameters.
-\end{itemize}
-
-\subsubsection{Triggering a note}\label{triggering-a-note}
-
-For the first iteration the data from the sensor would simply be used to
-test when a threshold had been passed. On passing the threshold a
-message would be sent to trigger the playback of a corresponding note.
-
-The message format covered in \textbackslash section
-
-\section{Related Work and Motivation}\label{related-work-and-motivation}
-
-\begin{itemize}
-\tightlist
-\item
-  \cite{HapticKey}
-\item
-  \cite{mcPherson1} \cite{mcPherson2}
-\item
-  \cite{Mudd2013}
-\item
-  \cite{FritzClaudia2017Leon}
-\end{itemize}
-
-\section{Hardware Design}\label{hardware-design}
-
-The hardware was designed with the xx in mind.
-
-Using a similalr apporach to @HapticKey the first step was to fabricate
-or procure a model of a single working key. San Colombano provided a
-model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
-3-key model was used a test-bed for potential sensors while designs were
-drawn for a full-scale model. The cross-sectional nature of the model
-meant it was easy to fit and remove sensors. Small size meant that it
-would also be easy to transfer between workshops.
-
-3-key model also had a sliding jack rail which allowed for easy
-disengagement of the physical mechanical response, which allowed for
-some preliminary tests on the presence and absence of mechanism.
-
-\subsection{Failed sensors}\label{failed-sensors}
-
-A number of sensors were tried before the final system was decided upon.
-An inertial measurment unit (IMU) was mounted to a single key which
-contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
-a consistent set of conditions to trigger sound, but this was not easily
-differntiable when other keys were played. The fitting time and cost of
-a single IMU could b reduced when purchased in bulk but was far greater
-than other systems considered.
-
-A reed sensor and a hall sensor were mounted and magnets embedded within
-the jacks. Adjusting the hieight of the sensor provided a reliable means
-of detecting a threshold, inparticular the hall sensors which could be
-partially calibrarted physically and also in softwrae. The time taken to
-calibarte a single key was time consuming. Ideally each sensor would be
-placed on a thread for adjustment, but given the constraints on the
-internal space it was decided any adjustment of heights would likely
-mean an unnacceptable length of time to install and calibrate. Embedding
-magnets into the jacks was also did not satisfy the criteria of being
-scalable and non-invasive.
-
-Light dependant resistor mounted to the jack rail. as jacks were pressed
-less light would be let through. Usable data but very dependant on
-lighting conditions. Modifying the jack also did not satisfy the
-non-intrusive criteria and the placement outside the internal chamber of
-the model meant hiding electronics would be more difficult.
-
-Finally, force sensitive resistors placed under the jack of each key was
-considered but was never implemented. The sensors were to be placed
-under each jack either above or below the padding on which the bottom of
-the jack rests. Further discussion of this apporach is in
-\textbackslash section.
-
-\section{Electronics}\label{electronics}
-
-Design of electronics carried out in tandem with construction of full
-model. Since the Electronics were designed at a distanced to the final
-full model some flexbility was required in the design. Design decisions
-that were not as optimal but would allow for change in wiring and
-fitting should there be any issues during installation.
-
-Electronics designed so that they would not require any equipment
-outside what can commonly be found in a maker space at a univeristy.
-
-The equipment required including soldering suitable for throughole
-technology (THT) components and 3D printing.
-
-Laser cutting and CNC routing fascilities were used during the research
-for the project but not required for the final implementation.
-
-The idea behind this was to allow for the project to be reasonably
-recreated without procurement of any additional specialist equipment.
-
-NEMUS lab under construction during the first section of the project.
-Facilities available at the maker spaces of Bologna (Alma Labor) and
-Edinburgh (uCreate) used for what was deemed reasonable assumption in
-terms of available equipment.
-
-\begin{itemize}
-\tightlist
-\item
-  Over arching themes
-
-  \begin{itemize}
-  \tightlist
-  \item
-    simplicity of design
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Limited resources
-    \item
-      easily reproducible
-    \end{itemize}
-  \item
-    availability and lead times
-  \item
-    iteration and design from distance
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{QRE1113}\label{qre1113}
-
-Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
-a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
-combination infrared LED and phototransistor sensitive to IR light. The
-phototransistor of QRE1113 provides data on how much light is present
-and in particular how much light is being relfected from a surface.
-Since refelected light will be proportional to the distance of a surfcae
-it is a good, close proximity, distance sensor. Distance is the way in
-which the instruments in \citeMcpherson1\&2 function as well the Moog
-Piano Bar from which it too inspiration.
-
-The limitation for taking a distance approach is that there is not one
-but 2 jacks per key. For each jack to be measured indepeedntly the data
-would need tp com fropm the jacks themselves.
-
-They \emph{could} be used and individual trigger points set in
-calibration, but this creates new problems.
-
-\begin{itemize}
-\tightlist
-\item
-  The triggering would need to be bespoke for every key as fluxuations
-  in the dimensions of every piece mean that the threshold would be
-  different.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Moving a problem into additional time in the calibtraryoin setup
-    phase.
-  \end{itemize}
-\item
-  Remebering the constraint that system isshould noyt fix problem, there
-  would be no easy way to tell if a jack had nbot reseated.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    In playing a harpihchord when the jack does not go to its rest
-    position, not note would play. This fault should be carried over
-    intp the final instrument.
-  \end{itemize}
-\end{itemize}
-
-Another common application of the QRE1113 is within line following
-robots, where an array of QRE1113s ar used in conjunction with black
-tape along which some motorised system can following
-
-Applied to the jacks was a greyscale gradient printed the length full
-travel range of the jack.
-
-For the first iteration this was performed on an inkjet printer and
-double-sided tape. For the final iteration the stickers were printed on
-(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
-µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
-cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
-reduced the manual labour for cutting stickers meaning the process would
-scale from the 3-key to the 49-key model.
-
-An initial sticker design was used that contained only the gradient. It
-was found that the jacks wer sticking in their slots andthe edges on the
-stickers easily peeled around the printing The sticker was redesigned to
-be the full length of the jack which avoided the problem of peeling. The
-larger are also mean the stickers adhered betterto the jack Some
-trimming was still required of each sticker to avoid them adding extraa
-resistence to the jack's movement.
-
-During testing it was found that their was a considerable amount
-cross-talk bwteen adjacent sensors. This meant the reading from key
-would be affected by the movement of the key next to it. The cross-talk
-was assumed to be as a result of light reflecting from other jacks.
-Given that IR is used this was difficult to confirm by sight. A set of
-baffles was designed for the pcb that would wrap round the jack and stop
-light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
-printers using PLA plastic. A dark pigmented PLA was used to avoid IR
-light simply being diffused. After the baffles were fitted the
-cross-talk was entirely eliminated.
-
-\begin{itemize}
-\tightlist
-\item
-  previous use
-\item
-  problem of what to sense
-\item
-  satisfied both criteria and design restrictions
-\item
-  usage
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Sweet spot is around 6mm distance
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{sensor surface}\label{sensor-surface}
-
-\begin{itemize}
-\tightlist
-\item
-  linear gradient
-\item
-  on Jack
-\item
-  printed and cut on vinyl {[}find name{]}
-\item
-  cut to travel size
-\item
-  then to jack size
-\item
-  jacks varnished on side
-\item
-  alternatives
-
-  \begin{itemize}
-  \tightlist
-  \item
-    CNC plywood and mdf
-
-    \begin{itemize}
-    \tightlist
-    \item
-      problems
-    \end{itemize}
-  \item
-    laser cut plywood and mdf
-
-    \begin{itemize}
-    \tightlist
-    \item
-      problems
-    \end{itemize}
-  \item
-    3D printed jack body
-
-    \begin{itemize}
-    \tightlist
-    \item
-      problems
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
-\subsection{Arduino}\label{arduino}
-
-\subsubsection{Nano BLE}\label{nano-ble}
-
-Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
-appealing as it would allow for easy changes between chipsets without
-requiring any rewiring and thus no new PCBs would need to be created if
-a better alternative was found later in the project. Another benefit to
-the Arduino Nano boards was that typically had native-USB meaning the
-could be programmed as a hardware USB MIDI device directly.
-
-The Nano BLE specifically was used in the initial stages for the
-integrated IMU.
-
-When it was finalised that the prokect would proceed using the QRE1113 a
-number of microcontroller units (MCU) were condsidered.
-
-The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
-Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
-
-A rough performance test was later carried out using an STM32
-NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
-more performative. To execute a single sensor read cycle of the firmware
-the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
-on average. The Nano ESP32 had high variability so it was decided to
-continue with the Nano BLE. Further discussion of using ESP32 continues
-in
-
-\section{Going forward}
-
-BLE functionality also meant if the cable connection between the nano
-and computer running audio synthesis was impractical, BLE MIDI was
-available as a fallback. The Nano BLE also has a 12-bit DAC, which
-provided a contingency should the default 10-bit not be sensitive enough
-for the signal from the sensors.
-
-Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
-functionality, but it was found that 2 ADC channels were in fact not
-useable and would have required additional multiplexers.
-
-\begin{itemize}
-\tightlist
-\item
-  Nano form factor small
-\item
-  allows for easy changes in chipset
-\item
-  ``Native USB''
-  https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
-\item
-  BLE Model
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Performative
-  \item
-    BLE capabilities for MIDI BLE transmission
-  \item
-    All analog pins available
-  \item
-    available at the time
-  \item
-    12-bit DAC
-  \end{itemize}
-\end{itemize}
-
-\subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
-
-There would need to be the ability to store values on non-volatile
-memory so that the electroincs could be powered-down without losing
-data.
-
-Adjusting the values through firmware would not have been flexible or
-practical. Instead an Ferroelectric RAM unit was chosen and connected
-using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
-was primarily chosen for it's ready availability.
-
-Most microcontrollers have some variety of non-volatile program memory
-that can be read and writ accessed but in most cases this memory is
-cleared during compilation and re-flashing. The small cost was far
-outweighed by the potential for data loss if using onboard RAM.
-
-\emph{It is worth noting that the FRAM model chosen stopped being
-available during the project, meaning any future iteration would need to
-adjust for another board in firmware}
-
-The firmware was structure in a way that the interface could be changed
-so long as the code required for any other board would comply with
-function arguments and outputs. Worth noting the fragility and
-volatility when designing this kind of project and how easily it is for
-the project to become broken.
-
-SD card would provide an easier way to interact with calibration data.
-The posiiton of the controller board did not permit easy access.
-reliability and speed was preffered over storage capacity.
-
-\begin{itemize}
-\tightlist
-\item
-  On board available, but volatile
-\item
-  retained across firmware changes
-\item
-  small,
-\item
-  could have been sd card or anything else
-\item
-  available
-\item
-  interfaced via rotary encoder
-\end{itemize}
-
-\subsection{RGB LEDs}\label{rgb-leds}
-
-for problems of scale:
-
-Addressing the problem of scale of identifying which key's threshold is
-currently being edited.
-
-The LEDs also allowed for easy identification for keys that were outside
-of expected values and when a key was triggered. This allowed for visual
-identification of malfunctioning or uncalibrated keys.
-
-The LEDs are on the same PCB as the sensors and are hidden when the
-nameboard is in place.
-
-Would sugggest that this is a necessity of this kind to provide a means
-to identify and diagnose LEDs with state.
-
-TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
-to assembling, but were avoid in favour of Addressable LEDs with
-intergrated driver. The intergrated LEDs share the same data line which
-has the limitation where a fualty LED will stop all following LEDs from
-working. An individual PWM driver per PCB would allow avoid situations
-where failutre cascades, but would add complexity in assembly as the
-parts are typically SMD. Also, given the LEDs are fixed to the board and
-could not be repaired in situ, the solution would be to replace the PCB
-regardless of which approach was taken. Given the LEDs were only used in
-debugging and calibration and otehrwise would not be visible, a simpler
-design approach was taken.
-
-\begin{itemize}
-\tightlist
-\item
-  locating keys
-\item
-  displaying key state
-\end{itemize}
-
-\subsection{Multiplexing}\label{multiplexing}
-
-CD4051BE Texas Instruments
-
-problem of scale
-
-For the 3-key model, individual jacks could be connected to one of the 8
-available ADC pins.
-
-The BLE Nano has 8 ADC pins connected to one ADC channel via a
-multiplexer
-
-Multiplexer added to expand to the
-
-Transistor / Op-amp circuit could have been used similar to the one in
-\cite{mcpherson}.
-
-Benefits were keeping with THT based designed for easy assembly.
-
-An 8-channel multiplexer was chosen as the model has 49-keys and a
-division of 7 PCBs each containing 7 sensors
-
-The multiplexer reduced 7 signals to a single pin. Three pins are
-required pins for addressing the muliplexewrs and all PCBs are addressed
-simulatenously.
-
-\begin{itemize}
-\tightlist
-\item
-  limited ADC channels
-\item
-  ADC channels already multiplexed
-\item
-  simple
-\item
-  8 channel
-\item
-  THT
-\item
-  7 into 49
-
-  \begin{itemize}
-  \tightlist
-  \item
-    7 PCBS
-  \item
-    each with 7 sensors
-  \end{itemize}
-\end{itemize}
-
-\subsection{PCB design}\label{pcb-design}
-
-\begin{itemize}
-\tightlist
-\item
-  EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
-  deprecated EAGLE and the programme will no longer be supported from
-  2026 Project files will need to be transitioned to another standard,
-  likely the KiCAD programme.
-\end{itemize}
-
-trade off between limiting number of boards required and allowing
-compensation for variation in pitch.
-
-Smaller PCBs means that spacing could be adjusted to accomodate the
-changes in pitch between jacks. It also allowed for a modular approach
-where any PCB could swapped for minimal cost.
-
-\subsubsection{Sensor board}\label{sensor-board}
-
-The 49 QRE1113s are divided across 7 PCBs each containing
-
-\begin{itemize}
-\item
-  x7 QRE1113
-\item
-  x7 100R
-\item
-  x7 10kR
-\item
-  x1 CD4051BE Multipexer
-\item
-  jack / sensor pitch by board
-\item
-  7 boards adjust for pitch/spacing error
-\item
-  modular, replaceable
-\item
-  orientation problems
-
-  \begin{itemize}
-  \tightlist
-  \item
-    legs break easily
-  \item
-    end mount like A.P.Mc. time consuming
-  \item
-    problem
-
-    \begin{itemize}
-    \tightlist
-    \item
-      milling out keys
-    \end{itemize}
-  \end{itemize}
-\item
-  designed from distance
-\item
-  striped lines across pcbs
-\item
-  problems
-
-  \begin{itemize}
-  \tightlist
-  \item
-    additional terminals uneeded resulted in lots of additional work
-  \item
-    long cabling increase risk of breaking
-  \end{itemize}
-\item
-  solutions
-
-  \begin{itemize}
-  \tightlist
-  \item
-    striping all adc lines with FFC
-  \item
-    solder pads to select channel
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{controller board}\label{controller-board}
-
-Controller board designed around Arduino nano form factor. The Nano is
-connected through headers to allow easy switching between chips. Should
-another chip in a Nano for factor be found it could supplant the current
-Nano without any need for a redesign.
-
-The first version of the board esclusively used 2.54mm pitch headers.
-The reasons for this were flexibility when creating cable looms and to
-leave open the possiility of rewuring the PCB.
-
-This was the case for the LEDs as they initially wer to be powered on
-5V. It was discovered that the 3v regulator of the nano was sufficient
-to power all LEDs at a lower brightness while reducing current draw.
-Since brightness was not importnat.
-
-The 2.54mm pitch headers had two major drawbacks. First, the orientation
-of the cables could be reversed, which in the best case scenario would
-simply mean the system did not work correctly and in the worst case it
-could potentially damaged some components.
-
-Secondly, the headers do not connect securely and wait of the cabling
-was enough for components to become disconnected as the harpsichord was
-handled.
-
-To address this, once the wiring for the controller board was finalised
-a new design was drafted this time using JST-PH connectors. No
-commercial cables were immediately available in the configuration and
-length required, as such a cable loom needed to be made which meant
-crimping cables by hand. A crimping tool from JST is prohibitively
-expensive compared to the cost of other components in the project*. The
-crimping tool also appeared to be uncommon and was not readily available
-in the workshops at the projects disposal. Eventually a tool was found,
-but an alternative approach was also researched in order to maintain the
-criteria of accessibility.
-
-A locator was adapted from a creative commons stl model
-https://www.thingiverse.com/thing:1646016 that allowed a generic
-crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
-locator was printed using a Bambu X1 3D printer.
-
-\emph{** A second alterntaive was assembling cables from other sets that
-were at hand.} \emph{* from RS
-https://uk.rs-online.com/web/p/crimp-tools/6880877 £508 inc VAT} \emph{†
-model available in the git repository}
-
-\begin{itemize}
-\tightlist
-\item
-  mount for nano to allow switching
-\item
-  initially push fit 2.54mm header
-
-  \begin{itemize}
-  \tightlist
-  \item
-    not a secure fit
-  \end{itemize}
-\item
-  JST-PH used
-
-  \begin{itemize}
-  \tightlist
-  \item
-    crimping required custom cables
-
-    \begin{itemize}
-    \tightlist
-    \item
-      could be assembled from other cables
-    \item
-      length limited
-    \end{itemize}
-  \item
-    prohibitevly expensive to crimpers
-  \item
-    3D print helper to allow for crimping with generic crimpers
-  \end{itemize}
-\end{itemize}
-
-\subsection{3D printing}\label{d-printing}
-
-Access to a 3D printer is was vital to the success of the project. Given
-the proliferation of 3D printing, particular in maker spaces in
-universities it was deemed an unreasonable requirement. 3D printed parts
-include the support brackets for PCBs, the baffles for the sensors,
-locator for crimping JST-PH crimps and a mount for the MacMini used for
-audio synthesis.
-
-The printer used throughout the project was a Bambu X1-Carbon using
-recycled PLA filament.
-
-Support brackets and baffles were made for the project while the MacMini
-bracket and JST locator were sourced from creative commons models.
-
-Any colour of filament can be used, though it is advised to use a dark
-coloured filament for the baffles.
-
-See each corresponding section for more detail on 3D printed components.
-
-\begin{itemize}
-\item
-  https://www.thingiverse.com/thing:1646016
-\item
-  https://www.thingiverse.com/thing:2886460
-\item
-  Supports
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Push fit
-  \item
-    adjustable for best average placement
-  \end{itemize}
-\item
-  Blinkers
-
-  \begin{itemize}
-  \tightlist
-  \item
-    reduce noise from other senors
-  \end{itemize}
-\item
-  mounts
-
-  \begin{itemize}
-  \tightlist
-  \item
-    MacMini
-  \end{itemize}
-\end{itemize}
-
-\subsection{Power}\label{power}
-
-All electronics draws around 1.1amps of current at 5V with some
-fluctuation during boot.
-
-Arduino BLE has some functionality disabled to limit current draw.
-
-Ideally the unit could be bus powered from USB.
-
-The sensors are in a state of constantly drawing power. It is possible
-that powering only when data is required could be reduce power enough.
-
-Current designs have a separate MIDI device for each jack row. Combining
-the devices and also reducing current draw to below the common 0.5 amps
-for a USB port would certainly be a challenge.
-
-\section{Firmware Design}\label{firmware-design}
-
-Arduino platform chosen for wide adoption, relative ease of use and
-library support for components.
-
-Firmare checks components are connected then reads the FRAM to see if
-threshold data is already present. The rotary encoder and it's momentary
-switch are polled for any change in state. A key is selected with the
-rotary encoder while the momentary switch is used for changing between
-states for key selection and threshold adjustment A double click of the
-rotary encoder writes the current threshold values to FRAM. The sensors
-for the current multiplexer channels are polled. If any sensor reads
-above the threshold when it was previously below a MIDI On message is
-sent. For the reverse case where the value was below the threshold when
-it was previously above, a MIDI Note Off message ois sent.
-
-\begin{itemize}
-\tightlist
-\item
-  Arduino platform
-
-  \begin{itemize}
-  \tightlist
-  \item
-    library support and portability
-  \item
-    dependancy management
-  \item
-    wide spread use
-  \end{itemize}
-\item
-  core algorithm
-
-  \begin{itemize}
-  \tightlist
-  \item
-    check components connected
-  \item
-    READ data from FRAM
-  \item
-    select key with rotary
-
-    \begin{itemize}
-    \tightlist
-    \item
-      click to edit, double to save
-    \end{itemize}
-  \item
-    current threshold and reading sent over serial and plotted
-  \item
-    when over threshold send MIDI note
-  \end{itemize}
-\end{itemize}
-
-\subsection{Project structure}\label{project-structure}
-
-The project is separated into three repositories encapsulating
-
-\begin{itemize}
-\tightlist
-\item
-  Firmware
-\item
-  Keyboard CAD Data
-\item
-  PCB CAD Data
-\end{itemize}
-
-\subsubsection{Firmware}\label{firmware}
-
-Firmware repository contains all firmware for the full and smaller
-models and components.
-
-\subsubsection{Keyboard}\label{keyboard-1}
-
-The Keyboard repository contains all plans for fabricating the keyboard,
-measurements and 3D models of the jacks.
-
-\subsubsection{PCB}\label{pcb}
-
-PCB repository contains the EAGLE files for creating the sensor and
-controller boards.
-
-Given the regular spacing of components, scripts for automatic placement
-of components based on parameter of jack pitch spacing and number of
-sensors per jack.
-
-\section{Implementation}\label{implementation}
-
-\subsection{Installation}\label{installation}
-
-The electronic components were partly pre-assembled at the university of
-Edinburgh. Though all measurments were known ahead, enough flexibiliyty
-was needed to allow for proboems to solved in situ.
-
-The final assembly took place at the NEMUS lab at the univerity of
-bologna over the final week in October 2024 for delivery at San
-Colombano for exhibition in the first week of November in 2024.
-
-A misreading in the original clearance between the keys and the PCB
-meant that the keys wer block entirely from moving. A 10mm by 15mm
-section was milled out of each lke \FIGURE to create extra clearance.
-The milling required the removal, shortening and refitting of the
-leather pads for the jacks \FIGURE. The extra work added an extra day to
-the full install.
-
-A redesign of the PCBs meant that removing the material would no longer
-be necessary.
-
-The PCBs shared power and data lines for components with the exception
-of the output from the multiplexer. Ribbon cable was cut to size to
-allow for some room to readjust the distance of the sensors from the
-jacks. The data signals from each PCB were connected using a separate
-cable loom. Soldering cables and inspecting each join required another
-day work. This did allow for readjustments to the positon of the
-controller board Measurements of the rear chamber did not match with the
-MacMini that was to be used a different layout was decided for fitting
-the controller board/ Cutting cables to size allowed for a great deal of
-felxibility but at the cost of time. After the first install it was
-clearer what constraints were in place for fittig and an alternative
-wiring system using flat flex cable was implement (discussed in
-\textbackslash section)
-
-Fitting PCBs required the removal of all keys. The support brackets were
-assembled and screwed into the the underside. With the harpsichord on
-it's back the sensors were tested to ensure an expected range.
-Self-tapping round head screws were used to to attach the brackets. The
-screws were not over tightened to allow for brackets to move. The
-brackets were not easily adjusted once the key were replaced. Adjusted
-to a best average with subsequent adjustment carried out on the
-microcontroller.
-
-As the PCBs were fitted the jacks had the gradient stickers applied to
-them. The jacks had been coated with a varnish on the short side to
-provide a better surface for the stikcers to adhere to. After the
-stickers were applied excess material was trimmed from each jack. The
-jack were then check that they moved freely in their slot.
-
-After fitting the PCBs preliminary calibration was carried out. Before
-the keys were reffitted the jacks wer put back in place. The data from
-each key was confirmed to be in a similar range. Any key showing
-anomalous dat had it's jack removed and re-inspected. The PCB fitting
-process and refitting required another day as faults were found in
-wiring or misalignement of the stickers. After the sensors were
-confirmed to be functioning the keys were put back.
-
-The calibration workflow was with Dr.~Craig Webb who was unfamiliar with
-the system. Adjustments were made to the firmware to to make it more
-intuitive for the user. Addressable LEDs became vital as it allowed for
-easy identification of the key being calibrated. They were also used to
-identify any keys whose current reading was beyond what was expected.
-
-The rotary encoder was used as the interface to select a key, adjust
-it's threshold and update the Ferroelectric RAM (FRAM) with new values.
-
-A process was implemented where the key was selected by playing, but
-this caused problems in situations where the data was unusual due to
-misalignement of the sensors
-
-This would likely be a better system for when the keyboard was fully
-operational
-
-The thresholds were calibrated visually utilising the serial plotter of
-the Arduino IDE and MIDI was confirmed using the open source MIDI
-Monitor (https://github.com/krevis/MIDIApps).
-
-\begin{itemize}
-\tightlist
-\item
-  partial pre-assembly
-\item
-  handsolder components
-
-  \begin{itemize}
-  \tightlist
-  \item
-    analog lines
-  \end{itemize}
-\item
-  connecting supports
-\end{itemize}
-
-\subsection{Context}\label{context}
-
-Designed as part of an exhibition for the museum San Colombano in
-Bologna. Instruments is to be played by general public as a means of
-interacting with items in the collection that are no longer in playing
-condition. Audience of one who listens via headphones.
-
-The strings of the interface are damped with felt strips. The damping
-did change the tension and thereforew the feel of the strings so
-adjustments needed to be made to the tuning to accomodate for this.
-
-\begin{itemize}
-\tightlist
-\item
-  in museum context
-\item
-  strings damped
-\item
-  headphones
-\item
-  interfacing with kontakt spitfire library
-\item
-  display
-
-  \begin{itemize}
-  \tightlist
-  \item
-    secure storage and access
-  \item
-    interface to the public
-  \end{itemize}
-\end{itemize}
-
-\subsection{Calibration}\label{calibration}
-
-Finer calibration carried out with the guidance of Catalina Vincens as
-an expert harpsichord performer
-
-\begin{itemize}
-\tightlist
-\item
-  sensor surface
-
-  \begin{itemize}
-  \tightlist
-  \item
-    printed gradient
-  \end{itemize}
-\end{itemize}
-
-\section{Conclusion}\label{conclusion}
-
-\begin{itemize}
-\tightlist
-\item
-  scale
-\item
-  time scale
-\item
-  working at distance
-\item
-  limitations
-
-  \begin{itemize}
-  \tightlist
-  \item
-    latency
-  \item
-    deprecation
-
-    \begin{itemize}
-    \tightlist
-    \item
-      software
-
-      \begin{itemize}
-      \tightlist
-      \item
-        EAGLE
-      \end{itemize}
-    \item
-      hardware
-
-      \begin{itemize}
-      \tightlist
-      \item
-        FRAM
-      \end{itemize}
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
-\section{Going forward}\label{going-forward}
-
-\begin{itemize}
-\tightlist
-\item
-  Velocity, aftertouch
-\item
-  second jack row
-\item
-  reducing latency
-\item
-  alternative sensor surfaces
-\end{itemize}
-
-A limitation in the design is the variablility in positioning of tage,
-especially when they need to be reapplied. A jig was designed for
-applying tags, but could not be printed in time for installation. Even
-with a jig, some tags had to be radjusted by hand regardless. Long term
-this hasd impliccation on how to maintain the interface.
-
-Taken a different approach to what surface is used. Preliminary test
-were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
-cutting results in a loss of material that has too much ariablity. The
-edge o f the jack body will be charrred. The material also needs to be
-reliable. on the advice of two luthiers, roberto livi and Jonathan Santa
-Maria Bouquet at Edinburgh, Plywood and MDF have problems with
-variations in humidity, tear-out and delamination.
-
-Some tests were carried out using a notched jack body that was 3D
-printed. The notch means that as the jack raises, the forward material
-acts like a shutter. No reasons that this could not be CNC though a
-variety of materials would be needed. Traditional material used is xx
-though need to test of the material is naturally reflective enough or if
-a coating would beed to be applied.
-
-Outputs of the NEMUS project will also include numerically simulated
-non-linear plucked string models with which the device can interact. A
-bespoke sound synthesis engine to leverage data from both jack rows
-which will also be open-sourced. Design for instrument to be compatible
-with standard MIDI instruments as well software designed for the full
-data output.
-
-A second interface designed more closely to the Trasuntino style
-instrument commisioned by the NEMUS project with Roberto Livi. Using
-lessons learned for the first iteration the second can make better
-acommodations for internal electronics. Interface to be used for a
-performer study and used for performance with a numerically simulated
-sound syntheiss engine.
-
-Testing current sensors in conjunction with force sensitive resistors to
-research if a impulse response for the pluck can be extrapolated.
-
-Future novel designs for an interface that would utilise velocity and
-aftertouch messages in software. This would be applicable for numericla
-simulations of strings. A bespoke piec e of sowftare would also allow
-for direct mapping of jack rows to registers. There is enough data to
-implemenet functionality for velocity and aftertouch MIDI messages and
-simply needs to be activated in the firmware.
-
-Evaluation
-
-\subsection{improvements}\label{improvements}
-
-\begin{itemize}
-\tightlist
-\item
-  ESP-IDF
-
-  \begin{itemize}
-  \tightlist
-  \item
-    latency half
-  \item
-    workload high
-  \end{itemize}
-\end{itemize}
-
-A current limitation is the latency in triggering MIDI output. Currently
-the latency is around 20 ms, but should ideally be closer to 5ms.
-
-An obvious approach would be merely changing the microcontroller. A test
-was carried out using an Arduino Nano ESP32 which is a NORA-W106, a
-module with a ESP32-S3. The benefit of this microcontroller is that it
-would allow simply the board to be swapped and the rest of the PCB
-designs would be unaffected.
-
-Using the Espressif development framework ESP-IDF and follwong the same
-logic as the original firmware source, the latency was reduced to around
-10ms. ESP-IDF allows access to functionality of the ESP32 that is not
-exposed in the Arduino libraries. It is possible that the 10ms latency
-could be improved upon but there would be a cost in the form of
-develoment time. Consideration would need to be made on how to port
-existing code or whether to completely re-write the firmware.
-
-\begin{itemize}
-\tightlist
-\item
-  avoid soldering cabling
-\item
-  power indicators on pcbs
-
-  \begin{itemize}
-  \tightlist
-  \item
-    debug if problem is power related Add power small power LEDs to the
-    each PCB. It was not common for a board to not be functioning and
-    thye result to be that the power weas not correctly connected.
-    Especially when the ower is connected via somethgin such as an FFC
-    cable where it is easy to plug-in the wrong way esepcially when
-    fitting is to be carried out by someone unfmailiar wiyth the system.
-    Power LEDs would quickly confirm whether the problem was power
-    related. They would have to be small enough and low enough current
-    draw to avoid glare or being mistaken with other indcator LEDs To
-    continue using THT components, the choice would have to be moade
-    carefully.
-  \end{itemize}
-\end{itemize}
-
-\section{Acknowledgments}\label{acknowledgments}
-
-A project of this breadth developed across 3 institutes and 2 countries
-cannot be carried without the support and technical knowledge of many
-skilled and passionate people.
-
-We acknowledge the invaluable contributions of
-
-\begin{itemize}
-\tightlist
-\item
-  Unibo
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Alma Labor
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
-      of bologna for their 3D printing expertise and generosity with
-      electronics equipment
-    \end{itemize}
-  \item
-    Fisica Tecnica
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
-      Costantini
-    \end{itemize}
-  \end{itemize}
-\item
-  San Colombano
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Catalina Vincens
-  \item
-    Roberto Livi
-  \end{itemize}
-\item
-  Edinburgh
-
-  \begin{itemize}
-  \tightlist
-  \item
-    uCreate
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
-      Edinburgh for their support with soldering facilities.
-    \end{itemize}
-  \item
-    ECA
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Richard Collins and the Digital Making Staff at the Edinburgh
-      College of Art for 3D printing instruction and CNC support.
-    \item
-      Joe Hathaway of Digital Development at the College of Art for
-      providing electronics components
-    \item
-      St Cecilia's Jenny Nex
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
-\section{Reference}\label{reference}
-
-\subsection{Example Trasuntino}\label{example-trasuntino}
-
-\begin{itemize}
-\tightlist
-\item
-  {[}1538{]}
-  https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
-\item
-  {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
-\end{itemize}
+% \subsubsection{Sensor Criteria}\label{sensor-criteria}
+
+% A set of criteria was drawn to which the final sensor system would have
+% to satisfy. This criteria was used as a means of evaluating the
+% feasability throughout the prototyping phase.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Non-invasive: The sensor system should not require major modifications
+%   to the interface on which it is installed.
+% \item
+%   Low Latency: Reading sensors and then triggering an output should have
+%   a low latency. Using {[}paper on latency{]} a target was set as X ms
+%   with an upper limit set to 25ms beyond which the sensors would be
+%   deemed unusable
+% \item
+%   Reliable: Data obtained should be reliable and accurate enough not to
+%   result in false positives or false negatives without extensive
+%   filtering
+% \item
+%   Scalable: The system should be scalable both economically and
+%   temporally. Given the open soucre nature of the porject, it could not
+%   be prohibitvely expensive to carry out. Also any system that would
+%   work on a single key would have to scale financially to potentially 50
+%   or more. The time required to install and calibrate the sensors would
+%   also have to scale.
+% \item
+%   Expandable: Any sensor system should be functionally expandable to
+%   allow for the usage of the data for control over other parameters.
+% \end{itemize}
+
+% \subsubsection{Triggering a note}\label{triggering-a-note}
+
+% For the first iteration the data from the sensor would simply be used to
+% test when a threshold had been passed. On passing the threshold a
+% message would be sent to trigger the playback of a corresponding note.
+
+% The message format covered in \textbackslash section
+
+% \section{Related Work and Motivation}\label{related-work-and-motivation}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   \cite{HapticKey}
+% \item
+%   \cite{mcPherson1} \cite{mcPherson2}
+% \item
+%   \cite{Mudd2013}
+% \item
+%   \cite{FritzClaudia2017Leon}
+% \end{itemize}
+
+% \section{Hardware Design}\label{hardware-design}
+
+% The hardware was designed with the xx in mind.
+
+% Using a similalr apporach to @HapticKey the first step was to fabricate
+% or procure a model of a single working key. San Colombano provided a
+% model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
+% 3-key model was used a test-bed for potential sensors while designs were
+% drawn for a full-scale model. The cross-sectional nature of the model
+% meant it was easy to fit and remove sensors. Small size meant that it
+% would also be easy to transfer between workshops.
+
+% 3-key model also had a sliding jack rail which allowed for easy
+% disengagement of the physical mechanical response, which allowed for
+% some preliminary tests on the presence and absence of mechanism.
+
+% \subsection{Failed sensors}\label{failed-sensors}
+
+% A number of sensors were tried before the final system was decided upon.
+% An inertial measurment unit (IMU) was mounted to a single key which
+% contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
+% a consistent set of conditions to trigger sound, but this was not easily
+% differntiable when other keys were played. The fitting time and cost of
+% a single IMU could b reduced when purchased in bulk but was far greater
+% than other systems considered.
+
+% A reed sensor and a hall sensor were mounted and magnets embedded within
+% the jacks. Adjusting the hieight of the sensor provided a reliable means
+% of detecting a threshold, inparticular the hall sensors which could be
+% partially calibrarted physically and also in softwrae. The time taken to
+% calibarte a single key was time consuming. Ideally each sensor would be
+% placed on a thread for adjustment, but given the constraints on the
+% internal space it was decided any adjustment of heights would likely
+% mean an unnacceptable length of time to install and calibrate. Embedding
+% magnets into the jacks was also did not satisfy the criteria of being
+% scalable and non-invasive.
+
+% Light dependant resistor mounted to the jack rail. as jacks were pressed
+% less light would be let through. Usable data but very dependant on
+% lighting conditions. Modifying the jack also did not satisfy the
+% non-intrusive criteria and the placement outside the internal chamber of
+% the model meant hiding electronics would be more difficult.
+
+% Finally, force sensitive resistors placed under the jack of each key was
+% considered but was never implemented. The sensors were to be placed
+% under each jack either above or below the padding on which the bottom of
+% the jack rests. Further discussion of this apporach is in
+% \textbackslash section.
+
+% \section{Electronics}\label{electronics}
+
+% Design of electronics carried out in tandem with construction of full
+% model. Since the Electronics were designed at a distanced to the final
+% full model some flexbility was required in the design. Design decisions
+% that were not as optimal but would allow for change in wiring and
+% fitting should there be any issues during installation.
+
+% Electronics designed so that they would not require any equipment
+% outside what can commonly be found in a maker space at a univeristy.
+
+% The equipment required including soldering suitable for throughole
+% technology (THT) components and 3D printing.
+
+% Laser cutting and CNC routing fascilities were used during the research
+% for the project but not required for the final implementation.
+
+% The idea behind this was to allow for the project to be reasonably
+% recreated without procurement of any additional specialist equipment.
+
+% NEMUS lab under construction during the first section of the project.
+% Facilities available at the maker spaces of Bologna (Alma Labor) and
+% Edinburgh (uCreate) used for what was deemed reasonable assumption in
+% terms of available equipment.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Over arching themes
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     simplicity of design
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       Limited resources
+%     \item
+%       easily reproducible
+%     \end{itemize}
+%   \item
+%     availability and lead times
+%   \item
+%     iteration and design from distance
+%   \end{itemize}
+% \end{itemize}
+
+% \subsubsection{QRE1113}\label{qre1113}
+
+% Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
+% a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
+% combination infrared LED and phototransistor sensitive to IR light. The
+% phototransistor of QRE1113 provides data on how much light is present
+% and in particular how much light is being relfected from a surface.
+% Since refelected light will be proportional to the distance of a surfcae
+% it is a good, close proximity, distance sensor. Distance is the way in
+% which the instruments in \citeMcpherson1\&2 function as well the Moog
+% Piano Bar from which it too inspiration.
+
+% The limitation for taking a distance approach is that there is not one
+% but 2 jacks per key. For each jack to be measured indepeedntly the data
+% would need tp com fropm the jacks themselves.
+
+% They \emph{could} be used and individual trigger points set in
+% calibration, but this creates new problems.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   The triggering would need to be bespoke for every key as fluxuations
+%   in the dimensions of every piece mean that the threshold would be
+%   different.
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Moving a problem into additional time in the calibtraryoin setup
+%     phase.
+%   \end{itemize}
+% \item
+%   Remebering the constraint that system isshould noyt fix problem, there
+%   would be no easy way to tell if a jack had nbot reseated.
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     In playing a harpihchord when the jack does not go to its rest
+%     position, not note would play. This fault should be carried over
+%     intp the final instrument.
+%   \end{itemize}
+% \end{itemize}
+
+% Another common application of the QRE1113 is within line following
+% robots, where an array of QRE1113s ar used in conjunction with black
+% tape along which some motorised system can following
+
+% Applied to the jacks was a greyscale gradient printed the length full
+% travel range of the jack.
+
+% For the first iteration this was performed on an inkjet printer and
+% double-sided tape. For the final iteration the stickers were printed on
+% (Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
+% µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
+% cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
+% reduced the manual labour for cutting stickers meaning the process would
+% scale from the 3-key to the 49-key model.
+
+% An initial sticker design was used that contained only the gradient. It
+% was found that the jacks wer sticking in their slots andthe edges on the
+% stickers easily peeled around the printing The sticker was redesigned to
+% be the full length of the jack which avoided the problem of peeling. The
+% larger are also mean the stickers adhered betterto the jack Some
+% trimming was still required of each sticker to avoid them adding extraa
+% resistence to the jack's movement.
+
+% During testing it was found that their was a considerable amount
+% cross-talk bwteen adjacent sensors. This meant the reading from key
+% would be affected by the movement of the key next to it. The cross-talk
+% was assumed to be as a result of light reflecting from other jacks.
+% Given that IR is used this was difficult to confirm by sight. A set of
+% baffles was designed for the pcb that would wrap round the jack and stop
+% light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
+% printers using PLA plastic. A dark pigmented PLA was used to avoid IR
+% light simply being diffused. After the baffles were fitted the
+% cross-talk was entirely eliminated.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   previous use
+% \item
+%   problem of what to sense
+% \item
+%   satisfied both criteria and design restrictions
+% \item
+%   usage
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Sweet spot is around 6mm distance
+%   \end{itemize}
+% \end{itemize}
+
+% \subsubsection{sensor surface}\label{sensor-surface}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   linear gradient
+% \item
+%   on Jack
+% \item
+%   printed and cut on vinyl {[}find name{]}
+% \item
+%   cut to travel size
+% \item
+%   then to jack size
+% \item
+%   jacks varnished on side
+% \item
+%   alternatives
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     CNC plywood and mdf
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       problems
+%     \end{itemize}
+%   \item
+%     laser cut plywood and mdf
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       problems
+%     \end{itemize}
+%   \item
+%     3D printed jack body
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       problems
+%     \end{itemize}
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{Arduino}\label{arduino}
+
+% \subsubsection{Nano BLE}\label{nano-ble}
+
+% Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
+% appealing as it would allow for easy changes between chipsets without
+% requiring any rewiring and thus no new PCBs would need to be created if
+% a better alternative was found later in the project. Another benefit to
+% the Arduino Nano boards was that typically had native-USB meaning the
+% could be programmed as a hardware USB MIDI device directly.
+
+% The Nano BLE specifically was used in the initial stages for the
+% integrated IMU.
+
+% When it was finalised that the prokect would proceed using the QRE1113 a
+% number of microcontroller units (MCU) were condsidered.
+
+% The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
+% Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
+
+% A rough performance test was later carried out using an STM32
+% NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
+% more performative. To execute a single sensor read cycle of the firmware
+% the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
+% on average. The Nano ESP32 had high variability so it was decided to
+% continue with the Nano BLE. Further discussion of using ESP32 continues
+% in
+
+% \section{Going forward}
+
+% BLE functionality also meant if the cable connection between the nano
+% and computer running audio synthesis was impractical, BLE MIDI was
+% available as a fallback. The Nano BLE also has a 12-bit DAC, which
+% provided a contingency should the default 10-bit not be sensitive enough
+% for the signal from the sensors.
+
+% Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
+% functionality, but it was found that 2 ADC channels were in fact not
+% useable and would have required additional multiplexers.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Nano form factor small
+% \item
+%   allows for easy changes in chipset
+% \item
+%   ``Native USB''
+%   https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
+% \item
+%   BLE Model
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Performative
+%   \item
+%     BLE capabilities for MIDI BLE transmission
+%   \item
+%     All analog pins available
+%   \item
+%     available at the time
+%   \item
+%     12-bit DAC
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
+
+% There would need to be the ability to store values on non-volatile
+% memory so that the electroincs could be powered-down without losing
+% data.
+
+% Adjusting the values through firmware would not have been flexible or
+% practical. Instead an Ferroelectric RAM unit was chosen and connected
+% using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
+% was primarily chosen for it's ready availability.
+
+% Most microcontrollers have some variety of non-volatile program memory
+% that can be read and writ accessed but in most cases this memory is
+% cleared during compilation and re-flashing. The small cost was far
+% outweighed by the potential for data loss if using onboard RAM.
+
+% \emph{It is worth noting that the FRAM model chosen stopped being
+% available during the project, meaning any future iteration would need to
+% adjust for another board in firmware}
+
+% The firmware was structure in a way that the interface could be changed
+% so long as the code required for any other board would comply with
+% function arguments and outputs. Worth noting the fragility and
+% volatility when designing this kind of project and how easily it is for
+% the project to become broken.
+
+% SD card would provide an easier way to interact with calibration data.
+% The posiiton of the controller board did not permit easy access.
+% reliability and speed was preffered over storage capacity.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   On board available, but volatile
+% \item
+%   retained across firmware changes
+% \item
+%   small,
+% \item
+%   could have been sd card or anything else
+% \item
+%   available
+% \item
+%   interfaced via rotary encoder
+% \end{itemize}
+
+% \subsection{RGB LEDs}\label{rgb-leds}
+
+% for problems of scale:
+
+% Addressing the problem of scale of identifying which key's threshold is
+% currently being edited.
+
+% The LEDs also allowed for easy identification for keys that were outside
+% of expected values and when a key was triggered. This allowed for visual
+% identification of malfunctioning or uncalibrated keys.
+
+% The LEDs are on the same PCB as the sensors and are hidden when the
+% nameboard is in place.
+
+% Would sugggest that this is a necessity of this kind to provide a means
+% to identify and diagnose LEDs with state.
+
+% TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
+% to assembling, but were avoid in favour of Addressable LEDs with
+% intergrated driver. The intergrated LEDs share the same data line which
+% has the limitation where a fualty LED will stop all following LEDs from
+% working. An individual PWM driver per PCB would allow avoid situations
+% where failutre cascades, but would add complexity in assembly as the
+% parts are typically SMD. Also, given the LEDs are fixed to the board and
+% could not be repaired in situ, the solution would be to replace the PCB
+% regardless of which approach was taken. Given the LEDs were only used in
+% debugging and calibration and otehrwise would not be visible, a simpler
+% design approach was taken.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   locating keys
+% \item
+%   displaying key state
+% \end{itemize}
+
+% \subsection{Multiplexing}\label{multiplexing}
+
+% CD4051BE Texas Instruments
+
+% problem of scale
+
+% For the 3-key model, individual jacks could be connected to one of the 8
+% available ADC pins.
+
+% The BLE Nano has 8 ADC pins connected to one ADC channel via a
+% multiplexer
+
+% Multiplexer added to expand to the
+
+% Transistor / Op-amp circuit could have been used similar to the one in
+% \cite{mcpherson}.
+
+% Benefits were keeping with THT based designed for easy assembly.
+
+% An 8-channel multiplexer was chosen as the model has 49-keys and a
+% division of 7 PCBs each containing 7 sensors
+
+% The multiplexer reduced 7 signals to a single pin. Three pins are
+% required pins for addressing the muliplexewrs and all PCBs are addressed
+% simulatenously.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   limited ADC channels
+% \item
+%   ADC channels already multiplexed
+% \item
+%   simple
+% \item
+%   8 channel
+% \item
+%   THT
+% \item
+%   7 into 49
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     7 PCBS
+%   \item
+%     each with 7 sensors
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{PCB design}\label{pcb-design}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
+%   deprecated EAGLE and the programme will no longer be supported from
+%   2026 Project files will need to be transitioned to another standard,
+%   likely the KiCAD programme.
+% \end{itemize}
+
+% trade off between limiting number of boards required and allowing
+% compensation for variation in pitch.
+
+% Smaller PCBs means that spacing could be adjusted to accomodate the
+% changes in pitch between jacks. It also allowed for a modular approach
+% where any PCB could swapped for minimal cost.
+
+% \subsubsection{Sensor board}\label{sensor-board}
+
+% The 49 QRE1113s are divided across 7 PCBs each containing
+
+% \begin{itemize}
+% \item
+%   x7 QRE1113
+% \item
+%   x7 100R
+% \item
+%   x7 10kR
+% \item
+%   x1 CD4051BE Multipexer
+% \item
+%   jack / sensor pitch by board
+% \item
+%   7 boards adjust for pitch/spacing error
+% \item
+%   modular, replaceable
+% \item
+%   orientation problems
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     legs break easily
+%   \item
+%     end mount like A.P.Mc. time consuming
+%   \item
+%     problem
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       milling out keys
+%     \end{itemize}
+%   \end{itemize}
+% \item
+%   designed from distance
+% \item
+%   striped lines across pcbs
+% \item
+%   problems
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     additional terminals uneeded resulted in lots of additional work
+%   \item
+%     long cabling increase risk of breaking
+%   \end{itemize}
+% \item
+%   solutions
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     striping all adc lines with FFC
+%   \item
+%     solder pads to select channel
+%   \end{itemize}
+% \end{itemize}
+
+% \subsubsection{controller board}\label{controller-board}
+
+% Controller board designed around Arduino nano form factor. The Nano is
+% connected through headers to allow easy switching between chips. Should
+% another chip in a Nano for factor be found it could supplant the current
+% Nano without any need for a redesign.
+
+% The first version of the board esclusively used 2.54mm pitch headers.
+% The reasons for this were flexibility when creating cable looms and to
+% leave open the possiility of rewuring the PCB.
+
+% This was the case for the LEDs as they initially wer to be powered on
+% 5V. It was discovered that the 3v regulator of the nano was sufficient
+% to power all LEDs at a lower brightness while reducing current draw.
+% Since brightness was not importnat.
+
+% The 2.54mm pitch headers had two major drawbacks. First, the orientation
+% of the cables could be reversed, which in the best case scenario would
+% simply mean the system did not work correctly and in the worst case it
+% could potentially damaged some components.
+
+% Secondly, the headers do not connect securely and wait of the cabling
+% was enough for components to become disconnected as the harpsichord was
+% handled.
+
+% To address this, once the wiring for the controller board was finalised
+% a new design was drafted this time using JST-PH connectors. No
+% commercial cables were immediately available in the configuration and
+% length required, as such a cable loom needed to be made which meant
+% crimping cables by hand. A crimping tool from JST is prohibitively
+% expensive compared to the cost of other components in the project*. The
+% crimping tool also appeared to be uncommon and was not readily available
+% in the workshops at the projects disposal. Eventually a tool was found,
+% but an alternative approach was also researched in order to maintain the
+% criteria of accessibility.
+
+% A locator was adapted from a creative commons stl model
+% https://www.thingiverse.com/thing:1646016 that allowed a generic
+% crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
+% locator was printed using a Bambu X1 3D printer.
+
+% \emph{** A second alterntaive was assembling cables from other sets that
+% were at hand.} \emph{* from RS
+% https://uk.rs-online.com/web/p/crimp-tools/6880877 £508 inc VAT} \emph{†
+% model available in the git repository}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   mount for nano to allow switching
+% \item
+%   initially push fit 2.54mm header
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     not a secure fit
+%   \end{itemize}
+% \item
+%   JST-PH used
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     crimping required custom cables
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       could be assembled from other cables
+%     \item
+%       length limited
+%     \end{itemize}
+%   \item
+%     prohibitevly expensive to crimpers
+%   \item
+%     3D print helper to allow for crimping with generic crimpers
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{3D printing}\label{d-printing}
+
+% Access to a 3D printer is was vital to the success of the project. Given
+% the proliferation of 3D printing, particular in maker spaces in
+% universities it was deemed an unreasonable requirement. 3D printed parts
+% include the support brackets for PCBs, the baffles for the sensors,
+% locator for crimping JST-PH crimps and a mount for the MacMini used for
+% audio synthesis.
+
+% The printer used throughout the project was a Bambu X1-Carbon using
+% recycled PLA filament.
+
+% Support brackets and baffles were made for the project while the MacMini
+% bracket and JST locator were sourced from creative commons models.
+
+% Any colour of filament can be used, though it is advised to use a dark
+% coloured filament for the baffles.
+
+% See each corresponding section for more detail on 3D printed components.
+
+% \begin{itemize}
+% \item
+%   https://www.thingiverse.com/thing:1646016
+% \item
+%   https://www.thingiverse.com/thing:2886460
+% \item
+%   Supports
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Push fit
+%   \item
+%     adjustable for best average placement
+%   \end{itemize}
+% \item
+%   Blinkers
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     reduce noise from other senors
+%   \end{itemize}
+% \item
+%   mounts
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     MacMini
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{Power}\label{power}
+
+% All electronics draws around 1.1amps of current at 5V with some
+% fluctuation during boot.
+
+% Arduino BLE has some functionality disabled to limit current draw.
+
+% Ideally the unit could be bus powered from USB.
+
+% The sensors are in a state of constantly drawing power. It is possible
+% that powering only when data is required could be reduce power enough.
+
+% Current designs have a separate MIDI device for each jack row. Combining
+% the devices and also reducing current draw to below the common 0.5 amps
+% for a USB port would certainly be a challenge.
+
+% \section{Firmware Design}\label{firmware-design}
+
+% Arduino platform chosen for wide adoption, relative ease of use and
+% library support for components.
+
+% Firmare checks components are connected then reads the FRAM to see if
+% threshold data is already present. The rotary encoder and it's momentary
+% switch are polled for any change in state. A key is selected with the
+% rotary encoder while the momentary switch is used for changing between
+% states for key selection and threshold adjustment A double click of the
+% rotary encoder writes the current threshold values to FRAM. The sensors
+% for the current multiplexer channels are polled. If any sensor reads
+% above the threshold when it was previously below a MIDI On message is
+% sent. For the reverse case where the value was below the threshold when
+% it was previously above, a MIDI Note Off message ois sent.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Arduino platform
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     library support and portability
+%   \item
+%     dependancy management
+%   \item
+%     wide spread use
+%   \end{itemize}
+% \item
+%   core algorithm
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     check components connected
+%   \item
+%     READ data from FRAM
+%   \item
+%     select key with rotary
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       click to edit, double to save
+%     \end{itemize}
+%   \item
+%     current threshold and reading sent over serial and plotted
+%   \item
+%     when over threshold send MIDI note
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{Project structure}\label{project-structure}
+
+% The project is separated into three repositories encapsulating
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Firmware
+% \item
+%   Keyboard CAD Data
+% \item
+%   PCB CAD Data
+% \end{itemize}
+
+% \subsubsection{Firmware}\label{firmware}
+
+% Firmware repository contains all firmware for the full and smaller
+% models and components.
+
+% \subsubsection{Keyboard}\label{keyboard-1}
+
+% The Keyboard repository contains all plans for fabricating the keyboard,
+% measurements and 3D models of the jacks.
+
+% \subsubsection{PCB}\label{pcb}
+
+% PCB repository contains the EAGLE files for creating the sensor and
+% controller boards.
+
+% Given the regular spacing of components, scripts for automatic placement
+% of components based on parameter of jack pitch spacing and number of
+% sensors per jack.
+
+% \section{Implementation}\label{implementation}
+
+% \subsection{Installation}\label{installation}
+
+% The electronic components were partly pre-assembled at the university of
+% Edinburgh. Though all measurments were known ahead, enough flexibiliyty
+% was needed to allow for proboems to solved in situ.
+
+% The final assembly took place at the NEMUS lab at the univerity of
+% bologna over the final week in October 2024 for delivery at San
+% Colombano for exhibition in the first week of November in 2024.
+
+% A misreading in the original clearance between the keys and the PCB
+% meant that the keys wer block entirely from moving. A 10mm by 15mm
+% section was milled out of each lke \FIGURE to create extra clearance.
+% The milling required the removal, shortening and refitting of the
+% leather pads for the jacks \FIGURE. The extra work added an extra day to
+% the full install.
+
+% A redesign of the PCBs meant that removing the material would no longer
+% be necessary.
+
+% The PCBs shared power and data lines for components with the exception
+% of the output from the multiplexer. Ribbon cable was cut to size to
+% allow for some room to readjust the distance of the sensors from the
+% jacks. The data signals from each PCB were connected using a separate
+% cable loom. Soldering cables and inspecting each join required another
+% day work. This did allow for readjustments to the positon of the
+% controller board Measurements of the rear chamber did not match with the
+% MacMini that was to be used a different layout was decided for fitting
+% the controller board/ Cutting cables to size allowed for a great deal of
+% felxibility but at the cost of time. After the first install it was
+% clearer what constraints were in place for fittig and an alternative
+% wiring system using flat flex cable was implement (discussed in
+% \textbackslash section)
+
+% Fitting PCBs required the removal of all keys. The support brackets were
+% assembled and screwed into the the underside. With the harpsichord on
+% it's back the sensors were tested to ensure an expected range.
+% Self-tapping round head screws were used to to attach the brackets. The
+% screws were not over tightened to allow for brackets to move. The
+% brackets were not easily adjusted once the key were replaced. Adjusted
+% to a best average with subsequent adjustment carried out on the
+% microcontroller.
+
+% As the PCBs were fitted the jacks had the gradient stickers applied to
+% them. The jacks had been coated with a varnish on the short side to
+% provide a better surface for the stikcers to adhere to. After the
+% stickers were applied excess material was trimmed from each jack. The
+% jack were then check that they moved freely in their slot.
+
+% After fitting the PCBs preliminary calibration was carried out. Before
+% the keys were reffitted the jacks wer put back in place. The data from
+% each key was confirmed to be in a similar range. Any key showing
+% anomalous dat had it's jack removed and re-inspected. The PCB fitting
+% process and refitting required another day as faults were found in
+% wiring or misalignement of the stickers. After the sensors were
+% confirmed to be functioning the keys were put back.
+
+% The calibration workflow was with Dr.~Craig Webb who was unfamiliar with
+% the system. Adjustments were made to the firmware to to make it more
+% intuitive for the user. Addressable LEDs became vital as it allowed for
+% easy identification of the key being calibrated. They were also used to
+% identify any keys whose current reading was beyond what was expected.
+
+% The rotary encoder was used as the interface to select a key, adjust
+% it's threshold and update the Ferroelectric RAM (FRAM) with new values.
+
+% A process was implemented where the key was selected by playing, but
+% this caused problems in situations where the data was unusual due to
+% misalignement of the sensors
+
+% This would likely be a better system for when the keyboard was fully
+% operational
+
+% The thresholds were calibrated visually utilising the serial plotter of
+% the Arduino IDE and MIDI was confirmed using the open source MIDI
+% Monitor (https://github.com/krevis/MIDIApps).
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   partial pre-assembly
+% \item
+%   handsolder components
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     analog lines
+%   \end{itemize}
+% \item
+%   connecting supports
+% \end{itemize}
+
+% \subsection{Context}\label{context}
+
+% Designed as part of an exhibition for the museum San Colombano in
+% Bologna. Instruments is to be played by general public as a means of
+% interacting with items in the collection that are no longer in playing
+% condition. Audience of one who listens via headphones.
+
+% The strings of the interface are damped with felt strips. The damping
+% did change the tension and thereforew the feel of the strings so
+% adjustments needed to be made to the tuning to accomodate for this.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   in museum context
+% \item
+%   strings damped
+% \item
+%   headphones
+% \item
+%   interfacing with kontakt spitfire library
+% \item
+%   display
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     secure storage and access
+%   \item
+%     interface to the public
+%   \end{itemize}
+% \end{itemize}
+
+% \subsection{Calibration}\label{calibration}
+
+% Finer calibration carried out with the guidance of Catalina Vincens as
+% an expert harpsichord performer
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   sensor surface
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     printed gradient
+%   \end{itemize}
+% \end{itemize}
+
+% \section{Conclusion}\label{conclusion}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   scale
+% \item
+%   time scale
+% \item
+%   working at distance
+% \item
+%   limitations
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     latency
+%   \item
+%     deprecation
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       software
+
+%       \begin{itemize}
+%       \tightlist
+%       \item
+%         EAGLE
+%       \end{itemize}
+%     \item
+%       hardware
+
+%       \begin{itemize}
+%       \tightlist
+%       \item
+%         FRAM
+%       \end{itemize}
+%     \end{itemize}
+%   \end{itemize}
+% \end{itemize}
+
+% \section{Going forward}\label{going-forward}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Velocity, aftertouch
+% \item
+%   second jack row
+% \item
+%   reducing latency
+% \item
+%   alternative sensor surfaces
+% \end{itemize}
+
+% A limitation in the design is the variablility in positioning of tage,
+% especially when they need to be reapplied. A jig was designed for
+% applying tags, but could not be printed in time for installation. Even
+% with a jig, some tags had to be radjusted by hand regardless. Long term
+% this hasd impliccation on how to maintain the interface.
+
+% Taken a different approach to what surface is used. Preliminary test
+% were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
+% cutting results in a loss of material that has too much ariablity. The
+% edge o f the jack body will be charrred. The material also needs to be
+% reliable. on the advice of two luthiers, roberto livi and Jonathan Santa
+% Maria Bouquet at Edinburgh, Plywood and MDF have problems with
+% variations in humidity, tear-out and delamination.
+
+% Some tests were carried out using a notched jack body that was 3D
+% printed. The notch means that as the jack raises, the forward material
+% acts like a shutter. No reasons that this could not be CNC though a
+% variety of materials would be needed. Traditional material used is xx
+% though need to test of the material is naturally reflective enough or if
+% a coating would beed to be applied.
+
+% Outputs of the NEMUS project will also include numerically simulated
+% non-linear plucked string models with which the device can interact. A
+% bespoke sound synthesis engine to leverage data from both jack rows
+% which will also be open-sourced. Design for instrument to be compatible
+% with standard MIDI instruments as well software designed for the full
+% data output.
+
+% A second interface designed more closely to the Trasuntino style
+% instrument commisioned by the NEMUS project with Roberto Livi. Using
+% lessons learned for the first iteration the second can make better
+% acommodations for internal electronics. Interface to be used for a
+% performer study and used for performance with a numerically simulated
+% sound syntheiss engine.
+
+% Testing current sensors in conjunction with force sensitive resistors to
+% research if a impulse response for the pluck can be extrapolated.
+
+% Future novel designs for an interface that would utilise velocity and
+% aftertouch messages in software. This would be applicable for numericla
+% simulations of strings. A bespoke piec e of sowftare would also allow
+% for direct mapping of jack rows to registers. There is enough data to
+% implemenet functionality for velocity and aftertouch MIDI messages and
+% simply needs to be activated in the firmware.
+
+% Evaluation
+
+% \subsection{improvements}\label{improvements}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   ESP-IDF
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     latency half
+%   \item
+%     workload high
+%   \end{itemize}
+% \end{itemize}
+
+% A current limitation is the latency in triggering MIDI output. Currently
+% the latency is around 20 ms, but should ideally be closer to 5ms.
+
+% An obvious approach would be merely changing the microcontroller. A test
+% was carried out using an Arduino Nano ESP32 which is a NORA-W106, a
+% module with a ESP32-S3. The benefit of this microcontroller is that it
+% would allow simply the board to be swapped and the rest of the PCB
+% designs would be unaffected.
+
+% Using the Espressif development framework ESP-IDF and follwong the same
+% logic as the original firmware source, the latency was reduced to around
+% 10ms. ESP-IDF allows access to functionality of the ESP32 that is not
+% exposed in the Arduino libraries. It is possible that the 10ms latency
+% could be improved upon but there would be a cost in the form of
+% develoment time. Consideration would need to be made on how to port
+% existing code or whether to completely re-write the firmware.
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   avoid soldering cabling
+% \item
+%   power indicators on pcbs
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     debug if problem is power related Add power small power LEDs to the
+%     each PCB. It was not common for a board to not be functioning and
+%     thye result to be that the power weas not correctly connected.
+%     Especially when the ower is connected via somethgin such as an FFC
+%     cable where it is easy to plug-in the wrong way esepcially when
+%     fitting is to be carried out by someone unfmailiar wiyth the system.
+%     Power LEDs would quickly confirm whether the problem was power
+%     related. They would have to be small enough and low enough current
+%     draw to avoid glare or being mistaken with other indcator LEDs To
+%     continue using THT components, the choice would have to be moade
+%     carefully.
+%   \end{itemize}
+% \end{itemize}
+
+% \section{Acknowledgments}\label{acknowledgments}
+
+% A project of this breadth developed across 3 institutes and 2 countries
+% cannot be carried without the support and technical knowledge of many
+% skilled and passionate people.
+
+% We acknowledge the invaluable contributions of
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   Unibo
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Alma Labor
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
+%       of bologna for their 3D printing expertise and generosity with
+%       electronics equipment
+%     \end{itemize}
+%   \item
+%     Fisica Tecnica
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
+%       Costantini
+%     \end{itemize}
+%   \end{itemize}
+% \item
+%   San Colombano
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     Catalina Vincens
+%   \item
+%     Roberto Livi
+%   \end{itemize}
+% \item
+%   Edinburgh
+
+%   \begin{itemize}
+%   \tightlist
+%   \item
+%     uCreate
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
+%       Edinburgh for their support with soldering facilities.
+%     \end{itemize}
+%   \item
+%     ECA
+
+%     \begin{itemize}
+%     \tightlist
+%     \item
+%       Richard Collins and the Digital Making Staff at the Edinburgh
+%       College of Art for 3D printing instruction and CNC support.
+%     \item
+%       Joe Hathaway of Digital Development at the College of Art for
+%       providing electronics components
+%     \item
+%       St Cecilia's Jenny Nex
+%     \end{itemize}
+%   \end{itemize}
+% \end{itemize}
+
+% \section{Reference}\label{reference}
+
+% \subsection{Example Trasuntino}\label{example-trasuntino}
+
+% \begin{itemize}
+% \tightlist
+% \item
+%   {[}1538{]}
+%   https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
+% \item
+%   {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
+% \end{itemize}

### 2025-01-15 src/article.tex
cancel cmds

--- 2025-01-15 src/article.tex
@@
 (from Luigi Ferdinando Tagliavini), a collection of approximately 70
-historical keyboard instruments in working order \cite{Tagliavini2009}
+historical keyboard instruments in working order
+\textbackslash cite\{Tagliavini2009\}
 (https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
@@
 instruments in the Tagliavini collection is a 1547 Harpsichord by
-Alessandro Trasuntino \cite{Treccani2019}, which was used as the basis
-for an interactive exhibition commisioned by San Colombano.
+Alessandro Trasuntino \textbackslash cite\{Treccani2019\}, which was
+used as the basis for an interactive exhibition commisioned by San
+Colombano.
 
@@
 
-The paper by McAlpine \cite{McAlpine2014} also has very similar
-motivations to the project presented here. Both were started from a
-musical instrument museum, housing a collection of historic instruments
-gifted by a collector (Benton Fletcher / Taglivaini) who wished for
-collection to remain playable, which is restricted growing fragility of
-some items in the respective collections. Bothe projects center around
-the creation of a custioom MIDI controller for interfacing with sound
-synthesis engine. The keyboard consisted of two modified commercial MIDI
-keyboard controllers in a two manual harpsichord arrangement. Weherease
-the McAlpine keyboard used weighted keys, this interface extends the
-idea by using the mechanism and strings under tension to provide the
-haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
-but before taking such a leap we look at what affect a more literal
-haptic keyboard may provide to the performer experience.
-
-The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
-looked at reconstructing a full tromba marina to be used as a digital
-interface. The one of the limitations flagged by McAlpine was the
-oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
-Moderna ``attempted to avoid this problem by constructing a fully
-functioning replica of the tromba marina'' \cite{Baldwin}. This project
-applies the same approach to the harpsichord.m where the projects
-deviate is a questions of scale.
+The paper by McAlpine \textbackslash cite\{McAlpine2014\} also has very
+similar motivations to the project presented here. Both were started
+from a musical instrument museum, housing a collection of historic
+instruments gifted by a collector (Benton Fletcher / Taglivaini) who
+wished for collection to remain playable, which is restricted growing
+fragility of some items in the respective collections. Bothe projects
+center around the creation of a custioom MIDI controller for interfacing
+with sound synthesis engine. The keyboard consisted of two modified
+commercial MIDI keyboard controllers in a two manual harpsichord
+arrangement. Weherease the McAlpine keyboard used weighted keys, this
+interface extends the idea by using the mechanism and strings under
+tension to provide the haptic feedback. McAlpine posits a fully
+paremterised haptic keyboard, but before taking such a leap we look at
+what affect a more literal haptic keyboard may provide to the performer
+experience.
+
+The previous NIME project ``Tromba Moderna''
+\textbackslash cite\{Baldwin2016\}, which looked at reconstructing a
+full tromba marina to be used as a digital interface. The one of the
+limitations flagged by McAlpine was the oneffectiveness of weighted keys
+in mimicing the harpischord. The Tromba Moderna ``attempted to avoid
+this problem by constructing a fully functioning replica of the tromba
+marina'' \textbackslash cite\{Baldwin\}. This project applies the same
+approach to the harpsichord.m where the projects deviate is a questions
+of scale.
 
@@
 
-Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
-sourcing of hardware and software for the project.
+Extension to \textbackslash cite\{McAlpine2014\} and
+\textbackslash cite\{Baldwin2016\} are the open sourcing of hardware and
+software for the project.
 
@@
 Discussion of spatial constraints and affects on electyronics design and
-installation is covered in
-
-\section{setup}
+installation is covered in \textbackslash section\{setup\}
 
@@
 The keyboard is fully enclosed and internally has two chambers, front
-and rear, in which the electronics could be installed \figure(underside
-49-key).
+and rear, in which the electronics could be installed
+\textbackslash figure(underside 49-key).
 
@@
 space was added into the design of the internal chamber of the model
-\FIGURE, but there was still a limit given the structural and aesthatic
-requirements.
+\textbackslash FIGURE, but there was still a limit given the structural
+and aesthatic requirements.
 
@@
 \item
-  \cite{HapticKey}
-\item
-  \cite{mcPherson1} \cite{mcPherson2}
-\item
-  \cite{Mudd2013}
-\item
-  \cite{FritzClaudia2017Leon}
+  \textbackslash cite\{HapticKey\}
+\item
+  \textbackslash cite\{mcPherson1\} \textbackslash cite\{mcPherson2\}
+\item
+  \textbackslash cite\{Mudd2013\}
+\item
+  \textbackslash cite\{FritzClaudia2017Leon\}
 \end{itemize}
@@
 or procure a model of a single working key. San Colombano provided a
-model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
-3-key model was used a test-bed for potential sensors while designs were
-drawn for a full-scale model. The cross-sectional nature of the model
-meant it was easy to fit and remove sensors. Small size meant that it
-would also be easy to transfer between workshops.
+model 3-key of a harposchord mechanism by Graziano Bandini
+\textbackslash FIGURE. The 3-key model was used a test-bed for potential
+sensors while designs were drawn for a full-scale model. The
+cross-sectional nature of the model meant it was easy to fit and remove
+sensors. Small size meant that it would also be easy to transfer between
+workshops.
 
@@
 
-Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
-a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
-combination infrared LED and phototransistor sensitive to IR light. The
+Following from the work in \textbackslash cite\{McPherson1\}
+\textbackslash cite\{McPherson2\} we tested a system using the Fairchild
+QRE1113 \textbackslash datasheetfigue. The QRE1113 is a combination
+infrared LED and phototransistor sensitive to IR light. The
 phototransistor of QRE1113 provides data on how much light is present
@@
 it is a good, close proximity, distance sensor. Distance is the way in
-which the instruments in \citeMcpherson1\&2 function as well the Moog
-Piano Bar from which it too inspiration.
+which the instruments in \textbackslash citeMcpherson1\&2 function as
+well the Moog Piano Bar from which it too inspiration.
 
@@
 continue with the Nano BLE. Further discussion of using ESP32 continues
-in
-
-\section{Going forward}
+in \textbackslash section\{Going forward\}
 
@@
 
-TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
-to assembling, but were avoid in favour of Addressable LEDs with
-intergrated driver. The intergrated LEDs share the same data line which
-has the limitation where a fualty LED will stop all following LEDs from
-working. An individual PWM driver per PCB would allow avoid situations
-where failutre cascades, but would add complexity in assembly as the
-parts are typically SMD. Also, given the LEDs are fixed to the board and
-could not be repaired in situ, the solution would be to replace the PCB
-regardless of which approach was taken. Given the LEDs were only used in
-debugging and calibration and otehrwise would not be visible, a simpler
-design approach was taken.
+TLC5940 or other PWM driver such as in \textbackslash cite\{mcpherson\}
+added complexity to assembling, but were avoid in favour of Addressable
+LEDs with intergrated driver. The intergrated LEDs share the same data
+line which has the limitation where a fualty LED will stop all following
+LEDs from working. An individual PWM driver per PCB would allow avoid
+situations where failutre cascades, but would add complexity in assembly
+as the parts are typically SMD. Also, given the LEDs are fixed to the
+board and could not be repaired in situ, the solution would be to
+replace the PCB regardless of which approach was taken. Given the LEDs
+were only used in debugging and calibration and otehrwise would not be
+visible, a simpler design approach was taken.
 
@@
 Transistor / Op-amp circuit could have been used similar to the one in
-\cite{mcpherson}.
+\textbackslash cite\{mcpherson\}.
 
@@
 https://www.thingiverse.com/thing:1646016 that allowed a generic
-crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
-locator was printed using a Bambu X1 3D printer.
+crimping tool to be adapted for use with JST-PH crimps†
+\textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
+printer.
 
@@
 meant that the keys wer block entirely from moving. A 10mm by 15mm
-section was milled out of each lke \FIGURE to create extra clearance.
-The milling required the removal, shortening and refitting of the
-leather pads for the jacks \FIGURE. The extra work added an extra day to
-the full install.
+section was milled out of each lke \textbackslash FIGURE to create extra
+clearance. The milling required the removal, shortening and refitting of
+the leather pads for the jacks \textbackslash FIGURE. The extra work
+added an extra day to the full install.
 

### 2025-01-15 src/article.tex
add ethical standards template

@@
 
+\section{Ethical Standards}\label{ethical-standards}
+
+To ensure objectivity and transparency in research and to ensure that
+accepted principles of ethical and professional conduct have been
+followed, authors must include a section ``Ethical Standards'' before
+the References. This section should include (if relevant): information
+regarding sources of funding, potential conflicts of interest (financial
+or non-financial), informed consent if the research involved human
+participants, statement on welfare of animals if the research involved
+animals or any other information or context that helps ethically situate
+your research. For help with the ethics section, feel free to ask on the
+NIME forum: \url{https://forum.nime.org}.
+
+\subsection{Example Trasuntino}\label{example-trasuntino}
+
+\begin{itemize}
+\tightlist
+\item
+  {[}1538{]}
+  https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
+\item
+  {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
+\end{itemize}
+
 \section{Acknowledgments}\label{acknowledgments}
@@
 \end{itemize}
-
-\section{Reference}\label{reference}
-
-\subsection{Example Trasuntino}\label{example-trasuntino}
-
-\begin{itemize}
-\tightlist
-\item
-  {[}1538{]}
-  https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
-\item
-  {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
-\end{itemize}

### 2025-01-15 src/main.tex

@@
     Bib\TeX}}}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{images/sampleteaser.pdf}
+  \includegraphics[width=\textwidth]{src/images/side_w-o_sensors.jpeg}
   \caption{Two synthesisers on a table, 2019.}
@@
 
-\section{Introduction}
-ACM's consolidated article template, introduced in 2017, provides a
-consistent \LaTeX\ style for use across ACM publications, and
-incorporates accessibility and metadata-extraction functionality
-necessary for future Digital Library endeavors. Numerous ACM and
-SIG-specific \LaTeX\ templates have been examined, and their unique
-features incorporated into this single new template.
-
-If you are new to publishing with ACM, this document is a valuable
-guide to the process of preparing your work for publication. If you
-have published with ACM before, this document provides insight and
-instruction into more recent changes to the article template.
-
-The ``\verb|nimeart|'' document class can be used to prepare articles
-for any ACM publication --- conference or journal, and for any stage
-of publication, from review to final ``camera-ready'' copy, to the
-author's own version, with {\itshape very} few changes to the source.
-
-\section{Template Overview}
-As noted in the introduction, the ``\verb|nimeart|'' document class can
-be used to prepare many different kinds of documentation --- a
-double-anonymous initial submission of a full-length technical paper, a
-two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
-journal article, a SIGCHI Extended Abstract, and more --- all by
-selecting the appropriate {\itshape template style} and {\itshape
-  template parameters}.
-
-This document will explain the major features of the document
-class. For further information, the {\itshape \LaTeX\ User's Guide} is
-available from
-\url{https://www.acm.org/publications/proceedings-template}.
-
-\subsection{Template Styles}
-
-The primary parameter given to the ``\verb|nimeart|'' document class is
-the {\itshape template style} which corresponds to the kind of publication
-or SIG publishing the work. This parameter is enclosed in square
-brackets and is a part of the {\verb|documentclass|} command:
-\begin{verbatim}
-  \documentclass[STYLE]{nimeart}
-\end{verbatim}
-
-Journals use one of three template styles. All but three ACM journals
-use the {\verb|acmsmall|} template style:
-\begin{itemize}
-\item {\texttt{acmsmall}}: The default journal template style.
-\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
-\item {\texttt{acmtog}}: Used by TOG.
-\end{itemize}
-
-The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
-\begin{itemize}
-\item {\texttt{sigconf}}: The default proceedings template style.
-\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
-\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
-\end{itemize}
-
-\subsection{Template Parameters}
-
-In addition to specifying the {\itshape template style} to be used in
-formatting your work, there are a number of {\itshape template parameters}
-which modify some part of the applied template style. A complete list
-of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}
-
-Frequently-used parameters, or combinations of parameters, include:
-\begin{itemize}
-\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
-  conference submission. Anonymizes the work and includes line
-  numbers. Use with the \texttt{\string\acmSubmissionID} command to print the
-  submission's unique ID on each page of the work.
-\item{\texttt{authorversion}}: Produces a version of the work suitable
-  for posting by the author.
-\item{\texttt{screen}}: Produces colored hyperlinks.
-\end{itemize}
-
-This document uses the following string as the first command in the
-source file:
-\begin{verbatim}
-\documentclass[sigconf]{nimeart}
-\end{verbatim}
-
-\section{Modifications}
-
-Modifying the template --- including but not limited to: adjusting
-margins, typeface sizes, line spacing, paragraph and list definitions,
-and the use of the \verb|\vspace| command to manually adjust the
-vertical spacing between elements of your work --- is not allowed.
-
-{\bfseries Your document will be returned to you for revision if
-  modifications are discovered.}
-
-\section{Typefaces}
-
-The ``\verb|nimeart|'' document class requires the use of the
-``Libertine'' typeface family. Your \TeX\ installation should include
-this set of packages. Please do not substitute other typefaces. The
-``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
-as they will override the built-in typeface families.
-
-\section{Title Information}
-
-The title of your work should use capital letters appropriately -
-\url{https://capitalizemytitle.com/} has useful rules for
-capitalization. Use the {\verb|title|} command to define the title of
-your work. If your work has a subtitle, define it with the
-{\verb|subtitle|} command.  Do not insert line breaks in your title.
-
-If your title is lengthy, you must define a short version to be used
-in the page headers, to prevent overlapping text. The \verb|title|
-command has a ``short title'' parameter:
-\begin{verbatim}
-  \title[short title]{full title}
-\end{verbatim}
-
-\section{Authors and Affiliations}
-
-Each author must be defined separately for accurate metadata
-identification.  As an exception, multiple authors may share one
-affiliation. Authors' names should not be abbreviated; use full first
-names wherever possible. Include authors' e-mail addresses whenever
-possible.
-
-Grouping authors' names or e-mail addresses, or providing an ``e-mail
-alias,'' as shown below, is not acceptable:
-\begin{verbatim}
-  \author{Brooke Aster, David Mehldau}
-  \email{dave,judy,steve@university.edu}
-  \email{firstname.lastname@phillips.org}
-\end{verbatim}
-
-The \verb|authornote| and \verb|authornotemark| commands allow a note
-to apply to multiple authors --- for example, if the first two authors
-of an article contributed equally to the work.
-
-If your author list is lengthy, you must define a shortened version of
-the list of authors to be used in the page headers, to prevent
-overlapping text. The following command should be placed just after
-the last \verb|\author{}| definition:
-\begin{verbatim}
-  \renewcommand{\shortauthors}{McCartney, et al.}
-\end{verbatim}
-Omitting this command will force the use of a concatenated list of all
-of the authors' names, which may result in overlapping text in the
-page headers.
-
-The article template's documentation, available at
-\url{https://www.acm.org/publications/proceedings-template}, has a
-complete explanation of these commands and tips for their effective
-use.
-
-Note that authors' addresses are mandatory for journal articles.
-
-\section{Anonymisation and Blind Review}
-Anonymisation is a \textbf{crucial} component for NIME paper submissions. We kindly ask authors to ensure that all information that could introduce bias in the review process is anonymised.
-
-To anonymise author names, please specify \verb|anonymous,review| in square brackets when submitting your paper.
-\begin{verbatim}
-  \documentclass[sigchi,anonymous,review]{nimepaper}
-\end{verbatim}
-
-You can also hide text in the paper like this:
-
-\begin{verbatim}
-% Both examples from acmart manual.
-\begin{anonsuppress}
-  This is the continuation of the previous work by the author
-  \cite{prev1, prev2}.
-\end{anonsuppress}
-\end{verbatim}
-
-or
-\begin{verbatim}
-This work was performed at \anon{NSA}.
-This work was performed at \anon[No Such Agency]{NSA}.
-\end{verbatim}
-
-\section{Rights Information}
-
-Authors of any work published by ACM will need to complete a rights
-form. Depending on the kind of work, and the rights management choice
-made by the author, this may be copyright transfer, permission,
-license, or an OA (open access) agreement.
-
-Regardless of the rights management choice, the author will receive a
-copy of the completed rights form once it has been submitted. This
-form contains \LaTeX\ commands that must be copied into the source
-document. When the document source is compiled, these commands and
-their parameters add formatted text to several areas of the final
-document:
-\begin{itemize}
-\item the ``ACM Reference Format'' text on the first page.
-\item the ``rights management'' text on the first page.
-\item the conference information in the page header(s).
-\end{itemize}
-
-Rights information is unique to the work; if you are preparing several
-works for an event, make sure to use the correct set of commands with
-each of the works.
-
-The ACM Reference Format text is required for all articles over one
-page in length, and is optional for one-page articles (abstracts).
-
-\section{CCS Concepts and User-Defined Keywords}
-
-Two elements of the ``nimeart'' document class provide powerful
-taxonomic tools for you to help readers find your work in an online
-search.
-
-The ACM Computing Classification System ---
-\url{https://www.acm.org/publications/class-2012} --- is a set of
-classifiers and concepts that describe the computing
-discipline. Authors can select entries from this classification
-system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
-commands to be included in the \LaTeX\ source.
-
-User-defined keywords are a comma-separated list of words and phrases
-of the authors' choosing, providing a more flexible way of describing
-the research being presented.
-
-CCS concepts and user-defined keywords are required for for all
-articles over two pages in length, and are optional for one- and
-two-page articles (or abstracts).
-
-\section{Sectioning Commands}
-
-Your work should use standard \LaTeX\ sectioning commands:
-\verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
-\verb|\paragraph|, and \verb|\subparagraph|. The sectioning levels up to
-\verb|\subsusection| should be numbered; do not remove the numbering
-from the commands.
-
-Simulating a sectioning command by setting the first word or words of
-a paragraph in boldface or italicized text is {\bfseries not allowed.}
-
-Below are examples of sectioning commands.
-
-\subsection{Subsection}
-\label{sec:subsection}
-
-This is a subsection.
-
-\subsubsection{Subsubsection}
-\label{sec:subsubsection}
-
-This is a subsubsection.
-
-\paragraph{Paragraph}
-
-This is a paragraph.
-
-\subparagraph{Subparagraph}
-
-This is a subparagraph.
-
-\section{Tables}
-
-The ``\verb|nimeart|'' document class includes the ``\verb|booktabs|''
-package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
-high-quality tables.
-
-Table captions are placed {\itshape above} the table.
-
-Because tables cannot be split across pages, the best placement for
-them is typically the top of the page nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and the
-table caption.  The contents of the table itself must go in the
-\textbf{tabular} environment, to be aligned properly in rows and
-columns, with the desired horizontal and vertical rules.  Again,
-detailed instructions on \textbf{tabular} material are found in the
-\textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table~\ref{tab:freq} is included in the input file; compare the
-placement of the table here with the table in the printed output of
-this document.
-
-\begin{table}
-  \caption{Frequency of Special Characters}
-  \label{tab:freq}
-  \begin{tabular}{ccl}
-    \toprule
-    Non-English or Math&Frequency&Comments\\
-    \midrule
-    \O & 1 in 1,000& For Swedish names\\
-    $\pi$ & 1 in 5& Common in math\\
-    \$ & 4 in 5 & Used in business\\
-    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of the page's
-live area, use the environment \textbf{table*} to enclose the table's
-contents and the table caption.  As with a single-column table, this
-wide table will ``float'' to a location deemed more
-desirable. Immediately following this sentence is the point at which
-Table~\ref{tab:commands} is included in the input file; again, it is
-instructive to compare the placement of the table here with the table
-in the printed output of this document.
-
-\begin{table*}
-  \caption{Some Typical Commands}
-  \label{tab:commands}
-  \begin{tabular}{ccl}
-    \toprule
-    Command &A Number & Comments\\
-    \midrule
-    \texttt{{\char'134}author} & 100& Author \\
-    \texttt{{\char'134}table}& 300 & For tables\\
-    \texttt{{\char'134}table*}& 400& For wider tables\\
-    \bottomrule
-  \end{tabular}
-\end{table*}
-
-Always use midrule to separate table header rows from data rows, and
-use it only for this purpose. This enables assistive technologies to
-recognise table headers and support their users in navigating tables
-more easily.
-
-\section{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of the three are
-discussed in the next sections.
-
-\subsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an inline or
-in-text formula.  It is produced by the \textbf{math} environment,
-which can be invoked with the usual
-\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
-the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
-and structures, from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
-examples of in-text equations in context. Notice how this equation:
-\begin{math}
-  \lim_{n\rightarrow \infty}x=0
-\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsection{Display Equations}
-A numbered display equation---one set off by vertical space from the
-text and centered horizontally---is produced by the \textbf{equation}
-environment. An unnumbered display equation is produced by the
-\textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols and
-structures available in \LaTeX\@; this section will just give a couple
-of examples of display equations in context.  First, consider the
-equation, shown as an inline equation above:
-\begin{equation}
-  \lim_{n\rightarrow \infty}x=0
-\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}
-  \sum_{i=0}^{\infty} x + 1
-\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}
-  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\section{Figures}
-
-The ``\verb|figure|'' environment should be used for figures. One or
-more images can be placed within a figure. If your figure contains
-third-party material, you must clearly identify it as such, as shown
-in the example below.
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{images/sample-bela.jpg}
-  \caption{A Bela board in action at a workshop (2017).}
-  \Description{Somebody manipulating sensors on a Bela board.}
-\end{figure}
-
-Your figures should contain a caption which describes the figure to
-the reader.
-
-Figure captions are placed {\itshape below} the figure.
-
-Every figure should also have a figure description unless it is purely
-decorative. These descriptions convey what’s in the image to someone
-who cannot see it. They are also used by search engine crawlers for
-indexing images, and when images cannot be loaded.
-
-A figure description must be unformatted plain text less than 2000
-characters long (including spaces).  {\bfseries Figure descriptions
-  should not repeat the figure caption – their purpose is to capture
-  important information that is not already provided in the caption or
-  the main text of the paper.} For figures that convey important and
-complex new information, a short text description may not be
-adequate. More complex alternative descriptions can be placed in an
-appendix and referenced in a short figure description. For example,
-provide a data table capturing the information in a bar chart, or a
-structured list representing a graph.  For additional information
-regarding how best to write figure descriptions and why doing this is
-so important, please see
-\url{https://www.acm.org/publications/taps/describing-figures/}.
-
-\subsection{The ``Teaser Figure''}
-
-A ``teaser figure'' is an image, or set of images in one figure, that
-are placed after all author and affiliation information, and before
-the body of the article, spanning the page. If you wish to have such a
-figure in your article, place the command immediately before the
-\verb|\maketitle| command:
-\begin{verbatim}
-  \begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{figure caption}
-    \Description{figure description}
-  \end{teaserfigure}
-\end{verbatim}
-
-\section{Citations and Bibliographies}
-
-The use of \BibTeX\ for the preparation and formatting of one's
-references is strongly recommended. Authors' names should be complete
---- use full first names (``Donald E. Knuth'') not initials
-(``D. E. Knuth'') --- and the salient identifying features of a
-reference should be included: title, year, volume, number, pages,
-article DOI, etc.
-
-The bibliography is included in your source document with these two
-commands, placed just before the \verb|\end{document}| command:
-\begin{verbatim}
-  \bibliographystyle{ACM-Reference-Format}
-  \bibliography{bibfile}
-\end{verbatim}
-where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-suffix, of the \BibTeX\ file.
-
-Citations and references are numbered by default. A small number of
-ACM publications have citations and references formatted in the
-``author year'' style; for these exceptions, please include this
-command in the {\bfseries preamble} (before the command
-``\verb|\begin{document}|'') of your \LaTeX\ source:
-\begin{verbatim}
-  \citestyle{acmauthoryear}
-\end{verbatim}
-
-  Some examples.  A paginated journal article \cite{Abril07}, an
-  enumerated journal article \cite{Cohen07}, a reference to an entire
-  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
-  monograph/whole book in a series (see 2a in spec. document)
-  \cite{Harel79}, a divisible-book such as an anthology or compilation
-  \cite{Editor00} followed by the same example, however we only output
-  the series if the volume number is given \cite{Editor00a} (so
-  Editor00a's series should NOT be present since it has no vol. no.),
-  a chapter in a divisible book \cite{Spector90}, a chapter in a
-  divisible book in a series \cite{Douglass98}, a multi-volume work as
-  book \cite{Knuth97}, a couple of articles in a proceedings (of a
-  conference, symposium, workshop for example) (paginated proceedings
-  article) \cite{Andler79, Hagerup1993}, a proceedings article with
-  all possible elements \cite{Smith10}, an example of an enumerated
-  proceedings article \cite{VanGundy07}, an informally published work
-  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
-    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
-  master's thesis: \cite{anisi03}, an online document / world wide web
-  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
-  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
-  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
-  publication \cite{rous08}, 'YYYYb'-test for prolific author
-  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
-  contain 'duplicate' DOI and URLs (some SIAM articles)
-  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
-  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
-  couple of citations with DOIs:
-  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
-  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
-  Artifacts: \cite{R} and \cite{UMassCitations}.
-
-\section{Acknowledgments}
-
-Identification of funding sources and other support, and thanks to
-individuals and groups that assisted in the research and the
-preparation of the work should be included in an acknowledgment
-section, which is placed just before the reference section in your
-document.
-
-This section has a special environment:
-\begin{verbatim}
-  \begin{acks}
-  ...
-  \end{acks}
-\end{verbatim}
-so that the information contained therein can be more easily collected
-during the article metadata extraction phase, and to ensure
-consistency in the spelling of the section heading.
-
-Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.
-
-\section{Appendices}
-
-If your work needs an appendix, add it before the
-``\verb|\end{document}|'' command at the conclusion of your source
-document.
-
-Start the appendix with the ``\verb|appendix|'' command:
-\begin{verbatim}
-  \appendix
-\end{verbatim}
-and note that in the appendix, sections are lettered, not
-numbered. This document has two appendices, demonstrating the section
-and subsection identification method.
-
-%% The Ethical Standards section is mandatory for all NIME submissions.
-\section{Ethical Standards}
-
-To ensure objectivity and transparency in research and to ensure that accepted principles of ethical and professional conduct have been followed, authors must include a section “Ethical Standards” before the References.
-This section should include (if relevant): information regarding sources of funding, potential conflicts of interest (financial or non-financial), informed consent if the research involved human participants, statement on welfare of animals if the research involved animals or any other information or context that helps ethically situate your research.
-For help with the ethics section, feel free to ask on the NIME forum: \url{https://forum.nime.org}.
+\include{article.tex}
 
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-references}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-
-\section{Research Methods}
-
-\subsection{Part One}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
-malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
-sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
-vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
-lacinia dolor. Integer ultricies commodo sem nec semper.
-
-\subsection{Part Two}
-
-Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
-ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
-ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
-eros. Vivamus non purus placerat, scelerisque diam eu, cursus
-ante. Etiam aliquam tortor auctor efficitur mattis.
-
-\section{Online Resources}
-
-Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
-pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
-enim maximus. Vestibulum gravida massa ut felis suscipit
-congue. Quisque mattis elit a risus ultrices commodo venenatis eget
-dui. Etiam sagittis eleifend elementum.
-
-Nam interdum magna at lectus dignissim, ac dignissim lorem
-rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
-massa et mattis lacinia.
+\bibliography{main}
 

### 2025-01-15 src/main.tex
authors

@@
 %% used to denote shared contribution to the research.
-\author{Ben Trovato}
-\authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
+\author{Matthew Hamilton}
+% \authornote{Both authors contributed equally to this research.}
+\email{matthew.hamilton2@unibo.it}
 \orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
+\author{Michele Ducceschi}
+% \authornotemark[1]
+\email{michele.ducceschi@unibo.it}
 \affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \city{Dublin}
-  \state{Ohio}
-  \country{USA}
+  \institution{Università di Bologna}
+  \city{Bologna}  
+  \country{Italy}
 }
 
-\author{Lars Th{\o}rv{\"a}ld}
+\author{Roberto Livi}
+\author{Catalina Vincens}
 \affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
+  \institution{Museo San Colombano}
+  \city{Bologna}
+  \country{Italy}}
 
-\author{Valerie B\'eranger}
+\author{Andrew McPherson}
 \affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \city{San Antonio}
-  \state{Texas}
-  \country{USA}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \city{Hekla}
-  \country{Iceland}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{%
-  \institution{The Kumquat Consortium}
-  \city{New York}
-  \country{USA}}
-\email{jpkumquat@consortium.net}
+ \institution{Imperial College}
+ \city{London}
+ \country{United Kingdom}}
 
@@
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{src/images/side_w-o_sensors.jpeg}
-  \caption{Two synthesisers on a table, 2019.}
-  \Description{Enjoying looking at two synthesisers.
-  A Korg Monotron and a Korg Volca FM.}
+  \includegraphics[width=\textwidth]{images/side_w-o_sensors.jpeg}
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
+  \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
   \label{fig:teaser}

### 2025-01-15 src/main.tex
abstract

@@
 \begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. This uses ``nimeart'', a fork of the ``acmart''
-  documentclass with additions for the NIME conference. This
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
+Presented is a design for an electronically augmented replica of a 16th century
+harpsichord interface, in the style of a Trasuntino from 1547. The interface was
+designed, fabricated and exhibited in a collaboration between the NEMUS project
+at the Università di Bologna and the Museo San Colombano, which hosts the
+Tagliavini Collection of playable historical keyboard instruments. The interface
+was commissioned as part of an exhibition and as a means of interacting with
+items in the collection that are no longer in playing condition. The keyboard is
+a replica of the mechanism found in harpsichords in the Trasuntino style, with
+un-pitched strings under tension to provide a comparable mechanical resistance
+as found on an original. A system of infra-red LED / phototransistor pairs were
+connected to a microcontroller and  installed inside the interface from which a
+stream of data can be generated for each jack. The paper covers the design
+criteria, process, open-source development and execution of installation for
+exhibition at San Colombano.
 \end{abstract}

### 2025-01-16 src/main.tex
Updates from Overleaf

@@
 \begin{acks}
-To Robert, for the bagels and explaining CMYK and color spaces.
+A project of this breadth developed across 3 institutes and 2 countries
+cannot be carried without the support and technical knowledge of many
+skilled and passionate people.
+
+We acknowledge the invaluable contributions of
+
+\begin{itemize}
+\tightlist
+\item
+  Unibo
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Alma Labor
+
+    \begin{itemize}
+    \tightlist
+    \item
+      Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
+      of bologna for their 3D printing expertise and generosity with
+      electronics equipment
+    \end{itemize}
+  \item
+    Fisica Tecnica
+
+    \begin{itemize}
+    \tightlist
+    \item
+      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
+      Costantini
+    \end{itemize}
+  \end{itemize}
+\item
+  San Colombano
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Catalina Vincens
+  \item
+    Roberto Livi
+  \end{itemize}
+\item
+  Edinburgh
+
+  \begin{itemize}
+  \tightlist
+  \item
+    uCreate
+
+    \begin{itemize}
+    \tightlist
+    \item
+      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
+      Edinburgh for their support with soldering facilities.
+    \end{itemize}
+  \item
+    ECA
+
+    \begin{itemize}
+    \tightlist
+    \item
+      Richard Collins and the Digital Making Staff at the Edinburgh
+      College of Art for 3D printing instruction and CNC support.
+    \item
+      Joe Hathaway of Digital Development at the College of Art for
+      providing electronics components
+    \item
+      St Cecilia's Jenny Nex
+    \end{itemize}
+  \end{itemize}
+\end{itemize}
 \end{acks}

### 2025-01-16 src/main.tex
Updates from Overleaf
abstract wording

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The NIME of the Title Is Hello}
+\title{Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction}
 
@@
 \email{matthew.hamilton2@unibo.it}
-\orcid{1234-5678-9012}
+% \orcid{1234-5678-9012}
 \author{Michele Ducceschi}
@@
 Presented is a design for an electronically augmented replica of a 16th century
-harpsichord interface, in the style of a Trasuntino from 1547. The interface was
-designed, fabricated and exhibited in a collaboration between the NEMUS project
-at the Università di Bologna and the Museo San Colombano, which hosts the
-Tagliavini Collection of playable historical keyboard instruments. The interface
-was commissioned as part of an exhibition and as a means of interacting with
-items in the collection that are no longer in playing condition. The keyboard is
-a replica of the mechanism found in harpsichords in the Trasuntino style, with
-un-pitched strings under tension to provide a comparable mechanical resistance
-as found on an original. A system of infra-red LED / phototransistor pairs were
-connected to a microcontroller and installed inside the interface from which a
-stream of data can be generated for each jack. The paper covers the design
-criteria, process, open-source development and execution of installation for
-exhibition at San Colombano.
+harpsichord interface, in the style of a 1547 Trasuntino harpsichord. The
+interface was designed, fabricated, and exhibited in a collaboration between the
+NEMUS project at the Università di Bologna and the Museo San Colombano, which
+hosts the Tagliavini Collection of playable historical keyboard instruments. The
+interface was commissioned as part of an exhibition and as a means of
+interacting with items in the collection that are no longer in playing
+condition. The keyboard is a replica of the mechanism found in harpsichords in
+the Trasuntino style, with un-pitched strings under tension to provide a
+comparable mechanical resistance as found on an original. A system of infra-red
+LED / phototransistor pairs was connected to a microcontroller and installed
+inside the interface from which a stream of data can be generated for each jack.
+The paper covers the design criteria, process, open source development, and
+execution of installation for exhibition at San Colombano.
 \end{abstract}
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
-  Your, Paper}
+\keywords{interactive exhibit, augmented instrument, optical sensing, harpsichord}
+
 %% A "teaser" image appears between the author and affiliation
@@
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{images/side_w-o_sensors.jpeg}
+  \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
   \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}

### 2025-01-17 nime-paper-template.tex

%%
%% This is file `sample-nime-paper.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-nime-paper.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% History of NIME Template Updaters:
%% Modified by Florent Berthaut, Charles Martin, Yichen Wang, 21 November 2024
%% Modified by Adnan Marquez-Borbon 30 November 2022
%% Modified by Courtney Reed 28 November 2022
%% Modified by Joe Wright 14 December 2019
%% Modified by Niccolò Granieri 10 October 2018
%% Modified by Angelo Fraietta 23 December 2018
%% Modified by Angelo Fraietta 22 November 2018
%% Modified by Rodrigo Schramm on 22 September 2018
%% Modified by Luke Dahl on 17 October 2-17
%% Modified by Cumhur Erkut on <2016-10-11 Tue>
%% Modified by Edgar Berdahl on 5 November 2014
%% Modified by Baptiste Caramiaux on 25 November 2013
%% Modified by Kyogu Lee on 7 October 2012
%% Modified by Georg Essl on 7 November 2011
%%
%% And here's the template:
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{nimeart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{nimeart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,anonymous,review]{nimeart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{cc}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{}
%% These commands are for a PROCEEDINGS abstract or paper.
%% Make sure this is up to date with the correct edition of NIME.
\acmConference[NIME '25]{International Conference on New Interfaces for Musical Expression}{June 24--27,
  2025}{Canberra, Australia}
%% This suppresses the ACM Reference Format printing.
\settopmatter{printacmref=false}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The NIME of the Title Is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \city{San Antonio}
  \state{Texas}
  \country{USA}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. This uses ``nimeart'', a fork of the ``acmart''
  documentclass with additions for the NIME conference. This
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{images/sampleteaser.pdf}
  \caption{Two synthesisers on a table, 2019.}
  \Description{Enjoying looking at two synthesisers.
  A Korg Monotron and a Korg Volca FM.}
  \label{fig:teaser}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|nimeart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|nimeart|'' document class can
be used to prepare many different kinds of documentation --- a
double-anonymous initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|nimeart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{nimeart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\texttt{acmsmall}}: The default journal template style.
\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
\item {\texttt{acmtog}}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\texttt{sigconf}}: The default proceedings template style.
\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \texttt{\string\acmSubmissionID} command to print the
  submission's unique ID on each page of the work.
\item{\texttt{authorversion}}: Produces a version of the work suitable
  for posting by the author.
\item{\texttt{screen}}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{nimeart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|nimeart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification.  As an exception, multiple authors may share one
affiliation. Authors' names should not be abbreviated; use full first
names wherever possible. Include authors' e-mail addresses whenever
possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Anonymisation and Blind Review}
Anonymisation is a \textbf{crucial} component for NIME paper submissions. We kindly ask authors to ensure that all information that could introduce bias in the review process is anonymised.

To anonymise author names, please specify \verb|anonymous,review| in square brackets when submitting your paper.
\begin{verbatim}
  \documentclass[sigchi,anonymous,review]{nimepaper}
\end{verbatim}

You can also hide text in the paper like this:

\begin{verbatim}
% Both examples from acmart manual.
\begin{anonsuppress}
  This is the continuation of the previous work by the author
  \cite{prev1, prev2}.
\end{anonsuppress}
\end{verbatim}

% Both examples from acmart manual.
Begin anonymous thing
\begin{anonsuppress}
  This is the continuation of the previous work by the author
\end{anonsuppress}
End anonymous thing
or
\begin{verbatim}
This work was performed at \anon{NSA}.
This work was performed at \anon[No Such Agency]{NSA}.
\end{verbatim}

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``nimeart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
\verb|\paragraph|, and \verb|\subparagraph|. The sectioning levels up to
\verb|\subsusection| should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

Below are examples of sectioning commands.

\subsection{Subsection}
\label{sec:subsection}

This is a subsection.

\subsubsection{Subsubsection}
\label{sec:subsubsection}

This is a subsubsection.

\paragraph{Paragraph}

This is a paragraph.

\subparagraph{Subparagraph}

This is a subparagraph.

\section{Tables}

The ``\verb|nimeart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{images/sample-bela.jpg}
  \caption{A Bela board in action at a workshop (2017).}
  \Description{Somebody manipulating sensors on a Bela board.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
  Artifacts: \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

%% The Ethical Standards section is mandatory for all NIME submissions.
\section{Ethical Standards}

To ensure objectivity and transparency in research and to ensure that accepted principles of ethical and professional conduct have been followed, authors must include a section “Ethical Standards” before the References.
This section should include (if relevant): information regarding sources of funding, potential conflicts of interest (financial or non-financial), informed consent if the research involved human participants, statement on welfare of animals if the research involved animals or any other information or context that helps ethically situate your research.
For help with the ethics section, feel free to ask on the NIME forum: \url{https://forum.nime.org}.

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-references}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-nime-paper.tex'.

### 2025-01-17 src/main.tex
abstract

@@
 %%
-\documentclass[sigconf]{nimeart}
+\documentclass[sigconf,anonymous]{nimeart}
 %%
@@
 \begin{abstract}
-Presented is a design for an electronically augmented replica of a 16th century
-harpsichord interface, in the style of a 1547 Trasuntino harpsichord. The
-interface was designed, fabricated, and exhibited in a collaboration between the
-NEMUS project at the Università di Bologna and the Museo San Colombano, which
-hosts the Tagliavini Collection of playable historical keyboard instruments. The
-interface was commissioned as part of an exhibition and as a means of
-interacting with items in the collection that are no longer in playing
-condition. The keyboard is a replica of the mechanism found in harpsichords in
-the Trasuntino style, with un-pitched strings under tension to provide a
-comparable mechanical resistance as found on an original. A system of infra-red
-LED / phototransistor pairs was connected to a microcontroller and installed
-inside the interface from which a stream of data can be generated for each jack.
-The paper covers the design criteria, process, open source development, and
-execution of installation for exhibition at San Colombano.
+This work describes the design of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout.
+The keyboard is a collaboration between the authors and the Museo di San Colombano, which houses the Tagliavini Collection of playable historical keyboard instruments.
+The keyboard was commissioned for exhibition purposes and to enhance the experiential authenticity of the collection.
+The keyboard emulates the mechanics of a two-register harpsichord featuring a single manual.
+Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass.
+An array of infrared LED/phototransistor pairs, connected to a microcontroller, was installed within the interface to generate a continuous stream of data corresponding to the jacks' motion.
+This paper details the design principles, development process, open-source implementation, and realisation of the installation for the San Colombano exhibition.
 \end{abstract}
@@
 %% consistent spelling of the heading.
+
+\section{Ethical Standards}\label{ethical-standards}
+
+% \begin{anonsuppress}
+
+This work was performed at \anon{San Colombano}
+To ensure objectivity and transparency in research and to ensure that
+accepted principles of ethical and professional conduct have been
+followed, authors must include a section ``Ethical Standards'' before
+the References. This section should include (if relevant): information
+regarding sources of funding, potential conflicts of interest (financial
+or non-financial), informed consent if the research involved human
+participants, statement on welfare of animals if the research involved
+animals or any other information or context that helps ethically situate
+your research. For help with the ethics section, feel free to ask on the
+NIME forum: \url{https://forum.nime.org}.
+% \end{anonsuppress}
+
 \begin{acks}
-A project of this breadth developed across 3 institutes and 2 countries
-cannot be carried without the support and technical knowledge of many
-skilled and passionate people.
 
+% \begin{anonsuppress}
+% This is the continuation of the previous work by the author
+% A project of this breadth developed across 3 institutes and 2 countries
+% cannot be carried without the support and technical knowledge of many
+% skilled and passionate people.
+% \end{anonsuppress}
+
+% \begin{anonsuppress}
+% A project of this breadth developed across 3 institutes and 2 countries
+% cannot be carried without the support and technical knowledge of many
+% skilled and passionate people.
+% \end{anonsuppress}
+
+\begin{anonsuppress}
 We acknowledge the invaluable contributions of
@@
 \end{itemize}
+
+\end{anonsuppress}
 \end{acks}

### 2025-01-17 src/main.tex
move acknowledgments to main
rm tightlist

@@
     Bib\TeX}}}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 %% consistent spelling of the heading.
+\begin{acks}
+A project of this breadth developed across 3 institutes and 2 countries
+cannot be carried out without the support and technical knowledge of many
+skilled and passionate people.
 
-\section{Ethical Standards}\label{ethical-standards}
-
-% \begin{anonsuppress}
-
-This work was performed at \anon{San Colombano}
-To ensure objectivity and transparency in research and to ensure that
-accepted principles of ethical and professional conduct have been
-followed, authors must include a section ``Ethical Standards'' before
-the References. This section should include (if relevant): information
-regarding sources of funding, potential conflicts of interest (financial
-or non-financial), informed consent if the research involved human
-participants, statement on welfare of animals if the research involved
-animals or any other information or context that helps ethically situate
-your research. For help with the ethics section, feel free to ask on the
-NIME forum: \url{https://forum.nime.org}.
-% \end{anonsuppress}
-
-\begin{acks}
-
-% \begin{anonsuppress}
-% This is the continuation of the previous work by the author
-% A project of this breadth developed across 3 institutes and 2 countries
-% cannot be carried without the support and technical knowledge of many
-% skilled and passionate people.
-% \end{anonsuppress}
-
-% \begin{anonsuppress}
-% A project of this breadth developed across 3 institutes and 2 countries
-% cannot be carried without the support and technical knowledge of many
-% skilled and passionate people.
-% \end{anonsuppress}
-
-\begin{anonsuppress}
 We acknowledge the invaluable contributions of
@@
 \begin{itemize}
-\tightlist
 \item
   Unibo
-
   \begin{itemize}
-  \tightlist
   \item
     Alma Labor
-
     \begin{itemize}
-    \tightlist
     \item
@@
     Fisica Tecnica
-
     \begin{itemize}
-    \tightlist
     \item
@@
   San Colombano
-
   \begin{itemize}
-  \tightlist
   \item
@@
   Edinburgh
-
   \begin{itemize}
-  \tightlist
   \item
     uCreate
-
     \begin{itemize}    
-    \tightlist
     \item
@@
   \item
-    ECA
-
+    University of Edinburgh
     \begin{itemize}
-    \tightlist
     \item
@@
 
-\end{anonsuppress}
 \end{acks}

### 2025-01-18 src/main.tex
format abstract

@@
 %%
-\documentclass[sigconf,anonymous]{nimeart}
+\documentclass[sigconf]{nimeart}
 %%
@@
 \begin{abstract}
-This work describes the design of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout.
-The keyboard is a collaboration between the authors and the Museo di San Colombano, which houses the Tagliavini Collection of playable historical keyboard instruments.
-The keyboard was commissioned for exhibition purposes and to enhance the experiential authenticity of the collection.
-The keyboard emulates the mechanics of a two-register harpsichord featuring a single manual.
-Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass.
-An array of infrared LED/phototransistor pairs, connected to a microcontroller, was installed within the interface to generate a continuous stream of data corresponding to the jacks' motion.
-This paper details the design principles, development process, open-source implementation, and realisation of the installation for the San Colombano exhibition.
+This work describes the design of an electronically augmented replica of a
+16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
+keyboard is a collaboration between the authors and the Museo di San Colombano,
+which houses the Tagliavini Collection of playable historical keyboard
+instruments. The keyboard was commissioned for exhibition purposes and to
+enhance the experiential authenticity of the collection. The keyboard emulates
+the mechanics of a two-register harpsichord featuring a single manual. Unpitched
+strings under tension replicate the mechanical resistance characteristic of the
+original compass. An array of infrared LED/phototransistor pairs, connected to a
+microcontroller, was installed within the interface to generate a continuous
+stream of data corresponding to the jacks' motion. This paper details the design
+principles, the development, open source implementation, and execution of
+the installation for the San Colombano exhibition.
 \end{abstract}
@@
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
-  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
-  \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
+  \includegraphics[width=\textwidth]{images/49-key-front.png}
+  \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+  \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
   \label{fig:teaser}
 \end{teaserfigure}
+
+% \begin{teaserfigure}
+%   \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
+%   \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
+%   \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
+%   \label{fig:teaser}
+% \end{teaserfigure}
 

### 2025-01-18 src/article.tex
Update article.tex
Updates from Overleaf
rm tightlist

@@
-\section{Abstract}\label{abstract}
+\section{Introduction}\label{introduction}
 
@@
 (from Luigi Ferdinando Tagliavini), a collection of approximately 70
-historical keyboard instruments in working order
-\textbackslash cite\{Tagliavini2009\}
+historical keyboard instruments in working order \cite{Tagliavini2007}
 (https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
@@
 instruments in the Tagliavini collection is a 1547 Harpsichord by
-Alessandro Trasuntino \textbackslash cite\{Treccani2019\}, which was
-used as the basis for an interactive exhibition commisioned by San
-Colombano.
+Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
+for an interactive exhibition commisioned by San Colombano.
 
@@
 project can be improved and developed further.
-
-\section{Introduction}\label{introduction}
-
-With the advent of digital musical instruments in the 20th Century there
-has been a decoupling between interface and sound generation when
-playing. Consequently, aspects of the sensory experience of playing have
-been lost, namely the sensation of vibration and mechanical impedance
-that a traditional musical instrument provides.
 
@@
 develop further the methodologies implemented in musical haptics
-projects {[}3{]} such as {[}2{]}. The final outcome will be to create a
-framework for digital conservation of historic musical instruments such
-as the 1638 Ioannes Ruckers and the 1646 Ioannes Couchet double-manual
-harpsichords as referenced in the NEMUS project {[}4{]}.
+projects \cite{MusicalHaptics} such as \cite{Timmermans2020}.
 
@@
 \begin{itemize}
-\tightlist
 \item
@@
   \begin{itemize}
-  \tightlist
-  \item
-    Yamaha Disklaviare
-  \item
-    Bosendorfer
-
-    \begin{itemize}
-    \tightlist
+  \item
+    Yamaha Disklavier used as a basis for exploration into haptics and
+    the piano keyabord
+    \cite{MusicalHaptics2018_04, MusicalHaptics2018_05,
+    MusicalHaptics2018_13}
+  \item
+    Bosendorfer: LED photo transistor pair \cite{Moog1990} for obtaining
+    performance data \cite{MusicalHaptics2018_05}
+
+    \begin{itemize}
     \item
@@
   \item
-    PNOscan MIDI 9 QRS Music
-
-    \begin{itemize}
-    \tightlist
+    PNOscan MIDI 9 QRS Music as MIDI Piano system \cite{McPherson2013}
+
+    \begin{itemize}
     \item
@@
 
-The paper by McAlpine \textbackslash cite\{McAlpine2014\} also has very
-similar motivations to the project presented here. Both were started
-from a musical instrument museum, housing a collection of historic
-instruments gifted by a collector (Benton Fletcher / Taglivaini) who
-wished for collection to remain playable, which is restricted growing
-fragility of some items in the respective collections. Bothe projects
-center around the creation of a custioom MIDI controller for interfacing
-with sound synthesis engine. The keyboard consisted of two modified
-commercial MIDI keyboard controllers in a two manual harpsichord
-arrangement. Weherease the McAlpine keyboard used weighted keys, this
-interface extends the idea by using the mechanism and strings under
-tension to provide the haptic feedback. McAlpine posits a fully
-paremterised haptic keyboard, but before taking such a leap we look at
-what affect a more literal haptic keyboard may provide to the performer
-experience.
-
-The previous NIME project ``Tromba Moderna''
-\textbackslash cite\{Baldwin2016\}, which looked at reconstructing a
-full tromba marina to be used as a digital interface. The one of the
-limitations flagged by McAlpine was the oneffectiveness of weighted keys
-in mimicing the harpischord. The Tromba Moderna ``attempted to avoid
-this problem by constructing a fully functioning replica of the tromba
-marina'' \textbackslash cite\{Baldwin\}. This project applies the same
-approach to the harpsichord.m where the projects deviate is a questions
-of scale.
+The paper by McAlpine \cite{McAlpine2014} also has very similar
+motivations to the project presented here. Both were started from a
+musical instrument museum, housing a collection of historic instruments
+gifted by a collector (Benton Fletcher / Taglivaini) who wished for
+collection to remain playable, which is restricted growing fragility of
+some items in the respective collections. Bothe projects center around
+the creation of a custioom MIDI controller for interfacing with sound
+synthesis engine. The keyboard consisted of two modified commercial MIDI
+keyboard controllers in a two manual harpsichord arrangement. Weherease
+the McAlpine keyboard used weighted keys, this interface extends the
+idea by using the mechanism and strings under tension to provide the
+haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
+but before taking such a leap we look at what affect a more literal
+haptic keyboard may provide to the performer experience.
+
+The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
+looked at reconstructing a full tromba marina to be used as a digital
+interface. The one of the limitations flagged by McAlpine was the
+oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
+Moderna ``attempted to avoid this problem by constructing a fully
+functioning replica of the tromba marina'' \cite{Baldwin2016}. This
+project applies the same approach to the harpsichord.m where the
+projects deviate is a questions of scale.
 
@@
 
-Extension to \textbackslash cite\{McAlpine2014\} and
-\textbackslash cite\{Baldwin2016\} are the open sourcing of hardware and
-software for the project.
+Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
+sourcing of hardware and software for the project.
 
@@
   \begin{itemize}
-  \tightlist
   \item
@@
 \item
-  parts
-
-  \begin{itemize}
-  \tightlist
+  parts of harpsichord
+
+  \begin{itemize}
   \item
@@
     \begin{itemize}
-    \tightlist
     \item
@@
 \begin{itemize}
-\tightlist
-\item
-  \textbackslash cite\{HapticKey\}
-\item
-  \textbackslash cite\{mcPherson1\} \textbackslash cite\{mcPherson2\}
-\item
-  \textbackslash cite\{Mudd2013\}
-\item
-  \textbackslash cite\{FritzClaudia2017Leon\}
+\item
+  \cite{Timmermans2020}
+\item
+  \cite{McPherson2013} \cite{McPherson2019}
+\item
+  \cite{Mudd2013}
+\item
+  \cite{Fritz2017}
 \end{itemize}
@@
 
-Following from the work in \textbackslash cite\{McPherson1\}
-\textbackslash cite\{McPherson2\} we tested a system using the Fairchild
-QRE1113 \textbackslash datasheetfigue. The QRE1113 is a combination
-infrared LED and phototransistor sensitive to IR light. The
-phototransistor of QRE1113 provides data on how much light is present
-and in particular how much light is being relfected from a surface.
-Since refelected light will be proportional to the distance of a surfcae
-it is a good, close proximity, distance sensor. Distance is the way in
-which the instruments in \textbackslash citeMcpherson1\&2 function as
-well the Moog Piano Bar from which it too inspiration.
+Following from the work in \cite{McPherson2013, McPherson2019} we tested
+a system using the Fairchild QRE1113 \textbackslash datasheetfigue. The
+QRE1113 is a combination infrared LED and phototransistor sensitive to
+IR light. The phototransistor of QRE1113 provides data on how much light
+is present and in particular how much light is being relfected from a
+surface. Since refelected light will be proportional to the distance of
+a surfcae it is a good, close proximity, distance sensor. Distance is
+the way in which the instruments in \cite{McPherson2013, McPherson2019}
+function as well the Moog Piano Bar from which it too inspiration.
 
@@
 The LEDs are on the same PCB as the sensors and are hidden when the
-nameboard is in place.
+nameboard is in place.2024
 
@@
 
-TLC5940 or other PWM driver such as in \textbackslash cite\{mcpherson\}
-added complexity to assembling, but were avoid in favour of Addressable
-LEDs with intergrated driver. The intergrated LEDs share the same data
-line which has the limitation where a fualty LED will stop all following
-LEDs from working. An individual PWM driver per PCB would allow avoid
+TLC5940 or other PWM driver such as in \cite{McPherson2013} added
+complexity to assembling, but were avoid in favour of Addressable LEDs
+with intergrated driver. The intergrated LEDs share the same data line
+which has the limitation where a fualty LED will stop all following LEDs
+from working. An individual PWM driver per PCB would allow avoid
 situations where failutre cascades, but would add complexity in assembly
@@
 Transistor / Op-amp circuit could have been used similar to the one in
-\textbackslash cite\{mcpherson\}.
+\cite{McPherson2013}.
 
@@
       \begin{itemize}
-      \tightlist
       \item
@@
 \end{itemize}
-
-\section{Acknowledgments}\label{acknowledgments}
-
-A project of this breadth developed across 3 institutes and 2 countries
-cannot be carried without the support and technical knowledge of many
-skilled and passionate people.
-
-We acknowledge the invaluable contributions of
-
-\begin{itemize}
-\tightlist
-\item
-  Unibo
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Alma Labor
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
-      of bologna for their 3D printing expertise and generosity with
-      electronics equipment
-    \end{itemize}
-  \item
-    Fisica Tecnica
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
-      Costantini
-    \end{itemize}
-  \end{itemize}
-\item
-  San Colombano
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Catalina Vincens
-  \item
-    Roberto Livi
-  \end{itemize}
-\item
-  Edinburgh
-
-  \begin{itemize}
-  \tightlist
-  \item
-    uCreate
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
-      Edinburgh for their support with soldering facilities.
-    \end{itemize}
-  \item
-    ECA
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Richard Collins and the Digital Making Staff at the Edinburgh
-      College of Art for 3D printing instruction and CNC support.
-    \item
-      Joe Hathaway of Digital Development at the College of Art for
-      providing electronics components
-    \item
-      St Cecilia's Jenny Nex
-    \end{itemize}
-  \end{itemize}
-\end{itemize}

### 2025-01-18 src/main.tex

@@
 \item
-  Unibo
+  Università di Bologna
   \begin{itemize}
@@
     \item
-      Massimiliano Fraulini and the staff of Alma Labor at the Yiversità
-      of bologna for their 3D printing expertise and generosity with
+      Massimiliano Fraulini and the staff of Alma Labor at the Università
+      di Bologna for their 3D printing expertise and generosity with
       electronics equipment

### 2025-01-18 src/article.tex
footnote urls

@@
 historical keyboard instruments in working order \cite{Tagliavini2007}
-(https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
-items in the collection are not in playing condition or are too fragile
-to continue to be available for the public to play. One of those fragile
-instruments in the Tagliavini collection is a 1547 Harpsichord by
-Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
+\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}.
+Some items in the collection are not in playing condition or are too
+fragile to continue to be available for the public to play. One of those
+fragile instruments in the Tagliavini collection is a 1547 Harpsichord
+by Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
 for an interactive exhibition commisioned by San Colombano.
@@
 
-https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf
+\footnote{https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf}
 
@@
   ``Native USB''
-  https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
+  \footnote{https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards}
 \item
@@
 crimping cables by hand. A crimping tool from JST is prohibitively
-expensive compared to the cost of other components in the project*. The
-crimping tool also appeared to be uncommon and was not readily available
-in the workshops at the projects disposal. Eventually a tool was found,
-but an alternative approach was also researched in order to maintain the
-criteria of accessibility.
+expensive compared to the cost of other components in the project
+\footnote{from RS https://uk.rs-online.com/web/p/crimp-tools/6880877
+£508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
+not readily available in the workshops at the projects disposal.
+Eventually a tool was found, but an alternative approach was also
+researched in order to maintain the criteria of accessibility.
 
 A locator was adapted from a creative commons stl model
-https://www.thingiverse.com/thing:1646016 that allowed a generic
-crimping tool to be adapted for use with JST-PH crimps†
+\footnote{https://www.thingiverse.com/thing:1646016} that allowed a
+generic crimping tool to be adapted for use with JST-PH crimps†
 \textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
-printer.
-
-\emph{** A second alterntaive was assembling cables from other sets that
-were at hand.} \emph{* from RS
-https://uk.rs-online.com/web/p/crimp-tools/6880877 £508 inc VAT} \emph{†
-model available in the git repository}
+printer. A second alterntaive was assembling cables from other sets that
+were at hand.
 
@@
 the Arduino IDE and MIDI was confirmed using the open source MIDI
-Monitor (https://github.com/krevis/MIDIApps).
+Monitor \footnote{https://github.com/krevis/MIDIApps}.
 

### 2025-01-18 src/article.tex
update related work

@@
 
-The project look at restoring tactile and haptic response of 17th
-century harpsichords for an exhibition at the musical instrument museum
-San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
-(from Luigi Ferdinando Tagliavini), a collection of approximately 70
-historical keyboard instruments in working order \cite{Tagliavini2007}
-\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}.
-Some items in the collection are not in playing condition or are too
-fragile to continue to be available for the public to play. One of those
-fragile instruments in the Tagliavini collection is a 1547 Harpsichord
-by Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
-for an interactive exhibition commisioned by San Colombano.
+The project look at restoring tactile and haptic response of historical
+harpsichords for an exhibition at the musical instrument Museo di San
+Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
+in Bologna. San Colombano is host to the Tagliavini Collection, a
+collection of approximately 70 historical keyboard instruments in
+working order \cite{Tagliavini2007} gifted to the museum by\\
+from Luigi Ferdinando Tagliavini in 2010
+\cite{SanColombano2010,Carlino2010}.
+
+Some items among the Tagliavini collection are no longer in a playing
+condition or are too fragile to continue to be available by museum
+visitors. One of those fragile instruments is a 1547 Harpsichord in the
+italian style by Alessandro Trasuntino
+\cite{Wraight2024}\footnote{examples of similar instruments by
+Trasuntino available online in the [Royal College of Music Collection
+(1531)](https://museumcollections.rcm.ac.uk/collection/Details/collect/58) and
+the [Musée de la musique collection
+(1538)](https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView)
+}, which was used as the basis for an interactive exhibition commisioned
+by San Colombano.
 
@@
 
+\subsection{Motivations}\label{motivations}
+
+The project was carried out in collaboratoiun with San Colombano. The
+results of the project would be presented as an interactive exhibition
+for the public.
+
 \begin{itemize}
@@
 
-\section{Background}\label{background}
-
-Prvious projects looking at how to approach the problems of increasing
-engagement and retaining the interction with historical musical
-instruments in a museum scenario.
-
-The paper by McAlpine \cite{McAlpine2014} also has very similar
-motivations to the project presented here. Both were started from a
-musical instrument museum, housing a collection of historic instruments
-gifted by a collector (Benton Fletcher / Taglivaini) who wished for
-collection to remain playable, which is restricted growing fragility of
-some items in the respective collections. Bothe projects center around
-the creation of a custioom MIDI controller for interfacing with sound
-synthesis engine. The keyboard consisted of two modified commercial MIDI
-keyboard controllers in a two manual harpsichord arrangement. Weherease
-the McAlpine keyboard used weighted keys, this interface extends the
-idea by using the mechanism and strings under tension to provide the
-haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
-but before taking such a leap we look at what affect a more literal
-haptic keyboard may provide to the performer experience.
-
-The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
-looked at reconstructing a full tromba marina to be used as a digital
-interface. The one of the limitations flagged by McAlpine was the
-oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
-Moderna ``attempted to avoid this problem by constructing a fully
-functioning replica of the tromba marina'' \cite{Baldwin2016}. This
-project applies the same approach to the harpsichord.m where the
-projects deviate is a questions of scale.
-
-Where the tromba moderna considers a single data stream this project has
-had to address 98 and the problems that come with increasing scale.
+\section{Related Work}\label{related-work}
+
+Methodologies for increasing visitor engagmemnet or
+
+Museum staff want visitors to be engaged but limits in staffing and
+funding limit how vistors can interact with collections
+\cite{Templeton2018,
+McAlpine2014}. For musical instrument museums there is an additional
+difficulty in the balance of keeping historical instruments playable
+\cite{McAlpine2013}. The fragility and decay of musical instruments
+means that there is an inveitable point where instruments will no longer
+be in a playable condition \cite{McAlpine2014, Fritz2017}
+
+McAlpine outlines a similar situation to the Tagliavini collection in
+his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
+was a stipulation by Benton Fletcher that the instruments remain
+available to play when they were gifted to National Trust Fenton House.
+A custom MIDI interface arranged in a two manual harpischord style which
+triggered samples of the instruments in the collection using an
+appropraite recording strategy for each. One problem highlighted in user
+tests by McAlpine is that the weighted keys did no provide ``an
+experiential sense of interacting with a historical keyboard''
+\cite{McAlpine2014}. The MIDI interface created from commercially
+available weighted keys. McAlpine posits a haptic keyboard augemented
+with ``actuators to provide positionally-sensitive real-time force
+feedback at point-of-contact'' in a similar approach as the piano
+mechanism by Gillespie \cite{Gillespie1996}.
+
+Previous NIME project ``Tromba Moderna'' \cite{Baldwin2016} looked to
+avoid the complex engineering problem by simply recreating the tromba
+marina and augmenting it. In the case of the Tromba Modern, a piezo
+transducer was connected to a sound synthesis engine connected to a
+driver inside the instrument to simulate the vibration that would be
+expected of an historical tromba marina.
+
+The original design was to take a similar approach to the project by
+Timmermans \cite{Timmermans2020}, which was an extension of the system
+by Gillespie \cite{Gillespie1996}. The process laid out in Timmermans
+project was to begin with a single key model of the mechanism. We began
+with a 3-key Model Harpsichord Mechanism by Graziano Bandini
+\textbackslash figure\{\}. Before we continued with augementing a model
+with actuators it was decided that it we should first validate whether a
+recreation of the harpsichord mechanism would suffice in allowing
+visitors to suspend disbelief.
+
+Benefits of putting effort into the maintenance of one interface with a
+transferable sensor system
+
+An additonal factor not faced in the Tromba Moderna project is the
+problem of scale in managing 98 rather than a single data stream of
+data.
+
 Material costs mitigated by creating a scaled version of the sound
-board, provinding only the length of string necessary under comparable
+board, providing only the length of string necessary under comparable
 tension and thus resistance to the jack quill.
 
-Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
-sourcing of hardware and software for the project.
-
-\subsection{Why\ldots{}}\label{why}
-
-\subsubsection{\ldots a replica?}\label{a-replica}
-
-\subsubsection{\ldots an interface?}\label{an-interface}
-
-\begin{itemize}
-\item
-  maintain problems
-\item
-  probe for research
-
-  \begin{itemize}
-  \item
-    on haptics how it affects performance
-  \end{itemize}
-\end{itemize}
-
-\subsection{Motivations}\label{motivations}
-
-The project was carried out in collaboratoiun with San Colombano. The
-results of the project would be presented as an interactive exhibition
-for the public.
-
-\subsubsection{the problem}\label{the-problem}
+Finally, a further extension to the projects above was a commmitement to
+open sourcing all aspects of the project, hardware, software, and data.
+
+\section{Design}\label{design}
+
+\subsection{Design Problem}\label{design-problem}
 
@@
 
-\subsubsection{the solution}\label{the-solution}
-
 The approach taken was to recreate the interface of these instruments
@@
 
-\subsubsection{Criteria}\label{criteria}
-
-A keyboard interface that provides a comparable physical response to a
-harpsichord. Notes should be assigned per jack and not per key.
+The design criteria was to create a keyboard interface that provides a
+comparable physical response to a harpsichord.
+
+\subsubsection{Design Constraints}\label{design-constraints}
+
+\begin{itemize}
+\item
+  no electronics should be visible.
+\item
+  it should be robust and reliable, easily maintained by museum staff
+  without complex technical processes
+\item
+  it should not augment or fix the mechanics, but present them with
+  their original limitation
+\end{itemize}
+
+Another constraint was that the design of the sensor system and the
+model keyboard would have to carried out at a distance in separate
+workshops. Time constraints meant the keyboard was being fabricated for
+a sensor that did not yet exist.
+
+Accomodations were made in the design to allow for as much flexibility
+as possible. This also meant that designs for the eletronics would have
+to include a deal of flexibility in how they might be installed. Surplus
+space was added into the design of the internal chamber of the model
+\textbackslash FIGURE, but there was still a limit given the structural
+and aesthatic requirements.
 
@@
 
-\subsubsection{Design}\label{design}
+\subsubsection{Keyboard Construction}\label{keyboard-construction}
 
@@
 
+\subsection{Prototyping}\label{prototyping}
+
 \subsubsection{3-Key Prototype}\label{key-prototype}
@@
 \end{itemize}
-
-\subsubsection{Fig: Construction Design
-Diagram}\label{fig-construction-design-diagram}
 
@@
 \end{itemize}
-
-\subsubsection{Fig: Construction Design
-Diagram}\label{fig-construction-design-diagram-1}
-
-\subsubsection{Design Constraints}\label{design-constraints}
-
-\begin{itemize}
-\item
-  no electronics should be visible.
-\item
-  it should be robust and reliable, easily maintained by museum staff
-  without complex technical processes
-\item
-  it should not augment or fix the mechanics, but present them with
-  their original limitation
-\end{itemize}
-
-Another constraint was that the design of the sensor system and the
-model keyboard would have to carried out at a distance in separate
-workshops. Time constraints meant the keyboard was being fabricated for
-a sensor that did not yet exist.
-
-Accomodations were made in the design to allow for as much flexibility
-as possible. This also meant that designs for the eletronics would have
-to include a deal of flexibility in how they might be installed. Surplus
-space was added into the design of the internal chamber of the model
-\textbackslash FIGURE, but there was still a limit given the structural
-and aesthatic requirements.
 
@@
 implemenet functionality for velocity and aftertouch MIDI messages and
-simply needs to be activated in the firmware.
-
+simply needs to be activated in the firmware. (Hyperinstruments
+\cite{nime2024_20})
+
+seat vibrations \cite{MusicalHaptics2018_07}
+
+\begin{itemize}
+\item
+  ``there is a general connection between vibrations and the perceived
+  quality of music reproduction''
+\item
+  ``influence of vibrations on loudness perception at low frequencies''
   Evaluation
+\end{itemize}
 

### 2025-01-18 src/article.tex
urls

@@
 \cite{Wraight2024}\footnote{examples of similar instruments by
-Trasuntino available online in the [Royal College of Music Collection
-(1531)](https://museumcollections.rcm.ac.uk/collection/Details/collect/58) and
-the [Musée de la musique collection
-(1538)](https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView)
+Trasuntino available online in the Royal College of Music Collection
+(1531: \url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and
+the Musée de la musique collection
+(1538: \url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
 }, which was used as the basis for an interactive exhibition commisioned
@@
 develop further the methodologies implemented in musical haptics
-projects \cite{MusicalHaptics} such as \cite{Timmermans2020}.
+projects \cite{MusicalHaptics2018} such as \cite{Timmermans2020}.
 
@@
 difficulty in the balance of keeping historical instruments playable
-\cite{McAlpine2013}. The fragility and decay of musical instruments
+\cite{McAlpine2014}. The fragility and decay of musical instruments
 means that there is an inveitable point where instruments will no longer

### 2025-01-18 src/article.tex
replace nums with cites

@@
 this experience? This is central problem that is trying to be solved in
-{[}1, 2{]}.
+\cite{Nichols2002, Timmermans2020, McAlpine2014, Baldwin2016}.
 
@@
 NIME forum: \url{https://forum.nime.org}.
-
-\subsection{Example Trasuntino}\label{example-trasuntino}
-
-\begin{itemize}
-\item
-  {[}1538{]}
-  https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface\&module=collection\&objectId=106088\&viewType=detailView
-\item
-  {[}1531{]}https://museumcollections.rcm.ac.uk/collection/Details/collect/58
-\end{itemize}

### 2025-01-18 src/article.tex
test figure

@@
 \cite{Wraight2024}\footnote{examples of similar instruments by
-Trasuntino available online in the Royal College of Music Collection
-(1531: \url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and
-the Musée de la musique collection
-(1538: \url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
+Trasuntino available online in the Royal College of Music Collection (1531:
+\url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
+Musée de la musique collection (1538:
+\url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
 }, which was used as the basis for an interactive exhibition commisioned
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna
+  2023. A sidevew of the model which is a cross-section of what is a common
+mechanism for an italian style harpischord of the XVI century.} 
+\end{figure}
+
 \subsubsection{49-Key Prototype}\label{key-prototype-1}

### 2025-01-18 src/article.tex
start polishing introduction

@@
 
-The project look at restoring tactile and haptic response of historical
+The project looks at restoring tactile and haptic response of historical
 harpsichords for an exhibition at the musical instrument Museo di San
@@
 
-Project was also designed to allow others to easily recreate and iterate
-on the designs.
-
-First a small cross-sectional model was used as a proof of concept. Once
-a sensing system was decided upon, a larger 49-key scale model was
-commissioned with and then built by luthier Roberto Livi. The models are
-designed to provide only a tactile and haptic response for the player. A
-sensor system was installed into the model and messages were sent to
-computer that triggered sounds from a sample library which the player
-listens to through headphones. The paper looks at the design decisions,
+The project was also open sourced to allow for others to easily recreate
+and iterate on the designs.
+
+The haptic key project by Timmermans et al.~\cite{Timmermans2020} began
+with the model of a piano key. A 3-key cross-sectional model of a
+harpischord mechanism \figure{3key} was provided to the NEMUS project by
+San Colombano as a test bed for sensors and as a proof of concept for a
+larger scale interface. Each key of model has 2 sets of strings that are
+under tension but not at any specific musical pitch. The strings are
+there to provide same resistance to the jack plectra as would be found
+on an instrument. The mechanism is in the italian style of the XVI-XIX
+century with 2 jacks per key with corresponding strings tuned in unison
+\footnote{"8'-8' È la dispozione normale dei clavicembali
+italiani" \cite{Tagliavini1987}}. A suitable sensing system was found
+for reliably detecting when jack had plucked a string independently and
+sending out a corresponding MIDI message to a connected sound synthesis
+engine. A full scale 49-key scale model was commissioned with and then
+built by luthier Roberto Livi in the same style as the 1547 Trasuntino
+\figure{teaser}. The sensor system was installed into the full scale
+model and then setup for exhibition at San Colombano. Visitors are
+invited to play the interface which currently triggers samples on a
+connected sound engine. The paper looks at the design decisions,
 processes, and considerations to be made when undertaking a project of
 this kind.
+
+\begin{figure}
+  \label{}
+  \centering
+  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna.
+A sidevew of the model which is a cross-section of what is a common
+mechanism for an italian style harpischord of the XVI century.} 
+\label{fig:3key}
+\end{figure}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
-  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna
-  2023. A sidevew of the model which is a cross-section of what is a common
-mechanism for an italian style harpischord of the XVI century.} 
-\end{figure}
-
 \subsubsection{49-Key Prototype}\label{key-prototype-1}

### 2025-01-18 src/article.tex
double label
fig to ref

@@
 with the model of a piano key. A 3-key cross-sectional model of a
-harpischord mechanism \figure{3key} was provided to the NEMUS project by
+harpischord mechanism \ref{3key} was provided to the NEMUS project by
 San Colombano as a test bed for sensors and as a proof of concept for a
@@
 built by luthier Roberto Livi in the same style as the 1547 Trasuntino
-\figure{teaser}. The sensor system was installed into the full scale
+\ref{teaser}. The sensor system was installed into the full scale
 model and then setup for exhibition at San Colombano. Visitors are
@@
 \begin{figure}  
-  \label{}
   \centering

### 2025-01-18 src/article.tex
fix refs

--- 2025-01-18 src/article.tex
@@
 
-With the advent of digital musical instruments in the 20th Century there
-has been a decoupling between interface and sound generation when
-playing. Consequently, aspects of the sensory experience of playing have
-been lost, namely the sensation of vibration and mechanical impedance
-that a traditional musical instrument provides.
-
-The project look at restoring tactile and haptic response of historical
-harpsichords for an exhibition at the musical instrument Museo di San
+With the advent of digital musical instruments in the 20th century there has
+been a decoupling between interface and sound generation when playing.
+Consequently, aspects of the sensory experience of playing have been lost,
+namely the sensation of vibration and mechanical impedance that a traditional
+musical instrument provides.
+
+The project aims to restore the tactile and haptic response of historical
+harpsichords for an exhibition at the Museo di San
 Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
-in Bologna. San Colombano is host to the Tagliavini Collection, a
-collection of approximately 70 historical keyboard instruments in
-working order \cite{Tagliavini2007} gifted to the museum by\\
-from Luigi Ferdinando Tagliavini in 2010
-\cite{SanColombano2010,Carlino2010}.
-
-Some items among the Tagliavini collection are no longer in a playing
-condition or are too fragile to continue to be available by museum
-visitors. One of those fragile instruments is a 1547 Harpsichord in the
-italian style by Alessandro Trasuntino
-\cite{Wraight2024}\footnote{examples of similar instruments by
+in Bologna. San Colombano is home to the Tagliavini Collection, a collection of
+approximately 70 historical keyboard instruments in working order
+\cite{Tagliavini2007} gifted to the museum by the Museum.\\
+from Luigi Ferdinando Tagliavini in 2010 \cite{SanColombano2010,Carlino2010}.
+
+Some items in the Tagliavini collection are no longer in playing condition or
+are too fragile to continue to be available to museum visitors. One of those
+fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
+Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
 Trasuntino available online in the Royal College of Music Collection (1531:
@@
 \url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
-}, which was used as the basis for an interactive exhibition commisioned
-by San Colombano.
-
-Project was also designed to allow others to easily recreate and iterate
-on the designs.
-
-First a small cross-sectional model was used as a proof of concept. Once
-a sensing system was decided upon, a larger 49-key scale model was
-commissioned with and then built by luthier Roberto Livi. The models are
-designed to provide only a tactile and haptic response for the player. A
-sensor system was installed into the model and messages were sent to
-computer that triggered sounds from a sample library which the player
-listens to through headphones. The paper looks at the design decisions,
-processes, and considerations to be made when undertaking a project of
-this kind.
+}, which was used as the basis for an interactive exhibition organised by San
+Colombano.
+
+The project was also open-sourced to allow for others to easily recreate and
+iterate on the designs.
+
+The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
+the model of a piano key. A 3-key cross-sectional model of a harpischord
+mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
+test bed for sensors and as proof of concept for a larger scale interface. Each
+key of the model has 2 sets of strings that are under tension but not at any
+specific musical pitch. The strings are there to provide the same resistance to
+the jack plectra as would be found on an instrument. The mechanism is in the
+Italian style of the XVI-XIX century with 2 jacks per key and their
+corresponding strings tuned in unison \footnote{"8'-8' È la dispozione normale
+dei clavicembali italiani" \cite{Tagliavini1987}}.
+
+A suitable sensing system was found to reliably detect when a jack had plucked a
+string independently and then send a corresponding MIDI message to a connected
+audio engine. A full scale 49-key scale model was commissioned with
+and then built by luthier Roberto Livi in the same style as the 1547 Trasuntino
+\ref{fig:teaser}. The sensor system was installed into the full-scale model and
+then set up for exhibition at San Colombano. Visitors are invited to play on the
+interface, which currently triggers samples on a connected sound engine. The
+paper looks at the design decisions, processes, and considerations that need to
+be made when undertaking a project of this kind.
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} 
+  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
+mechanism for an Italian style harpsichord of the XVI century.} 
+  \label{fig:3key}
+\end{figure}
 
@@
 
-This is particularly pertinent as it relates to conservation of musical
+This is particularly pertinent as it relates to the conservation of musical
 instruments, as they represent both historic objects, but also tools for
 creating music. Historical musical instruments present an interesting
-problem for cultural heritage as they are both a physical object but
+problem for cultural heritage as they are both a physical object, but
 also a tool for making music. This begs the question, can we use haptics
@@
 
-The project was carried out in collaboratoiun with San Colombano. The
+The project was carried out in collaboration with San Colombano. The
 results of the project would be presented as an interactive exhibition

### 2025-01-18 src/article.tex
introduction wording

@@
 Musée de la musique collection (1538:
-\url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
+\url{https://carmentis.kmkg-mrah.be/eMP/eMuseumPlus?service=ExternalInterface&objectId=106088})
 }, which was used as the basis for an interactive exhibition organised by San
@@
 The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
-the model of a piano key. A 3-key cross-sectional model of a harpischord
+the model of a piano key. A 3-key cross-sectional model of a harpsichord
 mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
@@
 Italian style of the XVI-XIX century with 2 jacks per key and their
-corresponding strings tuned in unison \footnote{"8'-8' È la dispozione normale
-dei clavicembali italiani" \cite{Tagliavini1987}}.
+corresponding strings tuned in unison \footnote{``8'-8' È la disposizione normale
+dei clavicembali italiani'' \cite{Tagliavini1987}}.
 
@@
 available to play when they were gifted to National Trust Fenton House.
-A custom MIDI interface arranged in a two manual harpischord style which
+A custom MIDI interface arranged in a two manual harpsichord style which
 triggered samples of the instruments in the collection using an

### 2025-01-18 src/main.tex
introduction wording

@@
 \affiliation{%
-  \institution{Museo San Colombano}
+  \institution{Museo di San Colombano}
   \city{Bologna}
@@
 \affiliation{%
- \institution{Imperial College}
+ \institution{Imperial College London}
  \city{London}
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato et al.}
+% \renewcommand{\shortauthors}{Hamilton et al.}
 
@@
 strings under tension replicate the mechanical resistance characteristic of the
-original compass. An array of infrared LED/phototransistor pairs, connected to a
+original compass. An array of infrared LED / phototransistor pairs, connected to a
 microcontroller, was installed within the interface to generate a continuous
@@
 %% page.
+
+% % Option 1:
+% \begin{teaserfigure}
+%   \includegraphics[width=\textwidth]{images/49-key-front.png}
+%   \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+%   \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+%   \label{fig:teaser}
+% \end{teaserfigure}
+
+% Option 2:
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{images/49-key-front.png}
+ \hspace*{0.6in}
+  \includegraphics[width=0.8\textwidth]{src/images/49-key-front-top.png}
   \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
@@
 
+% % Option 3:
+% \begin{teaserfigure}
+%   \includegraphics[width=\textwidth]{src/images/sensors-without-keys-front.jpg}
+%   \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+%   \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+%   \label{fig:teaser}
+% \end{teaserfigure}
+
+%% Placeholder
 % \begin{teaserfigure}
@@
 
-\include{article.tex}
+\include{src/article}
 

### 2025-01-20 src/article.tex

@@
 
-With the advent of digital musical instruments in the 20th century there has
-been a decoupling between interface and sound generation when playing.
-Consequently, aspects of the sensory experience of playing have been lost,
-namely the sensation of vibration and mechanical impedance that a traditional
-musical instrument provides.
-
-The project aims to restore the tactile and haptic response of historical
-harpsichords for an exhibition at the Museo di San
-Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
-in Bologna. San Colombano is home to the Tagliavini Collection, a collection of
-approximately 70 historical keyboard instruments in working order
-\cite{Tagliavini2007} gifted to the museum by the Museum.\\
-from Luigi Ferdinando Tagliavini in 2010 \cite{SanColombano2010,Carlino2010}.
-
-Some items in the Tagliavini collection are no longer in playing condition or
-are too fragile to continue to be available to museum visitors. One of those
-fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
-Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
-Trasuntino available online in the Royal College of Music Collection (1531:
-\url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
-Musée de la musique collection (1538:
-\url{https://carmentis.kmkg-mrah.be/eMP/eMuseumPlus?service=ExternalInterface&objectId=106088})
-}, which was used as the basis for an interactive exhibition organised by San
-Colombano.
-
-The project was also open-sourced to allow for others to easily recreate and
-iterate on the designs.
-
-The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
-the model of a piano key. A 3-key cross-sectional model of a harpsichord
-mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
-test bed for sensors and as proof of concept for a larger scale interface. Each
-key of the model has 2 sets of strings that are under tension but not at any
-specific musical pitch. The strings are there to provide the same resistance to
-the jack plectra as would be found on an instrument. The mechanism is in the
-Italian style of the XVI-XIX century with 2 jacks per key and their
-corresponding strings tuned in unison \footnote{``8'-8' È la disposizione normale
-dei clavicembali italiani'' \cite{Tagliavini1987}}.
-
-A suitable sensing system was found to reliably detect when a jack had plucked a
-string independently and then send a corresponding MIDI message to a connected
-audio engine. A full scale 49-key scale model was commissioned with
-and then built by luthier Roberto Livi in the same style as the 1547 Trasuntino
-\ref{fig:teaser}. The sensor system was installed into the full-scale model and
-then set up for exhibition at San Colombano. Visitors are invited to play on the
-interface, which currently triggers samples on a connected sound engine. The
-paper looks at the design decisions, processes, and considerations that need to
-be made when undertaking a project of this kind.
+Technological advances have transformed how museums document, present and interpret their collections, fostering immersive experiences through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
+These technologies place an emphasis on experiential authenticity, enabling encounters that evoke the sensory, emotional, and intellectual essence of the past \cite{trant_Auth_1999}.
+However, as Pine and Gilmore note \cite{pinegilmore_2007}, achieving authenticity requires museums to navigate the delicate balance between preservation and meaningful engagement—a challenge that is particularly evident in the case of historical musical instrument collections \cite{McAlpine2014}.
+
+Musical instruments represent a distinctive fusion of form, function, and history.
+Their cultural value extends beyond their visual appeal to include the tactile and auditory dimensions of use \cite{Fritz2017}.
+Yet, preservation concerns often limit direct interaction, reducing these artefacts to static displays.
+This ``red velvet cord'' approach, as theorised by McAlpine \cite{McAlpine2014}, protects fragile mechanisms but diminishes the instruments’ functional identity, disconnecting visitors from the full richness of their historical and cultural context.
+
+The Tagliavini Collection in Bologna \cite{Tagliavini2007}, renowned for its historical keyboard instruments, exemplifies this dilemma.
+To address it, this project introduces a haptic keyboard, a digital interface designed to bridge the gap between preservation and accessibility.
+The project builds on Andrew McPherson’s earlier work on "Portable Measurement and Mapping of Continuous Piano Gesture" and aligns with principles outlined in The O in NIME, which emphasises the importance of extending the lifecycle of existing NIMEs (New Interfaces for Musical Expression).
+By reusing, updating, and complementing previous NIMEs, the project fosters sustainability and ensures their relevance beyond their original context.
+
+The haptic keyboard combines high-definition sound sampling, touch-sensitive technology, and ergonomic design, recreating the sound and tactile qualities of the Tagliavini instruments without compromising their physical integrity.
+
+The article is organised as follows: TBC
+
+% With the advent of digital musical instruments in the 20th century there has
+% been a decoupling between interface and sound generation when playing.
+% Consequently, aspects of the sensory experience of playing have been lost,
+% namely the sensation of vibration and mechanical impedance that a traditional
+% musical instrument provides.
+
+% The project aims to restore the tactile and haptic response of historical
+% harpsichords for an exhibition at the Museo di San
+% Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
+% in Bologna. San Colombano is home to the Tagliavini Collection, a collection of
+% approximately 70 historical keyboard instruments in working order
+% \cite{Tagliavini2007} gifted to the museum by Luigi Ferdinando
+% Tagliavini in 2010 \cite{SanColombano2010,Carlino2010}.
+
+% Some items in the Tagliavini collection are no longer in playing condition or
+% are too fragile to continue to be available to museum visitors. One of those
+% fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
+% Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
+% Trasuntino available online in the Royal College of Music Collection (1531:
+% \url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
+% Musée de la musique collection (1538:
+% \url{https://carmentis.kmkg-mrah.be/eMP/eMuseumPlus?service=ExternalInterface&objectId=106088})
+% }, which was used as the basis for an interactive exhibition arranged by San
+% Colombano. Part of the exhibit is a keyboard that replicates the mechanism of
+% the Trasuntino that aims to provide a similar tactile playing experience to that
+% of the original.
+
+% The project was also open-sourced to allow for others to easily recreate and
+% iterate on the designs.
+
+% The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
+% the model of a piano key. A 3-key cross-sectional model of a harpsichord
+% mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
+% test bed for sensors and as proof of concept for a larger scale interface. Each
+% key of the model has 2 sets of strings that are under tension but not at any
+% specific musical pitch. The strings are there to provide the same resistance to
+% the jack plectra as would be found on an instrument. The mechanism is in the
+% Italian style of the XVI-XIX century with 2 jacks per key and their
+% corresponding strings tuned in unison \footnote{``8'-8' È la disposizione normale
+% dei clavicembali italiani'' \cite{Tagliavini1987}}.
+
+% A suitable sensing system was found to reliably detect when a jack had plucked a
+% string independently and then send a corresponding MIDI message to a connected
+% audio engine. A full scale 49-key scale model was commissioned with and then
+% built by the luthier \anon{Roberto Livi} in the same style as the 1547 Trasuntino
+% \ref{fig:teaser}. The sensor system was installed into the full-scale model and
+% then set up for exhibition at San Colombano. Visitors are invited to play on the
+% interface, which currently triggers samples on a connected sound engine. The
+% paper looks at the design decisions, processes, and considerations that need to
+% be made when undertaking a project of this kind.
 
@@
 
-Overview of fabrication methods, techniques that can be employed.
-Finally a discussion of the limitations and areas in which the core
-project can be improved and developed further.
-
-In this context, a `traditional' instrument is one whose interface for
-generating sound and the sound generating mechanism are fundamentally
-coupled. A piano requires a key press which triggers a hammer, which
-strikes a string which vibrates a body and air cavity. This system is
-tied together and cannot be separated. A synthesizer, be it a theremin,
-a control voltage synthesizer, a MIDI controller, these are all
-instruments whose interface are decoupled from their sound generation.
-This decoupling presents a problem that then needs to be addressed. How
-do we re-introduce this dimension to the experience of playing a musical
-instrument? How do we, and is it possible to, parameterise and preserve
-this experience? This is central problem that is trying to be solved in
-\cite{Nichols2002, Timmermans2020, McAlpine2014, Baldwin2016}.
-
-This is particularly pertinent as it relates to the conservation of musical
-instruments, as they represent both historic objects, but also tools for
-creating music. Historical musical instruments present an interesting
-problem for cultural heritage as they are both a physical object, but
-also a tool for making music. This begs the question, can we use haptics
-to preserve experiences of the past? The proposed project intends to
-develop further the methodologies implemented in musical haptics
-projects \cite{MusicalHaptics2018} such as \cite{Timmermans2020}.
-
-Systems exist already for the conversion of piano keyboard. The Yamaha
-disklavier, the Don Buchla designed MOOG Piano Bar, Bosendorfer, PNOscan
-by QRS focus on reproduction. These same systems could be applied to a
-harpsichord, but the limitation is a single data stream per key.
-
-\subsection{Motivations}\label{motivations}
-
-The project was carried out in collaboration with San Colombano. The
-results of the project would be presented as an interactive exhibition
-for the public.
-
-\begin{itemize}
-\item
-  existing sysetms
-
-  \begin{itemize}
-  \item
-    Yamaha Disklavier used as a basis for exploration into haptics and
-    the piano keyabord
-    \cite{MusicalHaptics2018_04, MusicalHaptics2018_05,
-    MusicalHaptics2018_13}
-  \item
-    Bosendorfer: LED photo transistor pair \cite{Moog1990} for obtaining
-    performance data \cite{MusicalHaptics2018_05}
-
-    \begin{itemize}
-    \item
-      290SE
-    \item
-      CEUS
-    \end{itemize}
-  \item
-    Don Buchla / Moog Piano Bar
-  \item
-    PNOscan MIDI 9 QRS Music as MIDI Piano system \cite{McPherson2013}
-
-    \begin{itemize}
-    \item
-      PNOmation
-    \end{itemize}
-  \item
-    Piano Disc
-  \end{itemize}
-\end{itemize}
-
-\section{Related Work}\label{related-work}
-
-Methodologies for increasing visitor engagmemnet or
-
-Museum staff want visitors to be engaged but limits in staffing and
-funding limit how vistors can interact with collections
-\cite{Templeton2018,
-McAlpine2014}. For musical instrument museums there is an additional
-difficulty in the balance of keeping historical instruments playable
-\cite{McAlpine2014}. The fragility and decay of musical instruments
-means that there is an inveitable point where instruments will no longer
-be in a playable condition \cite{McAlpine2014, Fritz2017}
-
-McAlpine outlines a similar situation to the Tagliavini collection in
-his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
-was a stipulation by Benton Fletcher that the instruments remain
-available to play when they were gifted to National Trust Fenton House.
-A custom MIDI interface arranged in a two manual harpsichord style which
-triggered samples of the instruments in the collection using an
-appropraite recording strategy for each. One problem highlighted in user
-tests by McAlpine is that the weighted keys did no provide ``an
-experiential sense of interacting with a historical keyboard''
-\cite{McAlpine2014}. The MIDI interface created from commercially
-available weighted keys. McAlpine posits a haptic keyboard augemented
-with ``actuators to provide positionally-sensitive real-time force
-feedback at point-of-contact'' in a similar approach as the piano
-mechanism by Gillespie \cite{Gillespie1996}.
-
-Previous NIME project ``Tromba Moderna'' \cite{Baldwin2016} looked to
-avoid the complex engineering problem by simply recreating the tromba
-marina and augmenting it. In the case of the Tromba Modern, a piezo
-transducer was connected to a sound synthesis engine connected to a
-driver inside the instrument to simulate the vibration that would be
-expected of an historical tromba marina.
-
-The original design was to take a similar approach to the project by
-Timmermans \cite{Timmermans2020}, which was an extension of the system
-by Gillespie \cite{Gillespie1996}. The process laid out in Timmermans
-project was to begin with a single key model of the mechanism. We began
-with a 3-key Model Harpsichord Mechanism by Graziano Bandini
-\textbackslash figure\{\}. Before we continued with augementing a model
-with actuators it was decided that it we should first validate whether a
-recreation of the harpsichord mechanism would suffice in allowing
-visitors to suspend disbelief.
-
-Benefits of putting effort into the maintenance of one interface with a
-transferable sensor system
-
-An additonal factor not faced in the Tromba Moderna project is the
-problem of scale in managing 98 rather than a single data stream of
-data.
-
-Material costs mitigated by creating a scaled version of the sound
-board, providing only the length of string necessary under comparable
-tension and thus resistance to the jack quill.
-
-Finally, a further extension to the projects above was a commmitement to
-open sourcing all aspects of the project, hardware, software, and data.
+% Overview of fabrication methods, techniques that can be employed.
+% Finally a discussion of the limitations and areas in which the core
+% project can be improved and developed further.
+
+% In this context, a `traditional' instrument is one whose interface for
+% generating sound and the sound generating mechanism are fundamentally
+% coupled. A piano requires a key press which triggers a hammer, which
+% strikes a string which vibrates a body and air cavity. This system is
+% tied together and cannot be separated. A synthesizer, be it a theremin,
+% a control voltage synthesizer, a MIDI controller, these are all
+% instruments whose interface are decoupled from their sound generation.
+% This decoupling presents a problem that then needs to be addressed. How
+% do we re-introduce this dimension to the experience of playing a musical
+% instrument? How do we, and is it possible to, parameterise and preserve
+% this experience? This is central problem that is trying to be solved in
+% \cite{Nichols2002, Timmermans2020, McAlpine2014, Baldwin2016}.
+
+% This is particularly pertinent as it relates to the conservation of musical
+% instruments, as they represent both historic objects, but also tools for
+% creating music. Historical musical instruments present an interesting
+% problem for cultural heritage as they are both a physical object, but
+% also a tool for making music. This begs the question, can we use haptics
+% to preserve experiences of the past? The proposed project intends to
+% develop further the methodologies implemented in musical haptics
+% projects \cite{MusicalHaptics2018} such as \cite{Timmermans2020}.
+
+% Systems exist already for the conversion of piano keyboard. The Yamaha
+% disklavier, the Don Buchla designed MOOG Piano Bar, Bosendorfer, PNOscan
+% by QRS focus on reproduction. These same systems could be applied to a
+% harpsichord, but the limitation is a single data stream per key.
+
+% \subsection{Motivations}\label{motivations}
+
+% \begin{itemize}
+% \item
+%   existing sysetms
+%   \begin{itemize}
+%   \item
+%     Yamaha Disklavier used as a basis for exploration into haptics and
+%     the piano keyabord
+%     \cite{MusicalHaptics2018_04, MusicalHaptics2018_05,
+%     MusicalHaptics2018_13}
+%   \item
+%     Bosendorfer: LED photo transistor pair \cite{Moog1990} for obtaining
+%     performance data \cite{MusicalHaptics2018_05}
+
+%     \begin{itemize}
+%     \item
+%       290SE
+%     \item
+%       CEUS
+%     \end{itemize}
+%   \item
+%     Don Buchla / Moog Piano Bar
+%   \item
+%     PNOscan MIDI 9 QRS Music as MIDI Piano system \cite{McPherson2013}
+
+%     \begin{itemize}
+%     \item
+%       PNOmation
+%     \end{itemize}
+%   \item
+%     Piano Disc
+%   \end{itemize}
+% \end{itemize}
+
+\section{Related Work and Motivations}\label{related-work}
+
+% These systems mentioned could be adapted for a harpsichord, but they would not
+% easily capture the separate plucks from both jacks per key. We decided to create
+% our own system and focus on detecting the vertical displacement of the jacks
+% directly, rather than the position of the key.
+
+% Methodologies for increasing visitor engagmemnet or
+
+% Museum staff want visitors to be engaged but limits in staffing and
+% funding limit how vistors can interact with collections
+% \cite{Templeton2018,
+% McAlpine2014}. For musical instrument museums there is an additional
+% difficulty in the balance of keeping historical instruments playable
+% \cite{McAlpine2014}. The fragility and decay of musical instruments
+% means that there is an inveitable point where instruments will no longer
+% be in a playable condition \cite{McAlpine2014, Fritz2017}
+
+% McAlpine outlines a similar situation to the Tagliavini collection in
+% his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
+% was a stipulation by Benton Fletcher that the instruments remain
+% available to play when they were gifted to National Trust Fenton House.
+% A custom MIDI interface arranged in a two manual harpsichord style which
+% triggered samples of the instruments in the collection using an
+% appropraite recording strategy for each. One problem highlighted in user
+% tests by McAlpine is that the weighted keys did no provide ``an
+% experiential sense of interacting with a historical keyboard''
+% \cite{McAlpine2014}. The MIDI interface created from commercially
+% available weighted keys. McAlpine posits a haptic keyboard augemented
+% with ``actuators to provide positionally-sensitive real-time force
+% feedback at point-of-contact'' in a similar approach as the piano
+% mechanism by Gillespie \cite{Gillespie1996}.
+
+% Previous NIME project ``Tromba Moderna'' \cite{Baldwin2016} looked to
+% avoid the complex engineering problem by simply recreating the tromba
+% marina and augmenting it. In the case of the Tromba Modern, a piezo
+% transducer was connected to a sound synthesis engine connected to a
+% driver inside the instrument to simulate the vibration that would be
+% expected of an historical tromba marina.
+
+% The original design was to take a similar approach to the project by
+% Timmermans \cite{Timmermans2020}, which was an extension of the system
+% by Gillespie \cite{Gillespie1996}. The process laid out in Timmermans
+% project was to begin with a single key model of the mechanism. We began
+% with a 3-key Model Harpsichord Mechanism by Graziano Bandini
+% \textbackslash figure\{\}. Before we continued with augementing a model
+% with actuators it was decided that it we should first validate whether a
+% recreation of the harpsichord mechanism would suffice in allowing
+% visitors to suspend disbelief.
+
+% Benefits of putting effort into the maintenance of one interface with a
+% transferable sensor system
+
+% An additonal factor not faced in the Tromba Moderna project is the
+% problem of scale in managing 98 rather than a single data stream of
+% data.
+
+% Material costs mitigated by creating a scaled version of the sound
+% board, providing only the length of string necessary under comparable
+% tension and thus resistance to the jack quill.
+
+% Finally, a further extension to the projects above was a commmitement to
+% open sourcing all aspects of the project, hardware, software, and data.
+
+% There are commercial systems for augmenting piano keyboards generating MIDI data.
+
+% An additional goal of the project was to design an interface that could be used to research
+% the influence the mechanism and it's resistances have on a player's performance.
+% For example, the Yamaha Disklavier series of pianos have an embedded sensor
+% system that allows for the instruments to operate in a "silent mode", but still
+% output MIDI data when the keys are struck. This has been a rich avenue for
+% exploring the affect of vibration on performance,
+% \cite{MusicalHaptics2018_04,MusicalHaptics2018_13} perception in instrument
+% quality \cite{MusicalHaptics2018_05}.
+
+% As well as a trying to offer a more engaging experience to museum attendees, the
+% interface was also designed as a potential probe for further research into the
+% idiosyncrasies of the harpsichord and the influence the mechanism has on
+% performance.
+
+Museums often face significant challenges in engaging visitors due to limitations in staffing and funding, which restrict the ways in which visitors can interact with collections \cite{Templeton2018, McAlpine2014}.
+For musical instrument museums, these challenges are compounded by the difficulty of preserving historical instruments in a playable condition \cite{McAlpine2014}.
+The inherent fragility and gradual decay of such instruments inevitably result in a point where they can no longer be played.
+A significant cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
+Karp \cite{Karp1979,Karp1985} advocates for a deeper understanding of musical instruments so that enough knowledge is generated to make them as ``copyable'' as possible.
+
+McAlpine discusses a case similar to the Tagliavini collection in his examination of the Benton Fletcher Collection at National Trust Fenton House \cite{McAlpine2014}.
+When these instruments were donated, Benton Fletcher stipulated that they remain playable and should continue to be maintained for tuition and public performance.
+A large sampling campaign was conducted, and a custom MIDI interface was designed to fulfil this requirement while preserving the original instruments' integrity.
+The MIDI keyboard, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, were used by visitors to trigger the instrument samples recorded with tailored strategies for each.
+However, user tests identified a significant limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
+To address this, McAlpine proposed -- though did not follow up to -- a haptic keyboard enhanced with actuators capable of delivering positionally sensitive, real-time force feedback at the point of contact, inspired by Gillespie’s work on piano mechanisms \cite{Gillespie1996}.
+
+The ``Tromba Moderna'' project \cite{Baldwin2016}, a previous NIME initiative, approached this issue differently by recreating and augmenting a replica of a historical tromba marina.
+A piezo transducer was connected to a sound synthesis engine and a driver within the instrument to simulate the expected vibrations of a historical tromba marina.
+This approach avoided the complex engineering challenges posed by fully replicating the original mechanisms.
+
+This work inherits the same philosophy as the Tromba Moderna project.
+By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical musical instrument-building traditions.
+
+Beyond enhancing the visitor experience, the interface serves as a research tool to explore the unique characteristics of the harpsichord and its impact on performance.
 
 \section{Design}\label{design}
+
+The present project initially drew inspiration from Timmermans' extension of Gillespie’s system \cite{Timmermans2020, Gillespie1996}.
+Timmermans’ process began with a single-key model of the mechanism.
+Similarly, we started with a 3-key model harpsichord mechanism, crafted by Graziano Bandini.
+Before incorporating actuators into the model, we sought to validate whether a faithful recreation of the harpsichord mechanism could effectively enable visitors to suspend disbelief.
+
+A critical benefit of focusing on the development of a single interface with a transferable sensor system is the potential for scalability and adaptability.
+However, managing data streams presents a significant challenge.
+Unlike the Tromba Moderna project, which dealt with a single data stream, our design must manage 98 simultaneous data streams.
+This complexity necessitated cost-effective solutions, such as scaling down the soundboard to include only the necessary string length under comparable tension and resistance to the jack quill.
+
+Another distinguishing feature of this project is its commitment to open-sourcing all hardware, software, and data components.
+While commercial systems for augmenting piano keyboards and generating MIDI data exist, this project emphasizes accessibility and transparency.
+
+An additional objective was to design an interface that supports research into how the harpsichord mechanism and its resistances influence a player’s performance.
+For example, the Yamaha Disklavier series incorporates embedded sensors enabling "silent mode" operation while outputting MIDI data when keys are struck.
+This has facilitated research into the effects of vibration on performance \cite{MusicalHaptics2018_04, MusicalHaptics2018_13} and the perception of instrument quality \cite{MusicalHaptics2018_05}.
 
@@
 The approach taken was to recreate the interface of these instruments
-not for acoustics purposes be purely for the mechanical resoponse. As
+not for acoustics purposes be purely for the mechanical response. As
 exhibition for the general public in a musical instruments there were
@@
 
-\subsubsection{Anatomy of Harpsichord
-mechanics}\label{anatomy-of-harpsichord-mechanics}
-
-\begin{itemize}
-\item
-  simple
-\item
-  aesthetics
-\item
-  split workload
-\item
-  parts of harpsichord
-
-  \begin{itemize}
-  \item
-    jack
-
-    \begin{itemize}
-    \item
-      quill
-    \item
-      jack body
-    \item
-      pivot
-    \item
-      tongue
-    \end{itemize}
-  \item
-    key
-  \item
-    body
-  \end{itemize}
-\end{itemize}
-
 \subsubsection{Keyboard Construction}\label{keyboard-construction}
-
-The desicion was made to design the elctronic sna dthe model keyboard in
-tandem. Allow for quicker development time and iteration on the idea.
-Likely the first implementation would need re-designed and redesiginng
-the electronics to fit the form of the keyboard was more cost effective
-than building or carrying out major modifications to the model.
+% PARAGRAPH FROM ROBERTO
+The construction of the keyboard required three months of work to be fully completed
+and followed the same construction phases as a real harpsichord.
+
+The model was constructed using the materials of the Italian harpsichord
+tradition:
+
+\begin{itemize}
+\item
+  Walnut wood for the wrestplank
+\item
+  Chestnut wood for the key levers
+\item
+  Boxwood and ebony for the key caps.
+\item
+  Cypress wood for the soundboard, a wood characterized by a
+\item
+  Beech wood, a brass spring and a seagull feather plectrum for the 98 jacks
+\end{itemize}
+
+The body of the case is based on a rectangular poplar frame that leaves space
+for the installation of the electronic sensors.
+
+Deviations from the design of the Trasuntino are the rectangular, rather than
+logartithmic shape, removal of a bottom panel and ommision of a short octave,
+instead spanning 4 octaves beginning on C.
+
+The strings are Malcolm Rose yellow brass wire, which are secured to wrought
+iron pins. The tension of each string was adjusted to provide resistance to
+plucking close to that of a harpsichord. Three felt strips were woven between
+the strings so as to cancel the vibration of the strings.
+
+The result is an interface instrument that is placed between the mechanical
+action of the keys and the generation of a synthetic sound. The touch of the
+wooden keys, the friction of the natural feathers on the string, the noise of
+the percussion at the base of the key together produce the sensation of being in
+front of a real harpsichord.
+
+% \subsubsection{Anatomy of Harpsichord
+% mechanics}\label{anatomy-of-harpsichord-mechanics}
+
+% \begin{itemize}
+% \item
+%   simple
+% \item
+%   aesthetics
+% \item
+%   split workload
+% \item
+%   parts of harpsichord
+
+%   \begin{itemize}
+%   \item
+%     jack
+
+%     \begin{itemize}
+%     \item
+%       quill
+%     \item
+%       jack body
+%     \item
+%       pivot
+%     \item
+%       tongue
+%     \end{itemize}
+%   \item
+%     key
+%   \item
+%     body
+%   \end{itemize}
+% \end{itemize}
+
+% The decision was made to design the electronics and the model keyboard simultaneously. 
+% This meant development development time was quicker and iteration on the idea.
+% Likely the first implementation would need re-designed and redesiginng
+% the electronics to fit the form of the keyboard was more cost effective
+% than building or carrying out major modifications to the model.
 
@@
 
-Discussion of spatial constraints and affects on electyronics design and
-installation is covered in \textbackslash section\{setup\}
-
-The rear chamber was desigbned to house comonents for audio processing,
+% Discussion of spatial constraints and effects on electronics design and
+% installation is covered in \textbackslash section\{setup\}
+
+The rear chamber was designed to house components for audio processing,
 which is discussed further in \textbackslash section
 
-\footnote{https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf}
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{images/49-key-bottom.jpeg} 
+  \caption{Underside of the full model keyboard showing two chambers. Front chamber at the top and rear chamber at the bottom.} 
+  \Description{} 
+  \label{fig:49-key-bottom}
+\end{figure}
+
+% \footnote{https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf}
 
@@
 have provided a test for scaling the electronics, but internal
-measurments and layout would have been too different. A harpsichord that
+measurements and layout would have been too different. A harpsichord that
 was economic enough to be deconstructed would also have been of a more
@@
 
-\subsection{Building Process}\label{building-process}
-
-PARAGRAPH FROM ROBERTO
-
 \subsection{Prototyping}\label{prototyping}
@@
 
-\section{Related Work and Motivation}\label{related-work-and-motivation}
-
-\begin{itemize}
-\item
-  \cite{Timmermans2020}
-\item
-  \cite{McPherson2013} \cite{McPherson2019}
-\item
-  \cite{Mudd2013}
-\item
-  \cite{Fritz2017}
-\end{itemize}
+% \section{Related Work and Motivation}\label{related-work-and-motivation}
+
+% \begin{itemize}
+% \item
+%   \cite{Timmermans2020}
+% \item
+%   \cite{McPherson2013} \cite{McPherson2019}
+% \item
+%   \cite{Mudd2013}
+% \item
+%   \cite{Fritz2017}
+% \end{itemize}
 
@@
 
-Using a similalr apporach to @HapticKey the first step was to fabricate
-or procure a model of a single working key. San Colombano provided a
-model 3-key of a harposchord mechanism by Graziano Bandini
-\textbackslash FIGURE. The 3-key model was used a test-bed for potential
-sensors while designs were drawn for a full-scale model. The
-cross-sectional nature of the model meant it was easy to fit and remove
-sensors. Small size meant that it would also be easy to transfer between
-workshops.
-
-3-key model also had a sliding jack rail which allowed for easy
-disengagement of the physical mechanical response, which allowed for
-some preliminary tests on the presence and absence of mechanism.
+% Using a similar apporach to @HapticKey the first step was to fabricate
+% or procure a model of a single working key. San Colombano provided a
+% model 3-key of a harposchord mechanism by Graziano Bandini
+% \textbackslash FIGURE. The 3-key model was used a test-bed for potential
+% sensors while designs were drawn for a full-scale model. The
+% cross-sectional nature of the model meant it was easy to fit and remove
+% sensors. Small size meant that it would also be easy to transfer between
+% workshops.
+
+% 3-key model also had a sliding jack rail which allowed for easy
+% disengagement of the physical mechanical response, which allowed for
+% some preliminary tests on the presence and absence of mechanism.
 
@@
 A number of sensors were tried before the final system was decided upon.
-An inertial measurment unit (IMU) was mounted to a single key which
-contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
+An inertial measurement unit (IMU) was mounted to a single key which
+contained a 6-degrees-of-freedom accelerometer and gyroscope. There was
 a consistent set of conditions to trigger sound, but this was not easily
-differntiable when other keys were played. The fitting time and cost of
+differentiable when other keys were played. The fitting time and cost of
 a single IMU could b reduced when purchased in bulk but was far greater
@@
 A reed sensor and a hall sensor were mounted and magnets embedded within
-the jacks. Adjusting the hieight of the sensor provided a reliable means
-of detecting a threshold, inparticular the hall sensors which could be
-partially calibrarted physically and also in softwrae. The time taken to
-calibarte a single key was time consuming. Ideally each sensor would be
+the jacks. Adjusting the height of the sensor provided a reliable means
+of detecting a threshold, in particular the hall sensors which could be
+partially calibrated physically and also in software. The time taken to
+calibrate a single key was time consuming. Ideally each sensor would be
 placed on a thread for adjustment, but given the constraints on the
 internal space it was decided any adjustment of heights would likely
-mean an unnacceptable length of time to install and calibrate. Embedding
+mean an unacceptable length of time to install and calibrate. Embedding
 magnets into the jacks was also did not satisfy the criteria of being
@@
 under each jack either above or below the padding on which the bottom of
-the jack rests. Further discussion of this apporach is in
-\textbackslash section.
+the jack rests.
+
+% Further discussion of this approach in \textbackslash section.
 
@@
 
-\begin{itemize}
-\item
-  previous use
-\item
-  problem of what to sense
-\item
-  satisfied both criteria and design restrictions
-\item
-  usage
-
-  \begin{itemize}
-  \item
-    Sweet spot is around 6mm distance
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{sensor surface}\label{sensor-surface}
-
-\begin{itemize}
-\item
-  linear gradient
-\item
-  on Jack
-\item
-  printed and cut on vinyl {[}find name{]}
-\item
-  cut to travel size
-\item
-  then to jack size
-\item
-  jacks varnished on side
-\item
-  alternatives
-
-  \begin{itemize}
-  \item
-    CNC plywood and mdf
-
-    \begin{itemize}
-    \item
-      problems
-    \end{itemize}
-  \item
-    laser cut plywood and mdf
-
-    \begin{itemize}
-    \item
-      problems
-    \end{itemize}
-  \item
-    3D printed jack body
-
-    \begin{itemize}
-    \item
-      problems
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
 \subsection{Arduino}\label{arduino}
@@
 
-\begin{itemize}
-\item
-  Nano form factor small
-\item
-  allows for easy changes in chipset
-\item
-  ``Native USB''
-  \footnote{https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards}
-\item
-  BLE Model
-
-  \begin{itemize}
-  \item
-    Performative
-  \item
-    BLE capabilities for MIDI BLE transmission
-  \item
-    All analog pins available
-  \item
-    available at the time
-  \item
-    12-bit DAC
-  \end{itemize}
-\end{itemize}
-
 \subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
@@
 reliability and speed was preffered over storage capacity.
-
-\begin{itemize}
-\item
-  On board available, but volatile
-\item
-  retained across firmware changes
-\item
-  small,
-\item
-  could have been sd card or anything else
-\item
-  available
-\item
-  interfaced via rotary encoder
-\end{itemize}
 
@@
 
-\begin{itemize}
-\item
-  locating keys
-\item
-  displaying key state
-\end{itemize}
-
 \subsection{Multiplexing}\label{multiplexing}
@@
 The multiplexer reduced 7 signals to a single pin. Three pins are
-required pins for addressing the muliplexewrs and all PCBs are addressed
-simulatenously.
-
-\begin{itemize}
-\item
-  limited ADC channels
-\item
-  ADC channels already multiplexed
-\item
-  simple
-\item
-  8 channel
-\item
-  THT
-\item
-  7 into 49
-
-  \begin{itemize}
-  \item
-    7 PCBS
-  \item
-    each with 7 sensors
-  \end{itemize}
-\end{itemize}
+required pins for addressing the multiplexers and all PCBs are addressed
+simultaneously.
 
@@
 
-\begin{itemize}
-\item
 EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
@@
 likely the KiCAD programme.
-\end{itemize}
 
@@
   \item
-    striping all adc lines with FFC
+    striping all ADC lines with FFC
   \item
@@
 
-The first version of the board esclusively used 2.54mm pitch headers.
+The first version of the board exclusively used 2.54mm pitch headers.
 The reasons for this were flexibility when creating cable looms and to
-leave open the possiility of rewuring the PCB.
-
-This was the case for the LEDs as they initially wer to be powered on
+leave open the possibility of rewiring the PCB.
+
+This was the case for the LEDs as they initially were to be powered on
 5V. It was discovered that the 3v regulator of the nano was sufficient
 to power all LEDs at a lower brightness while reducing current draw.
-Since brightness was not importnat.
+Since brightness was not important.
 
@@
 expensive compared to the cost of other components in the project
-\footnote{from RS https://uk.rs-online.com/web/p/crimp-tools/6880877
+\footnote{from RS \url{https://uk.rs-online.com/web/p/crimp-tools/6880877}
 £508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
-not readily available in the workshops at the projects disposal.
+not readily available in the workshops at the project's disposal.
 Eventually a tool was found, but an alternative approach was also
@@
 
-A locator was adapted from a creative commons stl model
-\footnote{https://www.thingiverse.com/thing:1646016} that allowed a
-generic crimping tool to be adapted for use with JST-PH crimps†
-\textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
-printer. A second alterntaive was assembling cables from other sets that
-were at hand.
-
-\begin{itemize}
-\item
-  mount for nano to allow switching
-\item
-  initially push fit 2.54mm header
-
-  \begin{itemize}
-  \item
-    not a secure fit
-  \end{itemize}
-\item
-  JST-PH used
-
-  \begin{itemize}
-  \item
-    crimping required custom cables
-
-    \begin{itemize}
-    \item
-      could be assembled from other cables
-    \item
-      length limited
-    \end{itemize}
-  \item
-    prohibitevly expensive to crimpers
-  \item
-    3D print helper to allow for crimping with generic crimpers
-  \end{itemize}
-\end{itemize}
+A locator was adapted from a creative commons STL model
+\footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a
+generic crimping tool to be adapted for use with JST-PH crimps. 
+The locator was printed using a Bambu X1 3D printer. 
+A second alterntaive was assembling cables from other sets that were at hand.
 
@@
 See each corresponding section for more detail on 3D printed components.
-
-\begin{itemize}
-\item
-  https://www.thingiverse.com/thing:1646016
-\item
-  https://www.thingiverse.com/thing:2886460
-\item
-  Supports
-
-  \begin{itemize}
-  \item
-    Push fit
-  \item
-    adjustable for best average placement
-  \end{itemize}
-\item
-  Blinkers
-
-  \begin{itemize}
-  \item
-    reduce noise from other senors
-  \end{itemize}
-\item
-  mounts
-
-  \begin{itemize}
-  \item
-    MacMini
-  \end{itemize}
-\end{itemize}
 
@@
 
-\begin{itemize}
-\item
-  Arduino platform
-
-  \begin{itemize}
-  \item
-    library support and portability
-  \item
-    dependancy management
-  \item
-    wide spread use
-  \end{itemize}
-\item
-  core algorithm
-
-  \begin{itemize}
-  \item
-    check components connected
-  \item
-    READ data from FRAM
-  \item
-    select key with rotary
-
-    \begin{itemize}
-    \item
-      click to edit, double to save
-    \end{itemize}
-  \item
-    current threshold and reading sent over serial and plotted
-  \item
-    when over threshold send MIDI note
-  \end{itemize}
-\end{itemize}
-
-\subsection{Project structure}\label{project-structure}
-
-The project is separated into three repositories encapsulating
-
-\begin{itemize}
-\item
-  Firmware
-\item
-  Keyboard CAD Data
-\item
-  PCB CAD Data
-\end{itemize}
-
-\subsubsection{Firmware}\label{firmware}
-
-Firmware repository contains all firmware for the full and smaller
-models and components.
-
-\subsubsection{Keyboard}\label{keyboard-1}
-
-The Keyboard repository contains all plans for fabricating the keyboard,
-measurements and 3D models of the jacks.
-
-\subsubsection{PCB}\label{pcb}
-
-PCB repository contains the EAGLE files for creating the sensor and
-controller boards.
-
-Given the regular spacing of components, scripts for automatic placement
-of components based on parameter of jack pitch spacing and number of
-sensors per jack.
+% \subsection{Project structure}\label{project-structure}
+
+% The project is separated into three repositories encapsulating
+
+% \begin{itemize}
+% \item
+%   Firmware
+% \item
+%   Keyboard CAD Data
+% \item
+%   PCB CAD Data
+% \end{itemize}
+% 
+% \subsubsection{Firmware}\label{firmware}
+
+% Firmware repository contains all firmware for the full and smaller
+% models and components.
+
+% \subsubsection{Keyboard}\label{keyboard-1}
+
+% The Keyboard repository contains all plans for fabricating the keyboard,
+% measurements and 3D models of the jacks.
+
+% \subsubsection{PCB}\label{pcb}
+
+% PCB repository contains the EAGLE files for creating the sensor and
+% controller boards.
+
+% Given the regular spacing of components, scripts for automatic placement
+% of components based on parameter of jack pitch spacing and number of
+% sensors per jack.
 
@@
 The electronic components were partly pre-assembled at the university of
-Edinburgh. Though all measurments were known ahead, enough flexibiliyty
-was needed to allow for proboems to solved in situ.
-
-The final assembly took place at the NEMUS lab at the univerity of
-bologna over the final week in October 2024 for delivery at San
+Edinburgh. Though all measurements were known ahead, enough flexibility
+was needed to allow for problems to solved in situ.
+
+The final assembly took place at the NEMUS lab at the university of
+Bologna over the final week in October 2024 for delivery at San
 Colombano for exhibition in the first week of November in 2024.
@@
 A misreading in the original clearance between the keys and the PCB
-meant that the keys wer block entirely from moving. A 10mm by 15mm
-section was milled out of each lke \textbackslash FIGURE to create extra
+meant that the keys were blocked entirely from moving. A 10mm by 15mm
+section was milled out of each key to create extra
 clearance. The milling required the removal, shortening and refitting of
-the leather pads for the jacks \textbackslash FIGURE. The extra work
+the leather pads for the jacks. The extra work
 added an extra day to the full install.
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \caption{} 
+  \Description{} 
+  \label{fig:3key}
+\end{figure}
 

### 2025-01-20 src/main.tex

@@
 16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
-keyboard is a collaboration between the authors and the Museo di San Colombano,
+keyboard is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
 which houses the Tagliavini Collection of playable historical keyboard
@@
 stream of data corresponding to the jacks' motion. This paper details the design
-principles, the development, open source implementation, and execution of
-the installation for the San Colombano exhibition.
+principles, development, open-source implementation, and execution of the installation for the San Colombano exhibition.
 \end{abstract}
@@
   \includegraphics[width=0.8\textwidth]{src/images/49-key-front-top.png}
-  \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
-  \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
+  \caption{49-key Model Harpsichord Mechanism by \anon{Roberto Livi, Bologna 2024}}
+  \Description{49-key Model Harpsichord Mechanism by \anon{Roberto Livi, Bologna 2024}}
   \label{fig:teaser}

### 2025-01-20 src/article.tex
Updates from Overleaf

@@
 The Tagliavini Collection in Bologna \cite{Tagliavini2007}, renowned for its historical keyboard instruments, exemplifies this dilemma.
-To address it, this project introduces a haptic keyboard, a digital interface designed to bridge the gap between preservation and accessibility.
-The project builds on Andrew McPherson’s earlier work on "Portable Measurement and Mapping of Continuous Piano Gesture" and aligns with principles outlined in The O in NIME, which emphasises the importance of extending the lifecycle of existing NIMEs (New Interfaces for Musical Expression).
-By reusing, updating, and complementing previous NIMEs, the project fosters sustainability and ensures their relevance beyond their original context.
-
-The haptic keyboard combines high-definition sound sampling, touch-sensitive technology, and ergonomic design, recreating the sound and tactile qualities of the Tagliavini instruments without compromising their physical integrity.
+To address it, this project introduces an augmented replica of a historical harpsichord keyboard.
+% The project builds on an existing design borrowed from \cite{McPherson2013} and aligns with principles outlined in \cite{Masu_NIME_2023}, emphasising the importance of extending the lifecycle of existing NIMEs.
+The project fosters ``sustainability of the results'', as identified in a previous NIME contribution by Masu \emph{et al.}  \cite{Masu_NIME_2023}, by updating and complementing the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensuring the relevance of previous NIME designs beyond their original context.
 
@@
 % be made when undertaking a project of this kind.
-
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
-  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} 
-  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
-mechanism for an Italian style harpsichord of the XVI century.} 
-  \label{fig:3key}
-\end{figure}
 
@@
 By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical musical instrument-building traditions.
-
-Beyond enhancing the visitor experience, the interface serves as a research tool to explore the unique characteristics of the harpsichord and its impact on performance.
+Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
+As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to be as similar as possible to a real Italian harpsichord.
+However, this work complements and follows up on previous reflections regarding the \emph{sustainability of results} within the NIME community and identified as \emph{the O in NIME} \cite{Masu_NIME_2023}.
+The intended use of the keyboard through a meaningful bla bla is where the novelty of this work lies, rather than solely in its technological development.
+
+Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through upcoming projects such as the ERC-funded Synergy grant Rem@ke \cite{remake1}.
 
@@
 Before incorporating actuators into the model, we sought to validate whether a faithful recreation of the harpsichord mechanism could effectively enable visitors to suspend disbelief.
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} 
+  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
+mechanism for an Italian style harpsichord of the XVI century.} 
+  \label{fig:3key}
+\end{figure}
 
@@
 The front and back that was xxx by yyy by 370mm between the faceplate
-and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber though
-a small gap between the under side of the soundboard and the key guides.
+and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber through
+a small gap between the underside of the soundboard and the key guides.
 The key guide needs to be deep enough to allow for the full movement of
@@
 
-The rear chamber was designed to house components for audio processing,
-which is discussed further in \textbackslash section
+% The rear chamber was designed to house components for audio processing,
+% which is discussed further in \textbackslash section
 
@@
   \centering
-  \includegraphics[width=\linewidth]{images/49-key-bottom.jpeg} 
+  \includegraphics[width=\linewidth]{images/49-key-bottom-} 
   \caption{Underside of the full model keyboard showing two chambers. Front chamber at the top and rear chamber at the bottom.} 
@@
 For the final interface a 49-key (4-octave) was constructed by luthier
-Roberto Livi. The keyboard is in the same style† as the 3-key model with
+\anon{Roberto Livi}. The keyboard is in the same style† as the 3-key model with
 to jacks per key.
@@
 
-Timescale: from initial discussion to delivery of interface to the NEMUS
+Timescale: from initial discussion to delivery of interface to the \anon{NEMUS}
 lab was 8-months. Active assembly of the instrument was 3 months
@@
 to satisfy. This criteria was used as a means of evaluating the
-feasability throughout the prototyping phase.
+feasibility throughout the prototyping phase.
 
@@
 
-\section{Hardware Design}\label{hardware-design}
-
-The hardware was designed with the xx in mind.
+% \section{Hardware Design}\label{hardware-design}
+
+% The hardware was designed with the xx in mind.
 
@@
 
-\subsection{Failed sensors}\label{failed-sensors}
-
-A number of sensors were tried before the final system was decided upon.
-An inertial measurement unit (IMU) was mounted to a single key which
-contained a 6-degrees-of-freedom accelerometer and gyroscope. There was
-a consistent set of conditions to trigger sound, but this was not easily
-differentiable when other keys were played. The fitting time and cost of
-a single IMU could b reduced when purchased in bulk but was far greater
-than other systems considered.
-
-A reed sensor and a hall sensor were mounted and magnets embedded within
-the jacks. Adjusting the height of the sensor provided a reliable means
-of detecting a threshold, in particular the hall sensors which could be
-partially calibrated physically and also in software. The time taken to
-calibrate a single key was time consuming. Ideally each sensor would be
-placed on a thread for adjustment, but given the constraints on the
-internal space it was decided any adjustment of heights would likely
-mean an unacceptable length of time to install and calibrate. Embedding
-magnets into the jacks was also did not satisfy the criteria of being
-scalable and non-invasive.
-
-Light dependant resistor mounted to the jack rail. as jacks were pressed
-less light would be let through. Usable data but very dependant on
-lighting conditions. Modifying the jack also did not satisfy the
-non-intrusive criteria and the placement outside the internal chamber of
-the model meant hiding electronics would be more difficult.
-
-Finally, force sensitive resistors placed under the jack of each key was
-considered but was never implemented. The sensors were to be placed
-under each jack either above or below the padding on which the bottom of
-the jack rests.
+% \subsection{Failed sensors}\label{failed-sensors}
+
+% A number of sensors were tried before the final system was decided upon.
+% An inertial measurement unit (IMU) was mounted to a single key which
+% contained a 6-degrees-of-freedom accelerometer and gyroscope. There was
+% a consistent set of conditions to trigger sound, but this was not easily
+% differentiable when other keys were played. The fitting time and cost of
+% a single IMU could b reduced when purchased in bulk but was far greater
+% than other systems considered.
+
+% A reed sensor and a hall sensor were mounted and magnets embedded within
+% the jacks. Adjusting the height of the sensor provided a reliable means
+% of detecting a threshold, in particular the hall sensors which could be
+% partially calibrated physically and also in software. The time taken to
+% calibrate a single key was time consuming. Ideally each sensor would be
+% placed on a thread for adjustment, but given the constraints on the
+% internal space it was decided any adjustment of heights would likely
+% mean an unacceptable length of time to install and calibrate. Embedding
+% magnets into the jacks was also did not satisfy the criteria of being
+% scalable and non-invasive.
+
+% Light dependant resistor mounted to the jack rail. as jacks were pressed
+% less light would be let through. Usable data but very dependant on
+% lighting conditions. Modifying the jack also did not satisfy the
+% non-intrusive criteria and the placement outside the internal chamber of
+% the model meant hiding electronics would be more difficult.
+
+% Finally, force sensitive resistors placed under the jack of each key was
+% considered but was never implemented. The sensors were to be placed
+% under each jack either above or below the padding on which the bottom of
+% the jack rests.
 
@@
 model. Since the Electronics were designed at a distanced to the final
-full model some flexbility was required in the design. Design decisions
+full model, some flexibility was required in the design. Design decisions
 that were not as optimal but would allow for change in wiring and
@@
 Electronics designed so that they would not require any equipment
-outside what can commonly be found in a maker space at a univeristy.
-
-The equipment required including soldering suitable for throughole
+outside what can commonly be found in a maker space at a university.
+
+The equipment required including soldering suitable for through-hole
 technology (THT) components and 3D printing.
 
-Laser cutting and CNC routing fascilities were used during the research
+Laser cutting and CNC routing facilities were used during the research
 for the project but not required for the final implementation.
@@
 NEMUS lab under construction during the first section of the project.
-Facilities available at the maker spaces of Bologna (Alma Labor) and
-Edinburgh (uCreate) used for what was deemed reasonable assumption in
+Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
+\anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
 terms of available equipment.
 
-\begin{itemize}
-\item
-  Over arching themes
-
-  \begin{itemize}
-  \item
-    simplicity of design
-
-    \begin{itemize}
-    \item
-      Limited resources
-    \item
-      easily reproducible
-    \end{itemize}
-  \item
-    availability and lead times
-  \item
-    iteration and design from distance
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{QRE1113}\label{qre1113}
-
-Following from the work in \cite{McPherson2013, McPherson2019} we tested
-a system using the Fairchild QRE1113 \textbackslash datasheetfigue. The
-QRE1113 is a combination infrared LED and phototransistor sensitive to
-IR light. The phototransistor of QRE1113 provides data on how much light
-is present and in particular how much light is being relfected from a
-surface. Since refelected light will be proportional to the distance of
-a surfcae it is a good, close proximity, distance sensor. Distance is
-the way in which the instruments in \cite{McPherson2013, McPherson2019}
-function as well the Moog Piano Bar from which it too inspiration.
-
-The limitation for taking a distance approach is that there is not one
-but 2 jacks per key. For each jack to be measured indepeedntly the data
-would need tp com fropm the jacks themselves.
-
-They \emph{could} be used and individual trigger points set in
-calibration, but this creates new problems.
-
-\begin{itemize}
-\item
-  The triggering would need to be bespoke for every key as fluxuations
-  in the dimensions of every piece mean that the threshold would be
-  different.
-
-  \begin{itemize}
-  \item
-    Moving a problem into additional time in the calibtraryoin setup
-    phase.
-  \end{itemize}
-\item
-  Remebering the constraint that system isshould noyt fix problem, there
-  would be no easy way to tell if a jack had nbot reseated.
-
-  \begin{itemize}
-  \item
-    In playing a harpihchord when the jack does not go to its rest
-    position, not note would play. This fault should be carried over
-    intp the final instrument.
-  \end{itemize}
-\end{itemize}
-
-Another common application of the QRE1113 is within line following
-robots, where an array of QRE1113s ar used in conjunction with black
-tape along which some motorised system can following
-
-Applied to the jacks was a greyscale gradient printed the length full
-travel range of the jack.
-
-For the first iteration this was performed on an inkjet printer and
-double-sided tape. For the final iteration the stickers were printed on
-(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
-µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
-cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
-reduced the manual labour for cutting stickers meaning the process would
-scale from the 3-key to the 49-key model.
-
-An initial sticker design was used that contained only the gradient. It
-was found that the jacks wer sticking in their slots andthe edges on the
-stickers easily peeled around the printing The sticker was redesigned to
-be the full length of the jack which avoided the problem of peeling. The
-larger are also mean the stickers adhered betterto the jack Some
-trimming was still required of each sticker to avoid them adding extraa
-resistence to the jack's movement.
-
-During testing it was found that their was a considerable amount
-cross-talk bwteen adjacent sensors. This meant the reading from key
-would be affected by the movement of the key next to it. The cross-talk
-was assumed to be as a result of light reflecting from other jacks.
-Given that IR is used this was difficult to confirm by sight. A set of
-baffles was designed for the pcb that would wrap round the jack and stop
-light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
-printers using PLA plastic. A dark pigmented PLA was used to avoid IR
-light simply being diffused. After the baffles were fitted the
-cross-talk was entirely eliminated.
-
-\subsection{Arduino}\label{arduino}
-
-\subsubsection{Nano BLE}\label{nano-ble}
-
-Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
-appealing as it would allow for easy changes between chipsets without
-requiring any rewiring and thus no new PCBs would need to be created if
-a better alternative was found later in the project. Another benefit to
-the Arduino Nano boards was that typically had native-USB meaning the
-could be programmed as a hardware USB MIDI device directly.
-
-The Nano BLE specifically was used in the initial stages for the
-integrated IMU.
-
-When it was finalised that the prokect would proceed using the QRE1113 a
-number of microcontroller units (MCU) were condsidered.
-
-The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
-Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
-
-A rough performance test was later carried out using an STM32
-NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
-more performative. To execute a single sensor read cycle of the firmware
-the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
-on average. The Nano ESP32 had high variability so it was decided to
-continue with the Nano BLE. Further discussion of using ESP32 continues
-in \textbackslash section\{Going forward\}
-
-BLE functionality also meant if the cable connection between the nano
-and computer running audio synthesis was impractical, BLE MIDI was
-available as a fallback. The Nano BLE also has a 12-bit DAC, which
-provided a contingency should the default 10-bit not be sensitive enough
-for the signal from the sensors.
-
-Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
-functionality, but it was found that 2 ADC channels were in fact not
-useable and would have required additional multiplexers.
-
-\subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
-
-There would need to be the ability to store values on non-volatile
-memory so that the electroincs could be powered-down without losing
-data.
-
-Adjusting the values through firmware would not have been flexible or
-practical. Instead an Ferroelectric RAM unit was chosen and connected
-using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
-was primarily chosen for it's ready availability.
-
-Most microcontrollers have some variety of non-volatile program memory
-that can be read and writ accessed but in most cases this memory is
-cleared during compilation and re-flashing. The small cost was far
-outweighed by the potential for data loss if using onboard RAM.
-
-\emph{It is worth noting that the FRAM model chosen stopped being
-available during the project, meaning any future iteration would need to
-adjust for another board in firmware}
-
-The firmware was structure in a way that the interface could be changed
-so long as the code required for any other board would comply with
-function arguments and outputs. Worth noting the fragility and
-volatility when designing this kind of project and how easily it is for
-the project to become broken.
-
-SD card would provide an easier way to interact with calibration data.
-The posiiton of the controller board did not permit easy access.
-reliability and speed was preffered over storage capacity.
-
-\subsection{RGB LEDs}\label{rgb-leds}
-
-for problems of scale:
-
-Addressing the problem of scale of identifying which key's threshold is
-currently being edited.
-
-The LEDs also allowed for easy identification for keys that were outside
-of expected values and when a key was triggered. This allowed for visual
-identification of malfunctioning or uncalibrated keys.
-
-The LEDs are on the same PCB as the sensors and are hidden when the
-nameboard is in place.2024
-
-Would sugggest that this is a necessity of this kind to provide a means
-to identify and diagnose LEDs with state.
-
-TLC5940 or other PWM driver such as in \cite{McPherson2013} added
-complexity to assembling, but were avoid in favour of Addressable LEDs
-with intergrated driver. The intergrated LEDs share the same data line
-which has the limitation where a fualty LED will stop all following LEDs
-from working. An individual PWM driver per PCB would allow avoid
-situations where failutre cascades, but would add complexity in assembly
-as the parts are typically SMD. Also, given the LEDs are fixed to the
-board and could not be repaired in situ, the solution would be to
-replace the PCB regardless of which approach was taken. Given the LEDs
-were only used in debugging and calibration and otehrwise would not be
-visible, a simpler design approach was taken.
-
-\subsection{Multiplexing}\label{multiplexing}
-
-CD4051BE Texas Instruments
-
-problem of scale
-
-For the 3-key model, individual jacks could be connected to one of the 8
-available ADC pins.
-
-The BLE Nano has 8 ADC pins connected to one ADC channel via a
-multiplexer
-
-Multiplexer added to expand to the
-
-Transistor / Op-amp circuit could have been used similar to the one in
-\cite{McPherson2013}.
-
-Benefits were keeping with THT based designed for easy assembly.
-
-An 8-channel multiplexer was chosen as the model has 49-keys and a
-division of 7 PCBs each containing 7 sensors
-
-The multiplexer reduced 7 signals to a single pin. Three pins are
-required pins for addressing the multiplexers and all PCBs are addressed
-simultaneously.
-
-\subsection{PCB design}\label{pcb-design}
-
-EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
-deprecated EAGLE and the programme will no longer be supported from
-2026 Project files will need to be transitioned to another standard,
-likely the KiCAD programme.
-
-trade off between limiting number of boards required and allowing
-compensation for variation in pitch.
-
-Smaller PCBs means that spacing could be adjusted to accomodate the
-changes in pitch between jacks. It also allowed for a modular approach
-where any PCB could swapped for minimal cost.
-
-\subsubsection{Sensor board}\label{sensor-board}
-
-The 49 QRE1113s are divided across 7 PCBs each containing
-
-\begin{itemize}
-\item
-  x7 QRE1113
-\item
-  x7 100R
-\item
-  x7 10kR
-\item
-  x1 CD4051BE Multipexer
-\item
-  jack / sensor pitch by board
-\item
-  7 boards adjust for pitch/spacing error
-\item
-  modular, replaceable
-\item
-  orientation problems
-
-  \begin{itemize}
-  \item
-    legs break easily
-  \item
-    end mount like A.P.Mc. time consuming
-  \item
-    problem
-
-    \begin{itemize}
-    \item
-      milling out keys
-    \end{itemize}
-  \end{itemize}
-\item
-  designed from distance
-\item
-  striped lines across pcbs
-\item
-  problems
-
-  \begin{itemize}
-  \item
-    additional terminals uneeded resulted in lots of additional work
-  \item
-    long cabling increase risk of breaking
-  \end{itemize}
-\item
-  solutions
-
-  \begin{itemize}
-  \item
-    striping all ADC lines with FFC
-  \item
-    solder pads to select channel
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{controller board}\label{controller-board}
+\subsection{Controller Board}\label{controller-board}
 
@@
 simply mean the system did not work correctly and in the worst case it
-could potentially damaged some components.
+could potentially damage the components.
 
@@
 The locator was printed using a Bambu X1 3D printer. 
-A second alterntaive was assembling cables from other sets that were at hand.
+A second alternative was assembling cables from other sets that were at hand.
+
+\subsubsection{Arduino Nano BLE}\label{arduino-nano-ble}
+
+Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
+appealing as it would allow for easy changes between chipsets without
+requiring any rewiring and thus no new PCBs would need to be created if
+a better alternative was found later in the project. Another benefit to
+the Arduino Nano boards was that typically had native-USB meaning the
+could be programmed as a hardware USB MIDI device directly.
+
+The Nano BLE was used in the initial stages for its integrated inertial measurment unit.
+
+When it was finalised that the project would proceed using the QRE1113 a
+number of microcontroller units (MCU) were considered.
+
+The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
+Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
+
+A rough performance test was later carried out using an STM32
+NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
+more performative. To execute a single sensor read cycle of the firmware
+the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
+on average. The Nano ESP32 had high variability so it was decided to
+continue with the Nano BLE.
+
+% Further discussion of using ESP32 continues
+% in \textbackslash section\{Going forward\}
+
+BLE functionality also meant if the cable connection between the nano
+and computer running audio synthesis was impractical, BLE MIDI was
+available as a fallback. The Nano BLE also has a 12-bit DAC, which
+provided a contingency should the default 10-bit not be sensitive enough
+for the signal from the sensors.
+
+Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
+functionality, but it was found that 2 ADC channels were in fact not
+useable and would have required additional multiplexers.
+
+\subsubsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
+
+There would need to be the ability to store values on non-volatile
+memory so that the electronics could be powered-down without losing
+data.
+
+Adjusting the values through firmware would not have been flexible or
+practical. Instead an Ferroelectric RAM unit was chosen and connected
+using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
+was primarily chosen for it's ready availability.
+
+Most microcontrollers have some variety of non-volatile program memory
+that can be read and writ accessed but in most cases this memory is
+cleared during compilation and re-flashing. The small cost was far
+outweighed by the potential for data loss if using onboard RAM.
+
+% \emph{It is worth noting that the FRAM model chosen stopped being
+% available during the project, meaning any future iteration would need to
+% adjust for another board in firmware}
+
+% The firmware was structured in a way that the interface could be changed
+% so long as the code required for any other board would comply with
+% function arguments and outputs.
+
+% Worth noting the fragility and
+% volatility when designing this kind of project and how easily it is for
+% the project to become broken.
+
+SD card would provide an easier way to interact with calibration data.
+The position of the controller board did not permit easy access.
+reliability and speed was preferred over storage capacity.
+
+\subsection{Sensor board}\label{sensor-board}
+
+The 49 QRE1113s are divided across 7 PCBs each containing
+
+\subsubsection{QRE1113}\label{qre1113}
+
+A number of sensors were tried before the final system was decided upon.
+Sensors included, an 6-degrees-of-freedom inertial measurement unit,
+reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
+None of these sensors satisfied the sensor criteria described above.
+
+Following from the work in \cite{McPherson2013, McPherson2019}, we tested
+a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
+QRE1113 is a combination infrared LED and phototransistor sensitive to
+IR light. The phototransistor of QRE1113 provides data on how much light
+is present and in particular how much light is being reflected from a
+surface. Since reflected light will be proportional to the distance of
+a surface it is a good, close-proximity, distance sensor. Distance is
+the way in which the instruments in \cite{McPherson2013, McPherson2019}
+function as well the Moog Piano Bar from which it took inspiration.
+
+The limitation for taking a distance approach is that there is not one
+but 2 jacks per key. For each jack to be measured independently the data
+would need to come from the jacks themselves.
+
+They \emph{could} be used and individual trigger points set in
+calibration, but this creates new problems.
+
+Another common application of the QRE1113 is within line following
+robots, where an array of QRE1113s are used in conjunction with black
+tape or reflective. A motorised system can then use this data to follow the line of tape and adjust its position.
+
+Applied to the jacks was a greyscale gradient printed the length full
+travel range of the jack.
+
+For the first iteration this was performed on an inkjet printer and
+double-sided tape. For the final iteration the stickers were printed on
+(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
+µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
+cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
+reduced the manual labour for cutting stickers meaning the process would
+scale from the 3-key to the 49-key model.
+
+An initial sticker design was used that contained only the gradient. It
+was found that the jacks were sticking in their slots and the edges on the
+stickers easily peeled around the printing The sticker was redesigned to
+be the full length of the jack which avoided the problem of peeling. The
+larger are also mean the stickers adhered better to the jack Some
+trimming was still required of each sticker to avoid them adding extra
+resistance to the jack's movement.
+
+During testing it was found that their was a considerable amount
+cross-talk between adjacent sensors. This meant the reading from key
+would be affected by the movement of the key next to it. The cross-talk
+was assumed to be as a result of light reflecting from other jacks.
+Given that IR is used this was difficult to confirm by sight. A set of
+baffles was designed for the PCB that would wrap round the jack and stop
+light from other jacks or sensors. The baffles were printed on a BAMBU X1 3D
+printers using PLA plastic. A dark pigmented PLA was used to avoid IR
+light simply being diffused. After the baffles were fitted, the
+cross-talk was eliminated.
+
+\subsubsection{RGB LEDs}\label{rgb-leds}
+
+for problems of scale:
+
+Addressing the problem of scale of identifying which key's threshold is
+currently being edited.
+
+The LEDs also allowed for easy identification for keys that were outside
+of expected values and when a key was triggered. This allowed for visual
+identification of malfunctioning or un-calibrated keys.
+
+The LEDs are on the same PCB as the sensors and are hidden when the
+name board is in place.2024
+
+Would suggest that this is a necessity of this kind to provide a means
+to identify and diagnose LEDs with state.
+
+TLC5940 or other PWM driver such as in \cite{McPherson2013} added
+complexity to assembling, but were avoid in favour of Addressable LEDs
+with integrated driver. The integrated LEDs share the same data line
+which has the limitation where a faulty LED will stop all following LEDs
+from working. An individual PWM driver per PCB would allow avoid
+situations where failure cascades, but would add complexity in assembly
+as the parts are typically SMD. Also, given the LEDs are fixed to the
+board and could not be repaired in situ, the solution would be to
+replace the PCB regardless of which approach was taken. Given the LEDs
+were only used in debugging and calibration and otherwise would not be
+visible, a simpler design approach was taken.
+
+\subsubsection{Multiplexing}\label{multiplexing}
+
+The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
+We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
+An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
+The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
+Transistor / Op-amp circuit could have been used similar to the one in
+\cite{McPherson2013}. Given time constraints it was beneficial to continue using and entirely THT based designed for easy assembly.
+
+% \subsection{PCB design}\label{pcb-design}
+
+% EAGLE PCBs designed using EAGLE.
+
+% Worth noting that AutoCAD has
+% deprecated EAGLE and the programme will no longer be supported from
+% 2026 Project files will need to be transitioned to another standard,
+% likely the KiCAD programme.
+
+% trade off between limiting number of boards required and allowing
+% compensation for variation in pitch.
+
+% Smaller PCBs means that spacing could be adjusted to accomodate the
+% changes in pitch between jacks. It also allowed for a modular approach
+% where any PCB could swapped for minimal cost.
+
+% \begin{itemize}
+% \item
+%   x7 QRE1113
+% \item
+%   x7 100R
+% \item
+%   x7 10kR
+% \item
+%   x1 CD4051BE Multipexer
+% \item
+%   jack / sensor pitch by board
+% \item
+%   7 boards adjust for pitch/spacing error
+% \item
+%   modular, replaceable
+% \item
+%   orientation problems
+
+%   \begin{itemize}
+%   \item
+%     legs break easily
+%   \item
+%     end mount like A.P.Mc. time consuming
+%   \item
+%     problem
+
+%     \begin{itemize}
+%     \item
+%       milling out keys
+%     \end{itemize}
+%   \end{itemize}
+% \item
+%   designed from distance
+% \item
+%   striped lines across pcbs
+% \item
+%   problems
+
+%   \begin{itemize}
+%   \item
+%     additional terminals uneeded resulted in lots of additional work
+%   \item
+%     long cabling increase risk of breaking
+%   \end{itemize}
+% \item
+%   solutions
+
+%   \begin{itemize}
+%   \item
+%     striping all ADC lines with FFC
+%   \item
+%     solder pads to select channel
+%   \end{itemize}
+% \end{itemize}
 
@@
 
-Firmare checks components are connected then reads the FRAM to see if
+Firmware checks components are connected then reads the FRAM to see if
 threshold data is already present. The rotary encoder and it's momentary
@@
 sent. For the reverse case where the value was below the threshold when
-it was previously above, a MIDI Note Off message ois sent.
+it was previously above, a MIDI Note Off message is sent.
 
@@
 
-The final assembly took place at the NEMUS lab at the university of
+The final assembly took place at the NEMUS lab at the University of
 Bologna over the final week in October 2024 for delivery at San
@@
   \centering
-  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
+  \includegraphics[width=\linewidth]{src/images/keys-milled.jpeg} 
   \caption{} 
   \Description{} 
-  \label{fig:3key}
+  \label{fig:milling}
 \end{figure}
@@
 cable loom. Soldering cables and inspecting each join required another
-day work. This did allow for readjustments to the positon of the
+day work. This did allow for readjustments to the position of the
 controller board Measurements of the rear chamber did not match with the
-MacMini that was to be used a different layout was decided for fitting
+Mac Mini that was to be used a different layout was decided for fitting
 the controller board/ Cutting cables to size allowed for a great deal of
-felxibility but at the cost of time. After the first install it was
-clearer what constraints were in place for fittig and an alternative
+flexibility but at the cost of time. After the first install it was
+clearer what constraints were in place for fitting and an alternative
 wiring system using flat flex cable was implement (discussed in
@@
 them. The jacks had been coated with a varnish on the short side to
-provide a better surface for the stikcers to adhere to. After the
+provide a better surface for the stickers to adhere to. After the
 stickers were applied excess material was trimmed from each jack. The
@@
 After fitting the PCBs preliminary calibration was carried out. Before
-the keys were reffitted the jacks wer put back in place. The data from
+the keys were refitted the jacks were put back in place. The data from
 each key was confirmed to be in a similar range. Any key showing
-anomalous dat had it's jack removed and re-inspected. The PCB fitting
+anomalous data had it's jack removed and re-inspected. The PCB fitting
 process and refitting required another day as faults were found in
-wiring or misalignement of the stickers. After the sensors were
+wiring or misalignment of the stickers. After the sensors were
 confirmed to be functioning the keys were put back.
 
-The calibration workflow was with Dr.~Craig Webb who was unfamiliar with
+The calibration workflow was with \anon{Dr. Craig Webb} who was unfamiliar with
 the system. Adjustments were made to the firmware to to make it more
@@
 this caused problems in situations where the data was unusual due to
-misalignement of the sensors
+misalignment of the sensors.
 
 This would likely be a better system for when the keyboard was fully
-operational
+operational.
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{images/serial_monitor.png} 
+  \caption{} 
+  \Description{} 
+  \label{fig:serial_monitor}
+\end{figure}
 
 The thresholds were calibrated visually utilising the serial plotter of
-the Arduino IDE and MIDI was confirmed using the open source MIDI
-Monitor \footnote{https://github.com/krevis/MIDIApps}.
+the Arduino IDE (Figure \ref{fig:serial_monitor}) and the open source MIDI
+Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}.
 
@@
 
-Finer calibration carried out with the guidance of Catalina Vincens as
+Finer calibration carried out with the guidance of \anon{Catalina Vincens} as
 an expert harpsichord performer
@@
 
-A limitation in the design is the variablility in positioning of tage,
+A limitation in the design is the variability in positioning of tag,
 especially when they need to be reapplied. A jig was designed for
 applying tags, but could not be printed in time for installation. Even
-with a jig, some tags had to be radjusted by hand regardless. Long term
-this hasd impliccation on how to maintain the interface.
+with a jig, some tags had to be readjusted by hand regardless. Long term
+this has implications on how to maintain the interface.
 
@@
 were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
-cutting results in a loss of material that has too much ariablity. The
-edge o f the jack body will be charrred. The material also needs to be
-reliable. on the advice of two luthiers, roberto livi and Jonathan Santa
-Maria Bouquet at Edinburgh, Plywood and MDF have problems with
+cutting results in a loss of material that has too much variability. The
+edge of the jack body will be charred. The material also needs to be
+reliable. on the advice of two luthiers, \anon{Roberto Livi} and \anon{Jonathan Santa
+Maria Bouquet} at \anon{Edinburgh}, Plywood and MDF have problems with
 variations in humidity, tear-out and delamination.
@@
 though need to test of the material is naturally reflective enough or if
-a coating would beed to be applied.
+a coating would need to be applied.
 
@@
 A second interface designed more closely to the Trasuntino style
-instrument commisioned by the NEMUS project with Roberto Livi. Using
+instrument commissioned by the \anon{NEMUS} project with \anon{Roberto Livi}. Using
 lessons learned for the first iteration the second can make better
-acommodations for internal electronics. Interface to be used for a
+accommodations for internal electronics. Interface to be used for a
 performer study and used for performance with a numerically simulated
-sound syntheiss engine.
+sound synthesis engine.
 
@@
 Future novel designs for an interface that would utilise velocity and
-aftertouch messages in software. This would be applicable for numericla
-simulations of strings. A bespoke piec e of sowftare would also allow
+aftertouch messages in software. This would be applicable for numerical
+simulations of strings. A bespoke piece of software would also allow
 for direct mapping of jack rows to registers. There is enough data to
-implemenet functionality for velocity and aftertouch MIDI messages and
+implement functionality for velocity and aftertouch MIDI messages and
 simply needs to be activated in the firmware. (Hyperinstruments
@@
 
-Using the Espressif development framework ESP-IDF and follwong the same
+Using the Espressif development framework ESP-IDF and following the same
 logic as the original firmware source, the latency was reduced to around
@@
 exposed in the Arduino libraries. It is possible that the 10ms latency
-could be improved upon but there would be a cost in the form of
-develoment time. Consideration would need to be made on how to port
+could be improved upon, but there would be a cost in the form of
+development time. Consideration would need to be made on how to port
 existing code or whether to completely re-write the firmware.
@@
     each PCB. It was not common for a board to not be functioning and
-    thye result to be that the power weas not correctly connected.
-    Especially when the ower is connected via somethgin such as an FFC
-    cable where it is easy to plug-in the wrong way esepcially when
-    fitting is to be carried out by someone unfmailiar wiyth the system.
+    thye result to be that the power was not correctly connected.
+    Especially when the power is connected via something such as an FFC
+    cable where it is easy to plug-in the wrong way especially when
+    fitting is to be carried out by someone unfamiliar with the system.
     Power LEDs would quickly confirm whether the problem was power
     related. They would have to be small enough and low enough current
-    draw to avoid glare or being mistaken with other indcator LEDs To
-    continue using THT components, the choice would have to be moade
+    draw to avoid glare or being mistaken with other indicator LEDs To
+    continue using THT components, the choice would have to be made
     carefully.

### 2025-01-20 src/article.tex

@@
 % The project builds on an existing design borrowed from \cite{McPherson2013} and aligns with principles outlined in \cite{Masu_NIME_2023}, emphasising the importance of extending the lifecycle of existing NIMEs.
-The project fosters ``sustainability of the results'', as identified in a previous NIME contribution by Masu \emph{et al.}  \cite{Masu_NIME_2023}, by updating and complementing the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensuring the relevance of previous NIME designs beyond their original context.
+The project fosters and complements the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensures the relevance of previous NIME designs beyond their original context \cite{Masu_NIME_2023}.
 
@@
 For musical instrument museums, these challenges are compounded by the difficulty of preserving historical instruments in a playable condition \cite{McAlpine2014}.
-The inherent fragility and gradual decay of such instruments inevitably result in a point where they can no longer be played.
+The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played.
 A significant cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
@@
 However, user tests identified a significant limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
-To address this, McAlpine proposed -- though did not follow up to -- a haptic keyboard enhanced with actuators capable of delivering positionally sensitive, real-time force feedback at the point of contact, inspired by Gillespie’s work on piano mechanisms \cite{Gillespie1996}.
-
-The ``Tromba Moderna'' project \cite{Baldwin2016}, a previous NIME initiative, approached this issue differently by recreating and augmenting a replica of a historical tromba marina.
+
+% To address this, McAlpine suggested designing a haptic keyboard enhanced with actuators capable of delivering positionally sensitive, real-time force feedback at the point of contact, inspired by previous projects such as Gillespie’s work on piano mechanisms \cite{Gillespie1996}.
+
+On the other hand, the ``Tromba Moderna'' project \cite{Baldwin2016}, a previous NIME initiative, approached the issue of musical heritage playability by recreating and augmenting a replica of a historical tromba marina.
 A piezo transducer was connected to a sound synthesis engine and a driver within the instrument to simulate the expected vibrations of a historical tromba marina.
-This approach avoided the complex engineering challenges posed by fully replicating the original mechanisms.
 
 This work inherits the same philosophy as the Tromba Moderna project.
-By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical musical instrument-building traditions.
+By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions.
 Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
-As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to be as similar as possible to a real Italian harpsichord.
+As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to adhere to longstanding harpsichord building traditions.
 However, this work complements and follows up on previous reflections regarding the \emph{sustainability of results} within the NIME community and identified as \emph{the O in NIME} \cite{Masu_NIME_2023}.
-The intended use of the keyboard through a meaningful bla bla is where the novelty of this work lies, rather than solely in its technological development.
+The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{images/49-key-bottom-} 
+  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
   \caption{Underside of the full model keyboard showing two chambers. Front chamber at the top and rear chamber at the bottom.} 
@@
 
+\begin{figure}  
+  \centering
+  \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
+  \caption{} 
+  \Description{} 
+  \label{fig:7-sensor-board}
+\end{figure}
+
 \subsubsection{QRE1113}\label{qre1113}
@@
 calibration, but this creates new problems.
+
+\begin{figure}  
+  \centering
+  \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
+  \caption{} 
+  \Description{} 
+  \label{fig:cross-section}
+\end{figure}
 

### 2025-01-20 src/article.tex
Updates from Overleaf

@@
 This work inherits the same philosophy as the Tromba Moderna project.
-By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions.
+By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics and controlling a MIDI-triggered sample library, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions.
 Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
@@
 
-Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through upcoming projects such as the ERC-funded Synergy grant Rem@ke \cite{remake1}.
+Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anona{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
 
@@
 
-Accomodations were made in the design to allow for as much flexibility
-as possible. This also meant that designs for the eletronics would have
+Accommodations were made in the design to allow for as much flexibility
+as possible. This also meant that designs for the electronics would have
 to include a deal of flexibility in how they might be installed. Surplus
@@
 \textbackslash FIGURE, but there was still a limit given the structural
-and aesthatic requirements.
+and aesthetic requirements.
 
@@
 Deviations from the design of the Trasuntino are the rectangular, rather than
-logartithmic shape, removal of a bottom panel and ommision of a short octave,
+'logarithmic' shape, removal of a bottom panel and omission of a short octave,
 instead spanning 4 octaves beginning on C.
@@
 plucking close to that of a harpsichord. Three felt strips were woven between
-the strings so as to cancel the vibration of the strings.
+the strings in order to dampen the vibration of the strings.
 
@@
   Low Latency: Reading sensors and then triggering an output should have
-  a low latency. Using {[}paper on latency{]} a target was set as X ms
+  a low latency. Following \cite{Jack2016} an latency was set as 10 ms
   with an upper limit set to 25ms beyond which the sensors would be
@@
   Scalable: The system should be scalable both economically and
-  temporally. Given the open soucre nature of the porject, it could not
-  be prohibitvely expensive to carry out. Also any system that would
+  temporally. Given the open source nature of the project, it could not
+  be prohibitively expensive to carry out. Also any system that would
   work on a single key would have to scale financially to potentially 50

### 2025-01-21 src/article.tex
Updates from Overleaf

@@
 For musical instrument museums, these challenges are compounded by the difficulty of preserving historical instruments in a playable condition \cite{McAlpine2014}.
-The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played.
+The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played, even when collections adhere to the strictest conservation protocols.
 A significant cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
@@
 
-Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anona{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
+Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
 
@@
 
-\subsection{Design Problem}\label{design-problem}
+\section{Design}
+
+\section{Constraints}\label{design-problem}
 
@@
 
-\subsubsection{Design Constraints}\label{design-constraints}
+% \subsection{Design Constraints}\label{design-constraints}
 
@@
 
-\subsection{Keyboard}\label{keyboard}
-
-\subsubsection{Keyboard Construction}\label{keyboard-construction}
+\subsection{Building the Model Keyboard}\label{keyboard}
+
 % PARAGRAPH FROM ROBERTO
 The construction of the keyboard required three months of work to be fully completed
-and followed the same construction phases as a real harpsichord.
+and followed the same construction phases of a real harpsichord.
 
@@
 \item
-  Cypress wood for the soundboard, a wood characterized by a
+  Cypress wood for the soundboard
 \item
@@
 
-The model was designed to have two internal chambers a front and rear.
-The front and back that was xxx by yyy by 370mm between the faceplate
-and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber through
-a small gap between the underside of the soundboard and the key guides.
-The key guide needs to be deep enough to allow for the full movement of
-the key and stop the key becoming unseated. The chamber then extends to
-a support wall which divides the front and rear.
+% The model was designed to have two internal chambers a front and rear.
+% The front and back that was xxx by yyy by 370mm between the faceplate
+% and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber through
+% a small gap between the underside of the soundboard and the key guides.
+% The key guide needs to be deep enough to allow for the full movement of
+% the key and stop the key becoming unseated. The chamber then extends to
+% a support wall which divides the front and rear.
 
@@
 
-\subsection{Prototyping}\label{prototyping}
-
-\subsubsection{3-Key Prototype}\label{key-prototype}
-
-\begin{itemize}
-\item
-  proof of concept
-\item
-  available on loan from museum
-\item
-  easy access to internal structure
-\end{itemize}
-
-\subsubsection{49-Key Prototype}\label{key-prototype-1}
-
-For the final interface a 49-key (4-octave) was constructed by luthier
-\anon{Roberto Livi}. The keyboard is in the same style† as the 3-key model with
-to jacks per key.
-
-The keyboard is fully enclosed and internally has two chambers, front
-and rear, in which the electronics could be installed
-\textbackslash figure(underside 49-key).
-
-Keyboard to be used for exhibition at san colombano where it would be
-available to play by general public.
-
-Timescale: from initial discussion to delivery of interface to the \anon{NEMUS}
-lab was 8-months. Active assembly of the instrument was 3 months
-
-Constructed From:
-
-\begin{itemize}
-\item
-  for exhibition
-\item
-  problems of scale
-\item
-  problems of timescale
-\end{itemize}
-
-\subsubsection{Sensor Criteria}\label{sensor-criteria}
+% \subsection{Prototyping}\label{prototyping}
+
+% \subsubsection{3-Key Prototype}\label{key-prototype}
+
+% \begin{itemize}
+% \item
+%   proof of concept
+% \item
+%   available on loan from museum
+% \item
+%   easy access to internal structure
+% \end{itemize}
+
+% \subsubsection{49-Key Prototype}\label{key-prototype-1}
+
+% For the final interface a 49-key (4-octave) was constructed by luthier
+% \anon{Roberto Livi}. The keyboard is in the same style† as the 3-key model with
+% to jacks per key.
+
+% The keyboard is fully enclosed and internally has two chambers, front
+% and rear, in which the electronics could be installed
+% \textbackslash figure(underside 49-key).
+
+% Keyboard to be used for exhibition at san colombano where it would be
+% available to play by general public.
+
+% Timescale: from initial discussion to delivery of interface to the \anon{NEMUS}
+% lab was 8-months. Active assembly of the instrument was 3 months
+
+% Constructed From:
+
+% \begin{itemize}
+% \item
+%   for exhibition
+% \item
+%   problems of scale
+% \item
+%   problems of timescale
+% \end{itemize}
+
+\section{Hardware Design}
+
+For the first iteration the data from the sensor would simply be used to
+test when a threshold had been passed. On passing the threshold a
+message would be sent to trigger the playback of a corresponding note.
+
+\subsection{Sensor Criteria}\label{sensor-criteria}
 
@@
 
-\subsubsection{Triggering a note}\label{triggering-a-note}
-
-For the first iteration the data from the sensor would simply be used to
-test when a threshold had been passed. On passing the threshold a
-message would be sent to trigger the playback of a corresponding note.
-
-The message format covered in \textbackslash section
+% \subsubsection{Triggering a note}\label{triggering-a-note}
+
+% The message format covered in \textbackslash section
 
@@
 
-\section{Electronics}\label{electronics}
+% \section{Electronics}\label{electronics}
 
@@
 \section{Implementation}\label{implementation}
+
+\begin{figure}  
+  \centering
+  \includegraphics[,width=\linewidth]{src/images/exhibition-user-1.jpeg} 
+  \caption{} 
+  \Description{} 
+  \label{fig:exhibition}
+\end{figure}
 

### 2025-01-21 src/article.tex
Updates from Overleaf

@@
   \centering
-  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
-  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} 
+  \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png} 
+  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023. 3-Key modelled used for initial prototyping. A: Wrestplank B: Key Lever C: Key cap D: soundboard E: Jack} 
   \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
@@
 
+The design criteria was to create a keyboard interface that provided a comparable tactile response to that of a harpsichord in the italian style.
+
+Accommodations were made in the design to allow for as much flexibility as possible. 
+This also meant that designs for the electronics would have to include a level of flexibility in how they might be installed. 
+Surplus space was added into the design of the internal chamber of the model Figure \ref{fig:49-key-bottom}, but there was still a limit given the structural
+and aesthetic requirements.
+
 \section{Constraints}\label{design-problem}
 
-Central philosophy of the Tagliavini Collection is that the instrument
-should remain playable to the public. The condition of some instruments
-is such that restoration would required a replacement of vast majority
-of the instrument, would be too costly, would risk the instrument given
-its fragility, or all of those in combination.
-
-This is not an uncommon problem faced by musical instrument museums
-worldwide. The central question the project tries to answer is ``how do
-we conserve musical instruments in order to continue interacting with
-them''
-
-The approach taken was to recreate the interface of these instruments
-not for acoustics purposes be purely for the mechanical response. As
-exhibition for the general public in a musical instruments there were
-aesthetic and functional constraints placed on the design of the
-interface.
-
-The design criteria was to create a keyboard interface that provides a
-comparable physical response to a harpsichord.
+% Central philosophy of the Tagliavini Collection is that the instrument
+% should remain playable to the public. The condition of some instruments
+% is such that restoration would required a replacement of vast majority
+% of the instrument, would be too costly, would risk the instrument given
+% its fragility, or all of those in combination.
+
+% This is not an uncommon problem faced by musical instrument museums
+% worldwide. The central question the project tries to answer is ``how do
+% we conserve musical instruments in order to continue interacting with
+% them''
+
+% The approach taken was to recreate the interface of these instruments
+% not for acoustics purposes be purely for the mechanical response. As
+% exhibition for the general public in a musical instruments there were
+% aesthetic and functional constraints placed on the design of the
+% interface.
 
@@
 
-\begin{itemize}
-\item
-  no electronics should be visible.
-\item
-  it should be robust and reliable, easily maintained by museum staff
-  without complex technical processes
-\item
-  it should not augment or fix the mechanics, but present them with
-  their original limitation
-\end{itemize}
-
-Another constraint was that the design of the sensor system and the
-model keyboard would have to carried out at a distance in separate
-workshops. Time constraints meant the keyboard was being fabricated for
-a sensor that did not yet exist.
-
-Accommodations were made in the design to allow for as much flexibility
-as possible. This also meant that designs for the electronics would have
-to include a deal of flexibility in how they might be installed. Surplus
-space was added into the design of the internal chamber of the model
-\textbackslash FIGURE, but there was still a limit given the structural
-and aesthetic requirements.
+There were constraints that were put in place by the Museum on the design of the exhibit.
+
+\begin{itemize}
+\item
+  No electronics should be visible \footnote{with the exception of a pair of headphones for listening and a display to change parameters of the audio engine}
+\item
+  It should be robust and reliable, easily maintained by museum staff without complex technical processes
+\item
+  it should not augment or fix the mechanics, but present them with their original limitation
+\end{itemize}
+
+% Another constraint was that the design of the sensor system and the
+% model keyboard would have to carried out at a distance in separate
+% workshops. Time constraints meant the keyboard was being fabricated for
+% a sensor that did not yet exist.
 
@@
 The model was constructed using the materials of the Italian harpsichord
-tradition:
+tradition. Parts are labelled in Figure \ref{fig:3key}.
 
@@
 
-For the initial prototyping stage, modifying an existing harpsichord was
-considered though ultimately discarded. An existing instrument would
-have provided a test for scaling the electronics, but internal
-measurements and layout would have been too different. A harpsichord that
-was economic enough to be deconstructed would also have been of a more
-modern mechanical design and not of the design that was being recreated
-with the model. Prototyping stage would likely be invasive and damage or
-heavy modification likely. As such it was not a good use of funds as a
-modern instrument fit for destruction did not offer benefits that the
-small model with easy internal access did not already provide.
-
-% \subsection{Prototyping}\label{prototyping}
+% A harpsichord that
+% was economic enough to be deconstructed would also have been of a more
+% modern mechanical design and not of the design that was being recreated
+% with the model. Prototyping stage would likely be invasive and damage or
+% heavy modification likely. As such it was not a good use of funds as a
+% modern instrument fit for destruction did not offer benefits that the
+% small model with easy internal access did not already provide.
 
@@
 
-For the first iteration the data from the sensor would simply be used to
-test when a threshold had been passed. On passing the threshold a
-message would be sent to trigger the playback of a corresponding note.
+For the first iteration the data from the sensor would simply be used to test when a threshold had been passed.
+On passing the threshold a message would be sent to trigger the playback of a corresponding note.
+
+For the initial prototyping stage, modifying an existing harpsichord was considered though ultimately discarded. An existing instrument would
+have provided a test for scaling the electronics, but internal measurements and layout would have been too different.
+
+Using a similar approach to Timmermans' Haptic Key project, \cite{Timmermans2020} our first step was to fabricate or procure a model of a single working key.
+San Colombano was able to provide a model 3-key of a harpischord mechanism by Graziano Bandini (Figure \ref{fig:3key}).
 
@@
 \end{itemize}
+
+% \subsection{Prototyping}\label{prototyping}
 
@@
 
-Design of electronics carried out in tandem with construction of full
-model. Since the Electronics were designed at a distanced to the final
-full model, some flexibility was required in the design. Design decisions
-that were not as optimal but would allow for change in wiring and
-fitting should there be any issues during installation.
-
-Electronics designed so that they would not require any equipment
-outside what can commonly be found in a maker space at a university.
-
-The equipment required including soldering suitable for through-hole
-technology (THT) components and 3D printing.
-
-Laser cutting and CNC routing facilities were used during the research
-for the project but not required for the final implementation.
-
-The idea behind this was to allow for the project to be reasonably
-recreated without procurement of any additional specialist equipment.
-
-NEMUS lab under construction during the first section of the project.
-Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
-\anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
-terms of available equipment.
+% Design of electronics carried out in tandem with construction of full
+% model. Since the Electronics were designed at a distanced to the final
+% full model, some flexibility was required in the design. Design decisions
+% that were not as optimal but would allow for change in wiring and
+% fitting should there be any issues during installation.
+
+The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.
+
+% The equipment required including soldering suitable for through-hole technology (THT) components and 3D printing.
+
+% Laser cutting and CNC routing facilities were used during the research
+% for the project but not required for the final implementation.
+
+This decisions was motivated by the project's goal to be open source and reasonably recreated without procurement of any additional specialist equipment.
+
+% NEMUS lab under construction during the first section of the project.
+% Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
+% \anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
+% terms of available equipment.
+
+The electronics system can be divided into two designs, the "sensor board" (Figure \ref{fig:7-sensor-board}) for components for detecting jack displacement and controller board for collecting and processing sensor signals.
+
+\begin{figure}  
+  \centering
+  \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
+  \caption{} 
+  \Description{} 
+  \label{fig:7-sensor-board}
+\end{figure}
+
+\subsection{Sensor board}\label{sensor-board}
+
+The 49 optical sensors are divided across 7 PCBs each containing
+
+\begin{itemize}
+\item
+  x7 QRE1113 optical sensors
+\item
+  x7 100 $\Omega$ resistors
+\item
+  x7 10 k$\Omega$ resistors \footnote{a update on the design showed these could be reduced to a single resistor.}
+\item
+  x1 CD4051BE Multipexer
+\end{itemize}
+
+A number of sensors were tried before the final system was decided upon. 
+Sensors tested for the design included, a 6-degrees-of-freedom inertial measurement unit, reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
+None of the above satisfied the sensor criteria described and were thus discounted.
+
+Following from the work in \cite{McPherson2013, McPherson2019}, we tested
+a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
+QRE1113 is a combination infrared LED and phototransistor sensitive to
+IR light. The phototransistor of QRE1113 provides data on how much light
+is present and in particular how much light is being reflected from a
+surface. Since reflected light will be proportional to the distance of
+a surface it is a good, close-proximity, distance sensor. Distance is
+the way in which the instruments in \cite{McPherson2013, McPherson2019}
+function as well the Moog Piano Bar from which it took inspiration.
+
+The limitation for taking a distance approach is that there is not one
+but 2 jacks per key. For each jack to be measured independently the data
+would need to come from the jacks themselves.
+
+% They \emph{could} be used and individual trigger points set in
+% calibration, but this creates new problems.
+
+\begin{figure}  
+  \centering
+  \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
+  \caption{} 
+  \Description{} 
+  \label{fig:cross-section}
+\end{figure}
+
+Another common application of the QRE1113 is within line following
+robots, where an array of QRE1113s are used in conjunction with black
+tape or reflective. A motorised system can then use this data to follow the line of tape and adjust its position.
+
+Applied to the jacks was a greyscale gradient printed the length full
+travel range of the jack.
+
+For the first iteration this was performed on an inkjet printer and
+double-sided tape. For the final iteration the stickers were printed on
+(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
+µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
+cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
+reduced the manual labour for cutting stickers meaning the process would
+scale from the 3-key to the 49-key model.
+
+% An initial sticker design was used that contained only the gradient. It
+% was found that the jacks were sticking in their slots and the edges on the
+% stickers easily peeled around the printing The sticker was redesigned to
+% be the full length of the jack which avoided the problem of peeling. The
+% larger are also mean the stickers adhered better to the jack Some
+% trimming was still required of each sticker to avoid them adding extra
+% resistance to the jack's movement.
+
+During testing it was found that their was a considerable amount
+cross-talk between adjacent sensors. This meant the reading from key
+would be affected by the movement of the key next to it. The cross-talk
+was assumed to be as a result of light reflecting from other jacks.
+Given that IR is used this was difficult to confirm by sight. A set of
+baffles was designed for the PCB that would wrap round the jack and stop
+light from other jacks or sensors. The baffles were printed on a BAMBU X1 3D
+printers using PLA plastic. A dark pigmented PLA was used to avoid IR
+light simply being diffused. After the baffles were fitted, the
+cross-talk was eliminated.
+
+\subsubsection{RGB LEDs}\label{rgb-leds}
+
+for problems of scale:
+
+Addressing the problem of scale of identifying which key's threshold is
+currently being edited.
+
+The LEDs also allowed for easy identification for keys that were outside
+of expected values and when a key was triggered. This allowed for visual
+identification of malfunctioning or un-calibrated keys.
+
+The LEDs are on the same PCB as the sensors and are hidden when the
+name board is in place.2024
+
+Would suggest that this is a necessity of this kind to provide a means
+to identify and diagnose LEDs with state.
+
+TLC5940 or other PWM driver such as in \cite{McPherson2013} added
+complexity to assembling, but were avoid in favour of Addressable LEDs
+with integrated driver. The integrated LEDs share the same data line
+which has the limitation where a faulty LED will stop all following LEDs
+from working. An individual PWM driver per PCB would allow avoid
+situations where failure cascades, but would add complexity in assembly
+as the parts are typically SMD. Also, given the LEDs are fixed to the
+board and could not be repaired in situ, the solution would be to
+replace the PCB regardless of which approach was taken. Given the LEDs
+were only used in debugging and calibration and otherwise would not be
+visible, a simpler design approach was taken.
+
+\subsubsection{Multiplexing}\label{multiplexing}
+
+The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
+We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
+An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
+The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
+Transistor / Op-amp circuit could have been used similar to the one in
+\cite{McPherson2013}. Given time constraints it was beneficial to continue using and entirely THT based designed for easy assembly.
+
+% \subsection{PCB design}\label{pcb-design}
+
+% EAGLE PCBs designed using EAGLE.
+
+% Worth noting that AutoCAD has
+% deprecated EAGLE and the programme will no longer be supported from
+% 2026 Project files will need to be transitioned to another standard,
+% likely the KiCAD programme.
+
+% trade off between limiting number of boards required and allowing
+% compensation for variation in pitch.
+
+% Smaller PCBs means that spacing could be adjusted to accomodate the
+% changes in pitch between jacks. It also allowed for a modular approach
+% where any PCB could swapped for minimal cost.
+
+% \begin{itemize}
+% \item
+%   x7 QRE1113
+% \item
+%   x7 100R
+% \item
+%   x7 10kR
+% \item
+%   x1 CD4051BE Multipexer
+% \item
+%   jack / sensor pitch by board
+% \item
+%   7 boards adjust for pitch/spacing error
+% \item
+%   modular, replaceable
+% \item
+%   orientation problems
+
+%   \begin{itemize}
+%   \item
+%     legs break easily
+%   \item
+%     end mount like A.P.Mc. time consuming
+%   \item
+%     problem
+
+%     \begin{itemize}
+%     \item
+%       milling out keys
+%     \end{itemize}
+%   \end{itemize}
+% \item
+%   designed from distance
+% \item
+%   striped lines across pcbs
+% \item
+%   problems
+
+%   \begin{itemize}
+%   \item
+%     additional terminals uneeded resulted in lots of additional work
+%   \item
+%     long cabling increase risk of breaking
+%   \end{itemize}
+% \item
+%   solutions
+
+%   \begin{itemize}
+%   \item
+%     striping all ADC lines with FFC
+%   \item
+%     solder pads to select channel
+%   \end{itemize}
+% \end{itemize}
 
@@
 reliability and speed was preferred over storage capacity.
-
-\subsection{Sensor board}\label{sensor-board}
-
-The 49 QRE1113s are divided across 7 PCBs each containing
-
-\begin{figure}  
-  \centering
-  \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
-  \caption{} 
-  \Description{} 
-  \label{fig:7-sensor-board}
-\end{figure}
-
-\subsubsection{QRE1113}\label{qre1113}
-
-A number of sensors were tried before the final system was decided upon.
-Sensors included, an 6-degrees-of-freedom inertial measurement unit,
-reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
-None of these sensors satisfied the sensor criteria described above.
-
-Following from the work in \cite{McPherson2013, McPherson2019}, we tested
-a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
-QRE1113 is a combination infrared LED and phototransistor sensitive to
-IR light. The phototransistor of QRE1113 provides data on how much light
-is present and in particular how much light is being reflected from a
-surface. Since reflected light will be proportional to the distance of
-a surface it is a good, close-proximity, distance sensor. Distance is
-the way in which the instruments in \cite{McPherson2013, McPherson2019}
-function as well the Moog Piano Bar from which it took inspiration.
-
-The limitation for taking a distance approach is that there is not one
-but 2 jacks per key. For each jack to be measured independently the data
-would need to come from the jacks themselves.
-
-They \emph{could} be used and individual trigger points set in
-calibration, but this creates new problems.
-
-\begin{figure}  
-  \centering
-  \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
-  \caption{} 
-  \Description{} 
-  \label{fig:cross-section}
-\end{figure}
-
-Another common application of the QRE1113 is within line following
-robots, where an array of QRE1113s are used in conjunction with black
-tape or reflective. A motorised system can then use this data to follow the line of tape and adjust its position.
-
-Applied to the jacks was a greyscale gradient printed the length full
-travel range of the jack.
-
-For the first iteration this was performed on an inkjet printer and
-double-sided tape. For the final iteration the stickers were printed on
-(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
-µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
-cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
-reduced the manual labour for cutting stickers meaning the process would
-scale from the 3-key to the 49-key model.
-
-An initial sticker design was used that contained only the gradient. It
-was found that the jacks were sticking in their slots and the edges on the
-stickers easily peeled around the printing The sticker was redesigned to
-be the full length of the jack which avoided the problem of peeling. The
-larger are also mean the stickers adhered better to the jack Some
-trimming was still required of each sticker to avoid them adding extra
-resistance to the jack's movement.
-
-During testing it was found that their was a considerable amount
-cross-talk between adjacent sensors. This meant the reading from key
-would be affected by the movement of the key next to it. The cross-talk
-was assumed to be as a result of light reflecting from other jacks.
-Given that IR is used this was difficult to confirm by sight. A set of
-baffles was designed for the PCB that would wrap round the jack and stop
-light from other jacks or sensors. The baffles were printed on a BAMBU X1 3D
-printers using PLA plastic. A dark pigmented PLA was used to avoid IR
-light simply being diffused. After the baffles were fitted, the
-cross-talk was eliminated.
-
-\subsubsection{RGB LEDs}\label{rgb-leds}
-
-for problems of scale:
-
-Addressing the problem of scale of identifying which key's threshold is
-currently being edited.
-
-The LEDs also allowed for easy identification for keys that were outside
-of expected values and when a key was triggered. This allowed for visual
-identification of malfunctioning or un-calibrated keys.
-
-The LEDs are on the same PCB as the sensors and are hidden when the
-name board is in place.2024
-
-Would suggest that this is a necessity of this kind to provide a means
-to identify and diagnose LEDs with state.
-
-TLC5940 or other PWM driver such as in \cite{McPherson2013} added
-complexity to assembling, but were avoid in favour of Addressable LEDs
-with integrated driver. The integrated LEDs share the same data line
-which has the limitation where a faulty LED will stop all following LEDs
-from working. An individual PWM driver per PCB would allow avoid
-situations where failure cascades, but would add complexity in assembly
-as the parts are typically SMD. Also, given the LEDs are fixed to the
-board and could not be repaired in situ, the solution would be to
-replace the PCB regardless of which approach was taken. Given the LEDs
-were only used in debugging and calibration and otherwise would not be
-visible, a simpler design approach was taken.
-
-\subsubsection{Multiplexing}\label{multiplexing}
-
-The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
-We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
-An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
-The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
-Transistor / Op-amp circuit could have been used similar to the one in
-\cite{McPherson2013}. Given time constraints it was beneficial to continue using and entirely THT based designed for easy assembly.
-
-% \subsection{PCB design}\label{pcb-design}
-
-% EAGLE PCBs designed using EAGLE.
-
-% Worth noting that AutoCAD has
-% deprecated EAGLE and the programme will no longer be supported from
-% 2026 Project files will need to be transitioned to another standard,
-% likely the KiCAD programme.
-
-% trade off between limiting number of boards required and allowing
-% compensation for variation in pitch.
-
-% Smaller PCBs means that spacing could be adjusted to accomodate the
-% changes in pitch between jacks. It also allowed for a modular approach
-% where any PCB could swapped for minimal cost.
-
-% \begin{itemize}
-% \item
-%   x7 QRE1113
-% \item
-%   x7 100R
-% \item
-%   x7 10kR
-% \item
-%   x1 CD4051BE Multipexer
-% \item
-%   jack / sensor pitch by board
-% \item
-%   7 boards adjust for pitch/spacing error
-% \item
-%   modular, replaceable
-% \item
-%   orientation problems
-
-%   \begin{itemize}
-%   \item
-%     legs break easily
-%   \item
-%     end mount like A.P.Mc. time consuming
-%   \item
-%     problem
-
-%     \begin{itemize}
-%     \item
-%       milling out keys
-%     \end{itemize}
-%   \end{itemize}
-% \item
-%   designed from distance
-% \item
-%   striped lines across pcbs
-% \item
-%   problems
-
-%   \begin{itemize}
-%   \item
-%     additional terminals uneeded resulted in lots of additional work
-%   \item
-%     long cabling increase risk of breaking
-%   \end{itemize}
-% \item
-%   solutions
-
-%   \begin{itemize}
-%   \item
-%     striping all ADC lines with FFC
-%   \item
-%     solder pads to select channel
-%   \end{itemize}
-% \end{itemize}
 

### 2025-01-21 src/article.tex

@@
 
-% With the advent of digital musical instruments in the 20th century there has
-% been a decoupling between interface and sound generation when playing.
-% Consequently, aspects of the sensory experience of playing have been lost,
-% namely the sensation of vibration and mechanical impedance that a traditional
-% musical instrument provides.
-
-% The project aims to restore the tactile and haptic response of historical
-% harpsichords for an exhibition at the Museo di San
-% Colombano\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}
-% in Bologna. San Colombano is home to the Tagliavini Collection, a collection of
-% approximately 70 historical keyboard instruments in working order
-% \cite{Tagliavini2007} gifted to the museum by Luigi Ferdinando
-% Tagliavini in 2010 \cite{SanColombano2010,Carlino2010}.
-
-% Some items in the Tagliavini collection are no longer in playing condition or
-% are too fragile to continue to be available to museum visitors. One of those
-% fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
-% Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
-% Trasuntino available online in the Royal College of Music Collection (1531:
-% \url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
-% Musée de la musique collection (1538:
-% \url{https://carmentis.kmkg-mrah.be/eMP/eMuseumPlus?service=ExternalInterface&objectId=106088})
-% }, which was used as the basis for an interactive exhibition arranged by San
-% Colombano. Part of the exhibit is a keyboard that replicates the mechanism of
-% the Trasuntino that aims to provide a similar tactile playing experience to that
-% of the original.
-
-% The project was also open-sourced to allow for others to easily recreate and
-% iterate on the designs.
-
-% The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
-% the model of a piano key. A 3-key cross-sectional model of a harpsichord
-% mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
-% test bed for sensors and as proof of concept for a larger scale interface. Each
-% key of the model has 2 sets of strings that are under tension but not at any
-% specific musical pitch. The strings are there to provide the same resistance to
-% the jack plectra as would be found on an instrument. The mechanism is in the
-% Italian style of the XVI-XIX century with 2 jacks per key and their
-% corresponding strings tuned in unison \footnote{``8'-8' È la disposizione normale
-% dei clavicembali italiani'' \cite{Tagliavini1987}}.
-
-% A suitable sensing system was found to reliably detect when a jack had plucked a
-% string independently and then send a corresponding MIDI message to a connected
-% audio engine. A full scale 49-key scale model was commissioned with and then
-% built by the luthier \anon{Roberto Livi} in the same style as the 1547 Trasuntino
-% \ref{fig:teaser}. The sensor system was installed into the full-scale model and
-% then set up for exhibition at San Colombano. Visitors are invited to play on the
-% interface, which currently triggers samples on a connected sound engine. The
-% paper looks at the design decisions, processes, and considerations that need to
-% be made when undertaking a project of this kind.
-
-% Overview of fabrication methods, techniques that can be employed.
-% Finally a discussion of the limitations and areas in which the core
-% project can be improved and developed further.
-
-% In this context, a `traditional' instrument is one whose interface for
-% generating sound and the sound generating mechanism are fundamentally
-% coupled. A piano requires a key press which triggers a hammer, which
-% strikes a string which vibrates a body and air cavity. This system is
-% tied together and cannot be separated. A synthesizer, be it a theremin,
-% a control voltage synthesizer, a MIDI controller, these are all
-% instruments whose interface are decoupled from their sound generation.
-% This decoupling presents a problem that then needs to be addressed. How
-% do we re-introduce this dimension to the experience of playing a musical
-% instrument? How do we, and is it possible to, parameterise and preserve
-% this experience? This is central problem that is trying to be solved in
-% \cite{Nichols2002, Timmermans2020, McAlpine2014, Baldwin2016}.
-
-% This is particularly pertinent as it relates to the conservation of musical
-% instruments, as they represent both historic objects, but also tools for
-% creating music. Historical musical instruments present an interesting
-% problem for cultural heritage as they are both a physical object, but
-% also a tool for making music. This begs the question, can we use haptics
-% to preserve experiences of the past? The proposed project intends to
-% develop further the methodologies implemented in musical haptics
-% projects \cite{MusicalHaptics2018} such as \cite{Timmermans2020}.
-
-% Systems exist already for the conversion of piano keyboard. The Yamaha
-% disklavier, the Don Buchla designed MOOG Piano Bar, Bosendorfer, PNOscan
-% by QRS focus on reproduction. These same systems could be applied to a
-% harpsichord, but the limitation is a single data stream per key.
-
-% \subsection{Motivations}\label{motivations}
-
-% \begin{itemize}
-% \item
-%   existing sysetms
-%   \begin{itemize}
-%   \item
-%     Yamaha Disklavier used as a basis for exploration into haptics and
-%     the piano keyabord
-%     \cite{MusicalHaptics2018_04, MusicalHaptics2018_05,
-%     MusicalHaptics2018_13}
-%   \item
-%     Bosendorfer: LED photo transistor pair \cite{Moog1990} for obtaining
-%     performance data \cite{MusicalHaptics2018_05}
-
-%     \begin{itemize}
-%     \item
-%       290SE
-%     \item
-%       CEUS
-%     \end{itemize}
-%   \item
-%     Don Buchla / Moog Piano Bar
-%   \item
-%     PNOscan MIDI 9 QRS Music as MIDI Piano system \cite{McPherson2013}
-
-%     \begin{itemize}
-%     \item
-%       PNOmation
-%     \end{itemize}
-%   \item
-%     Piano Disc
-%   \end{itemize}
-% \end{itemize}
-
 \section{Related Work and Motivations}\label{related-work}
 
-% These systems mentioned could be adapted for a harpsichord, but they would not
-% easily capture the separate plucks from both jacks per key. We decided to create
-% our own system and focus on detecting the vertical displacement of the jacks
-% directly, rather than the position of the key.
-
-% Methodologies for increasing visitor engagmemnet or
-
-% Museum staff want visitors to be engaged but limits in staffing and
-% funding limit how vistors can interact with collections
-% \cite{Templeton2018,
-% McAlpine2014}. For musical instrument museums there is an additional
-% difficulty in the balance of keeping historical instruments playable
-% \cite{McAlpine2014}. The fragility and decay of musical instruments
-% means that there is an inveitable point where instruments will no longer
-% be in a playable condition \cite{McAlpine2014, Fritz2017}
-
-% McAlpine outlines a similar situation to the Tagliavini collection in
-% his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
-% was a stipulation by Benton Fletcher that the instruments remain
-% available to play when they were gifted to National Trust Fenton House.
-% A custom MIDI interface arranged in a two manual harpsichord style which
-% triggered samples of the instruments in the collection using an
-% appropraite recording strategy for each. One problem highlighted in user
-% tests by McAlpine is that the weighted keys did no provide ``an
-% experiential sense of interacting with a historical keyboard''
-% \cite{McAlpine2014}. The MIDI interface created from commercially
-% available weighted keys. McAlpine posits a haptic keyboard augemented
-% with ``actuators to provide positionally-sensitive real-time force
-% feedback at point-of-contact'' in a similar approach as the piano
-% mechanism by Gillespie \cite{Gillespie1996}.
-
-% Previous NIME project ``Tromba Moderna'' \cite{Baldwin2016} looked to
-% avoid the complex engineering problem by simply recreating the tromba
-% marina and augmenting it. In the case of the Tromba Modern, a piezo
-% transducer was connected to a sound synthesis engine connected to a
-% driver inside the instrument to simulate the vibration that would be
-% expected of an historical tromba marina.
-
-% The original design was to take a similar approach to the project by
-% Timmermans \cite{Timmermans2020}, which was an extension of the system
-% by Gillespie \cite{Gillespie1996}. The process laid out in Timmermans
-% project was to begin with a single key model of the mechanism. We began
-% with a 3-key Model Harpsichord Mechanism by Graziano Bandini
-% \textbackslash figure\{\}. Before we continued with augementing a model
-% with actuators it was decided that it we should first validate whether a
-% recreation of the harpsichord mechanism would suffice in allowing
-% visitors to suspend disbelief.
-
-% Benefits of putting effort into the maintenance of one interface with a
-% transferable sensor system
-
-% An additonal factor not faced in the Tromba Moderna project is the
-% problem of scale in managing 98 rather than a single data stream of
-% data.
-
-% Material costs mitigated by creating a scaled version of the sound
-% board, providing only the length of string necessary under comparable
-% tension and thus resistance to the jack quill.
-
-% Finally, a further extension to the projects above was a commmitement to
-% open sourcing all aspects of the project, hardware, software, and data.
-
-% There are commercial systems for augmenting piano keyboards generating MIDI data.
-
-% An additional goal of the project was to design an interface that could be used to research
-% the influence the mechanism and it's resistances have on a player's performance.
-% For example, the Yamaha Disklavier series of pianos have an embedded sensor
-% system that allows for the instruments to operate in a "silent mode", but still
-% output MIDI data when the keys are struck. This has been a rich avenue for
-% exploring the affect of vibration on performance,
-% \cite{MusicalHaptics2018_04,MusicalHaptics2018_13} perception in instrument
-% quality \cite{MusicalHaptics2018_05}.
-
-% As well as a trying to offer a more engaging experience to museum attendees, the
-% interface was also designed as a potential probe for further research into the
-% idiosyncrasies of the harpsichord and the influence the mechanism has on
-% performance.
-
-Museums often face significant challenges in engaging visitors due to limitations in staffing and funding, which restrict the ways in which visitors can interact with collections \cite{Templeton2018, McAlpine2014}.
+Museums often face significant challenges in engaging visitors due to limitations in staffing and funding, which restrict how visitors can interact with collections \cite{Templeton2018, McAlpine2014}.
 For musical instrument museums, these challenges are compounded by the difficulty of preserving historical instruments in a playable condition \cite{McAlpine2014}.
-The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played, even when collections adhere to the strictest conservation protocols.
+The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played, even when collections adhere to the strictest conservation protocols \cite{NYT_strad}.
 A significant cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
@@
 A large sampling campaign was conducted, and a custom MIDI interface was designed to fulfil this requirement while preserving the original instruments' integrity.
-The MIDI keyboard, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, were used by visitors to trigger the instrument samples recorded with tailored strategies for each.
+The MIDI keyboard, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each.
 However, user tests identified a significant limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
-
-% To address this, McAlpine suggested designing a haptic keyboard enhanced with actuators capable of delivering positionally sensitive, real-time force feedback at the point of contact, inspired by previous projects such as Gillespie’s work on piano mechanisms \cite{Gillespie1996}.
 
@@
 
-\section{Design}\label{design}
+\section{Design Principles}\label{design}
 
-The present project initially drew inspiration from Timmermans' extension of Gillespie’s system \cite{Timmermans2020, Gillespie1996}.
-Timmermans’ process began with a single-key model of the mechanism.
-Similarly, we started with a 3-key model harpsichord mechanism, crafted by Graziano Bandini.
-Before incorporating actuators into the model, we sought to validate whether a faithful recreation of the harpsichord mechanism could effectively enable visitors to suspend disbelief.
+\begin{figure*}
+\includegraphics[width=\linewidth]{src/images/details.jpg}
+\caption{Details of the replica, including the key slots and sides, keys, and jacks with plectra.}\label{fig:details}
+\end{figure*}
 
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png} 
-  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023. 3-Key modelled used for initial prototyping. A: Wrestplank B: Key Lever C: Key cap D: soundboard E: Jack} 
-  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
-mechanism for an Italian style harpsichord of the XVI century.} 
-  \label{fig:3key}
-\end{figure}
+The design of the augmented replica keyboard for the Tagliavini Collection was guided by three principal constraints stipulated by the museum:
+\begin{itemize}
+\item \emph{No visible electronics}.
+To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
+\item \emph{Robustness and reliability}.
+The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
+\item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation, keeping artefacts to a minimum. 
+\end{itemize}
+These constraints demanded a non-invasive, robust design that was respectful of the historical authenticity of the instruments.
+The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
+While actuators can generate considerable force, they cannot replicate the wideband signals characteristic of the harpsichord's sharp and responsive haptics (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?), which differ significantly from those of a piano.
 
-A critical benefit of focusing on the development of a single interface with a transferable sensor system is the potential for scalability and adaptability.
-However, managing data streams presents a significant challenge.
-Unlike the Tromba Moderna project, which dealt with a single data stream, our design must manage 98 simultaneous data streams.
-This complexity necessitated cost-effective solutions, such as scaling down the soundboard to include only the necessary string length under comparable tension and resistance to the jack quill.
+\subsection{Summary of the Finalised Design}
 
-Another distinguishing feature of this project is its commitment to open-sourcing all hardware, software, and data components.
-While commercial systems for augmenting piano keyboards and generating MIDI data exist, this project emphasizes accessibility and transparency.
-
-An additional objective was to design an interface that supports research into how the harpsichord mechanism and its resistances influence a player’s performance.
-For example, the Yamaha Disklavier series incorporates embedded sensors enabling "silent mode" operation while outputting MIDI data when keys are struck.
-This has facilitated research into the effects of vibration on performance \cite{MusicalHaptics2018_04, MusicalHaptics2018_13} and the perception of instrument quality \cite{MusicalHaptics2018_05}.
-
-\section{Design}
-
-The design criteria was to create a keyboard interface that provided a comparable tactile response to that of a harpsichord in the italian style.
-
-Accommodations were made in the design to allow for as much flexibility as possible. 
-This also meant that designs for the electronics would have to include a level of flexibility in how they might be installed. 
-Surplus space was added into the design of the internal chamber of the model Figure \ref{fig:49-key-bottom}, but there was still a limit given the structural
-and aesthetic requirements.
-
-\section{Constraints}\label{design-problem}
-
-% Central philosophy of the Tagliavini Collection is that the instrument
-% should remain playable to the public. The condition of some instruments
-% is such that restoration would required a replacement of vast majority
-% of the instrument, would be too costly, would risk the instrument given
-% its fragility, or all of those in combination.
-
-% This is not an uncommon problem faced by musical instrument museums
-% worldwide. The central question the project tries to answer is ``how do
-% we conserve musical instruments in order to continue interacting with
-% them''
-
-% The approach taken was to recreate the interface of these instruments
-% not for acoustics purposes be purely for the mechanical response. As
-% exhibition for the general public in a musical instruments there were
-% aesthetic and functional constraints placed on the design of the
-% interface.
-
-% \subsection{Design Constraints}\label{design-constraints}
-
-There were constraints that were put in place by the Museum on the design of the exhibit.
+The final design is a 49-key harpsichord replica equipped with an optical sensor system to reproduce the tactile characteristics of the original instrument, visible in Figure \ref{fig:teaser}.
+Several key features were implemented to address the project’s requirements:
 
 \begin{itemize}
-\item
-  No electronics should be visible \footnote{with the exception of a pair of headphones for listening and a display to change parameters of the audio engine}
-\item
-  It should be robust and reliable, easily maintained by museum staff without complex technical processes
-\item
-  it should not augment or fix the mechanics, but present them with their original limitation
+\item \emph{QRE1113 infrared sensors}.
+These sensors were selected for their non-invasive properties and precision, enabling the detection of jack displacement without requiring structural modifications to the keyboard mechanism.
+\item \emph{Gradient stickers}.
+Greyscale gradient stickers were applied to each jack to allow the infrared sensors to make precise readings.
+This approach made calibration easier and enabled scalability from a 3-key prototype to the final 49-key design.
+\item \emph{Jack baffles}.
+Custom-designed baffles, 3D printed from dark pigmented PLA to minimise infrared reflection, were installed around the jacks to prevent cross-talk between adjacent sensors and ensure reliable data capture.
 \end{itemize}
-
-% Another constraint was that the design of the sensor system and the
-% model keyboard would have to carried out at a distance in separate
-% workshops. Time constraints meant the keyboard was being fabricated for
-% a sensor that did not yet exist.
-
-\subsection{Building the Model Keyboard}\label{keyboard}
-
-% PARAGRAPH FROM ROBERTO
-The construction of the keyboard required three months of work to be fully completed
-and followed the same construction phases of a real harpsichord.
-
-The model was constructed using the materials of the Italian harpsichord
-tradition. Parts are labelled in Figure \ref{fig:3key}.
-
-\begin{itemize}
-\item
-  Walnut wood for the wrestplank
-\item
-  Chestnut wood for the key levers
-\item
-  Boxwood and ebony for the key caps.
-\item
-  Cypress wood for the soundboard
-\item
-  Beech wood, a brass spring and a seagull feather plectrum for the 98 jacks
-\end{itemize}
-
-The body of the case is based on a rectangular poplar frame that leaves space
-for the installation of the electronic sensors.
-
-Deviations from the design of the Trasuntino are the rectangular, rather than
-'logarithmic' shape, removal of a bottom panel and omission of a short octave,
-instead spanning 4 octaves beginning on C.
-
-The strings are Malcolm Rose yellow brass wire, which are secured to wrought
-iron pins. The tension of each string was adjusted to provide resistance to
-plucking close to that of a harpsichord. Three felt strips were woven between
-the strings in order to dampen the vibration of the strings.
-
-The result is an interface instrument that is placed between the mechanical
-action of the keys and the generation of a synthetic sound. The touch of the
-wooden keys, the friction of the natural feathers on the string, the noise of
-the percussion at the base of the key together produce the sensation of being in
-front of a real harpsichord.
-
-% \subsubsection{Anatomy of Harpsichord
-% mechanics}\label{anatomy-of-harpsichord-mechanics}
-
-% \begin{itemize}
-% \item
-%   simple
-% \item
-%   aesthetics
-% \item
-%   split workload
-% \item
-%   parts of harpsichord
-
-%   \begin{itemize}
-%   \item
-%     jack
-
-%     \begin{itemize}
-%     \item
-%       quill
-%     \item
-%       jack body
-%     \item
-%       pivot
-%     \item
-%       tongue
-%     \end{itemize}
-%   \item
-%     key
-%   \item
-%     body
-%   \end{itemize}
-% \end{itemize}
-
-% The decision was made to design the electronics and the model keyboard simultaneously. 
-% This meant development development time was quicker and iteration on the idea.
-% Likely the first implementation would need re-designed and redesiginng
-% the electronics to fit the form of the keyboard was more cost effective
-% than building or carrying out major modifications to the model.
-
-% The model was designed to have two internal chambers a front and rear.
-% The front and back that was xxx by yyy by 370mm between the faceplate
-% and the jacks. The TONGUE THAT GOES INTO SLOT divides the chamber through
-% a small gap between the underside of the soundboard and the key guides.
-% The key guide needs to be deep enough to allow for the full movement of
-% the key and stop the key becoming unseated. The chamber then extends to
-% a support wall which divides the front and rear.
-
-% Discussion of spatial constraints and effects on electronics design and
-% installation is covered in \textbackslash section\{setup\}
-
-% The rear chamber was designed to house components for audio processing,
-% which is discussed further in \textbackslash section
 
@@
   \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
-  \caption{Underside of the full model keyboard showing two chambers. Front chamber at the top and rear chamber at the bottom.} 
+  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
   \Description{} 
@@
 
-% \footnote{https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf}
+A modular system of PCBs was designed to manage the sensors and process their output.
+The 49 QRE1113 sensors were distributed across seven boards, containing 7 sensors each.
+The PCBs were secured to the underside of the keyboard mechanism, allowing them to be adjusted during installation.
+Ribbon cables connected the PCBs, providing flexibility during assembly while maintaining a compact form factor.
+Additional modifications, including baffles and adhesive improvements, were made to optimise the reliability of the sensor system during calibration and use.
 
-% A harpsichord that
-% was economic enough to be deconstructed would also have been of a more
-% modern mechanical design and not of the design that was being recreated
-% with the model. Prototyping stage would likely be invasive and damage or
-% heavy modification likely. As such it was not a good use of funds as a
-% modern instrument fit for destruction did not offer benefits that the
-% small model with easy internal access did not already provide.
+The project built upon earlier NIME research on generating MIDI messages from piano keystrokes \cite{McPherson2013}, adapting it to address the unique characteristics of harpsichords and the specified constraints.
+Whereas the previous design emphasised continuous gesture tracking, this implementation required discrete key-triggered data to align with the needs of MIDI-triggered audio playback.
+The adoption of gradient stickers in place of continuous optical tracking embodies this adaptation, striking a balance between precision, cost-efficiency, and practicality.
 
-% \subsubsection{3-Key Prototype}\label{key-prototype}
+\subsection{Materials and Construction}
 
-% \begin{itemize}
-% \item
-%   proof of concept
-% \item
-%   available on loan from museum
-% \item
-%   easy access to internal structure
-% \end{itemize}
+The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord.
+Traditional materials were used, including walnut for the wrestplank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard.
+The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra.
+The design was inspired by Venetian harpsichords, particularly the Alessandro Trasuntino instrument at the San Colombano Museum.
+Details are visible in Figures \ref{fig:teaser} and \ref{fig:details}.
 
-% \subsubsection{49-Key Prototype}\label{key-prototype-1}
+The rectangular poplar frame deviates from the traditional logarithmic form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
+Two string orders, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
+Felt strips were added to dampen vibrations.
 
-% For the final interface a 49-key (4-octave) was constructed by luthier
-% \anon{Roberto Livi}. The keyboard is in the same style† as the 3-key model with
-% to jacks per key.
-
-% The keyboard is fully enclosed and internally has two chambers, front
-% and rear, in which the electronics could be installed
-% \textbackslash figure(underside 49-key).
-
-% Keyboard to be used for exhibition at san colombano where it would be
-% available to play by general public.
-
-% Timescale: from initial discussion to delivery of interface to the \anon{NEMUS}
-% lab was 8-months. Active assembly of the instrument was 3 months
-
-% Constructed From:
-
-% \begin{itemize}
-% \item
-%   for exhibition
-% \item
-%   problems of scale
-% \item
-%   problems of timescale
-% \end{itemize}
+The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile and auditory qualities.
 
@@
   a low latency. Following \cite{Jack2016} an latency was set as 10 ms
-  with an upper limit set to 25ms beyond which the sensors would be
+  with an upper limit set to 25 ms beyond which the sensors would be
   deemed unusable
@@
 
-% \subsection{Prototyping}\label{prototyping}
-
-% \subsubsection{Triggering a note}\label{triggering-a-note}
-
-% The message format covered in \textbackslash section
-
-% \section{Related Work and Motivation}\label{related-work-and-motivation}
-
-% \begin{itemize}
-% \item
-%   \cite{Timmermans2020}
-% \item
-%   \cite{McPherson2013} \cite{McPherson2019}
-% \item
-%   \cite{Mudd2013}
-% \item
-%   \cite{Fritz2017}
-% \end{itemize}
-
-% \section{Hardware Design}\label{hardware-design}
-
-% The hardware was designed with the xx in mind.
-
-% Using a similar apporach to @HapticKey the first step was to fabricate
-% or procure a model of a single working key. San Colombano provided a
-% model 3-key of a harposchord mechanism by Graziano Bandini
-% \textbackslash FIGURE. The 3-key model was used a test-bed for potential
-% sensors while designs were drawn for a full-scale model. The
-% cross-sectional nature of the model meant it was easy to fit and remove
-% sensors. Small size meant that it would also be easy to transfer between
-% workshops.
-
-% 3-key model also had a sliding jack rail which allowed for easy
-% disengagement of the physical mechanical response, which allowed for
-% some preliminary tests on the presence and absence of mechanism.
-
-% \subsection{Failed sensors}\label{failed-sensors}
-
-% A number of sensors were tried before the final system was decided upon.
-% An inertial measurement unit (IMU) was mounted to a single key which
-% contained a 6-degrees-of-freedom accelerometer and gyroscope. There was
-% a consistent set of conditions to trigger sound, but this was not easily
-% differentiable when other keys were played. The fitting time and cost of
-% a single IMU could b reduced when purchased in bulk but was far greater
-% than other systems considered.
-
-% A reed sensor and a hall sensor were mounted and magnets embedded within
-% the jacks. Adjusting the height of the sensor provided a reliable means
-% of detecting a threshold, in particular the hall sensors which could be
-% partially calibrated physically and also in software. The time taken to
-% calibrate a single key was time consuming. Ideally each sensor would be
-% placed on a thread for adjustment, but given the constraints on the
-% internal space it was decided any adjustment of heights would likely
-% mean an unacceptable length of time to install and calibrate. Embedding
-% magnets into the jacks was also did not satisfy the criteria of being
-% scalable and non-invasive.
-
-% Light dependant resistor mounted to the jack rail. as jacks were pressed
-% less light would be let through. Usable data but very dependant on
-% lighting conditions. Modifying the jack also did not satisfy the
-% non-intrusive criteria and the placement outside the internal chamber of
-% the model meant hiding electronics would be more difficult.
-
-% Finally, force sensitive resistors placed under the jack of each key was
-% considered but was never implemented. The sensors were to be placed
-% under each jack either above or below the padding on which the bottom of
-% the jack rests.
-
-% Further discussion of this approach in \textbackslash section.
-
-% \section{Electronics}\label{electronics}
-
-% Design of electronics carried out in tandem with construction of full
-% model. Since the Electronics were designed at a distanced to the final
-% full model, some flexibility was required in the design. Design decisions
-% that were not as optimal but would allow for change in wiring and
-% fitting should there be any issues during installation.
-
 The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.
 
-% The equipment required including soldering suitable for through-hole technology (THT) components and 3D printing.
-
-% Laser cutting and CNC routing facilities were used during the research
-% for the project but not required for the final implementation.
-
 This decisions was motivated by the project's goal to be open source and reasonably recreated without procurement of any additional specialist equipment.
-
-% NEMUS lab under construction during the first section of the project.
-% Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
-% \anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
-% terms of available equipment.
 
@@
 
-Another common application of the QRE1113 is within line following
-robots, where an array of QRE1113s are used in conjunction with black
-tape or reflective. A motorised system can then use this data to follow the line of tape and adjust its position.
+The optical sensors were directed at a greyscale gradient printed on a sticker and applied to the side of the jack.
+This was first trialled using a generic inkjet printer and double-sided tape. 
+For the final iteration, the stickers were printed on Coala 1D 100 Gloss PG 100 $\mu$m white gloss monomeric PVC with permanent grey acrylic dispersion adhesive \footnote{Datasheet: \url{https://www.antalis.co.uk/mediashare/g4media/pdf/TS_EN_COALA_1D_100_GLOSS_P_00_ISS_15022023.pdf}} using a HP Latex 115 and then cut with a Summa 150 roll cutter.
+The printer and roll cutter greatly reduced the manual labour for cutting stickers meaning the process would scale from the 6 jacks of the 3-key to the 98 jacks of the 49-key model.
 
-Applied to the jacks was a greyscale gradient printed the length full
-travel range of the jack.
+\begin{figure}  
+  \centering
+  \includegraphics[angle=90,width=\linewidth]{src/images/jack-w-tag.jpeg} 
+  \caption{} 
+  \Description{} 
+  \label{fig:jack-w-tag}
+\end{figure}
 
-For the first iteration this was performed on an inkjet printer and
-double-sided tape. For the final iteration the stickers were printed on
-(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
-µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
-cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
-reduced the manual labour for cutting stickers meaning the process would
-scale from the 3-key to the 49-key model.
+During testing it was found that their was a considerable amount cross-talk between adjacent sensors. 
+This meant the reading from key would be affected by the movement of the key next to it. 
+The cross-talk was assumed to be as a result of light reflecting from other jacks.
+Given that IR is used this was difficult to confirm by sight. 
+A set of baffles was designed for the PCB that would wrap round the jack and stop light from other jacks or sensors \ref{fig:baffles}. 
+The baffles were printed on a BAMBU X1 3D printers using PLA plastic. A dark pigmented PLA was used to avoid IR light simply being diffused. 
+After the baffles were fitted, the cross-talk was eliminated.
 
-% An initial sticker design was used that contained only the gradient. It
-% was found that the jacks were sticking in their slots and the edges on the
-% stickers easily peeled around the printing The sticker was redesigned to
-% be the full length of the jack which avoided the problem of peeling. The
-% larger are also mean the stickers adhered better to the jack Some
-% trimming was still required of each sticker to avoid them adding extra
-% resistance to the jack's movement.
+\begin{figure}  
+  \centering
+  \includegraphics[,width=\linewidth]{src/images/baffles.png} 
+  \caption{} 
+  \Description{} 
+  \label{fig:baffles}
+\end{figure}
 
-During testing it was found that their was a considerable amount
-cross-talk between adjacent sensors. This meant the reading from key
-would be affected by the movement of the key next to it. The cross-talk
-was assumed to be as a result of light reflecting from other jacks.
-Given that IR is used this was difficult to confirm by sight. A set of
-baffles was designed for the PCB that would wrap round the jack and stop
-light from other jacks or sensors. The baffles were printed on a BAMBU X1 3D
-printers using PLA plastic. A dark pigmented PLA was used to avoid IR
-light simply being diffused. After the baffles were fitted, the
-cross-talk was eliminated.
+Calibration to pluck thresholds were to be carried out manually.
+To allow for easy identification of the sensor being addressed during the calibration process, RGB LEDs were integrated into the design.
+The LEDs were also used in the identification of any malfunctioning or un-calibrated sensor.
 
-\subsubsection{RGB LEDs}\label{rgb-leds}
-
-for problems of scale:
-
-Addressing the problem of scale of identifying which key's threshold is
-currently being edited.
-
-The LEDs also allowed for easy identification for keys that were outside
-of expected values and when a key was triggered. This allowed for visual
-identification of malfunctioning or un-calibrated keys.
-
-The LEDs are on the same PCB as the sensors and are hidden when the
-name board is in place.2024
-
-Would suggest that this is a necessity of this kind to provide a means
-to identify and diagnose LEDs with state.
-
-TLC5940 or other PWM driver such as in \cite{McPherson2013} added
-complexity to assembling, but were avoid in favour of Addressable LEDs
-with integrated driver. The integrated LEDs share the same data line
-which has the limitation where a faulty LED will stop all following LEDs
-from working. An individual PWM driver per PCB would allow avoid
-situations where failure cascades, but would add complexity in assembly
-as the parts are typically SMD. Also, given the LEDs are fixed to the
-board and could not be repaired in situ, the solution would be to
-replace the PCB regardless of which approach was taken. Given the LEDs
-were only used in debugging and calibration and otherwise would not be
-visible, a simpler design approach was taken.
-
-\subsubsection{Multiplexing}\label{multiplexing}
+TLC5940 or other PWM driver such as in \cite{McPherson2013} add robustness should an LED fail, but complexity to assembly process. 
+Instead, the design uses addressable LEDs with integrated driver that share the same data line. 
+This has the limitation where a faulty LED will stop all subsequent LEDs in the line from working.
 
@@
 The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
-Transistor / Op-amp circuit could have been used similar to the one in
-\cite{McPherson2013}. Given time constraints it was beneficial to continue using and entirely THT based designed for easy assembly.
-
-% \subsection{PCB design}\label{pcb-design}
-
-% EAGLE PCBs designed using EAGLE.
-
-% Worth noting that AutoCAD has
-% deprecated EAGLE and the programme will no longer be supported from
-% 2026 Project files will need to be transitioned to another standard,
-% likely the KiCAD programme.
-
-% trade off between limiting number of boards required and allowing
-% compensation for variation in pitch.
-
-% Smaller PCBs means that spacing could be adjusted to accomodate the
-% changes in pitch between jacks. It also allowed for a modular approach
-% where any PCB could swapped for minimal cost.
-
-% \begin{itemize}
-% \item
-%   x7 QRE1113
-% \item
-%   x7 100R
-% \item
-%   x7 10kR
-% \item
-%   x1 CD4051BE Multipexer
-% \item
-%   jack / sensor pitch by board
-% \item
-%   7 boards adjust for pitch/spacing error
-% \item
-%   modular, replaceable
-% \item
-%   orientation problems
-
-%   \begin{itemize}
-%   \item
-%     legs break easily
-%   \item
-%     end mount like A.P.Mc. time consuming
-%   \item
-%     problem
-
-%     \begin{itemize}
-%     \item
-%       milling out keys
-%     \end{itemize}
-%   \end{itemize}
-% \item
-%   designed from distance
-% \item
-%   striped lines across pcbs
-% \item
-%   problems
-
-%   \begin{itemize}
-%   \item
-%     additional terminals uneeded resulted in lots of additional work
-%   \item
-%     long cabling increase risk of breaking
-%   \end{itemize}
-% \item
-%   solutions
-
-%   \begin{itemize}
-%   \item
-%     striping all ADC lines with FFC
-%   \item
-%     solder pads to select channel
-%   \end{itemize}
-% \end{itemize}
+Transistor / Op-amp circuit could have been used similar to the one in \cite{McPherson2013}.
 
@@
 
-The first version of the board exclusively used 2.54mm pitch headers.
+The first version of the board exclusively used 2.54 mm pitch headers.
 The reasons for this were flexibility when creating cable looms and to
@@
 This was the case for the LEDs as they initially were to be powered on
-5V. It was discovered that the 3v regulator of the nano was sufficient
+5V. It was discovered that the 3V regulator of the nano was sufficient
 to power all LEDs at a lower brightness while reducing current draw.
@@
 
-The 2.54mm pitch headers had two major drawbacks. First, the orientation
+The 2.54 mm pitch headers had two major drawbacks. First, the orientation
 of the cables could be reversed, which in the best case scenario would
@@
 
-A locator was adapted from a creative commons STL model
-\footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a
-generic crimping tool to be adapted for use with JST-PH crimps. 
-The locator was printed using a Bambu X1 3D printer. 
-A second alternative was assembling cables from other sets that were at hand.
+A locator was adapted from a creative commons STL model \footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a generic crimping tool to be adapted for use with JST-PH crimps.
 
@@
 more performative. To execute a single sensor read cycle of the firmware
-the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
+the STM32 took 40 ms, and both the Nano ESP32 and Nano BLE required 11 ms
 on average. The Nano ESP32 had high variability so it was decided to
 continue with the Nano BLE.
-
-% Further discussion of using ESP32 continues
-% in \textbackslash section\{Going forward\}
 
@@
 
-% \emph{It is worth noting that the FRAM model chosen stopped being
-% available during the project, meaning any future iteration would need to
-% adjust for another board in firmware}
-
-% The firmware was structured in a way that the interface could be changed
-% so long as the code required for any other board would comply with
-% function arguments and outputs.
-
-% Worth noting the fragility and
-% volatility when designing this kind of project and how easily it is for
-% the project to become broken.
-
 SD card would provide an easier way to interact with calibration data.
 The position of the controller board did not permit easy access.
-reliability and speed was preferred over storage capacity.
-
-\subsection{3D printing}\label{d-printing}
-
-Access to a 3D printer is was vital to the success of the project. Given
-the proliferation of 3D printing, particular in maker spaces in
-universities it was deemed an unreasonable requirement. 3D printed parts
-include the support brackets for PCBs, the baffles for the sensors,
-locator for crimping JST-PH crimps and a mount for the MacMini used for
-audio synthesis.
-
-The printer used throughout the project was a Bambu X1-Carbon using
-recycled PLA filament.
-
-Support brackets and baffles were made for the project while the MacMini
-bracket and JST locator were sourced from creative commons models.
-
-Any colour of filament can be used, though it is advised to use a dark
-coloured filament for the baffles.
-
-See each corresponding section for more detail on 3D printed components.
+Reliability and speed was preferred over storage capacity.
 
@@
 
-All electronics draws around 1.1amps of current at 5V with some
-fluctuation during boot.
+All electronics combined draw around 1.1 A of current at 5V with some fluctuation during start-up.
 
@@
 
-The sensors are in a state of constantly drawing power. It is possible
-that powering only when data is required could be reduce power enough.
+The sensors are in a state of constantly drawing power. 
+It is possible that powering only when data is required could be reduce power enough.
 
-Current designs have a separate MIDI device for each jack row. Combining
-the devices and also reducing current draw to below the common 0.5 amps
-for a USB port would certainly be a challenge.
+Current designs have a separate MIDI device for each jack row. 
+Combining the devices and also reducing current draw to below the common 0.5 amps for a USB port would certainly be a challenge.
 
@@
 it was previously above, a MIDI Note Off message is sent.
-
-% \subsection{Project structure}\label{project-structure}
-
-% The project is separated into three repositories encapsulating
-
-% \begin{itemize}
-% \item
-%   Firmware
-% \item
-%   Keyboard CAD Data
-% \item
-%   PCB CAD Data
-% \end{itemize}
-% 
-% \subsubsection{Firmware}\label{firmware}
-
-% Firmware repository contains all firmware for the full and smaller
-% models and components.
-
-% \subsubsection{Keyboard}\label{keyboard-1}
-
-% The Keyboard repository contains all plans for fabricating the keyboard,
-% measurements and 3D models of the jacks.
-
-% \subsubsection{PCB}\label{pcb}
-
-% PCB repository contains the EAGLE files for creating the sensor and
-% controller boards.
-
-% Given the regular spacing of components, scripts for automatic placement
-% of components based on parameter of jack pitch spacing and number of
-% sensors per jack.
 
@@
 A misreading in the original clearance between the keys and the PCB
-meant that the keys were blocked entirely from moving. A 10mm by 15mm
+meant that the keys were blocked entirely from moving. A 10 mm by 15 mm
 section was milled out of each key to create extra
@@
   \centering
-  \includegraphics[width=\linewidth]{src/images/keys-milled.jpeg} 
+  \includegraphics[width=\linewidth]{src/images/milled-keys.jpg} 
   \caption{} 
@@
 A current limitation is the latency in triggering MIDI output. Currently
-the latency is around 20 ms, but should ideally be closer to 5ms.
+the latency is around 20 ms, but should ideally be closer to 5 ms.
 
@@
 logic as the original firmware source, the latency was reduced to around
-10ms. ESP-IDF allows access to functionality of the ESP32 that is not
-exposed in the Arduino libraries. It is possible that the 10ms latency
+10 ms. ESP-IDF allows access to functionality of the ESP32 that is not
+exposed in the Arduino libraries. It is possible that the 10 ms latency
 could be improved upon, but there would be a cost in the form of

### 2025-01-22 src/article.tex
Updates from Overleaf

@@
 Technological advances have transformed how museums document, present and interpret their collections, fostering immersive experiences through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
-These technologies place an emphasis on experiential authenticity, enabling encounters that evoke the sensory, emotional, and intellectual essence of the past \cite{trant_Auth_1999}.
+These technologies emphasise experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.
 However, as Pine and Gilmore note \cite{pinegilmore_2007}, achieving authenticity requires museums to navigate the delicate balance between preservation and meaningful engagement—a challenge that is particularly evident in the case of historical musical instrument collections \cite{McAlpine2014}.
@@
 \includegraphics[width=\linewidth]{src/images/details.jpg}
-\caption{Details of the replica, including the key slots and sides, keys, and jacks with plectra.}\label{fig:details}
+\caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}\label{fig:details}
 \end{figure*}
@@
 
-The final design is a 49-key harpsichord replica equipped with an optical sensor system to reproduce the tactile characteristics of the original instrument, visible in Figure \ref{fig:teaser}.
+The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord replica equipped with an optical sensor system reading the amount of light reflected by gradient stickers applied to the jacks' sides.
 Several key features were implemented to address the project’s requirements:
@@
 
-A modular system of PCBs was designed to manage the sensors and process their output.
+A modular system of printed circuit boards (PCBs) was designed to manage the sensors and process their output.
 The 49 QRE1113 sensors were distributed across seven boards, containing 7 sensors each.
@@
 Whereas the previous design emphasised continuous gesture tracking, this implementation required discrete key-triggered data to align with the needs of MIDI-triggered audio playback.
-The adoption of gradient stickers in place of continuous optical tracking embodies this adaptation, striking a balance between precision, cost-efficiency, and practicality.
 
@@
 The design was inspired by Venetian harpsichords, particularly the Alessandro Trasuntino instrument at the San Colombano Museum.
-Details are visible in Figures \ref{fig:teaser} and \ref{fig:details}.
 
@@
 Felt strips were added to dampen vibrations.
-
 The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile and auditory qualities.
 
-\section{Hardware Design}
+Here’s the extended **Hardware Design** section in LaTeX, revised to meet a target of approximately 700 words while maintaining a coherent and logical structure:
 
-For the first iteration the data from the sensor would simply be used to test when a threshold had been passed.
-On passing the threshold a message would be sent to trigger the playback of a corresponding note.
+\section{Hardware Design}\label{hardware-design}
 
-For the initial prototyping stage, modifying an existing harpsichord was considered though ultimately discarded. An existing instrument would
-have provided a test for scaling the electronics, but internal measurements and layout would have been too different.
+The hardware design aimed to fulfil the constraints identified earlier, including non-invasiveness, reliability, and scalability, while ensuring a faithful recreation of the tactile response and auditory characteristics of historical harpsichords.
+To achieve these goals, the system evolved through iterative prototyping, beginning with simple threshold-based testing and culminating in a fully functional multi-sensor interface capable of triggering MIDI events.
 
-Using a similar approach to Timmermans' Haptic Key project, \cite{Timmermans2020} our first step was to fabricate or procure a model of a single working key.
-San Colombano was able to provide a model 3-key of a harpischord mechanism by Graziano Bandini (Figure \ref{fig:3key}).
+\subsection{Prototyping and Key Model}
+
+The initial stage of development focused on testing whether sensor data could reliably trigger MIDI playback.
+Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant discrepancies in internal measurements and layout.
+Instead, the team employed a custom 3-key harpsichord mechanism, crafted by Graziano Bandini and provided by San Colombano (Figure \ref{fig:3key}), as a foundation for prototyping.
+This approach followed a methodology similar to that used in Timmermans' Haptic Key project \cite{Timmermans2020}, enabling incremental validation of hardware and sensor integration.
+
+The 3-key model facilitated iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances.
+The model’s compact form allowed for rapid experimentation while preserving the physical properties of a full-scale harpsichord.
 
@@
 
-A set of criteria was drawn to which the final sensor system would have
-to satisfy. This criteria was used as a means of evaluating the
-feasibility throughout the prototyping phase.
+To ensure the system's suitability for a museum context, the following criteria were established to guide sensor selection and integration:
 
 \begin{itemize}
-\item
-  Non-invasive: The sensor system should not require major modifications
-  to the interface on which it is installed.
-\item
-  Low Latency: Reading sensors and then triggering an output should have
-  a low latency. Following \cite{Jack2016} an latency was set as 10 ms
-  with an upper limit set to 25 ms beyond which the sensors would be
-  deemed unusable
-\item
-  Reliable: Data obtained should be reliable and accurate enough not to
-  result in false positives or false negatives without extensive
-  filtering
-\item
-  Scalable: The system should be scalable both economically and
-  temporally. Given the open source nature of the project, it could not
-  be prohibitively expensive to carry out. Also any system that would
-  work on a single key would have to scale financially to potentially 50
-  or more. The time required to install and calibrate the sensors would
-  also have to scale.
-\item
-  Expandable: Any sensor system should be functionally expandable to
-  allow for the usage of the data for control over other parameters.
+    \item \textbf{Non-invasive:} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
+    \item \textbf{Low Latency:} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable \cite{Jack2016}.
+    \item \textbf{Reliability:} Sensor data had to be accurate, minimising false positives and negatives without excessive filtering.
+    \item \textbf{Scalability:} The design needed to be cost-effective and adaptable for up to 50 or more keys while maintaining consistent performance.
+    \item \textbf{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
 \end{itemize}
 
-The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.
+These criteria ensured that the design aligned with the open-source ethos of the project, allowing assembly in standard maker spaces without specialised equipment.
 
-This decisions was motivated by the project's goal to be open source and reasonably recreated without procurement of any additional specialist equipment.
+\subsection{Sensor Board Design}\label{sensor-board}
 
-The electronics system can be divided into two designs, the "sensor board" (Figure \ref{fig:7-sensor-board}) for components for detecting jack displacement and controller board for collecting and processing sensor signals.
+The final sensor system utilised QRE1113 infrared sensors, known for their precision and suitability for short-distance detection \cite{McPherson2013, McPherson2019}.
+The sensors were distributed across seven printed circuit boards (PCBs), each responsible for seven keys.
+Each PCB contained the following components:
+
+\begin{itemize}
+    \item 7 QRE1113 infrared sensors.
+    \item 7 100 $\Omega$ resistors and 7 10 k$\Omega$ resistors (later reduced to one resistor per board).
+    \item 1 Texas Instruments CD4051BE multiplexer for signal aggregation.
+\end{itemize}
+
+The QRE1113 sensors function by detecting infrared light reflected from nearby surfaces.
+Gradient stickers were affixed to each jack to provide a consistent reflective surface for the sensors, enabling precise tracking of jack displacement.
+This approach was chosen for its simplicity and scalability, allowing the design to progress from the 3-key prototype to a 49-key model without significant modifications.
+
+To eliminate cross-talk between adjacent sensors, 3D-printed baffles were installed on the PCBs (Figure \ref{fig:baffles}).
+These baffles, fabricated from dark pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
+Testing confirmed that this measure significantly improved data reliability.
 
@@
   \centering
-  \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
-  \caption{} 
+  \includegraphics[width=\linewidth]{src/images/simple-schematic-} 
+  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
   \Description{} 
-  \label{fig:7-sensor-board}
+  \label{fig:simple-schematic}
 \end{figure}
 
-\subsection{Sensor board}\label{sensor-board}
+\subsection{Controller Board}\label{controller-board}
 
-The 49 optical sensors are divided across 7 PCBs each containing
+The controller board, designed around the Arduino Nano BLE platform, served as the central hub for processing sensor data and triggering MIDI messages.
+The Nano BLE was selected for its compact form factor, integrated BLE MIDI functionality, and native USB support.
+These features simplified development while ensuring compatibility with existing MIDI-based synthesis systems.
 
-\begin{itemize}
-\item
-  x7 QRE1113 optical sensors
-\item
-  x7 100 $\Omega$ resistors
-\item
-  x7 10 k$\Omega$ resistors \footnote{a update on the design showed these could be reduced to a single resistor.}
-\item
-  x1 CD4051BE Multipexer
-\end{itemize}
+Initial iterations used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
+Custom cable looms were constructed using adapted crimping tools, ensuring robust connections without compromising accessibility.
 
-A number of sensors were tried before the final system was decided upon. 
-Sensors tested for the design included, a 6-degrees-of-freedom inertial measurement unit, reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
-None of the above satisfied the sensor criteria described and were thus discounted.
+Non-volatile memory, implemented using Ferroelectric RAM (FRAM), provided a reliable means of storing calibration data.
+This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in a museum setting.
 
-Following from the work in \cite{McPherson2013, McPherson2019}, we tested
-a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
-QRE1113 is a combination infrared LED and phototransistor sensitive to
-IR light. The phototransistor of QRE1113 provides data on how much light
-is present and in particular how much light is being reflected from a
-surface. Since reflected light will be proportional to the distance of
-a surface it is a good, close-proximity, distance sensor. Distance is
-the way in which the instruments in \cite{McPherson2013, McPherson2019}
-function as well the Moog Piano Bar from which it took inspiration.
+\subsection{Calibration and Power Management}\label{calibration}
 
-The limitation for taking a distance approach is that there is not one
-but 2 jacks per key. For each jack to be measured independently the data
-would need to come from the jacks themselves.
+Calibration was a critical component of the hardware design.
+Each sensor's threshold was manually adjusted using RGB LEDs integrated into the system.
+These LEDs provided visual feedback during calibration, enabling the identification of malfunctioning sensors and simplifying the alignment process.
+The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software (Figure \ref{fig:serial_monitor}).
 
-% They \emph{could} be used and individual trigger points set in
-% calibration, but this creates new problems.
+Power requirements for the system were estimated at 1.1 A at 5 V, with fluctuations during startup.
+While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors based on usage.
+
+\subsection{Final Assembly and Integration}
+
+The final 49-key harpsichord replica incorporated the sensor and controller boards into a rectangular frame, deviating slightly from traditional harpsichord designs to accommodate electronic components.
+Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
+
+The assembly process, conducted at the NEMUS lab in Bologna, involved meticulous adjustments to key clearances and wiring layouts.
+Modifications included milling sections of the keys to ensure adequate clearance for the PCBs, as well as varnishing the jacks to improve sticker adhesion.
+Subsequent iterations addressed these challenges, refining the system’s scalability and robustness.
 
@@
     \centering
-  \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
-  \caption{} 
-  \Description{} 
+    \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
+    \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+    \label{fig:3key}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
+    \caption{Cross-section illustrating sensor placement relative to the jack mechanism.}
     \label{fig:cross-section}
@@
 
-The optical sensors were directed at a greyscale gradient printed on a sticker and applied to the side of the jack.
-This was first trialled using a generic inkjet printer and double-sided tape. 
-For the final iteration, the stickers were printed on Coala 1D 100 Gloss PG 100 $\mu$m white gloss monomeric PVC with permanent grey acrylic dispersion adhesive \footnote{Datasheet: \url{https://www.antalis.co.uk/mediashare/g4media/pdf/TS_EN_COALA_1D_100_GLOSS_P_00_ISS_15022023.pdf}} using a HP Latex 115 and then cut with a Summa 150 roll cutter.
-The printer and roll cutter greatly reduced the manual labour for cutting stickers meaning the process would scale from the 6 jacks of the 3-key to the 98 jacks of the 49-key model.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/baffles.png}
+    \caption{3D-printed baffles designed to prevent cross-talk between adjacent sensors.}
+    \label{fig:baffles}
+\end{figure}
 
@@
     \centering
-  \includegraphics[angle=90,width=\linewidth]{src/images/jack-w-tag.jpeg} 
-  \caption{} 
-  \Description{} 
-  \label{fig:jack-w-tag}
+    \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
+    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.}
+    \label{fig:serial_monitor}
 \end{figure}
 
-During testing it was found that their was a considerable amount cross-talk between adjacent sensors. 
-This meant the reading from key would be affected by the movement of the key next to it. 
-The cross-talk was assumed to be as a result of light reflecting from other jacks.
-Given that IR is used this was difficult to confirm by sight. 
-A set of baffles was designed for the PCB that would wrap round the jack and stop light from other jacks or sensors \ref{fig:baffles}. 
-The baffles were printed on a BAMBU X1 3D printers using PLA plastic. A dark pigmented PLA was used to avoid IR light simply being diffused. 
-After the baffles were fitted, the cross-talk was eliminated.
+% \section{Hardware Design}
 
-\begin{figure}  
-  \centering
-  \includegraphics[,width=\linewidth]{src/images/baffles.png} 
-  \caption{} 
-  \Description{} 
-  \label{fig:baffles}
-\end{figure}
+% For the first iteration the data from the sensor would simply be used to test when a threshold had been passed.
+On passing the threshold, a message will be sent to trigger the playback of a corresponding note.
 
-Calibration to pluck thresholds were to be carried out manually.
+% For the initial prototyping stage, modifying an existing harpsichord was considered though ultimately discarded. An existing instrument would
+% have provided a test for scaling the electronics, but internal measurements and layout would have been too different.
+
+% Using a similar approach to Timmermans' Haptic Key project, \cite{Timmermans2020}, our first step was to fabricate or procure a model of a single working key.
+San Colombano provided a model 3-key of a harpsichord mechanism by Graziano Bandini (Figure \ref{fig:3key}).
+
+% \subsection{Sensor Criteria}\label{sensor-criteria}
+
+% A set of criteria was drawn to which the final sensor system would have to satisfy. These criteria were used to evaluate the feasibility throughout the prototyping phase.
+
+% \begin{itemize}
+% \item
+%   Non-invasive: The sensor system should not require major modifications
+%   to the interface on which it is installed.
+% \item
+%   Low Latency: Reading sensors and then triggering an output should have
+%   a low latency. Following \cite{Jack2016} an latency was set as 10 ms
+%   with an upper limit set to 25 ms beyond which the sensors would be
+%   deemed unusable
+% \item
+%   Reliable: Data obtained should be reliable and accurate enough not to
+%   result in false positives or false negatives without extensive
+%   filtering
+% \item
+%   Scalable: The system should be scalable both economically and
+%   temporally. Given the open source nature of the project, it could not
+%   be prohibitively expensive to carry out. Also any system that would
+%   work on a single key would have to scale financially to potentially 50
+%   or more. The time required to install and calibrate the sensors would
+%   also have to scale.
+% \item
+%   Expandable: Any sensor system should be functionally expandable to
+%   allow for the usage of the data for control over other parameters.
+% \end{itemize}
+
+% The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.
+This decision was motivated by the project's goal to be open source and reasonably recreated without procuring additional specialist equipment.
+
+% The electronics system can be divided into two designs, the "sensor board" (Figure \ref{fig:7-sensor-board}) for components for detecting jack displacement and controller board for collecting and processing sensor signals.
+
+% \begin{figure}  
+%   \centering
+%   \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:7-sensor-board}
+% \end{figure}
+
+% \subsection{Sensor board}\label{sensor-board}
+
+% The 49 optical sensors are divided across 7 PCBs each containing
+
+% \begin{itemize}
+% \item
+%   x7 QRE1113 optical sensors
+% \item
+%   x7 100 $\Omega$ resistors
+% \item
+%   x7 10 k$\Omega$ resistors \footnote{a update on the design showed these could be reduced to a single resistor.}
+% \item
+%   x1 CD4051BE Multipexer
+% \end{itemize}
+
+% A number of sensors were tried before the final system was decided upon. 
+% Sensors tested for the design included, a 6-degrees-of-freedom inertial measurement unit, reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
+% None of the above satisfied the sensor criteria described and were thus discounted.
+
+% Following from the work in \cite{McPherson2013, McPherson2019}, we tested
+% a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
+% QRE1113 is a combination infrared LED and phototransistor sensitive to
+% IR light. The phototransistor of QRE1113 provides data on how much light
+% is present and in particular how much light is being reflected from a
+% surface. Since reflected light will be proportional to the distance of
+% a surface it is a good, close-proximity, distance sensor. Distance is
+% the way in which the instruments in \cite{McPherson2013, McPherson2019}
+% function as well the Moog Piano Bar from which it took inspiration.
+
+% The limitation for taking a distance approach is that there is not one
+% but 2 jacks per key. For each jack to be measured independently the data
+% would need to come from the jacks themselves.
+
+% % They \emph{could} be used and individual trigger points set in
+% % calibration, but this creates new problems.
+
+% \begin{figure}  
+%   \centering
+%   \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:cross-section}
+% \end{figure}
+
+% The optical sensors were directed at a greyscale gradient printed on a sticker and applied to the side of the jack.
+% This was first trialled using a generic inkjet printer and double-sided tape. 
+% For the final iteration, the stickers were printed on Coala 1D 100 Gloss PG 100 $\mu$m white gloss monomeric PVC with permanent grey acrylic dispersion adhesive \footnote{Datasheet: \url{https://www.antalis.co.uk/mediashare/g4media/pdf/TS_EN_COALA_1D_100_GLOSS_P_00_ISS_15022023.pdf}} using a HP Latex 115 and then cut with a Summa 150 roll cutter.
+% The printer and roll cutter greatly reduced the manual labour for cutting stickers meaning the process would scale from the 6 jacks of the 3-key to the 98 jacks of the 49-key model.
+
+% \begin{figure}  
+%   \centering
+%   \includegraphics[angle=90,width=\linewidth]{src/images/jack-w-tag.jpeg} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:jack-w-tag}
+% \end{figure}
+
+% During testing it was found that their was a considerable amount cross-talk between adjacent sensors. 
+% This meant the reading from key would be affected by the movement of the key next to it. 
+% The cross-talk was assumed to be as a result of light reflecting from other jacks.
+% Given that IR is used this was difficult to confirm by sight. 
+% A set of baffles was designed for the PCB that would wrap round the jack and stop light from other jacks or sensors \ref{fig:baffles}. 
+% The baffles were printed on a BAMBU X1 3D printers using PLA plastic. A dark pigmented PLA was used to avoid IR light simply being diffused. 
+% After the baffles were fitted, the cross-talk was eliminated.
+
+% \begin{figure}  
+%   \centering
+%   \includegraphics[,width=\linewidth]{src/images/baffles.png} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:baffles}
+% \end{figure}
+
+% Calibration to pluck thresholds were to be carried out manually.
 To allow for easy identification of the sensor being addressed during the calibration process, RGB LEDs were integrated into the design.
@@
 
-TLC5940 or other PWM driver such as in \cite{McPherson2013} add robustness should an LED fail, but complexity to assembly process. 
-Instead, the design uses addressable LEDs with integrated driver that share the same data line. 
-This has the limitation where a faulty LED will stop all subsequent LEDs in the line from working.
+% TLC5940 or other PWM driver such as in \cite{McPherson2013} add robustness should an LED fail, but complexity to assembly process. 
+% Instead, the design uses addressable LEDs with integrated driver that share the same data line. 
+% This has the limitation where a faulty LED will stop all subsequent LEDs in the line from working.
 
-The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
-We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
-An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
-The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
-Transistor / Op-amp circuit could have been used similar to the one in \cite{McPherson2013}.
+% The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
+% We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
+% An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
+% The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
+% Transistor / Op-amp circuit could have been used similar to the one in \cite{McPherson2013}.
 
-\subsection{Controller Board}\label{controller-board}
+% \subsection{Controller Board}\label{controller-board}
 
-Controller board designed around Arduino nano form factor. The Nano is
-connected through headers to allow easy switching between chips. Should
-another chip in a Nano for factor be found it could supplant the current
-Nano without any need for a redesign.
+% Controller board designed around Arduino nano form factor. The Nano is
+% connected through headers to allow easy switching between chips. Should
+% another chip in a Nano for factor be found it could supplant the current
+% Nano without any need for a redesign.
 
-The first version of the board exclusively used 2.54 mm pitch headers.
-The reasons for this were flexibility when creating cable looms and to
-leave open the possibility of rewiring the PCB.
+% The first version of the board exclusively used 2.54 mm pitch headers.
+% The reasons for this were flexibility when creating cable looms and to
+% leave open the possibility of rewiring the PCB.
 
-This was the case for the LEDs as they initially were to be powered on
-5V. It was discovered that the 3V regulator of the nano was sufficient
-to power all LEDs at a lower brightness while reducing current draw.
-Since brightness was not important.
+% This was the case for the LEDs as they initially were to be powered on
+% 5V. It was discovered that the 3V regulator of the nano was sufficient
+% to power all LEDs at a lower brightness while reducing current draw.
+% Since brightness was not important.
 
-The 2.54 mm pitch headers had two major drawbacks. First, the orientation
-of the cables could be reversed, which in the best case scenario would
-simply mean the system did not work correctly and in the worst case it
-could potentially damage the components.
+% The 2.54 mm pitch headers had two major drawbacks. First, the orientation
+% of the cables could be reversed, which in the best case scenario would
+% simply mean the system did not work correctly and in the worst case it
+% could potentially damage the components.
 
-Secondly, the headers do not connect securely and wait of the cabling
-was enough for components to become disconnected as the harpsichord was
-handled.
+% Secondly, the headers do not connect securely and wait of the cabling
+% was enough for components to become disconnected as the harpsichord was
+% handled.
 
-To address this, once the wiring for the controller board was finalised
-a new design was drafted this time using JST-PH connectors. No
-commercial cables were immediately available in the configuration and
-length required, as such a cable loom needed to be made which meant
-crimping cables by hand. A crimping tool from JST is prohibitively
-expensive compared to the cost of other components in the project
-\footnote{from RS \url{https://uk.rs-online.com/web/p/crimp-tools/6880877}
-£508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
-not readily available in the workshops at the project's disposal.
-Eventually a tool was found, but an alternative approach was also
-researched in order to maintain the criteria of accessibility.
+% To address this, once the wiring for the controller board was finalised
+% a new design was drafted this time using JST-PH connectors. No
+% commercial cables were immediately available in the configuration and
+% length required, as such a cable loom needed to be made which meant
+% crimping cables by hand. A crimping tool from JST is prohibitively
+% expensive compared to the cost of other components in the project
+% \footnote{from RS \url{https://uk.rs-online.com/web/p/crimp-tools/6880877}
+% £508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
+% not readily available in the workshops at the project's disposal.
+% Eventually a tool was found, but an alternative approach was also
+% researched in order to maintain the criteria of accessibility.
 
-A locator was adapted from a creative commons STL model \footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a generic crimping tool to be adapted for use with JST-PH crimps.
+% A locator was adapted from a creative commons STL model \footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a generic crimping tool to be adapted for use with JST-PH crimps.
 
-\subsubsection{Arduino Nano BLE}\label{arduino-nano-ble}
+% \subsubsection{Arduino Nano BLE}\label{arduino-nano-ble}
 
-Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
-appealing as it would allow for easy changes between chipsets without
-requiring any rewiring and thus no new PCBs would need to be created if
-a better alternative was found later in the project. Another benefit to
-the Arduino Nano boards was that typically had native-USB meaning the
-could be programmed as a hardware USB MIDI device directly.
+% Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
+% appealing as it would allow for easy changes between chipsets without
+% requiring any rewiring and thus no new PCBs would need to be created if
+% a better alternative was found later in the project. Another benefit to
+% the Arduino Nano boards was that typically had native-USB meaning the
+% could be programmed as a hardware USB MIDI device directly.
 
-The Nano BLE was used in the initial stages for its integrated inertial measurment unit.
+% The Nano BLE was used in the initial stages for its integrated inertial measurment unit.
 
-When it was finalised that the project would proceed using the QRE1113 a
-number of microcontroller units (MCU) were considered.
+% When it was finalised that the project would proceed using the QRE1113 a
+% number of microcontroller units (MCU) were considered.
 
-The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
-Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
+% The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
+% Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
 
-A rough performance test was later carried out using an STM32
-NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
-more performative. To execute a single sensor read cycle of the firmware
-the STM32 took 40 ms, and both the Nano ESP32 and Nano BLE required 11 ms
-on average. The Nano ESP32 had high variability so it was decided to
-continue with the Nano BLE.
+% A rough performance test was later carried out using an STM32
+% NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
+% more performative. To execute a single sensor read cycle of the firmware
+% the STM32 took 40 ms, and both the Nano ESP32 and Nano BLE required 11 ms
+% on average. The Nano ESP32 had high variability so it was decided to
+% continue with the Nano BLE.
 
-BLE functionality also meant if the cable connection between the nano
-and computer running audio synthesis was impractical, BLE MIDI was
-available as a fallback. The Nano BLE also has a 12-bit DAC, which
-provided a contingency should the default 10-bit not be sensitive enough
-for the signal from the sensors.
+% BLE functionality also meant if the cable connection between the nano
+% and computer running audio synthesis was impractical, BLE MIDI was
+% available as a fallback. The Nano BLE also has a 12-bit DAC, which
+% provided a contingency should the default 10-bit not be sensitive enough
+% for the signal from the sensors.
 
-Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
-functionality, but it was found that 2 ADC channels were in fact not
-useable and would have required additional multiplexers.
+% Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
+% functionality, but it was found that 2 ADC channels were in fact not
+% useable and would have required additional multiplexers.
 
-\subsubsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
+% \subsubsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
 
-There would need to be the ability to store values on non-volatile
-memory so that the electronics could be powered-down without losing
-data.
+% There would need to be the ability to store values on non-volatile
+% memory so that the electronics could be powered-down without losing
+% data.
 
-Adjusting the values through firmware would not have been flexible or
-practical. Instead an Ferroelectric RAM unit was chosen and connected
-using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
-was primarily chosen for it's ready availability.
+% Adjusting the values through firmware would not have been flexible or
+% practical. Instead an Ferroelectric RAM unit was chosen and connected
+% using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
+% was primarily chosen for it's ready availability.
 
-Most microcontrollers have some variety of non-volatile program memory
-that can be read and writ accessed but in most cases this memory is
-cleared during compilation and re-flashing. The small cost was far
-outweighed by the potential for data loss if using onboard RAM.
+% Most microcontrollers have some variety of non-volatile program memory
+% that can be read and writ accessed but in most cases this memory is
+% cleared during compilation and re-flashing. The small cost was far
+% outweighed by the potential for data loss if using onboard RAM.
 
-SD card would provide an easier way to interact with calibration data.
-The position of the controller board did not permit easy access.
-Reliability and speed was preferred over storage capacity.
+% SD card would provide an easier way to interact with calibration data.
+% The position of the controller board did not permit easy access.
+% Reliability and speed was preferred over storage capacity.
 
-\subsection{Power}\label{power}
+% \subsection{Power}\label{power}
 
-All electronics combined draw around 1.1 A of current at 5V with some fluctuation during start-up.
+% All electronics combined draw around 1.1 A of current at 5V with some fluctuation during start-up.
 
-Arduino BLE has some functionality disabled to limit current draw.
+% Arduino BLE has some functionality disabled to limit current draw.
 
-Ideally the unit could be bus powered from USB.
+% Ideally the unit could be bus powered from USB.
 
-The sensors are in a state of constantly drawing power. 
-It is possible that powering only when data is required could be reduce power enough.
+% The sensors are in a state of constantly drawing power. 
+% It is possible that powering only when data is required could be reduce power enough.
 
-Current designs have a separate MIDI device for each jack row. 
-Combining the devices and also reducing current draw to below the common 0.5 amps for a USB port would certainly be a challenge.
+% Current designs have a separate MIDI device for each jack row. 
+% Combining the devices and also reducing current draw to below the common 0.5 amps for a USB port would certainly be a challenge.
 
-\section{Firmware Design}\label{firmware-design}
+% \section{Firmware Design}\label{firmware-design}
 
-Arduino platform chosen for wide adoption, relative ease of use and
-library support for components.
+% Arduino platform chosen for wide adoption, relative ease of use and
+% library support for components.
 
-Firmware checks components are connected then reads the FRAM to see if
-threshold data is already present. The rotary encoder and it's momentary
-switch are polled for any change in state. A key is selected with the
-rotary encoder while the momentary switch is used for changing between
-states for key selection and threshold adjustment A double click of the
-rotary encoder writes the current threshold values to FRAM. The sensors
-for the current multiplexer channels are polled. If any sensor reads
-above the threshold when it was previously below a MIDI On message is
-sent. For the reverse case where the value was below the threshold when
-it was previously above, a MIDI Note Off message is sent.
+% Firmware checks components are connected then reads the FRAM to see if
+% threshold data is already present. The rotary encoder and it's momentary
+% switch are polled for any change in state. A key is selected with the
+% rotary encoder while the momentary switch is used for changing between
+% states for key selection and threshold adjustment A double click of the
+% rotary encoder writes the current threshold values to FRAM. The sensors
+% for the current multiplexer channels are polled. If any sensor reads
+% above the threshold when it was previously below a MIDI On message is
+% sent. For the reverse case where the value was below the threshold when
+% it was previously above, a MIDI Note Off message is sent.
 
-\section{Implementation}\label{implementation}
+% \section{Implementation}\label{implementation}
 
-\begin{figure}  
-  \centering
-  \includegraphics[,width=\linewidth]{src/images/exhibition-user-1.jpeg} 
-  \caption{} 
-  \Description{} 
-  \label{fig:exhibition}
-\end{figure}
+% \begin{figure}  
+%   \centering
+%   \includegraphics[,width=\linewidth]{src/images/exhibition-user-1.jpeg} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:exhibition}
+% \end{figure}
 
-\subsection{Installation}\label{installation}
+% \subsection{Installation}\label{installation}
 
-The electronic components were partly pre-assembled at the university of
-Edinburgh. Though all measurements were known ahead, enough flexibility
-was needed to allow for problems to solved in situ.
+% The electronic components were partly pre-assembled at the university of
+% Edinburgh. Though all measurements were known ahead, enough flexibility
+% was needed to allow for problems to solved in situ.
 
-The final assembly took place at the NEMUS lab at the University of
-Bologna over the final week in October 2024 for delivery at San
-Colombano for exhibition in the first week of November in 2024.
+% The final assembly took place at the NEMUS lab at the University of
+% Bologna over the final week in October 2024 for delivery at San
+% Colombano for exhibition in the first week of November in 2024.
 
-A misreading in the original clearance between the keys and the PCB
-meant that the keys were blocked entirely from moving. A 10 mm by 15 mm
-section was milled out of each key to create extra
-clearance. The milling required the removal, shortening and refitting of
-the leather pads for the jacks. The extra work
-added an extra day to the full install.
+% A misreading in the original clearance between the keys and the PCB
+% meant that the keys were blocked entirely from moving. A 10 mm by 15 mm
+% section was milled out of each key to create extra
+% clearance. The milling required the removal, shortening and refitting of
+% the leather pads for the jacks. The extra work
+% added an extra day to the full install.
 
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/milled-keys.jpg} 
-  \caption{} 
-  \Description{} 
-  \label{fig:milling}
-\end{figure}
+% \begin{figure}  
+%   \centering
+%   \includegraphics[width=\linewidth]{src/images/milled-keys.jpg} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:milling}
+% \end{figure}
 
-A redesign of the PCBs meant that removing the material would no longer
-be necessary.
+% A redesign of the PCBs meant that removing the material would no longer
+% be necessary.
 
-The PCBs shared power and data lines for components with the exception
-of the output from the multiplexer. Ribbon cable was cut to size to
-allow for some room to readjust the distance of the sensors from the
-jacks. The data signals from each PCB were connected using a separate
-cable loom. Soldering cables and inspecting each join required another
-day work. This did allow for readjustments to the position of the
-controller board Measurements of the rear chamber did not match with the
-Mac Mini that was to be used a different layout was decided for fitting
-the controller board/ Cutting cables to size allowed for a great deal of
-flexibility but at the cost of time. After the first install it was
-clearer what constraints were in place for fitting and an alternative
-wiring system using flat flex cable was implement (discussed in
-\textbackslash section)
+% The PCBs shared power and data lines for components with the exception
+% of the output from the multiplexer. Ribbon cable was cut to size to
+% allow for some room to readjust the distance of the sensors from the
+% jacks. The data signals from each PCB were connected using a separate
+% cable loom. Soldering cables and inspecting each join required another
+% day work. This did allow for readjustments to the position of the
+% controller board Measurements of the rear chamber did not match with the
+% Mac Mini that was to be used a different layout was decided for fitting
+% the controller board/ Cutting cables to size allowed for a great deal of
+% flexibility but at the cost of time. After the first install it was
+% clearer what constraints were in place for fitting and an alternative
+% wiring system using flat flex cable was implement (discussed in
+% \textbackslash section)
 
-Fitting PCBs required the removal of all keys. The support brackets were
-assembled and screwed into the the underside. With the harpsichord on
-it's back the sensors were tested to ensure an expected range.
-Self-tapping round head screws were used to to attach the brackets. The
-screws were not over tightened to allow for brackets to move. The
-brackets were not easily adjusted once the key were replaced. Adjusted
-to a best average with subsequent adjustment carried out on the
-microcontroller.
+% Fitting PCBs required the removal of all keys. The support brackets were
+% assembled and screwed into the the underside. With the harpsichord on
+% it's back the sensors were tested to ensure an expected range.
+% Self-tapping round head screws were used to to attach the brackets. The
+% screws were not over tightened to allow for brackets to move. The
+% brackets were not easily adjusted once the key were replaced. Adjusted
+% to a best average with subsequent adjustment carried out on the
+% microcontroller.
 
-As the PCBs were fitted the jacks had the gradient stickers applied to
-them. The jacks had been coated with a varnish on the short side to
-provide a better surface for the stickers to adhere to. After the
-stickers were applied excess material was trimmed from each jack. The
-jack were then check that they moved freely in their slot.
+% As the PCBs were fitted the jacks had the gradient stickers applied to
+% them. The jacks had been coated with a varnish on the short side to
+% provide a better surface for the stickers to adhere to. After the
+% stickers were applied excess material was trimmed from each jack. The
+% jack were then check that they moved freely in their slot.
 
-After fitting the PCBs preliminary calibration was carried out. Before
-the keys were refitted the jacks were put back in place. The data from
-each key was confirmed to be in a similar range. Any key showing
-anomalous data had it's jack removed and re-inspected. The PCB fitting
-process and refitting required another day as faults were found in
-wiring or misalignment of the stickers. After the sensors were
-confirmed to be functioning the keys were put back.
+% After fitting the PCBs preliminary calibration was carried out. Before
+% the keys were refitted the jacks were put back in place. The data from
+% each key was confirmed to be in a similar range. Any key showing
+% anomalous data had it's jack removed and re-inspected. The PCB fitting
+% process and refitting required another day as faults were found in
+% wiring or misalignment of the stickers. After the sensors were
+% confirmed to be functioning the keys were put back.
 
-The calibration workflow was with \anon{Dr. Craig Webb} who was unfamiliar with
-the system. Adjustments were made to the firmware to to make it more
-intuitive for the user. Addressable LEDs became vital as it allowed for
-easy identification of the key being calibrated. They were also used to
-identify any keys whose current reading was beyond what was expected.
+% The calibration workflow was with \anon{Dr. Craig Webb} who was unfamiliar with
+% the system. Adjustments were made to the firmware to to make it more
+% intuitive for the user. Addressable LEDs became vital as it allowed for
+% easy identification of the key being calibrated. They were also used to
+% identify any keys whose current reading was beyond what was expected.
 
-The rotary encoder was used as the interface to select a key, adjust
-it's threshold and update the Ferroelectric RAM (FRAM) with new values.
+% The rotary encoder was used as the interface to select a key, adjust
+% it's threshold and update the Ferroelectric RAM (FRAM) with new values.
 
-A process was implemented where the key was selected by playing, but
-this caused problems in situations where the data was unusual due to
-misalignment of the sensors.
+% A process was implemented where the key was selected by playing, but
+% this caused problems in situations where the data was unusual due to
+% misalignment of the sensors.
 
-This would likely be a better system for when the keyboard was fully
-operational.
+% This would likely be a better system for when the keyboard was fully
+% operational.
 
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{images/serial_monitor.png} 
-  \caption{} 
-  \Description{} 
-  \label{fig:serial_monitor}
-\end{figure}
+% \begin{figure}  
+%   \centering
+%   \includegraphics[width=\linewidth]{images/serial_monitor.png} 
+%   \caption{} 
+%   \Description{} 
+%   \label{fig:serial_monitor}
+% \end{figure}
 
-The thresholds were calibrated visually utilising the serial plotter of
-the Arduino IDE (Figure \ref{fig:serial_monitor}) and the open source MIDI
-Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}.
+% The thresholds were calibrated visually utilising the serial plotter of
+% the Arduino IDE (Figure \ref{fig:serial_monitor}) and the open source MIDI
+% Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}.
 
-\begin{itemize}
-\item
-  partial pre-assembly
-\item
-  handsolder components
+% \begin{itemize}
+% \item
+%   partial pre-assembly
+% \item
+%   handsolder components
 
-  \begin{itemize}
-  \item
-    analog lines
-  \end{itemize}
-\item
-  connecting supports
-\end{itemize}
+%   \begin{itemize}
+%   \item
+%     analog lines
+%   \end{itemize}
+% \item
+%   connecting supports
+% \end{itemize}
 

### 2025-01-22 src/article.tex
Updates from Overleaf

@@
 
-Here’s the extended **Hardware Design** section in LaTeX, revised to meet a target of approximately 700 words while maintaining a coherent and logical structure:
-
 \section{Hardware Design}\label{hardware-design}
 
-The hardware design aimed to fulfil the constraints identified earlier, including non-invasiveness, reliability, and scalability, while ensuring a faithful recreation of the tactile response and auditory characteristics of historical harpsichords.
+The hardware design aimed to fulfil the earlier constraints, including non-invasiveness, reliability, and scalability.
 To achieve these goals, the system evolved through iterative prototyping, beginning with simple threshold-based testing and culminating in a fully functional multi-sensor interface capable of triggering MIDI events.
@@
 The initial stage of development focused on testing whether sensor data could reliably trigger MIDI playback.
-Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant discrepancies in internal measurements and layout.
-Instead, the team employed a custom 3-key harpsichord mechanism, crafted by Graziano Bandini and provided by San Colombano (Figure \ref{fig:3key}), as a foundation for prototyping.
+Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant internal measurement and layout discrepancies.
+Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was acquired from the museum and used as a foundation for prototyping.
 This approach followed a methodology similar to that used in Timmermans' Haptic Key project \cite{Timmermans2020}, enabling incremental validation of hardware and sensor integration.
@@
 The model’s compact form allowed for rapid experimentation while preserving the physical properties of a full-scale harpsichord.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
+    \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini at San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+    \label{fig:3key}
+\end{figure}
 
@@
 \begin{itemize}
-    \item \textbf{Non-invasive:} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
-    \item \textbf{Low Latency:} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable \cite{Jack2016}.
-    \item \textbf{Reliability:} Sensor data had to be accurate, minimising false positives and negatives without excessive filtering.
-    \item \textbf{Scalability:} The design needed to be cost-effective and adaptable for up to 50 or more keys while maintaining consistent performance.
-    \item \textbf{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
+    \item \emph{Non-invasive:} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
+    \item \emph{Low Latency:} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable \cite{Jack2016}.
+    \item \emph{Reliability:} Sensor data had to be accurate, minimising false positives and negatives without excessive filtering.
+    \item \emph{Scalability:} The design needed to be cost-effective and adaptable for up to 50 or more keys while maintaining consistent performance.
+    \item \emph{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
 \end{itemize}
@@
 \subsection{Sensor Board Design}\label{sensor-board}
+
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\linewidth]{src/images/bl}
+%     \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini at San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+%     \label{fig:3key}
+% \end{figure*}
 
@@
 The QRE1113 sensors function by detecting infrared light reflected from nearby surfaces.
-Gradient stickers were affixed to each jack to provide a consistent reflective surface for the sensors, enabling precise tracking of jack displacement.
+The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors, enabling precise tracking of jack displacement.
 This approach was chosen for its simplicity and scalability, allowing the design to progress from the 3-key prototype to a 49-key model without significant modifications.
@@
 To eliminate cross-talk between adjacent sensors, 3D-printed baffles were installed on the PCBs (Figure \ref{fig:baffles}).
-These baffles, fabricated from dark pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
+These baffles, fabricated from dark-pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
 Testing confirmed that this measure significantly improved data reliability.
@@
   \centering
-  \includegraphics[width=\linewidth]{src/images/simple-schematic-} 
-  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
+  \includegraphics[width=\linewidth]{src/images/simple-schematic-bw-.jpg} 
+  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
   \Description{} 
@@
 
-The controller board, designed around the Arduino Nano BLE platform, served as the central hub for processing sensor data and triggering MIDI messages.
+The controller board, designed around the Arduino Nano BLE platform, was the central hub for processing sensor data and triggering MIDI messages.
 The Nano BLE was selected for its compact form factor, integrated BLE MIDI functionality, and native USB support.
@@
 Non-volatile memory, implemented using Ferroelectric RAM (FRAM), provided a reliable means of storing calibration data.
-This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in a museum setting.
+This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in museums.
 
@@
 
-Calibration was a critical component of the hardware design.
 Each sensor's threshold was manually adjusted using RGB LEDs integrated into the system.
@@
 The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software (Figure \ref{fig:serial_monitor}).
+
+Finer calibration carried out with the guidance of \anon{Catalina Vincens} as an expert harpsichord performer.
 
@@
 
-The final 49-key harpsichord replica incorporated the sensor and controller boards into a rectangular frame, deviating slightly from traditional harpsichord designs to accommodate electronic components.
+The final 49-key harpsichord replica incorporated the sensor and controller boards into a rectangular frame, deviating slightly from traditional designs to accommodate electronic components.
 Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
 
-The assembly process, conducted at the NEMUS lab in Bologna, involved meticulous adjustments to key clearances and wiring layouts.
-Modifications included milling sections of the keys to ensure adequate clearance for the PCBs, as well as varnishing the jacks to improve sticker adhesion.
+The assembly process at the NEMUS lab in Bologna involved meticulous adjustments to key clearances and wiring layouts.
+Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
 Subsequent iterations addressed these challenges, refining the system’s scalability and robustness.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
-    \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
-    \label{fig:3key}
-\end{figure}
 
@@
     \includegraphics[width=\linewidth]{src/images/baffles.png}
-    \caption{3D-printed baffles designed to prevent cross-talk between adjacent sensors.}
+    \caption{STL Model of baffles designed to prevent cross-talk between adjacent sensors.}
     \label{fig:baffles}
@@
     \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
-    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.}
+    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
+\texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data.
+\texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two digit number denoting the sensor index (in this case sensor 32).}
     \label{fig:serial_monitor}
@@
 
-\subsection{Calibration}\label{calibration}
-
-Finer calibration carried out with the guidance of \anon{Catalina Vincens} as
-an expert harpsichord performer
-
-\begin{itemize}
-\item
-  sensor surface
-
-  \begin{itemize}
-  \item
-    printed gradient
-  \end{itemize}
-\end{itemize}
-
 \section{Conclusion}\label{conclusion}
 
-\begin{itemize}
-\item
-  scale
-\item
-  time scale
-\item
-  working at distance
-\item
-  limitations
-
-  \begin{itemize}
-  \item
-    latency
-  \item
-    deprecation
-
-    \begin{itemize}
-    \item
-      software
-
-      \begin{itemize}
-      \item
-        EAGLE
-      \end{itemize}
-    \item
-      hardware
-
-      \begin{itemize}
-      \item
-        FRAM
-      \end{itemize}
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
+Paper has presented the design for an interactive exhibition at a musical instrument museum. Design attempts to address the problem.
+Optical sensors read by a microcontroller have been shown to be a solution to tracking vertical displacement of harpsichord jacks and therefore a viable system to augmenting harpsichords.
+The interface's capabilities of the jack rails to disengage and to convert string plucks to MIDI data means this can be used as a probe to explore the how the ``double-pluck'' can be used as vector for expression during performance.
+The placement of the instrument as a museum exhibit means it also has potential to be used in a long form study.
 
@@
 
-\begin{itemize}
-\item
-  Velocity, aftertouch
-\item
-  second jack row
-\item
-  reducing latency
-\item
-  alternative sensor surfaces
-\end{itemize}
+The current interface is an advanced prototype, but there remain many avenues to explore and aspects of the design to improve.
 
@@
 NIME forum: \url{https://forum.nime.org}.
+
+% \begin{tikzpicture}[auto, node distance=2cm,>=latex']
+
+%     \node [input, name=input] {};
+%     \node [sum, right of=input] (sum) {};
+%     \node [block, right of=sum] (controller) {Controller};
+%     \node [block, right of=controller, pin={[pinstyle]above:D},
+%             node distance=3cm] (system) {System};
+
+%     \draw [->] (controller) -- node[name=u] {$u$} (system);
+%     \node [output, right of=system] (output) {};
+%     \node [block, below of=u] (measurements) {Measurements};
+
+%     \draw [draw,->] (input) -- node {$r$} (sum);
+%     \draw [->] (sum) -- node {$e$} (controller);
+%     \draw [->] (system) -- node [name=y] {$y$}(output);
+%     \draw [->] (y) |- (measurements);
+%     \draw [->] (measurements) -| node[pos=0.99] {$-$} 
+%         node [near end] {$y_m$} (sum);
+% \end{tikzpicture}

### 2025-01-22 src/article.tex
Updates from Overleaf

@@
 The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
-While actuators can generate considerable force, they cannot replicate the wideband signals characteristic of the harpsichord's sharp and responsive haptics (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?), which differ significantly from those of a piano.
+While actuators can generate considerable force, they cannot replicate the wideband signals characteristic of the harpsichord's sharp and responsive haptics (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?), which differ significantly from a piano.
+
+Here is a consolidated **Open-Source Design** subsection for your article:
+
+\subsection{Open-Source Design and Cost-Effectiveness}\label{susec:open-source}
+
+A central objective of this project was to ensure accessibility and reproducibility by committing to an open-source approach for all hardware, software, and data.
+The commitment to open-sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
+Cost-effectiveness was a central consideration.
+For example, the system was initially developed for a 3-key prototype and successfully scaled to 49 keys without significant increases in cost or complexity.
+Key components, such as QRE1113 sensors and CD4051BE multiplexers, were selected for their affordability and availability, while the modular PCB design ensured efficient replication and maintenance.
+
+The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open-source tools and its ability to support both USB and BLE MIDI.
+Ferroelectric RAM (FRAM) was employed for data storage, offering an economical yet robust solution for preserving calibration settings across power cycles.
+Calibration workflows were optimised using the Arduino IDE’s serial plotter and open-source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
+
+Scalability and ease of assembly were prioritised throughout the design process.
+The modular structure of the PCBs, combined with 3D-printed baffles and gradient stickers, ensured that the system could be scaled up without compromising performance or cost-efficiency.
+Installation and calibration times were reduced through iterative improvements, such as the integration of addressable LEDs for sensor identification during setup.
+These design choices align with the project’s open-source ethos, allowing assembly using standard tools commonly available in university maker spaces.
+
+By making the hardware and software openly available, the project invites contributions and adaptations from the broader community.
+This transparency supports critical evaluation and encourages iterative improvements, promoting sustainability and innovation in interactive musical interface design.
+Moreover, the emphasis on affordability and accessibility ensures that the system can be adopted by a wide range of users, from cultural heritage institutions to academic researchers, reinforcing the project’s commitment to democratising access to advanced technological solutions.
 
@@
     \centering
+    \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
+    \\
     \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
@@
 \end{figure}
+
+% \begin{figure}
+%     \centering
+%        \begin{subfigure}
+%          \centering
+%          \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
+%          \caption{Cross Section of 49-Key Model}
+%          % \label{fig:y equals x}
+%      \end{subfigure}\\
+%     \begin{subfigure}
+%          \centering
+%          \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
+%          \caption{3-Key Model: A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack}
+%          % \label{fig:y equals x}
+%      \end{subfigure}        
+%     \caption{Cross Sections of Models}
+%     \label{fig:3key}
+% \end{figure}
 
@@
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{src/images/bl}
-%     \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini at San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
-%     \label{fig:3key}
-% \end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/system-block-diagram-long.png}
+    \caption{Block diagram of PCB connection. Sensor and LED are routed through each sensor PCB}
+    \label{fig:system-block-diagram}
+\end{figure*}
 
@@
 
-Finer calibration carried out with the guidance of \anon{Catalina Vincens} as an expert harpsichord performer.
+As an expert harpsichord performer, finer calibration was carried out with the guidance of \anon{Catalina Vincens}.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
-    \caption{Cross-section illustrating sensor placement relative to the jack mechanism.}
-    \label{fig:cross-section}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
+%     \caption{Cross-section illustrating sensor placement relative to the jack mechanism.}
+%     \label{fig:cross-section}
+% \end{figure}
 
@@
     \centering
+    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \\
     \includegraphics[width=\linewidth]{src/images/baffles.png}
-    \caption{STL Model of baffles designed to prevent cross-talk between adjacent sensors.}
+    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
     \label{fig:baffles}
@@
 Optical sensors read by a microcontroller have been shown to be a solution to tracking vertical displacement of harpsichord jacks and therefore a viable system to augmenting harpsichords.
-The interface's capabilities of the jack rails to disengage and to convert string plucks to MIDI data means this can be used as a probe to explore the how the ``double-pluck'' can be used as vector for expression during performance.
+The interface's capabilities of the jack rails to disengage and to convert string plucks to MIDI data means this can be used as a probe to explore how the ``double-pluck'' can be used as vector for expression during performance.
 The placement of the instrument as a museum exhibit means it also has potential to be used in a long form study.
@@
 
-A limitation in the design is the variability in positioning of tag,
-especially when they need to be reapplied. A jig was designed for
-applying tags, but could not be printed in time for installation. Even
-with a jig, some tags had to be readjusted by hand regardless. Long term
-this has implications on how to maintain the interface.
+The stickers as a sensor surface have a few limitations that can be addressed.
+Firstly, the stickers need to be applied manually, which is laborious but also introduces variability.
+Secondly, the material for the stickers is rated at a 4 years lifespan before the adhesive will decay.
+The application is non-standard and it is likely there will be failure or drift in sensor readings over time, though to what extent is unknown.
+This presents a problem on how this approach can be maintained in the long term.
+Preliminary tests with 3D printed jack bodies incorporating the sensor surface were made.
+The jack body was notched and effectively acts a shutter against the optical sensor.
+More testing needs to be carried out before they are incorporated into the design.
+3D printed jacks would solve two problems simultaneously in that they would be easily reproducible, but also remove labour demands on the luthier.
 
-Taken a different approach to what surface is used. Preliminary test
-were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
-cutting results in a loss of material that has too much variability. The
-edge of the jack body will be charred. The material also needs to be
-reliable. on the advice of two luthiers, \anon{Roberto Livi} and \anon{Jonathan Santa
-Maria Bouquet} at \anon{Edinburgh}, Plywood and MDF have problems with
-variations in humidity, tear-out and delamination.
+Their are also plans to develop the current setup by incorporating force sensitive resistors (FSR). 
+FSRs would be used in conjunction with the current optical sensors to research if an impulse response for the pluck can be extrapolated.
 
-Some tests were carried out using a notched jack body that was 3D
-printed. The notch means that as the jack raises, the forward material
-acts like a shutter. No reasons that this could not be CNC though a
-variety of materials would be needed. Traditional material used is xx
-though need to test of the material is naturally reflective enough or if
-a coating would need to be applied.
+% Taken a different approach to what surface is used. Preliminary test
+% were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
+% cutting results in a loss of material that has too much variability. The
+% edge of the jack body will be charred. The material also needs to be
+% reliable. on the advice of two luthiers, \anon{Roberto Livi} and \anon{Jonathan Santa
+% Maria Bouquet} at \anon{Edinburgh}, Plywood and MDF have problems with
+% variations in humidity, tear-out and delamination.
 
-Outputs of the NEMUS project will also include numerically simulated
-non-linear plucked string models with which the device can interact. A
-bespoke sound synthesis engine to leverage data from both jack rows
-which will also be open-sourced. Design for instrument to be compatible
-with standard MIDI instruments as well software designed for the full
-data output.
+% Some tests were carried out using a notched jack body that was 3D
+% printed. The notch means that as the jack raises, the forward material
+% acts like a shutter. No reasons that this could not be CNC though a
+% variety of materials would be needed. Traditional material used is xx
+% though need to test of the material is naturally reflective enough or if
+% a coating would need to be applied.
 
-A second interface designed more closely to the Trasuntino style
-instrument commissioned by the \anon{NEMUS} project with \anon{Roberto Livi}. Using
-lessons learned for the first iteration the second can make better
-accommodations for internal electronics. Interface to be used for a
-performer study and used for performance with a numerically simulated
-sound synthesis engine.
+Space constraints meant that only one row of jacks could be sensed. A subsequent redesign in the dimensions of the PCB addressed this problem.
+The redesign included flat flex cable (FFC) instead of header pins. This reduced the the size of the PCBs as well as the labour on hand soldering.
+There was an additional added benefit in that the FFC has only a single orientation in which it conducts, meaning it cannot be incorrectly connected.
 
-Testing current sensors in conjunction with force sensitive resistors to
-research if a impulse response for the pluck can be extrapolated.
+Currently the instrument communicates with a Spitfire sample library through the Kontakt software sampler.
+As such, the full data from both rows of jacks cannot be utilised.
+which is limited.
+A future output of the \anon{NEMUS} project will be a bespoke sound synthesis engine to leverage data from both jack rows
+which will also be open-sourced. The intention is also to create a numerically simulated non-linear plucked string models with which the device can interact.
 
-Future novel designs for an interface that would utilise velocity and
-aftertouch messages in software. This would be applicable for numerical
-simulations of strings. A bespoke piece of software would also allow
-for direct mapping of jack rows to registers. There is enough data to
-implement functionality for velocity and aftertouch MIDI messages and
-simply needs to be activated in the firmware. (Hyperinstruments
-\cite{nime2024_20})
+% The design for instrument to be compatible with standard MIDI instruments as well software designed for the full data output.
 
-seat vibrations \cite{MusicalHaptics2018_07}
+A second interface designed more closely to the Trasuntino style instrument has been commissioned by the \anon{NEMUS} project with \anon{Roberto Livi}. 
+Using lessons learned for the first iteration the second can make better accommodations for internal electronics.
+Interface to be used for a performer study and used for performance with the aforementioned numerically simulated sound synthesis engine.
 
-\begin{itemize}
-\item
-  ``there is a general connection between vibrations and the perceived
-  quality of music reproduction''
-\item
-  ``influence of vibrations on loudness perception at low frequencies''
-  Evaluation
-\end{itemize}
+Future novel designs for an interface that would utilise available data, such as velocity and aftertouch, in software in the manner of a hyperinstrument.
+Currently the system looks for the cross over a threshold for triggering MIDI messages, but jack displacement is tracked continuously.
+Data for velocity, both of the pluck and the jack can be extracted from the available data, simply requiring the addition of state machine to make use of it.
+% This would be applicable for numerical simulations of strings.  
+% A bespoke piece of software would also allow for direct mapping of jack rows to registers. 
+% There is enough data to implement functionality for velocity and aftertouch MIDI messages and simply needs to be activated in the firmware. ()
 
-\subsection{improvements}\label{improvements}
+Seat vibrations \cite{MusicalHaptics2018_07} can also be explored to determine if they would create a more immersive experience in the exhibition context.
 
-\begin{itemize}
-\item
-  ESP-IDF
+% \begin{itemize}
+% \item
+%   ``there is a general connection between vibrations and the perceived
+%   quality of music reproduction''
+% \item
+%   ``influence of vibrations on loudness perception at low frequencies''
+%   Evaluation
+% \end{itemize}
 
-  \begin{itemize}
-  \item
-    latency half
-  \item
-    workload high
-  \end{itemize}
-\end{itemize}
+% \subsection{improvements}\label{improvements}
 
-A current limitation is the latency in triggering MIDI output. Currently
-the latency is around 20 ms, but should ideally be closer to 5 ms.
+% \begin{itemize}
+% \item
+%   ESP-IDF
 
-An obvious approach would be merely changing the microcontroller. A test
-was carried out using an Arduino Nano ESP32 which is a NORA-W106, a
-module with a ESP32-S3. The benefit of this microcontroller is that it
-would allow simply the board to be swapped and the rest of the PCB
+%   \begin{itemize}
+%   \item
+%     latency half
+%   \item
+%     workload high
+%   \end{itemize}
+% \end{itemize}
+
+A current limitation is the latency in triggering MIDI output. Currently the latency from pressing a key to triggering a note is around 20 ms. 
+The latency should ideally be closer to 10 ms and with no jitter so \cite{Jack2016}.
+
+One approach would be to change the microcontroller. A test
+was carried out using an Arduino Nano ESP32 which is a NORA-W106, a module containing an ESP32-S3. 
+The benefit of this microcontroller is that it would allow the microcontroller to be swapped and the rest of the PCB
 designs would be unaffected.
 
-Using the Espressif development framework ESP-IDF and following the same
-logic as the original firmware source, the latency was reduced to around
-10 ms. ESP-IDF allows access to functionality of the ESP32 that is not
-exposed in the Arduino libraries. It is possible that the 10 ms latency
-could be improved upon, but there would be a cost in the form of
-development time. Consideration would need to be made on how to port
-existing code or whether to completely re-write the firmware.
+Another approach would be to use a different microcontroller entirely. 
+Some research was carried during the course of the project using an ESP32 programmed with the Espressif development framework ESP-IDF. 
+The same logic as the original firmware was programmed and the latency was reduced to below 10 ms.
+ESP-IDF allows access to functionality of the ESP32 including regular processing of an internal audio buffer that is not exposed in the Arduino IDE.
+It is possible that the 10 ms latency could be improved upon with appropriate optimisation, but there would be a cost in the form of development time. 
+Consideration would need to be made on the time to port existing source code for the I2C FRAM chip or fall back on a simpler storage option.
 
-\begin{itemize}
-\item
-  avoid soldering cabling
-\item
-  power indicators on pcbs
+% \begin{itemize}
+% \item
+%   avoid soldering cabling
+% \item
+%   power indicators on pcbs
 
-  \begin{itemize}
-  \item
-    debug if problem is power related Add power small power LEDs to the
-    each PCB. It was not common for a board to not be functioning and
-    thye result to be that the power was not correctly connected.
-    Especially when the power is connected via something such as an FFC
-    cable where it is easy to plug-in the wrong way especially when
-    fitting is to be carried out by someone unfamiliar with the system.
-    Power LEDs would quickly confirm whether the problem was power
-    related. They would have to be small enough and low enough current
-    draw to avoid glare or being mistaken with other indicator LEDs To
-    continue using THT components, the choice would have to be made
-    carefully.
-  \end{itemize}
-\end{itemize}
+%   \begin{itemize}
+%   \item
+%     debug if problem is power related Add power small power LEDs to the
+%     each PCB. It was not common for a board to not be functioning and
+%     thye result to be that the power was not correctly connected.
+%     Especially when the power is connected via something such as an FFC
+%     cable where it is easy to plug-in the wrong way especially when
+%     fitting is to be carried out by someone unfamiliar with the system.
+%     Power LEDs would quickly confirm whether the problem was power
+%     related. They would have to be small enough and low enough current
+%     draw to avoid glare or being mistaken with other indicator LEDs To
+%     continue using THT components, the choice would have to be made
+%     carefully.
+%   \end{itemize}
+% \end{itemize}
 

### 2025-01-22 src/article.tex

@@
 
-Here is a consolidated **Open-Source Design** subsection for your article:
-
-\subsection{Open-Source Design and Cost-Effectiveness}\label{susec:open-source}
-
-A central objective of this project was to ensure accessibility and reproducibility by committing to an open-source approach for all hardware, software, and data.
-The commitment to open-sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
-Cost-effectiveness was a central consideration.
-For example, the system was initially developed for a 3-key prototype and successfully scaled to 49 keys without significant increases in cost or complexity.
-Key components, such as QRE1113 sensors and CD4051BE multiplexers, were selected for their affordability and availability, while the modular PCB design ensured efficient replication and maintenance.
-
-The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open-source tools and its ability to support both USB and BLE MIDI.
-Ferroelectric RAM (FRAM) was employed for data storage, offering an economical yet robust solution for preserving calibration settings across power cycles.
-Calibration workflows were optimised using the Arduino IDE’s serial plotter and open-source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
-
-Scalability and ease of assembly were prioritised throughout the design process.
-The modular structure of the PCBs, combined with 3D-printed baffles and gradient stickers, ensured that the system could be scaled up without compromising performance or cost-efficiency.
-Installation and calibration times were reduced through iterative improvements, such as the integration of addressable LEDs for sensor identification during setup.
-These design choices align with the project’s open-source ethos, allowing assembly using standard tools commonly available in university maker spaces.
-
-By making the hardware and software openly available, the project invites contributions and adaptations from the broader community.
-This transparency supports critical evaluation and encourages iterative improvements, promoting sustainability and innovation in interactive musical interface design.
-Moreover, the emphasis on affordability and accessibility ensures that the system can be adopted by a wide range of users, from cultural heritage institutions to academic researchers, reinforcing the project’s commitment to democratising access to advanced technological solutions.
-
 \subsection{Summary of the Finalised Design}
@@
 Whereas the previous design emphasised continuous gesture tracking, this implementation required discrete key-triggered data to align with the needs of MIDI-triggered audio playback.
+
+\subsection{Open-source considerations}
+A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an \emph{open-source} approach for all hardware, software, and data.
+The commitment to open-sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
+Cost-effectiveness was also a central consideration.
+The system was initially developed for a 3-key prototype and successfully scaled to 49 keys without significant increases in cost or complexity, and was designed to be easily built using resources typically available in a university-managed makers space.
+
+Key components, such as QRE1113 sensors and CD4051BE multiplexers, were selected for their affordability and availability, while the modular PCB design ensured easy replication and maintenance.
+
+The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open-source tools and ability to support both USB and BLE MIDI.
+Ferroelectric RAM (FRAM) was employed for data storage, offering an economical yet robust solution for preserving calibration settings across power cycles.
+Calibration workflows were optimised using the Arduino IDE’s serial plotter and open-source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
 

### 2025-01-22 src/article.tex
Updates from Overleaf

@@
 
-Technological advances have transformed how museums document, present and interpret their collections, fostering immersive experiences through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
-These technologies emphasise experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.
+Technological advances have transformed how museums document, present and interpret their collections.
+Immersive experiences are afforded through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
+These technologies form a kind of experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.
 However, as Pine and Gilmore note \cite{pinegilmore_2007}, achieving authenticity requires museums to navigate the delicate balance between preservation and meaningful engagement—a challenge that is particularly evident in the case of historical musical instrument collections \cite{McAlpine2014}.
 
-Musical instruments represent a distinctive fusion of form, function, and history.
+Musical instruments represent a peculiar fusion of form, function, and history.
 Their cultural value extends beyond their visual appeal to include the tactile and auditory dimensions of use \cite{Fritz2017}.
@@
 % The project builds on an existing design borrowed from \cite{McPherson2013} and aligns with principles outlined in \cite{Masu_NIME_2023}, emphasising the importance of extending the lifecycle of existing NIMEs.
-The project fosters and complements the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensures the relevance of previous NIME designs beyond their original context \cite{Masu_NIME_2023}.
+% The project fosters and complements the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensures the relevance of previous NIME designs beyond their original context \cite{Masu_NIME_2023}.
 
@@
 A large sampling campaign was conducted, and a custom MIDI interface was designed to fulfil this requirement while preserving the original instruments' integrity.
-The MIDI keyboard, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each.
-However, user tests identified a significant limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
+The MIDI controller, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each.
+However, user tests identified a limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
 
@@
 As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to adhere to longstanding harpsichord building traditions.
-However, this work complements and follows up on previous reflections regarding the \emph{sustainability of results} within the NIME community and identified as \emph{the O in NIME} \cite{Masu_NIME_2023}.
+However, this work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
+Furthermore, relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original papers and enables a form of sustainability.
 
-Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
+Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
 
@@
 
-\subsection{Open-source considerations}
+\subsection{Project Deployment}
 A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an \emph{open-source} approach for all hardware, software, and data.
@@
 Cost-effectiveness was also a central consideration.
-The system was initially developed for a 3-key prototype and successfully scaled to 49 keys without significant increases in cost or complexity, and was designed to be easily built using resources typically available in a university-managed makers space.
+The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity.
+Specifically, the system was designed to be easily assembled using resources typically available in a university-managed makers space.
 
-Key components, such as QRE1113 sensors and CD4051BE multiplexers, were selected for their affordability and availability, while the modular PCB design ensured easy replication and maintenance.
+Components, such as QRE1113 sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
 
 The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open-source tools and ability to support both USB and BLE MIDI.
-Ferroelectric RAM (FRAM) was employed for data storage, offering an economical yet robust solution for preserving calibration settings across power cycles.
+Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles.
 Calibration workflows were optimised using the Arduino IDE’s serial plotter and open-source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
+
+Reference repositories for this project can be found here. PLEASE FILL OUT AS APPROPRIATE (and anonymise).
 
@@
 
-The hardware design aimed to fulfil the earlier constraints, including non-invasiveness, reliability, and scalability.
-To achieve these goals, the system evolved through iterative prototyping, beginning with simple threshold-based testing and culminating in a fully functional multi-sensor interface capable of triggering MIDI events.
+Figure \ref{fig:system-block-diagram} shows a block diagram of the finalised hardware setup.
+The system evolved through iterative prototyping, beginning with simple threshold-based testing and ending in a fully functional multi-sensor interface capable of triggering MIDI events.
 
@@
 Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was acquired from the museum and used as a foundation for prototyping.
-This approach followed a methodology similar to that used in Timmermans' Haptic Key project \cite{Timmermans2020}, enabling incremental validation of hardware and sensor integration.
-
-The 3-key model facilitated iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances.
-The model’s compact form allowed for rapid experimentation while preserving the physical properties of a full-scale harpsichord.
+This approach followed a methodology similar to that used in Timmermans \emph{et al.}'s Haptic Key project \cite{Timmermans2020}, since the 3-key model enabled iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances, before upscaling.
 
@@
 
-% \begin{figure}
-%     \centering
-%        \begin{subfigure}
-%          \centering
-%          \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
-%          \caption{Cross Section of 49-Key Model}
-%          % \label{fig:y equals x}
-%      \end{subfigure}\\
-%     \begin{subfigure}
-%          \centering
-%          \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
-%          \caption{3-Key Model: A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack}
-%          % \label{fig:y equals x}
-%      \end{subfigure}        
-%     \caption{Cross Sections of Models}
-%     \label{fig:3key}
-% \end{figure}
-
 \subsection{Sensor Criteria}\label{sensor-criteria}
 
-To ensure the system's suitability for a museum context, the following criteria were established to guide sensor selection and integration:
+The following criteria were established to guide sensor selection and integration to make the system suitable for a museum context:
 
 \begin{itemize}
-    \item \emph{Non-invasive:} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
-    \item \emph{Low Latency:} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable \cite{Jack2016}.
+    \item \emph{Non-invasiveness.} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
+    \item \emph{Low Latency.} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable, following empirical criteria found in previous studies \cite{Jack2016}.
     \item \emph{Reliability:} Sensor data had to be accurate, minimising false positives and negatives without excessive filtering.
-    \item \emph{Scalability:} The design needed to be cost-effective and adaptable for up to 50 or more keys while maintaining consistent performance.
-    \item \emph{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
+    \item \emph{Scalability.} The design needed to be cost-effective and adaptable for up to 49 keys while maintaining consistent performance.
+    \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
 \end{itemize}
 
-These criteria ensured that the design aligned with the open-source ethos of the project, allowing assembly in standard maker spaces without specialised equipment.
+The open-source ethos of the project ensured assembly was made possible in standard maker spaces without specialised equipment and provided further guidance in design choices.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{src/images/system-block-diagram-long.png}
-    \caption{Block diagram of PCB connection. Sensor and LED are routed through each sensor PCB}
+    \includegraphics[width=\linewidth]{src/images/block-diagram.pdf}
+    \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
     \label{fig:system-block-diagram}
@@
 
-The QRE1113 sensors function by detecting infrared light reflected from nearby surfaces.
-The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors, enabling precise tracking of jack displacement.
-This approach was chosen for its simplicity and scalability, allowing the design to progress from the 3-key prototype to a 49-key model without significant modifications.
+The QRE1113 sensors detect infrared light reflected from nearby surfaces, as per the scheme in Figure \ref{fig:simple-schematic}.
+The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors, which were used to track jack displacement precisely.
+This easily scalable system allowed straightforward progress from the 3-key prototype to a 49-key model without significant modifications.
 
-To eliminate cross-talk between adjacent sensors, 3D-printed baffles were installed on the PCBs (Figure \ref{fig:baffles}).
-These baffles, fabricated from dark-pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
-Testing confirmed that this measure significantly improved data reliability.
-
-\begin{figure}  
+\begin{figure}[b]  
   \centering
@@
 
+3D-printed baffles were installed on the PCBs to eliminate cross-talk between adjacent sensors, as per Figure \ref{fig:baffles}.
+These baffles, fabricated from dark-pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
+Testing confirmed that this measure significantly improved data reliability.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \\
+    \includegraphics[width=\linewidth]{src/images/baffles.png}
+    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
+    \label{fig:baffles}
+\end{figure}
 \subsection{Controller Board}\label{controller-board}
@@
 The Nano BLE was selected for its compact form factor, integrated BLE MIDI functionality, and native USB support.
-These features simplified development while ensuring compatibility with existing MIDI-based synthesis systems.
 
@@
 These LEDs provided visual feedback during calibration, enabling the identification of malfunctioning sensors and simplifying the alignment process.
-The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software (Figure \ref{fig:serial_monitor}).
+The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software, as per Figure \ref{fig:serial_monitor}.
 
@@
 While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors based on usage.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
+    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
+\texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data.
+\texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two-digit number denoting the sensor index (in this case, sensor 32).}
+    \label{fig:serial_monitor}
+\end{figure}
 
@@
 Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
-Subsequent iterations addressed these challenges, refining the system’s scalability and robustness.
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
-%     \caption{Cross-section illustrating sensor placement relative to the jack mechanism.}
-%     \label{fig:cross-section}
-% \end{figure}
+\section{Context}\label{context}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
-    \\
-    \includegraphics[width=\linewidth]{src/images/baffles.png}
-    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
-    \label{fig:baffles}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
-    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
-\texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data.
-\texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two digit number denoting the sensor index (in this case sensor 32).}
-    \label{fig:serial_monitor}
-\end{figure}
-
-% \section{Hardware Design}
-
-% For the first iteration the data from the sensor would simply be used to test when a threshold had been passed.
-On passing the threshold, a message will be sent to trigger the playback of a corresponding note.
-
-% For the initial prototyping stage, modifying an existing harpsichord was considered though ultimately discarded. An existing instrument would
-% have provided a test for scaling the electronics, but internal measurements and layout would have been too different.
-
-% Using a similar approach to Timmermans' Haptic Key project, \cite{Timmermans2020}, our first step was to fabricate or procure a model of a single working key.
-San Colombano provided a model 3-key of a harpsichord mechanism by Graziano Bandini (Figure \ref{fig:3key}).
-
-% \subsection{Sensor Criteria}\label{sensor-criteria}
-
-% A set of criteria was drawn to which the final sensor system would have to satisfy. These criteria were used to evaluate the feasibility throughout the prototyping phase.
-
-% \begin{itemize}
-% \item
-%   Non-invasive: The sensor system should not require major modifications
-%   to the interface on which it is installed.
-% \item
-%   Low Latency: Reading sensors and then triggering an output should have
-%   a low latency. Following \cite{Jack2016} an latency was set as 10 ms
-%   with an upper limit set to 25 ms beyond which the sensors would be
-%   deemed unusable
-% \item
-%   Reliable: Data obtained should be reliable and accurate enough not to
-%   result in false positives or false negatives without extensive
-%   filtering
-% \item
-%   Scalable: The system should be scalable both economically and
-%   temporally. Given the open source nature of the project, it could not
-%   be prohibitively expensive to carry out. Also any system that would
-%   work on a single key would have to scale financially to potentially 50
-%   or more. The time required to install and calibrate the sensors would
-%   also have to scale.
-% \item
-%   Expandable: Any sensor system should be functionally expandable to
-%   allow for the usage of the data for control over other parameters.
-% \end{itemize}
-
-% The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.
-This decision was motivated by the project's goal to be open source and reasonably recreated without procuring additional specialist equipment.
-
-% The electronics system can be divided into two designs, the "sensor board" (Figure \ref{fig:7-sensor-board}) for components for detecting jack displacement and controller board for collecting and processing sensor signals.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[angle=-90,width=\linewidth]{src/images/7-sensor-board-v0.1.0.png} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:7-sensor-board}
-% \end{figure}
-
-% \subsection{Sensor board}\label{sensor-board}
-
-% The 49 optical sensors are divided across 7 PCBs each containing
-
-% \begin{itemize}
-% \item
-%   x7 QRE1113 optical sensors
-% \item
-%   x7 100 $\Omega$ resistors
-% \item
-%   x7 10 k$\Omega$ resistors \footnote{a update on the design showed these could be reduced to a single resistor.}
-% \item
-%   x1 CD4051BE Multipexer
-% \end{itemize}
-
-% A number of sensors were tried before the final system was decided upon. 
-% Sensors tested for the design included, a 6-degrees-of-freedom inertial measurement unit, reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
-% None of the above satisfied the sensor criteria described and were thus discounted.
-
-% Following from the work in \cite{McPherson2013, McPherson2019}, we tested
-% a system using the Fairchild QRE1113 \footnote{Fairchild, now ON Semiconductor, QRE1113 datasheet \url{https://www.onsemi.com/download/data-sheet/pdf/qre1113-d.pdf}}. The
-% QRE1113 is a combination infrared LED and phototransistor sensitive to
-% IR light. The phototransistor of QRE1113 provides data on how much light
-% is present and in particular how much light is being reflected from a
-% surface. Since reflected light will be proportional to the distance of
-% a surface it is a good, close-proximity, distance sensor. Distance is
-% the way in which the instruments in \cite{McPherson2013, McPherson2019}
-% function as well the Moog Piano Bar from which it took inspiration.
-
-% The limitation for taking a distance approach is that there is not one
-% but 2 jacks per key. For each jack to be measured independently the data
-% would need to come from the jacks themselves.
-
-% % They \emph{could} be used and individual trigger points set in
-% % calibration, but this creates new problems.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[,width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:cross-section}
-% \end{figure}
-
-% The optical sensors were directed at a greyscale gradient printed on a sticker and applied to the side of the jack.
-% This was first trialled using a generic inkjet printer and double-sided tape. 
-% For the final iteration, the stickers were printed on Coala 1D 100 Gloss PG 100 $\mu$m white gloss monomeric PVC with permanent grey acrylic dispersion adhesive \footnote{Datasheet: \url{https://www.antalis.co.uk/mediashare/g4media/pdf/TS_EN_COALA_1D_100_GLOSS_P_00_ISS_15022023.pdf}} using a HP Latex 115 and then cut with a Summa 150 roll cutter.
-% The printer and roll cutter greatly reduced the manual labour for cutting stickers meaning the process would scale from the 6 jacks of the 3-key to the 98 jacks of the 49-key model.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[angle=90,width=\linewidth]{src/images/jack-w-tag.jpeg} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:jack-w-tag}
-% \end{figure}
-
-% During testing it was found that their was a considerable amount cross-talk between adjacent sensors. 
-% This meant the reading from key would be affected by the movement of the key next to it. 
-% The cross-talk was assumed to be as a result of light reflecting from other jacks.
-% Given that IR is used this was difficult to confirm by sight. 
-% A set of baffles was designed for the PCB that would wrap round the jack and stop light from other jacks or sensors \ref{fig:baffles}. 
-% The baffles were printed on a BAMBU X1 3D printers using PLA plastic. A dark pigmented PLA was used to avoid IR light simply being diffused. 
-% After the baffles were fitted, the cross-talk was eliminated.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[,width=\linewidth]{src/images/baffles.png} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:baffles}
-% \end{figure}
-
-% Calibration to pluck thresholds were to be carried out manually.
-To allow for easy identification of the sensor being addressed during the calibration process, RGB LEDs were integrated into the design.
-The LEDs were also used in the identification of any malfunctioning or un-calibrated sensor.
-
-% TLC5940 or other PWM driver such as in \cite{McPherson2013} add robustness should an LED fail, but complexity to assembly process. 
-% Instead, the design uses addressable LEDs with integrated driver that share the same data line. 
-% This has the limitation where a faulty LED will stop all subsequent LEDs in the line from working.
-
-% The BLE Nano has 8 ADC pins connected to one ADC channel via a multiplexer. 
-% We avoided the complexity of communicating multiple microcontrollers and instead used a Texas Instruments CD4051B multiplexer on each sensor board.
-% An 8-channel multiplexer was chosen as the model has 49-keys and a division of 7 PCBs each containing 7 sensors.
-% The multiplexer reduced the 7 sensor signals to a single channel. The CD4051B is addressed with a 3-bit which necessitates 3 digital channels.
-% Transistor / Op-amp circuit could have been used similar to the one in \cite{McPherson2013}.
-
-% \subsection{Controller Board}\label{controller-board}
-
-% Controller board designed around Arduino nano form factor. The Nano is
-% connected through headers to allow easy switching between chips. Should
-% another chip in a Nano for factor be found it could supplant the current
-% Nano without any need for a redesign.
-
-% The first version of the board exclusively used 2.54 mm pitch headers.
-% The reasons for this were flexibility when creating cable looms and to
-% leave open the possibility of rewiring the PCB.
-
-% This was the case for the LEDs as they initially were to be powered on
-% 5V. It was discovered that the 3V regulator of the nano was sufficient
-% to power all LEDs at a lower brightness while reducing current draw.
-% Since brightness was not important.
-
-% The 2.54 mm pitch headers had two major drawbacks. First, the orientation
-% of the cables could be reversed, which in the best case scenario would
-% simply mean the system did not work correctly and in the worst case it
-% could potentially damage the components.
-
-% Secondly, the headers do not connect securely and wait of the cabling
-% was enough for components to become disconnected as the harpsichord was
-% handled.
-
-% To address this, once the wiring for the controller board was finalised
-% a new design was drafted this time using JST-PH connectors. No
-% commercial cables were immediately available in the configuration and
-% length required, as such a cable loom needed to be made which meant
-% crimping cables by hand. A crimping tool from JST is prohibitively
-% expensive compared to the cost of other components in the project
-% \footnote{from RS \url{https://uk.rs-online.com/web/p/crimp-tools/6880877}
-% £508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
-% not readily available in the workshops at the project's disposal.
-% Eventually a tool was found, but an alternative approach was also
-% researched in order to maintain the criteria of accessibility.
-
-% A locator was adapted from a creative commons STL model \footnote{\url{https://www.thingiverse.com/thing:1646016}} that allowed a generic crimping tool to be adapted for use with JST-PH crimps.
-
-% \subsubsection{Arduino Nano BLE}\label{arduino-nano-ble}
-
-% Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
-% appealing as it would allow for easy changes between chipsets without
-% requiring any rewiring and thus no new PCBs would need to be created if
-% a better alternative was found later in the project. Another benefit to
-% the Arduino Nano boards was that typically had native-USB meaning the
-% could be programmed as a hardware USB MIDI device directly.
-
-% The Nano BLE was used in the initial stages for its integrated inertial measurment unit.
-
-% When it was finalised that the project would proceed using the QRE1113 a
-% number of microcontroller units (MCU) were considered.
-
-% The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
-% Arduino Nano ESP32 and STM32 NUCLEO-L031K6.
-
-% A rough performance test was later carried out using an STM32
-% NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
-% more performative. To execute a single sensor read cycle of the firmware
-% the STM32 took 40 ms, and both the Nano ESP32 and Nano BLE required 11 ms
-% on average. The Nano ESP32 had high variability so it was decided to
-% continue with the Nano BLE.
-
-% BLE functionality also meant if the cable connection between the nano
-% and computer running audio synthesis was impractical, BLE MIDI was
-% available as a fallback. The Nano BLE also has a 12-bit DAC, which
-% provided a contingency should the default 10-bit not be sensitive enough
-% for the signal from the sensors.
-
-% Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
-% functionality, but it was found that 2 ADC channels were in fact not
-% useable and would have required additional multiplexers.
-
-% \subsubsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}
-
-% There would need to be the ability to store values on non-volatile
-% memory so that the electronics could be powered-down without losing
-% data.
-
-% Adjusting the values through firmware would not have been flexible or
-% practical. Instead an Ferroelectric RAM unit was chosen and connected
-% using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
-% was primarily chosen for it's ready availability.
-
-% Most microcontrollers have some variety of non-volatile program memory
-% that can be read and writ accessed but in most cases this memory is
-% cleared during compilation and re-flashing. The small cost was far
-% outweighed by the potential for data loss if using onboard RAM.
-
-% SD card would provide an easier way to interact with calibration data.
-% The position of the controller board did not permit easy access.
-% Reliability and speed was preferred over storage capacity.
-
-% \subsection{Power}\label{power}
-
-% All electronics combined draw around 1.1 A of current at 5V with some fluctuation during start-up.
-
-% Arduino BLE has some functionality disabled to limit current draw.
-
-% Ideally the unit could be bus powered from USB.
-
-% The sensors are in a state of constantly drawing power. 
-% It is possible that powering only when data is required could be reduce power enough.
-
-% Current designs have a separate MIDI device for each jack row. 
-% Combining the devices and also reducing current draw to below the common 0.5 amps for a USB port would certainly be a challenge.
-
-% \section{Firmware Design}\label{firmware-design}
-
-% Arduino platform chosen for wide adoption, relative ease of use and
-% library support for components.
-
-% Firmware checks components are connected then reads the FRAM to see if
-% threshold data is already present. The rotary encoder and it's momentary
-% switch are polled for any change in state. A key is selected with the
-% rotary encoder while the momentary switch is used for changing between
-% states for key selection and threshold adjustment A double click of the
-% rotary encoder writes the current threshold values to FRAM. The sensors
-% for the current multiplexer channels are polled. If any sensor reads
-% above the threshold when it was previously below a MIDI On message is
-% sent. For the reverse case where the value was below the threshold when
-% it was previously above, a MIDI Note Off message is sent.
-
-% \section{Implementation}\label{implementation}
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[,width=\linewidth]{src/images/exhibition-user-1.jpeg} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:exhibition}
-% \end{figure}
-
-% \subsection{Installation}\label{installation}
-
-% The electronic components were partly pre-assembled at the university of
-% Edinburgh. Though all measurements were known ahead, enough flexibility
-% was needed to allow for problems to solved in situ.
-
-% The final assembly took place at the NEMUS lab at the University of
-% Bologna over the final week in October 2024 for delivery at San
-% Colombano for exhibition in the first week of November in 2024.
-
-% A misreading in the original clearance between the keys and the PCB
-% meant that the keys were blocked entirely from moving. A 10 mm by 15 mm
-% section was milled out of each key to create extra
-% clearance. The milling required the removal, shortening and refitting of
-% the leather pads for the jacks. The extra work
-% added an extra day to the full install.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[width=\linewidth]{src/images/milled-keys.jpg} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:milling}
-% \end{figure}
-
-% A redesign of the PCBs meant that removing the material would no longer
-% be necessary.
-
-% The PCBs shared power and data lines for components with the exception
-% of the output from the multiplexer. Ribbon cable was cut to size to
-% allow for some room to readjust the distance of the sensors from the
-% jacks. The data signals from each PCB were connected using a separate
-% cable loom. Soldering cables and inspecting each join required another
-% day work. This did allow for readjustments to the position of the
-% controller board Measurements of the rear chamber did not match with the
-% Mac Mini that was to be used a different layout was decided for fitting
-% the controller board/ Cutting cables to size allowed for a great deal of
-% flexibility but at the cost of time. After the first install it was
-% clearer what constraints were in place for fitting and an alternative
-% wiring system using flat flex cable was implement (discussed in
-% \textbackslash section)
-
-% Fitting PCBs required the removal of all keys. The support brackets were
-% assembled and screwed into the the underside. With the harpsichord on
-% it's back the sensors were tested to ensure an expected range.
-% Self-tapping round head screws were used to to attach the brackets. The
-% screws were not over tightened to allow for brackets to move. The
-% brackets were not easily adjusted once the key were replaced. Adjusted
-% to a best average with subsequent adjustment carried out on the
-% microcontroller.
-
-% As the PCBs were fitted the jacks had the gradient stickers applied to
-% them. The jacks had been coated with a varnish on the short side to
-% provide a better surface for the stickers to adhere to. After the
-% stickers were applied excess material was trimmed from each jack. The
-% jack were then check that they moved freely in their slot.
-
-% After fitting the PCBs preliminary calibration was carried out. Before
-% the keys were refitted the jacks were put back in place. The data from
-% each key was confirmed to be in a similar range. Any key showing
-% anomalous data had it's jack removed and re-inspected. The PCB fitting
-% process and refitting required another day as faults were found in
-% wiring or misalignment of the stickers. After the sensors were
-% confirmed to be functioning the keys were put back.
-
-% The calibration workflow was with \anon{Dr. Craig Webb} who was unfamiliar with
-% the system. Adjustments were made to the firmware to to make it more
-% intuitive for the user. Addressable LEDs became vital as it allowed for
-% easy identification of the key being calibrated. They were also used to
-% identify any keys whose current reading was beyond what was expected.
-
-% The rotary encoder was used as the interface to select a key, adjust
-% it's threshold and update the Ferroelectric RAM (FRAM) with new values.
-
-% A process was implemented where the key was selected by playing, but
-% this caused problems in situations where the data was unusual due to
-% misalignment of the sensors.
-
-% This would likely be a better system for when the keyboard was fully
-% operational.
-
-% \begin{figure}  
-%   \centering
-%   \includegraphics[width=\linewidth]{images/serial_monitor.png} 
-%   \caption{} 
-%   \Description{} 
-%   \label{fig:serial_monitor}
-% \end{figure}
-
-% The thresholds were calibrated visually utilising the serial plotter of
-% the Arduino IDE (Figure \ref{fig:serial_monitor}) and the open source MIDI
-% Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}.
-
-% \begin{itemize}
-% \item
-%   partial pre-assembly
-% \item
-%   handsolder components
-
-%   \begin{itemize}
-%   \item
-%     analog lines
-%   \end{itemize}
-% \item
-%   connecting supports
-% \end{itemize}
-
-\subsection{Context}\label{context}
-
-Designed as part of an exhibition for the museum San Colombano in
-Bologna. Instruments is to be played by general public as a means of
-interacting with items in the collection that are no longer in playing
-condition. Audience of one who listens via headphones.
-
-The strings of the interface are damped with felt strips. The damping
-did change the tension and thereforew the feel of the strings so
+Designed as part of an exhibition for the museum San Colombano in Bologna.
+Instruments are to be played by the general public to interact with items in the collection that are no longer in playing condition.
+The audience of one who listens via headphones.
+The interface strings are damped with felt strips.
+The damping did change the tension and therefore the feel of the strings so
 adjustments needed to be made to the tuning to accomodate for this.
@@
 
-Their are also plans to develop the current setup by incorporating force sensitive resistors (FSR). 
+Their are also plans to develop the current setup by incorporating force-sensitive resistors (FSR). 
 FSRs would be used in conjunction with the current optical sensors to research if an impulse response for the pluck can be extrapolated.
-
-% Taken a different approach to what surface is used. Preliminary test
-% were taken using Laser Cutting and CNC to fabricate Jack bodies. Laser
-% cutting results in a loss of material that has too much variability. The
-% edge of the jack body will be charred. The material also needs to be
-% reliable. on the advice of two luthiers, \anon{Roberto Livi} and \anon{Jonathan Santa
-% Maria Bouquet} at \anon{Edinburgh}, Plywood and MDF have problems with
-% variations in humidity, tear-out and delamination.
-
-% Some tests were carried out using a notched jack body that was 3D
-% printed. The notch means that as the jack raises, the forward material
-% acts like a shutter. No reasons that this could not be CNC though a
-% variety of materials would be needed. Traditional material used is xx
-% though need to test of the material is naturally reflective enough or if
-% a coating would need to be applied.
 
@@
 
-% The design for instrument to be compatible with standard MIDI instruments as well software designed for the full data output.
-
 A second interface designed more closely to the Trasuntino style instrument has been commissioned by the \anon{NEMUS} project with \anon{Roberto Livi}. 
@@
 Data for velocity, both of the pluck and the jack can be extracted from the available data, simply requiring the addition of state machine to make use of it.
-% This would be applicable for numerical simulations of strings.  
-% A bespoke piece of software would also allow for direct mapping of jack rows to registers. 
-% There is enough data to implement functionality for velocity and aftertouch MIDI messages and simply needs to be activated in the firmware. ()
 
 Seat vibrations \cite{MusicalHaptics2018_07} can also be explored to determine if they would create a more immersive experience in the exhibition context.
-
-% \begin{itemize}
-% \item
-%   ``there is a general connection between vibrations and the perceived
-%   quality of music reproduction''
-% \item
-%   ``influence of vibrations on loudness perception at low frequencies''
-%   Evaluation
-% \end{itemize}
-
-% \subsection{improvements}\label{improvements}
-
-% \begin{itemize}
-% \item
-%   ESP-IDF
-
-%   \begin{itemize}
-%   \item
-%     latency half
-%   \item
-%     workload high
-%   \end{itemize}
-% \end{itemize}
 
@@
 Consideration would need to be made on the time to port existing source code for the I2C FRAM chip or fall back on a simpler storage option.
-
-% \begin{itemize}
-% \item
-%   avoid soldering cabling
-% \item
-%   power indicators on pcbs
-
-%   \begin{itemize}
-%   \item
-%     debug if problem is power related Add power small power LEDs to the
-%     each PCB. It was not common for a board to not be functioning and
-%     thye result to be that the power was not correctly connected.
-%     Especially when the power is connected via something such as an FFC
-%     cable where it is easy to plug-in the wrong way especially when
-%     fitting is to be carried out by someone unfamiliar with the system.
-%     Power LEDs would quickly confirm whether the problem was power
-%     related. They would have to be small enough and low enough current
-%     draw to avoid glare or being mistaken with other indicator LEDs To
-%     continue using THT components, the choice would have to be made
-%     carefully.
-%   \end{itemize}
-% \end{itemize}
 

### 2025-01-22 src/main.tex
Updates from Overleaf

@@
 \documentclass[sigconf]{nimeart}
+\usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+
 %%
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction}
+\title{Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions}
 
@@
 stream of data corresponding to the jacks' motion. This paper details the design
-principles, development, open-source implementation, and execution of the installation for the San Colombano exhibition.
+principles, development, open source implementation, and execution of the installation for the San Colombano exhibition.
 \end{abstract}

### 2025-01-22 src/article.tex
bibtex author names

@@
 Technological advances have transformed how museums document, present and interpret their collections.
-Immersive experiences are afforded through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
+Immersive experiences are realised through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
 These technologies form a kind of experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.

### 2025-01-23 src/article.tex
Updates from Overleaf

@@
 Technological advances have transformed how museums document, present and interpret their collections.
-Immersive experiences are realised through tools such as laser scanning, 3D printing, and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
+Immersive experiences are realised through tools such as 3D printing and and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
 These technologies form a kind of experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.
@@
 The Tagliavini Collection in Bologna \cite{Tagliavini2007}, renowned for its historical keyboard instruments, exemplifies this dilemma.
-To address it, this project introduces an augmented replica of a historical harpsichord keyboard.
-% The project builds on an existing design borrowed from \cite{McPherson2013} and aligns with principles outlined in \cite{Masu_NIME_2023}, emphasising the importance of extending the lifecycle of existing NIMEs.
-% The project fosters and complements the augmented piano keyboard design presented by McPherson \cite{McPherson2013} and ensures the relevance of previous NIME designs beyond their original context \cite{Masu_NIME_2023}.
+With more than seventy early keyboard instruments, including many unique Renaissance Italian harpsichords, spinets and virginals, the collection stands out as a valuable resource for musicologists, organologists and musicians alike.
+Ferdinando Tagliavini’s vision was to preserve the authenticity of the instruments.
+This guiding principle led him to collect only playable instruments, or instruments which could be restored to playing condition after minimal intervention.
 
-The article is organised as follows: TBC
+However, their delicate mechanisms and historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
+To enhance accessibility and engagement, the museum commissioned an electronically augmented keyboard built in the tradition of the Italian Renaissance.
+The keyboard is meant to be used by museum visitors without special permission.
+This work outlines the technological advancements involved in its construction and offers reflections on its role within the Tagliavini Collection’s context.
 
@@
 
-Museums often face significant challenges in engaging visitors due to limitations in staffing and funding, which restrict how visitors can interact with collections \cite{Templeton2018, McAlpine2014}.
+Museums face a constant tension between accessibility and preservation, restricting how visitors can interact with collections \cite{Templeton2018, McAlpine2014}.
 For musical instrument museums, these challenges are compounded by the difficulty of preserving historical instruments in a playable condition \cite{McAlpine2014}.
 The instruments' inherent fragility and gradual decay inevitably result in a point where they can no longer be played, even when collections adhere to the strictest conservation protocols \cite{NYT_strad}.
-A significant cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
+A marked cultural change has taken place in recent decades, shifting the focus from the playability of the originals to their conservation.
 Karp \cite{Karp1979,Karp1985} advocates for a deeper understanding of musical instruments so that enough knowledge is generated to make them as ``copyable'' as possible.
@@
 Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
-As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to adhere to longstanding harpsichord building traditions.
-However, this work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
+As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its tactile response designed to adhere to longstanding harpsichord building traditions.
+However, this work complements and follows up on previous reflections within the NIME community emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
-Furthermore, relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original papers and enables a form of sustainability.
+Furthermore, relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed in \cite{Masu_NIME_2023}.
 
@@
 
-\subsection{Summary of the Finalised Design}
+%\subsection{Summary of the Finalised Design}
 
 The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord replica equipped with an optical sensor system reading the amount of light reflected by gradient stickers applied to the jacks' sides.
-Several key features were implemented to address the project’s requirements:
 
-\begin{itemize}
-\item \emph{QRE1113 infrared sensors}.
+% Several key features were implemented to address the project’s requirements:
+
+% \begin{itemize}
+% \item \emph{QRE1113 infrared sensors}.
 These sensors were selected for their non-invasive properties and precision, enabling the detection of jack displacement without requiring structural modifications to the keyboard mechanism.
-\item \emph{Gradient stickers}.
+% \item \emph{Gradient stickers}.
 Greyscale gradient stickers were applied to each jack to allow the infrared sensors to make precise readings.
 This approach made calibration easier and enabled scalability from a 3-key prototype to the final 49-key design.
-\item \emph{Jack baffles}.
+% \item \emph{Jack baffles}.
 Custom-designed baffles, 3D printed from dark pigmented PLA to minimise infrared reflection, were installed around the jacks to prevent cross-talk between adjacent sensors and ensure reliable data capture.
-\end{itemize}
+% \end{itemize}
 
@@
 
-The project built upon earlier NIME research on generating MIDI messages from piano keystrokes \cite{McPherson2013}, adapting it to address the unique characteristics of harpsichords and the specified constraints.
+The project built upon earlier NIME research on generating MIDI messages from piano keystrokes \cite{McPherson2013}, adapting it to address the specific characteristics of harpsichords.
 Whereas the previous design emphasised continuous gesture tracking, this implementation required discrete key-triggered data to align with the needs of MIDI-triggered audio playback.
@@
 The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity.
-Specifically, the system was designed to be easily assembled using resources typically available in a university-managed makers space.
+Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space.
 
@@
 
-Reference repositories for this project can be found here. PLEASE FILL OUT AS APPROPRIATE (and anonymise).
+Reference repositories for this project can be found here. FILL OUT AS APPROPRIATE (and anonymise).
 
@@
 
-\subsection{Prototyping and Key Model}
+\subsection{Prototype Stage}
 
@@
 
-\subsection{Sensor Criteria}\label{sensor-criteria}
+%\subsection{Sensor Criteria}\label{sensor-criteria}
 
@@
 \begin{itemize}
-    \item \emph{Non-invasiveness.} Sensors must not require significant modifications to the harpsichord's mechanical structure or aesthetics.
+    \item \emph{Non-invasiveness.} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
     \item \emph{Low Latency.} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable, following empirical criteria found in previous studies \cite{Jack2016}.
@@
 
-The open-source ethos of the project ensured assembly was made possible in standard maker spaces without specialised equipment and provided further guidance in design choices.
+The hardware was designed to make assembly possible in standard maker spaces.
 
@@
 The final sensor system utilised QRE1113 infrared sensors, known for their precision and suitability for short-distance detection \cite{McPherson2013, McPherson2019}.
-The sensors were distributed across seven printed circuit boards (PCBs), each responsible for seven keys.
+The sensors were distributed across seven printed circuit boards, each responsible for seven keys.
 Each PCB contained the following components:
@@
 
-The QRE1113 sensors detect infrared light reflected from nearby surfaces, as per the scheme in Figure \ref{fig:simple-schematic}.
+The QRE1113 sensors detect infrared light reflected from nearby surfaces, as per Figure \ref{fig:simple-schematic}.
 The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors, which were used to track jack displacement precisely.
-This easily scalable system allowed straightforward progress from the 3-key prototype to a 49-key model without significant modifications.
 
@@
 
-\subsection{Calibration and Power Management}\label{calibration}
+%\subsection{Calibration and Power Management}\label{calibration}
 
@@
 The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software, as per Figure \ref{fig:serial_monitor}.
-
-As an expert harpsichord performer, finer calibration was carried out with the guidance of \anon{Catalina Vincens}.
+As an expert harpsichord performer, finer calibration was carried out with the guidance of \anon{Catalina Vicens}.
 
@@
 
-The final 49-key harpsichord replica incorporated the sensor and controller boards into a rectangular frame, deviating slightly from traditional designs to accommodate electronic components.
+The final 49-key harpsichord replica incorporated the sensor and controller boards into the rectangular frame.
 Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
@@
 NIME forum: \url{https://forum.nime.org}.
-
-% \begin{tikzpicture}[auto, node distance=2cm,>=latex']
-
-%     \node [input, name=input] {};
-%     \node [sum, right of=input] (sum) {};
-%     \node [block, right of=sum] (controller) {Controller};
-%     \node [block, right of=controller, pin={[pinstyle]above:D},
-%             node distance=3cm] (system) {System};
-
-%     \draw [->] (controller) -- node[name=u] {$u$} (system);
-%     \node [output, right of=system] (output) {};
-%     \node [block, below of=u] (measurements) {Measurements};
-
-%     \draw [draw,->] (input) -- node {$r$} (sum);
-%     \draw [->] (sum) -- node {$e$} (controller);
-%     \draw [->] (system) -- node [name=y] {$y$}(output);
-%     \draw [->] (y) |- (measurements);
-%     \draw [->] (measurements) -| node[pos=0.99] {$-$} 
-%         node [near end] {$y_m$} (sum);
-% \end{tikzpicture}

### 2025-01-24 src/article.tex
Updates from Overleaf

@@
 Technological advances have transformed how museums document, present and interpret their collections.
-Immersive experiences are realised through tools such as 3D printing and and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
+Immersive experiences are realised through tools such as 3D printing and virtual reality \cite{allard2005use,Wachowiak01082009,RCM_2024_3D,Kuzminsky_LaserScan_2012,Schaich_3D_2007}.
 These technologies form a kind of experiential authenticity, enabling encounters that evoke the past's sensory, emotional, and intellectual essence \cite{trant_Auth_1999}.
@@
 With more than seventy early keyboard instruments, including many unique Renaissance Italian harpsichords, spinets and virginals, the collection stands out as a valuable resource for musicologists, organologists and musicians alike.
-Ferdinando Tagliavini’s vision was to preserve the authenticity of the instruments.
+Preserving the instruments' authenticity was the cornerstone of Ferdinando Tagliavini’s vision.
 This guiding principle led him to collect only playable instruments, or instruments which could be restored to playing condition after minimal intervention.
 
-However, their delicate mechanisms and historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
-To enhance accessibility and engagement, the museum commissioned an electronically augmented keyboard built in the tradition of the Italian Renaissance.
-The keyboard is meant to be used by museum visitors without special permission.
-This work outlines the technological advancements involved in its construction and offers reflections on its role within the Tagliavini Collection’s context.
+However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
+To enhance accessibility and engagement, the museum commissioned the replica of a historical keyboard built in the tradition of the Italian Renaissance.
+The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
+The interface is currently connected to a software sampler loaded with the Spitfire sample library, though the ultimate goal is to use samples of instruments from the collection.
+Visitors to the muesum are invited to play the interface and listen through a pair of headphones.
+This work outlines the technological aspects of interface's construction and offers reflections on its role within the Tagliavini Collection and potential application within the wider musical instrument museum context.
 
@@
 
-This work inherits the same philosophy as the Tromba Moderna project.
-By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics and controlling a MIDI-triggered sample library, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions.
-Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
-As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its tactile response designed to adhere to longstanding harpsichord building traditions.
-However, this work complements and follows up on previous reflections within the NIME community emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
+The keyboard presented here inherits some aspects from the Tromba Moderna project.
+A goal of the project is to offer a tool enhancing the fruition of the Tagliavini collection while retaining a form of continuity with historical instrument-building traditions.
+However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile or auditory feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms.
+The design of the electronic systems borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
+Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions.
+This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
-Furthermore, relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed in \cite{Masu_NIME_2023}.
+Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.
 
-Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.
+Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \cite{NEMUS}, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.
 
@@
 
-The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord replica equipped with an optical sensor system reading the amount of light reflected by gradient stickers applied to the jacks' sides.
+The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor system reading the amount of light reflected by gradient stickers applied to the jacks' sides.
 
@@
 The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra.
-The design was inspired by Venetian harpsichords, particularly the Alessandro Trasuntino instrument at the San Colombano Museum.
+The design was inspired by Venetian harpsichords, particularly the 1547 Alessandro Trasuntino harpsichord at the San Colombano Museum.
+An exception was made for the short octave typically found in Italian harpsichords: this was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
 
-The rectangular poplar frame deviates from the traditional logarithmic form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
-Two string orders, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
+The rectangular poplar frame also deviates from the traditional logarithmic form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
+Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
 Felt strips were added to dampen vibrations.
-The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile and auditory qualities.
+The result is an interface that combines the mechanical action of keys with synthetic sound generation, preserving a real harpsichord's tactile qualities.
 
@@
     \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
-    \caption{3-Key Model Harpsichord Mechanism designed by Graziano Bandini at San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
     \label{fig:3key}
@@
 The QRE1113 sensors detect infrared light reflected from nearby surfaces, as per Figure \ref{fig:simple-schematic}.
-The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors, which were used to track jack displacement precisely.
+The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors (Figure \ref{fig:jack-tags}), which were used to track jack displacement precisely.
+
+\begin{figure}[b]  
+  \centering
+  \includegraphics[width=\linewidth]{src/images/tagging-jacks-3.jpg} 
+  \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
+  \Description{} 
+  \label{fig:jack-tags}
+\end{figure}
 
@@
     \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
+    \Description{}
     \label{fig:baffles}
@@
 The Nano BLE was selected for its compact form factor, integrated BLE MIDI functionality, and native USB support.
-
 Initial iterations used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
@@
 Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
-
 The assembly process at the NEMUS lab in Bologna involved meticulous adjustments to key clearances and wiring layouts.
 Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
+The sample library, purchased from Spitfire Audio and controlled via Native Instrument's Kontakt, was installed on a Mac Mini hosted inside the instrument's case.
+Special holes drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack.
+An iPad is used as the secondary monitor, displaying the library alone.
+Users interact with the library through it, while the main 12-inch display sits inside a custom-made drawer locked during usage.
 
-\section{Context}\label{context}
+\section{Discussion}\label{context}
 
-Designed as part of an exhibition for the museum San Colombano in Bologna.
-Instruments are to be played by the general public to interact with items in the collection that are no longer in playing condition.
-The audience of one who listens via headphones.
-The interface strings are damped with felt strips.
-The damping did change the tension and therefore the feel of the strings so
-adjustments needed to be made to the tuning to accomodate for this.
+In an effort to make the exhibition more accessible and authentic, the San Colombano Museum commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
+The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of testers.
+The keyboard shows significant promise in enhancing the museum experience.
+Testers unfamiliar with harpsichords were particularly impressed by its responsiveness.
+The keyboard is hosted in the \emph{Oratory} above the museum's main hall, an exceptional testimony to the Bolognese art, decorated by the finest students of the Carracci and presenting a series of frescoes covering most of the walls and ceiling.
+The room hosts unique examples of the Italian Renaissance building tradition, including the 1547 harpsichord and the 1540 spinet by Alessandro Trasuntino.
+Visitors often visit the room with the same caution and respect that is typical of worship spaces.
+Testers did not find that the exposed headphones and touchscreen affected such an experience negatively, though a more definitive answer will come as more feedback is collected after launch.
+Pictures of the keyboard hosted in the \emph{Oratory} are shown in Figure \ref{fig:oratory}.
 
-\begin{itemize}
-\item
-  in museum context
-\item
-  strings damped
-\item
-  headphones
-\item
-  interfacing with kontakt spitfire library
-\item
-  display
+\begin{figure}
+\centering
+\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
+\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}\label{fig:oratory}
+\end{figure}
 
-  \begin{itemize}
-  \item
-    secure storage and access
-  \item
-    interface to the public
-  \end{itemize}
-\end{itemize}
+Expert feedback has highlighted certain limitations, including imprecise key calibration, which creates a temporal disconnect between the tactile plucking sensation and sound onset.
+Additionally, the commercial sample library, while allowing the selection of registers such as 8', 4', and their combinations, restricts functionality to a single MIDI message per key, regardless of the number of string choirs controlled.
+Inspired by early Italian instruments, the keyboard's layout is designed to manage two 8' registers.
+However, one register was disengaged, and sensors were placed before a single set of jacks to accommodate these limitations.
+Technical improvements, such as refined calibration with hysteresis and developing a custom sample library, could address these constraints in future iterations.
+
+A longer discussion, but one going beyond the scope of this work, is whether the current setup or its future iterations may be effectively used to build legitimate replicas of historical musical instruments or even become a kind of \emph{new} musical instrument altogether.
+As such, the question is whether these designs may result in music being practised, performed and recorded with the instruments.
+This work supports an overarching narrative extending beyond its application in museum collections.
+The NEMUS project \cite{NEMUS}, focusing on developing advanced physical models simulating the nonlinear interaction between subcomponents, has commissioned a second keyboard to further explore the role of control interfaces in performance, exploiting the inherently bi-directional signal paths linking the mechanical world with the virtual, replicating the embodied relationships between performer and instrument and offering opportunities to modify or even disrupt these interactions.
+The upcoming Rem@ke project \cite{remake1} has also expressed an interest in engaging with the keyboard to explore meaningful embodied interactions between players and instruments.
 
@@
 
-Paper has presented the design for an interactive exhibition at a musical instrument museum. Design attempts to address the problem.
-Optical sensors read by a microcontroller have been shown to be a solution to tracking vertical displacement of harpsichord jacks and therefore a viable system to augmenting harpsichords.
-The interface's capabilities of the jack rails to disengage and to convert string plucks to MIDI data means this can be used as a probe to explore how the ``double-pluck'' can be used as vector for expression during performance.
-The placement of the instrument as a museum exhibit means it also has potential to be used in a long form study.
-
-\section{Going forward}\label{going-forward}
-
-The current interface is an advanced prototype, but there remain many avenues to explore and aspects of the design to improve.
-
-The stickers as a sensor surface have a few limitations that can be addressed.
-Firstly, the stickers need to be applied manually, which is laborious but also introduces variability.
-Secondly, the material for the stickers is rated at a 4 years lifespan before the adhesive will decay.
-The application is non-standard and it is likely there will be failure or drift in sensor readings over time, though to what extent is unknown.
-This presents a problem on how this approach can be maintained in the long term.
-Preliminary tests with 3D printed jack bodies incorporating the sensor surface were made.
-The jack body was notched and effectively acts a shutter against the optical sensor.
-More testing needs to be carried out before they are incorporated into the design.
-3D printed jacks would solve two problems simultaneously in that they would be easily reproducible, but also remove labour demands on the luthier.
-
-Their are also plans to develop the current setup by incorporating force-sensitive resistors (FSR). 
-FSRs would be used in conjunction with the current optical sensors to research if an impulse response for the pluck can be extrapolated.
-
-Space constraints meant that only one row of jacks could be sensed. A subsequent redesign in the dimensions of the PCB addressed this problem.
-The redesign included flat flex cable (FFC) instead of header pins. This reduced the the size of the PCBs as well as the labour on hand soldering.
-There was an additional added benefit in that the FFC has only a single orientation in which it conducts, meaning it cannot be incorrectly connected.
-
-Currently the instrument communicates with a Spitfire sample library through the Kontakt software sampler.
-As such, the full data from both rows of jacks cannot be utilised.
-which is limited.
-A future output of the \anon{NEMUS} project will be a bespoke sound synthesis engine to leverage data from both jack rows
-which will also be open-sourced. The intention is also to create a numerically simulated non-linear plucked string models with which the device can interact.
-
-A second interface designed more closely to the Trasuntino style instrument has been commissioned by the \anon{NEMUS} project with \anon{Roberto Livi}. 
-Using lessons learned for the first iteration the second can make better accommodations for internal electronics.
-Interface to be used for a performer study and used for performance with the aforementioned numerically simulated sound synthesis engine.
-
-Future novel designs for an interface that would utilise available data, such as velocity and aftertouch, in software in the manner of a hyperinstrument.
-Currently the system looks for the cross over a threshold for triggering MIDI messages, but jack displacement is tracked continuously.
-Data for velocity, both of the pluck and the jack can be extracted from the available data, simply requiring the addition of state machine to make use of it.
-
-Seat vibrations \cite{MusicalHaptics2018_07} can also be explored to determine if they would create a more immersive experience in the exhibition context.
-
-A current limitation is the latency in triggering MIDI output. Currently the latency from pressing a key to triggering a note is around 20 ms. 
-The latency should ideally be closer to 10 ms and with no jitter so \cite{Jack2016}.
-
-One approach would be to change the microcontroller. A test
-was carried out using an Arduino Nano ESP32 which is a NORA-W106, a module containing an ESP32-S3. 
-The benefit of this microcontroller is that it would allow the microcontroller to be swapped and the rest of the PCB
-designs would be unaffected.
-
-Another approach would be to use a different microcontroller entirely. 
-Some research was carried during the course of the project using an ESP32 programmed with the Espressif development framework ESP-IDF. 
-The same logic as the original firmware was programmed and the latency was reduced to below 10 ms.
-ESP-IDF allows access to functionality of the ESP32 including regular processing of an internal audio buffer that is not exposed in the Arduino IDE.
-It is possible that the 10 ms latency could be improved upon with appropriate optimisation, but there would be a cost in the form of development time. 
-Consideration would need to be made on the time to port existing source code for the I2C FRAM chip or fall back on a simpler storage option.
-
-\section{Ethical Standards}\label{ethical-standards}
-
-To ensure objectivity and transparency in research and to ensure that
-accepted principles of ethical and professional conduct have been
-followed, authors must include a section ``Ethical Standards'' before
-the References. This section should include (if relevant): information
-regarding sources of funding, potential conflicts of interest (financial
-or non-financial), informed consent if the research involved human
-participants, statement on welfare of animals if the research involved
-animals or any other information or context that helps ethically situate
-your research. For help with the ethics section, feel free to ask on the
-NIME forum: \url{https://forum.nime.org}.
+This study has presented the design and implementation of an electronically augmented keyboard for the Tagliavini Collection at the San Colombano Museum.
+By leveraging optical sensors and MIDI-controlled sample libraries, the keyboard allows users to explore a harpsichord's sonic and mechanical essence without compromising the integrity of the museum's historical artefacts.
+Preliminary feedback shows that the keyboard has the potential to enhance the museum experience, particularly for those unfamiliar with early keyboard instruments.
+However, technical limitations, such as calibration inconsistencies and restrictions imposed by the commercial sample library, highlight opportunities for refinement.
+These issues will be addressed in future iterations, including a second keyboard designed to control advanced physical models.
+This work opens new possibilities for engaging with heritage collections while respecting their authenticity.
+It also serves as a foundation for further research projects exploring embodied cognition in historical instruments, and it may provide a useful framework to investigate the links between physical modelling and digital musical interface design.

### 2025-01-24 src/main.tex

@@
 \documentclass[sigconf]{nimeart}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
+\usepackage{graphicx,subcaption}
+% \usepackage{tikz}
+% \usetikzlibrary{shapes,arrows}
 
@@
 % \authornotemark[1]
-\email{michele.ducceschi@unibo.it}
 \affiliation{%
@@
 \author{Roberto Livi}
+\affiliation{%
+  \institution{Harpsichord builder}
+  \city{Pesaro}
+  \country{Italy}
+  }
+
 \author{Catalina Vincens}
@@
 This work describes the design of an electronically augmented replica of a
-16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
-keyboard is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
-which houses the Tagliavini Collection of playable historical keyboard
-instruments. The keyboard was commissioned for exhibition purposes and to
-enhance the experiential authenticity of the collection. The keyboard emulates
-the mechanics of a two-register harpsichord featuring a single manual. Unpitched
-strings under tension replicate the mechanical resistance characteristic of the
-original compass. An array of infrared LED / phototransistor pairs, connected to a
-microcontroller, was installed within the interface to generate a continuous
-stream of data corresponding to the jacks' motion. This paper details the design
-principles, development, open source implementation, and execution of the installation for the San Colombano exhibition.
+16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a DMI. The
+project is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
+which houses the Tagliavini Collection of playable historical keyboard instruments. 
+The keyboard was commissioned for exhibition where it will act as an interface to  digitised instruments in the collection.
+
+The project aims to enhance the experiential authenticity of the collection and address how we can conserve the interaction with historical instruments when the interface is a limiting factor in the verisimilitude of the experience.
+
+The tactile feedback of the keyboard is created by incorporating the same mechanics of a two-register harpsichord with a single manual. 
+Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass. 
+An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous
+streams of data corresponding to the vertical motion of each jack. 
+This paper details the keyboard's design principles, technical development, and the motivations behind them. 
+% It also discusses the implications of its execution for installation for public exhibition. 
 \end{abstract}
@@
 \begin{teaserfigure}
- \hspace*{0.6in}
-  \includegraphics[width=0.8\textwidth]{src/images/49-key-front-top.png}
+ \centering
+  \includegraphics[width=0.58\textwidth]{src/images/49-key-front-top.png}
   \caption{49-key Model Harpsichord Mechanism by \anon{Roberto Livi, Bologna 2024}}
@@
 \begin{acks}
-A project of this breadth developed across 3 institutes and 2 countries
-cannot be carried out without the support and technical knowledge of many
-skilled and passionate people.
-
-We acknowledge the invaluable contributions of
-
+This work received funding from the European Commission's Horizon2020 programme, through grant ERC-StG 950084 NEMUS. We acknowledge the invaluable contributions of
 \begin{itemize}
@@
     \end{itemize}
-  \end{itemize}
-\item
-  San Colombano
-  \begin{itemize}
-  \item
-    Catalina Vincens
-  \item
-    Roberto Livi
   \end{itemize}

### 2025-01-24 src/article.tex
Updates from Overleaf

@@
 However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
-To enhance accessibility and engagement, the museum commissioned the replica of a historical keyboard built in the tradition of the Italian Renaissance.
+To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance.
 The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
@@
 
-\section{Design Principles}\label{design}
-
 \begin{figure*}
@@
 \end{figure*}
+
+\section{Design Principles}\label{design}
 
@@
 
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
-  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
-  \Description{} 
-  \label{fig:49-key-bottom}
-\end{figure}
-
 A modular system of printed circuit boards (PCBs) was designed to manage the sensors and process their output.
-The 49 QRE1113 sensors were distributed across seven boards, containing 7 sensors each.
+The system uses 49 QRE1113 sensors which are distributed across seven boards, containing 7 sensors each.
 The PCBs were secured to the underside of the keyboard mechanism, allowing them to be adjusted during installation.
@@
 \subsection{Project Deployment}
-A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an \emph{open-source} approach for all hardware, software, and data.
-The commitment to open-sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
+A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an \emph{open source} approach for all outputs of the project.
+The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
 Cost-effectiveness was also a central consideration.
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
+  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
+  \Description{} 
+  \label{fig:49-key-bottom}
+\end{figure}
+
 Components, such as QRE1113 sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
 
-The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open-source tools and ability to support both USB and BLE MIDI.
+The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open source tools and ability to support both USB and BLE MIDI.
 Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles.
-Calibration workflows were optimised using the Arduino IDE’s serial plotter and open-source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
+Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
 
-Reference repositories for this project can be found here. FILL OUT AS APPROPRIATE (and anonymise).
+Reference repositories for this project can be found here:
+\begin{anonsuppress}
+    \begin{itemize}
+        \item 
+        \emph{Firmware}: \url{https://github.com/Nemus-Project/haprsichord-interface-firmware}
+        \item 
+        \emph{PCB CAD}: \url{https://github.com/Nemus-Project/haprsichord-interface-cad}
+        \item 
+        \emph{Interface Models}: \url{https://github.com/Nemus-Project/haprsichord-interface-models}
+    \end{itemize}
+\end{anonsuppress}
+% FILL OUT AS APPROPRIATE (and anonymise).
 
@@
 
-\begin{figure}
+\begin{figure}[b]
     \centering
@@
 These LEDs provided visual feedback during calibration, enabling the identification of malfunctioning sensors and simplifying the alignment process.
-The calibration workflow utilised the Arduino IDE serial plotter and open-source MIDI monitoring software, as per Figure \ref{fig:serial_monitor}.
+The calibration workflow utilised the Arduino IDE serial plotter and open source MIDI monitoring software, as per Figure \ref{fig:serial_monitor}.
 As an expert harpsichord performer, finer calibration was carried out with the guidance of \anon{Catalina Vicens}.

### 2025-01-25 src/article.tex
full-read first draft

@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{src/images/details.jpg}
+\centering
+\includegraphics[width=0.8\linewidth]{src/images/details.jpg}
 \caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}\label{fig:details}
@@
 
-The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor system reading the amount of light reflected by gradient stickers applied to the jacks' sides.
+The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor in a voltage divider configuration (Figure \ref{fig:simple-schematic}) whose voltage output depends on the amount of light reflected by sensor surface.
+In this case, a sensor surface is a grey-scale gradient printed on a vinyl sticker and applied to the side of each jack (Figure \ref{fig:jack-tags}).
 
@@
 A modular system of printed circuit boards (PCBs) was designed to manage the sensors and process their output.
-The system uses 49 QRE1113 sensors which are distributed across seven boards, containing 7 sensors each.
+The system uses 49 optical sensors which are distributed across seven boards that contain 7 sensors each.
 The PCBs were secured to the underside of the keyboard mechanism, allowing them to be adjusted during installation.
@@
 
-The project built upon earlier NIME research on generating MIDI messages from piano keystrokes \cite{McPherson2013}, adapting it to address the specific characteristics of harpsichords.
+The project expanded upon earlier NIME research on generating MIDI messages from piano keystrokes \cite{McPherson2013}, adapting it to address the specific characteristics of harpsichords.
 Whereas the previous design emphasised continuous gesture tracking, this implementation required discrete key-triggered data to align with the needs of MIDI-triggered audio playback.
@@
 
-Components, such as QRE1113 sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
+Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
 
-The Arduino Nano BLE was chosen as the core microcontroller for its compatibility with open source tools and ability to support both USB and BLE MIDI.
+The Arduino Nano 33 BLE was chosen as the core microcontroller for its compatibility with open source tools and ability to support both USB and BLE MIDI.
 Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles.
@@
         \item 
-        \emph{Firmware}: \url{https://github.com/Nemus-Project/haprsichord-interface-firmware}
+        \emph{Firmware}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-firmware}}
         \item 
-        \emph{PCB CAD}: \url{https://github.com/Nemus-Project/haprsichord-interface-cad}
+        \emph{PCB CAD}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-cad}}
         \item 
-        \emph{Interface Models}: \url{https://github.com/Nemus-Project/haprsichord-interface-models}
+        \emph{Interface Models}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-models}}
     \end{itemize}
@@
     \centering
-    \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
-    \\
-    \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+    % \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
+    % \\
+    % \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
+    % \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
+    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
+    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
     \label{fig:3key}
@@
 
-The final sensor system utilised QRE1113 infrared sensors, known for their precision and suitability for short-distance detection \cite{McPherson2013, McPherson2019}.
+The final sensor system utilised QRE1113 optical sensors, known for their precision, low cost, and suitability for short-range distance detection \cite{McPherson2013, McPherson2019}.
 The sensors were distributed across seven printed circuit boards, each responsible for seven keys.
@@
 \begin{itemize}
-    \item 7 QRE1113 infrared sensors.
+    \item 7 QRE1113 optical sensors.
     \item 7 100 $\Omega$ resistors and 7 10 k$\Omega$ resistors (later reduced to one resistor per board).
@@
 
-The QRE1113 sensors detect infrared light reflected from nearby surfaces, as per Figure \ref{fig:simple-schematic}.
+The optical sensors detect infrared light reflected from nearby surfaces, as per Figure \ref{fig:simple-schematic}.
 The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors (Figure \ref{fig:jack-tags}), which were used to track jack displacement precisely.
@@
 
-The controller board, designed around the Arduino Nano BLE platform, was the central hub for processing sensor data and triggering MIDI messages.
-The Nano BLE was selected for its compact form factor, integrated BLE MIDI functionality, and native USB support.
-Initial iterations used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
+The controller board, designed around the Arduino Nano format, was the central hub for processing sensor data and triggering MIDI messages.
+The Arduino Nano was selected for its compact form factor and native USB support.
+The 33 BLE variationof the Nano had the additional benefit in its ability to be programmed as BLE MIDI device.
+Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
 Custom cable looms were constructed using adapted crimping tools, ensuring robust connections without compromising accessibility.
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
-    \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
+%     \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
 \texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data.
 \texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two-digit number denoting the sensor index (in this case, sensor 32).}
-    \label{fig:serial_monitor}
-\end{figure}
+%     \label{fig:serial_monitor}
+% \end{figure}
 
@@
 Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
-The assembly process at the NEMUS lab in Bologna involved meticulous adjustments to key clearances and wiring layouts.
+The assembly process at the \anon{NEMUS} lab in \anon{Bologna} involved meticulous adjustments to key clearances and wiring layouts.
 Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
@@
 
-In an effort to make the exhibition more accessible and authentic, the San Colombano Museum commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
+In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
 The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of testers.
@@
 Pictures of the keyboard hosted in the \emph{Oratory} are shown in Figure \ref{fig:oratory}.
-
-\begin{figure}
-\centering
-\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
-\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}\label{fig:oratory}
-\end{figure}
 
@@
 This work supports an overarching narrative extending beyond its application in museum collections.
-The NEMUS project \cite{NEMUS}, focusing on developing advanced physical models simulating the nonlinear interaction between subcomponents, has commissioned a second keyboard to further explore the role of control interfaces in performance, exploiting the inherently bi-directional signal paths linking the mechanical world with the virtual, replicating the embodied relationships between performer and instrument and offering opportunities to modify or even disrupt these interactions.
+The \anon{NEMUS project \cite{NEMUS}}, focusing on developing advanced physical models simulating the non-linear interaction between subcomponents, has commissioned a second keyboard to further explore the role of control interfaces in performance, exploiting the inherently bi-directional signal paths linking the mechanical world with the virtual, replicating the embodied relationships between performer and instrument and offering opportunities to modify or even disrupt these interactions.
 The upcoming Rem@ke project \cite{remake1} has also expressed an interest in engaging with the keyboard to explore meaningful embodied interactions between players and instruments.
+
+\begin{figure}[b]
+\centering
+\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
+\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}\label{fig:oratory}
+\end{figure}
 

### 2025-01-25 src/main.tex
full-read first draft

@@
 \begin{abstract}
-This work describes the design of an electronically augmented replica of a
-16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a DMI. The
-project is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
-which houses the Tagliavini Collection of playable historical keyboard instruments. 
-The keyboard was commissioned for exhibition where it will act as an interface to  digitised instruments in the collection.
+% This work describes the design of an electronically augmented replica of a
+% 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a DMI. The
+% project is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
+% which houses the Tagliavini Collection of playable historical keyboard instruments. 
+% The keyboard was commissioned for exhibition where it will act as an interface to digitised version of historical instruments that are not available to play by visitors.
 
-The project aims to enhance the experiential authenticity of the collection and address how we can conserve the interaction with historical instruments when the interface is a limiting factor in the verisimilitude of the experience.
+% The project aims to enhance the experiential authenticity of the collection and address how we can conserve the interaction with historical instruments when the interface is a limiting factor in the verisimilitude of the experience playing a digitised instrument.
 
-The tactile feedback of the keyboard is created by incorporating the same mechanics of a two-register harpsichord with a single manual. 
-Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass. 
-An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous
-streams of data corresponding to the vertical motion of each jack. 
-This paper details the keyboard's design principles, technical development, and the motivations behind them. 
-% It also discusses the implications of its execution for installation for public exhibition. 
+% The tactile feedback of the keyboard is created by incorporating the same mechanics of a two-register harpsichord with a single manual. 
+% Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass. 
+% An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous
+% streams of data corresponding to the vertical motion of each jack. 
+% This paper details the keyboard's design principles, technical development, and the motivations behind them.
+
+This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout in order to create a DMI. % [1]
+The keyboard was commissioned for exhibition in a musical instrument museum where it will act as an interface to digitised versions of historical instruments that are not in a playable condition.
+% [2]
+This project also addresses how we can interact with digitally conserve historic musical instruments when the interface is a limiting factor to the verisimilitude of the experience. % [2,3]
+This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
+
+The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass.
+% [1]
+An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
+We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
+
 \end{abstract}

### 2025-01-25 nime-paper-template.tex
abstract versions

@@
 %%
-\documentclass[sigconf,anonymous,review]{nimeart}
+\documentclass[sigconf]{nimeart}
 %%
@@
 features incorporated into this single new template.
+
+\begin{itemize}
+  \item
+  \emph{Repository one}: \url{https://www.github/com/organisationthatslong/everslightlytoo-long-repository-name-1}
+  \item
+  \emph{Repository two}: \url{https://www.github/com/organisationthatslong/everslightlytoo-long-repository-name-2}
+  \item
+  \emph{Repository three}: \url{https://www.github/com/organisationthatslong/everslightly-too-long-repository-name-3}
+\end{itemize}
 

### 2025-01-25 src/main.tex
abstract versions

@@
 \documentclass[sigconf]{nimeart}
+% \documentclass[sigconf,anonymous,review]{nimeart}
 \usepackage{graphicx,subcaption}
@@
 \begin{abstract}
-% This work describes the design of an electronically augmented replica of a
-% 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a DMI. The
-% project is a collaboration between the \anon{NEMUS project} and the Museo di San Colombano,
-% which houses the Tagliavini Collection of playable historical keyboard instruments. 
-% The keyboard was commissioned for exhibition where it will act as an interface to digitised version of historical instruments that are not available to play by visitors.
 
-% The project aims to enhance the experiential authenticity of the collection and address how we can conserve the interaction with historical instruments when the interface is a limiting factor in the verisimilitude of the experience playing a digitised instrument.
-
-% The tactile feedback of the keyboard is created by incorporating the same mechanics of a two-register harpsichord with a single manual. 
-% Unpitched strings under tension replicate the mechanical resistance characteristic of the original compass. 
-% An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous
-% streams of data corresponding to the vertical motion of each jack. 
-% This paper details the keyboard's design principles, technical development, and the motivations behind them.
-
-This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout in order to create a DMI. % [1]
-The keyboard was commissioned for exhibition in a musical instrument museum where it will act as an interface to digitised versions of historical instruments that are not in a playable condition.
+%% Version 1
+This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument.
+% [1]
+The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
 % [2]
@@
 We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
+
+%% Version 2
+% This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a vibrotactile digital musical instrument.
+% [1]
+% An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
+% The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
+
+% This project embodies the entanglement between musical instrument cultural heritage and its relationship with new and increasing accessible technologies. [3, 4]
+% Acknowledgment and conscientiousness to older designs and aesthetics have been used in service of, rather than obstruction to, richer interactions with digital musical instruments.[2,3,4]
+% This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
+% We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
 

### 2025-01-25 src/main.tex
move Ethical Standards

@@
 %% consistent spelling of the heading.
+
 \begin{acks}
-This work received funding from the European Commission's Horizon2020 programme, through grant ERC-StG 950084 NEMUS. We acknowledge the invaluable contributions of
+ We acknowledge the invaluable contributions of
 \begin{itemize}
@@
 
+\section{Ethical Standards}
+This work received funding from the European Commission's Horizon2020 programme, through grant \anon{ERC-StG 950084 NEMUS}
+
 %%

### 2025-01-27 src/main.tex

@@
   \item
-    Alma Labor
-    \begin{itemize}
-    \item
+    \emph{Alma Labor}:
       Massimiliano Fraulini and the staff of Alma Labor at the Università
@@
       electronics equipment
-    \end{itemize}
   \item
-    Fisica Tecnica
-    \begin{itemize}
-    \item
-      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
+    \emph{Fisica Tecnica}: Maurzio Chendi, Fabrizio Casarini, Giulio Severo and Barbara
       Costantini    
-    \end{itemize}
-  \end{itemize}
-\item
-  Edinburgh
-  \begin{itemize}
-  \item
-    uCreate
-    \begin{itemize}    
-    \item
-      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
-      Edinburgh for their support with soldering facilities.
   \end{itemize}
@@
   \item
-      Richard Collins and the Digital Making Staff at the Edinburgh
-      College of Art for 3D printing instruction and CNC support.
+    \emph{uCreate}:    
+      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
+      Edinburgh for their support with soldering facilities.    
   \item
-      Joe Hathaway of Digital Development at the College of Art for
-      providing electronics components
+    \emph{Edinburgh College of Art}: Richard Collins and the Digital Making Staff at the Edinburgh
+      College of Art for 3D printing instruction and CNC support. Digital Development Technician Joe Hathaway    
     \item  
-      St Cecilia's Jenny Nex
-    \end{itemize}
+      \emph{St Cecilia's}: Curator Dr. Jenny Nex
   \end{itemize}

### 2025-01-27 src/main.tex
Updates from Overleaf

@@
 %% Version 1
-This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument.
+This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a vibrotactile digital musical instrument.
 % [1]
-The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
-% [2]
-This project also addresses how we can interact with digitally conserve historic musical instruments when the interface is a limiting factor to the verisimilitude of the experience. % [2,3]
-This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
-
 The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass.
@@
 An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
+
+The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
+% [2]
+This project also addresses how we can interact with digitally conserved historic musical instruments when the interface is a limiting factor to the verisimilitude of the experience. % [2,3]
+This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
 We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
@@
 
-% This project embodies the entanglement between musical instrument cultural heritage and its relationship with new and increasing accessible technologies. [3, 4]
-% Acknowledgment and conscientiousness to older designs and aesthetics have been used in service of, rather than obstruction to, richer interactions with digital musical instruments.[2,3,4]
+% This project embodies the entanglement between musical instrument cultural heritage and its relationship with new and increasingly accessible technologies.% [3, 4]
+% Acknowledgment and conscientiousness of older designs and aesthetics have been used in service of, rather than obstruction to, richer interactions with digital musical instruments.%[2,3,4]
 %  This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]

### 2025-01-27 src/article.tex
abstract versions

@@
 
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
+  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
+  \Description{} 
+  \label{fig:49-key-bottom}
+\end{figure}
+
 The keyboard presented here inherits some aspects from the Tromba Moderna project.
@@
 
-% Several key features were implemented to address the project’s requirements:
-
-% \begin{itemize}
-% \item \emph{QRE1113 infrared sensors}.
-These sensors were selected for their non-invasive properties and precision, enabling the detection of jack displacement without requiring structural modifications to the keyboard mechanism.
-% \item \emph{Gradient stickers}.
-Greyscale gradient stickers were applied to each jack to allow the infrared sensors to make precise readings.
-This approach made calibration easier and enabled scalability from a 3-key prototype to the final 49-key design.
-% \item \emph{Jack baffles}.
-Custom-designed baffles, 3D printed from dark pigmented PLA to minimise infrared reflection, were installed around the jacks to prevent cross-talk between adjacent sensors and ensure reliable data capture.
-% \end{itemize}
+\subsubsection{Alternative}
+\textcolor{red}{The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
+The keyboard was consciously chosen to mimic that of the original XVI century Italian harpsichords.
+This design decision was made in order to use the propensity humans have to being influenced by visual components when musical making judgements \cite{Tsay2013} to our advantage.
+It is especially true with respect to musical instruments as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have clearly shown.
+Rather than avoiding this effect we have chosen to instead utilise it.
+This `musical instrument McGurk effect' is in part the reason behind hiding the electronics out of sight.
+The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument.
+With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component my provide the necessary push of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
+Such an effect is invaluable if visitors are expected to engage with the interface on an equal footing as the historic instruments that neighbour it in the museum.
+The keyboard in not merely a facsimile of a harpsichord, but rather by being equipped with optical sensors, it can provide modes of expression with a digital instrument which would otherwise be inaccessible to its analogue sibling.
+The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) whose voltage output depends on the amount of light reflected by sensor surface.
+In this case, a sensor surface is a grey-scale gradient printed on a vinyl sticker and applied to the side of each jack (Figure \ref{fig:jack-tags}).
+Though the current configuration simply sends MIDI note on and off at passing a threshold, the displacement of all jacks is available continuously.
+Touch with the key
+The double register means that there are two very distinct tactile `anchors' in the course of the keys travel that is championed by professional harpsichordists as invaluable vector for expression during performance.
+(GOING TO NEED A REFERENCE ON THIS BUT FAIRLY CONFIDENT I SAW IT IN THE TAGLIAVINI BOOK.
+MICHELE IF YOU HAVE A REFERENCE LEAVE IT HERE).}
+% "When working with an established interface like the keyboard, it is particularly important that new mappings do not interfere with familiar technique (McPherson, Gierakowski, and Stark 2013)."
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
-  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
-  \Description{} 
-  \label{fig:49-key-bottom}
-\end{figure}
-
 Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
@@
 
+\begin{anonsuppress}
 Reference repositories for this project can be found here:
-\begin{anonsuppress}
+
     \begin{itemize}

### 2025-01-27 src/article.tex
fix #28

@@
 
-\begin{figure}[b]
+\begin{figure}
   \centering
@@
 \subsubsection{Alternative}
-\textcolor{red}{The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
+\textcolor{red}{
+The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
 The keyboard was consciously chosen to mimic that of the original XVI century Italian harpsichords.
@@
 Though the current configuration simply sends MIDI note on and off at passing a threshold, the displacement of all jacks is available continuously.
-Touch with the key
-The double register means that there are two very distinct tactile `anchors' in the course of the keys travel that is championed by professional harpsichordists as invaluable vector for expression during performance.
-(GOING TO NEED A REFERENCE ON THIS BUT FAIRLY CONFIDENT I SAW IT IN THE TAGLIAVINI BOOK.
-MICHELE IF YOU HAVE A REFERENCE LEAVE IT HERE).}
+The staggering of pluck point of the jack means that there are two very distinct tactile `anchors' along the key dip.
+A combination of string tension, quill voicing and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
+}
+%
 % "When working with an established interface like the keyboard, it is particularly important that new mappings do not interfere with familiar technique (McPherson, Gierakowski, and Stark 2013)."
@@
 
-\begin{figure}[b]
+\begin{figure}
     \centering
@@
 
-\begin{figure}[b]  
+\begin{figure}  
   \centering
@@
 
-\begin{figure}[b]
+\begin{figure}
 \centering

### 2025-01-29 src/main.tex
fix #28
fix #29
v1.0.0

--- 2025-01-27 src/main.tex
@@
 % \authornote{Both authors contributed equally to this research.}
+\affiliation{%
+  \institution{Università di Bologna}
+  \city{Bologna}  
+  \country{Italy}
+}
 \email{matthew.hamilton2@unibo.it}
@@
 }
+\email{michele.ducceschi@unibo.it}
 
@@
 \affiliation{%
-  \institution{Harpsichord builder}
+  \institution{Master Harpsichord Maker}
   \city{Pesaro}
@@
   }
+\email{robertvs.livi@gmail.com}
 
-\author{Catalina Vincens}
+\author{Catalina Vicens}
 \affiliation{%
@@
   \country{Italy}}
+  \email{catalina.vicens@genusbononiae.it}
 
@@
  \country{United Kingdom}}
+ \email{andrew.mcpherson@imperial.ac.uk}
 
@@
 \begin{abstract}
+% This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument.
+% [1]
+% The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass.
+% [1]
+% An optical sensor system was fitted inside the body of the harpsichord, directed towards a gradient surface applied to the body of each jack, generating continuous streams of data corresponding that are used to trigger a sample library.
+% [1, ~4]
+% The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing means of accessing to instruments that are no longer in playing condition.
+% [2]
+% The details of the keyboard's design principles, technical development, and underlying motivations are described in detail. % [1]
+% % The design and mechanics of a harpsichord were replicated in the new keyboard as using modern plastic MIDI controllers has been shown to be an unsatisfying experience for users. % [3]
+% % This culminated in a design for a keyboard interface capable of extending the performance techniques of the harpsichord. % [4]
 
-%% Version 1
 This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument.
-% [1]
-The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
-% [2]
-This project also addresses how we can interact with digitally conserve historic musical instruments when the interface is a limiting factor to the verisimilitude of the experience. % [2,3]
-This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
-
-The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass.
-% [1]
-An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
-We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
-
-%% Version 2
-% This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a vibrotactile digital musical instrument.
-% [1]
-% An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
-% The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.
-
-% This project embodies the entanglement between musical instrument cultural heritage and its relationship with new and increasing accessible technologies. [3, 4]
-% Acknowledgment and conscientiousness to older designs and aesthetics have been used in service of, rather than obstruction to, richer interactions with digital musical instruments.[2,3,4]
-% This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
-% We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]
+The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
+The replica is designed to balance the competing demands of historical authenticity, public accessibility, and preservation. 
+It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques.
+Optical sensors integrated within the mechanism capture the jacks' motion data, enabling MIDI message generation.
+This work situates itself within broader discussions on the role of technology in museums. 
+A keyboard interface of this type offers an opportunity to enhance visitor interaction while safeguarding delicate artefacts. 
+The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.
 
@@
 \begin{acks}
+This work received funding from the European Commission's Horizon2020 programme, through grant \anon{ERC-StG 950084 NEMUS}
+
  We acknowledge the invaluable contributions of

### 2025-01-30 src/article.tex

@@
 However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
-To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance.
+To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance and is the subject of this paper.
 The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
 The interface is currently connected to a software sampler loaded with the Spitfire sample library, though the ultimate goal is to use samples of instruments from the collection.
-Visitors to the muesum are invited to play the interface and listen through a pair of headphones.
+Visitors to the museum are invited to play the interface and listen through a pair of headphones.
 This work outlines the technological aspects of interface's construction and offers reflections on its role within the Tagliavini Collection and potential application within the wider musical instrument museum context.
+% The paper outlines the design and motivations behind this replica keyboard interface.
 
@@
 
+\begin{figure}
+    \centering    
+    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
+    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
+    \label{fig:3key}
+\end{figure}
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{src/images/tagging-jacks-3.jpg} 
+  \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
+  \Description{} 
+  \label{fig:jack-tags}
+\end{figure}
+
+\begin{figure}  
+  \centering
+  \includegraphics[width=\linewidth]{src/images/simple-schematic-bw-.jpg} 
+  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
+  \Description{} 
+  \label{fig:simple-schematic}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \\
+    \includegraphics[width=\linewidth]{src/images/baffles.png}
+    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+\subsection{Controller Board}\label{controller-board}
+
+\begin{figure}
+\centering
+\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
+\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}
+\label{fig:oratory}
+\end{figure}
+
 The keyboard presented here inherits some aspects from the Tromba Moderna project.
-A goal of the project is to offer a tool enhancing the fruition of the Tagliavini collection while retaining a form of continuity with historical instrument-building traditions.
-However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile or auditory feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms.
-The design of the electronic systems borrows ideas from a previous NIME by McPherson \cite{McPherson2013}.
+The project aims to offer a tool to enhance the fruition of the Tagliavini collection while retaining a form of continuity with historical instrument-building traditions.
+However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms.
+The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}. 
 Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions.
@@
 
-Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \cite{NEMUS}, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.
+Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \anon{\cite{NEMUS}}, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.
 
@@
 
-The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor in a voltage divider configuration (Figure \ref{fig:simple-schematic}) whose voltage output depends on the amount of light reflected by sensor surface.
+The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor, whose voltage output depends on the amount of light reflected by sensor surface.
 In this case, a sensor surface is a grey-scale gradient printed on a vinyl sticker and applied to the side of each jack (Figure \ref{fig:jack-tags}).
@@
 It is especially true with respect to musical instruments as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have clearly shown.
-Rather than avoiding this effect we have chosen to instead utilise it.
+Rather than avoiding this effect we have chosen instead to utilise it.
 This `musical instrument McGurk effect' is in part the reason behind hiding the electronics out of sight.
@@
 With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component my provide the necessary push of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
-Such an effect is invaluable if visitors are expected to engage with the interface on an equal footing as the historic instruments that neighbour it in the museum.
-The keyboard in not merely a facsimile of a harpsichord, but rather by being equipped with optical sensors, it can provide modes of expression with a digital instrument which would otherwise be inaccessible to its analogue sibling.
+% Such an effect is invaluable if visitors are expected to engage with the interface on an equal footing as the historic instruments that neighbour it in the museum.
+% The keyboard in not merely a facsimile of a harpsichord, but rather by being equipped with optical sensors, it can provide modes of expression with a digital instrument which would otherwise be inaccessible to its analogue sibling.
 The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) whose voltage output depends on the amount of light reflected by sensor surface.
@@
 A combination of string tension, quill voicing and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
+Such tactile expectations provide fertile opportunities for extending the expressive capabilities of the interface.
+The current sensor system should for recording of traditional use and identify where those opportunities for new expression lie \cite{McPherson2013-2}.
 }
-%
-% "When working with an established interface like the keyboard, it is particularly important that new mappings do not interfere with familiar technique (McPherson, Gierakowski, and Stark 2013)."
+
+% "When working with an established interface like the keyboard, it is particularly important that new mappings do not interfere with familiar technique (McPherson, Gierakowski, and Stark 2013).
+The Space Between the Notes: Adding Expressive Pitch Control to the Piano Keyboard"
 
@@
 The system uses 49 optical sensors which are distributed across seven boards that contain 7 sensors each.
-The PCBs were secured to the underside of the keyboard mechanism, allowing them to be adjusted during installation.
+The PCBs were secured to the underside of wrest plank and above the keys (Figure \ref{fig:49-key-bottom}), allowing them to be adjusted during installation.
 Ribbon cables connected the PCBs, providing flexibility during assembly while maintaining a compact form factor.
@@
 \subsection{Project Deployment}
-A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an \emph{open source} approach for all outputs of the project.
+A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an open source approach for all outputs of the project.
 The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
@@
         \item 
-        \emph{Firmware}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-firmware}}
+        \emph{Firmware}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-firmware}}
         \item 
-        \emph{PCB CAD}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-cad}}
+        \emph{PCB CAD}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-cad}}
         \item 
-        \emph{Interface Models}: \anon{\url{https://github.com/Nemus-Project/haprsichord-interface-models}}
+        \emph{Interface Models}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-models}}
     \end{itemize}
@@
 The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord.
-Traditional materials were used, including walnut for the wrestplank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard.
+Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard.
 The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra.
 The design was inspired by Venetian harpsichords, particularly the 1547 Alessandro Trasuntino harpsichord at the San Colombano Museum.
-An exception was made for the short octave typically found in Italian harpsichords: this was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
+An exception was made for the short octave --- the assigning of common keys in the first octave over a chromatic scale --- typically found in Italian harpsichords: this was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
 
@@
 
-\begin{figure}
-    \centering
-    % \includegraphics[width=\linewidth]{src/images/CrossSectionSensorPlacement.jpg}
-    % \\
-    % \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png}
-    % \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023. Key components include A: Wrestplank, B: Key Lever, C: Key Cap, D: Soundboard, E: Jack.}
-    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
-    \label{fig:3key}
-\end{figure}
-
 %\subsection{Sensor Criteria}\label{sensor-criteria}
@@
     \item \emph{Non-invasiveness.} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
-    \item \emph{Low Latency.} Sensor response times were required to remain under 10 ms, with an upper limit of 25 ms deemed acceptable, following empirical criteria found in previous studies \cite{Jack2016}.
-    \item \emph{Reliability:} Sensor data had to be accurate, minimising false positives and negatives without excessive filtering.
-    \item \emph{Scalability.} The design needed to be cost-effective and adaptable for up to 49 keys while maintaining consistent performance.
-    \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
+    \item \emph{Low Latency.} The time required to process a single iteration of all sensor data should remain under 10 ms, with an upper limit of 25 ms deemed acceptable.
+Empirical criteria found in previous studies \cite{Jack2016} was used as a guide.
+    \item \emph{Reliability:} Sensor data should be dependable and consistent, with interference from anything external to the jack movement being absent or negligible.
+    \item \emph{Scalability.} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (figure \ref{fig:3key}) up to the 49 keys of the final design.
+    % \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
 \end{itemize}
@@
 
-The final sensor system utilised QRE1113 optical sensors, known for their precision, low cost, and suitability for short-range distance detection \cite{McPherson2013, McPherson2019}.
+The final sensor system utilised QRE1113 optical sensors, known for their small form factor, low cost, and suitability for short-range distance detection \cite{McPherson2013, McPherson2019}.
 The sensors were distributed across seven printed circuit boards, each responsible for seven keys.
@@
 
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/tagging-jacks-3.jpg} 
-  \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
-  \Description{} 
-  \label{fig:jack-tags}
-\end{figure}
-
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/simple-schematic-bw-.jpg} 
-  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
-  \Description{} 
-  \label{fig:simple-schematic}
-\end{figure}
-
 3D-printed baffles were installed on the PCBs to eliminate cross-talk between adjacent sensors, as per Figure \ref{fig:baffles}.
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
-    \\
-    \includegraphics[width=\linewidth]{src/images/baffles.png}
-    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
-    \Description{}
-    \label{fig:baffles}
-\end{figure}
-\subsection{Controller Board}\label{controller-board}
+The controller board, designed around the Arduino Nano format, was the central hub for processing sensor data and triggering MIDI messages.
+The in addition to solder terminals for sensor board channels, the controller board contained of:
 
-The controller board, designed around the Arduino Nano format, was the central hub for processing sensor data and triggering MIDI messages.
+\begin{itemize}
+    \item 1 Arduino Nano 33 IoT
+    \item 1 Fujitsu MB85RS64 SPI FRAM chip
+    \item 1 EC11 combined rotary encoder and tactile switch
+    % \item 1 DFROBOT DFR0785 Tactile Switch
+\end{itemize}
+
 The Arduino Nano was selected for its compact form factor and native USB support.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
-\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}\label{fig:oratory}
-\end{figure}
-
 \section{Conclusion}\label{conclusion}

### 2025-01-30 src/article.tex

@@
 
+% % Option 1
+\begin{figure}
+    \centering
+    \includesvg[inkscapelatex=false, height = 0.4\textheight]{images/log_harp_outline}
+    \caption{Logarithmic shape of an original Trasuntino harpsichord}
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+
+% % Option 2
+\begin{figure}
+    \centering
+     \begin{subfigure}[h]{0.5\linewidth}
+        \centering
+        \includesvg[inkscapelatex=false, width = \linewidth]{images/log_harp_outline}            
+        \caption{logarithmic shape}    
+        \label{fig:log-harp}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[h]{0.5\linewidth}
+        \centering
+        \includesvg[inkscapelatex=false, width = \linewidth]{images/49-key_outline}        
+        \caption{replica shape}
+        \label{fig:rect-harp}
+    \end{subfigure}
+    \caption{Logarithmic shape of an original Trasuntino harpsichord contrast with rectangular shape of the replica}    
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+
+% % Option 4
+
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{Logarithmic shape of an original Trasuntino harpsichord}
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+
+% % Option 3
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/trasuntino-log-harp.jpg}    
+    \caption{Logarithmic shape of an original Trasuntino harpsichord.
+Source: Image 3 Royal College of Music Collections Item RCM0002  \protect\footnote{\url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}}  }
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+
 \section{Hardware Design}\label{hardware-design}

### 2025-01-31 src/article.tex

--- 2025-01-30 src/article.tex
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
-  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
-  \Description{} 
-  \label{fig:49-key-bottom}
-\end{figure}
-
-\begin{figure}
-    \centering    
-    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
-    \label{fig:3key}
-\end{figure}
-
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/tagging-jacks-3.jpg} 
-  \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
-  \Description{} 
-  \label{fig:jack-tags}
-\end{figure}
-
-\begin{figure}  
-  \centering
-  \includegraphics[width=\linewidth]{src/images/simple-schematic-bw-.jpg} 
-  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
-  \Description{} 
-  \label{fig:simple-schematic}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
-    \\
-    \includegraphics[width=\linewidth]{src/images/baffles.png}
-    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
-    \Description{}
-    \label{fig:baffles}
-\end{figure}
-\subsection{Controller Board}\label{controller-board}
-
-\begin{figure}
-\centering
-\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
-\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}
-\label{fig:oratory}
-\end{figure}
-
 The keyboard presented here inherits some aspects from the Tromba Moderna project.
@@
 \end{itemize}
+
 These constraints demanded a non-invasive, robust design that was respectful of the historical authenticity of the instruments.
 The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
-While actuators can generate considerable force, they cannot replicate the wideband signals characteristic of the harpsichord's sharp and responsive haptics (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?), which differ significantly from a piano.
+While actuators can generate considerable force, it is difficult to achieve the free-motion and resistance of a jack without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
+A tactile realism of the harpsichord can already be achieved by the instrument itself.
+It should first be confirmed if mechanical recreation is sufficient as part of a digital musical instrument before elaborate electromechanical emulation.
+% (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?)
+, which differ significantly from a piano.
 
-%\subsection{Summary of the Finalised Design}
-
-The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key harpsichord keyboard replica equipped with an optical sensor, whose voltage output depends on the amount of light reflected by sensor surface.
-In this case, a sensor surface is a grey-scale gradient printed on a vinyl sticker and applied to the side of each jack (Figure \ref{fig:jack-tags}).
-
-\subsubsection{Alternative}
-\textcolor{red}{
 The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
@@
 The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument.
-With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component my provide the necessary push of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
-% Such an effect is invaluable if visitors are expected to engage with the interface on an equal footing as the historic instruments that neighbour it in the museum.
-% The keyboard in not merely a facsimile of a harpsichord, but rather by being equipped with optical sensors, it can provide modes of expression with a digital instrument which would otherwise be inaccessible to its analogue sibling.
-The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) whose voltage output depends on the amount of light reflected by sensor surface.
-In this case, a sensor surface is a grey-scale gradient printed on a vinyl sticker and applied to the side of each jack (Figure \ref{fig:jack-tags}).
+With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component may provide the necessary `push' of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
+
+\begin{figure}[!b]  
+  \centering
+  \includegraphics[width=0.9\linewidth]{src/images/tagging-jacks-3.jpg} 
+  \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
+  \Description{} 
+  \label{fig:jack-tags}
+\end{figure}
+
+Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected by a sensor surface.
+In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker \ref{fig:jack-tags} and applied to the side of each jack.
 Though the current configuration simply sends MIDI note on and off at passing a threshold, the displacement of all jacks is available continuously.
@@
 Such tactile expectations provide fertile opportunities for extending the expressive capabilities of the interface.
-The current sensor system should for recording of traditional use and identify where those opportunities for new expression lie \cite{McPherson2013-2}.
-}
+The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.
 
@@
 The Space Between the Notes: Adding Expressive Pitch Control to the Piano Keyboard"
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{src/images/49-key-bottom-sensors-no-keys.jpg} 
+  \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
+  \Description{} 
+  \label{fig:49-key-bottom}
+\end{figure}
 
@@
 \subsection{Project Deployment}
+
 A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an open source approach for all outputs of the project.
@@
 
+\begin{figure}[!b]
+    \centering    
+    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
+    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
+    \label{fig:3key}
+\end{figure}
+
 \subsection{Materials and Construction}
@@
 
-The rectangular poplar frame also deviates from the traditional logarithmic form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
+% % Option 2
+\begin{figure}
+    \centering
+     \begin{subfigure}[h]{0.4\linewidth}
+        \centering
+        \includesvg[inkscapelatex=false, width = \linewidth]{images/log_harp_outline}            
+        \caption{logarithmic shape}    
+        \label{fig:log-harp}
+    \end{subfigure}
+
+    \begin{subfigure}[h]{0.4\linewidth}
+        \centering
+        \includesvg[inkscapelatex=false, width = \linewidth]{images/49-key_outline}        
+        \caption{rectangular shape}
+        \label{fig:rect-harp}
+    \end{subfigure}
+    \caption{Logarithmic shape of an original Trasuntino harpsichord contrast with rectangular shape of the replica}    
+    \Description{}
+    \label{fig:log-harp-comp}
+\end{figure}
+
+The rectangular poplar frame \ref{fig:rect-harp} also deviates from the traditional logarithmic \ref{fig:log-harp} form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
 Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
@@
 
-%\subsection{Sensor Criteria}\label{sensor-criteria}
-
 The following criteria were established to guide sensor selection and integration to make the system suitable for a museum context:
@@
 
-\subsection{Sensor Board Design}\label{sensor-board}
+\subsection{Sensor Board}\label{sensor-board}
 
-\begin{figure*}
+\begin{figure}[!b] 
   \centering
-    \includegraphics[width=\linewidth]{src/images/block-diagram.pdf}
-    \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
-    \label{fig:system-block-diagram}
-\end{figure*}
+  \includegraphics[width=0.7\linewidth]{src/images/simple-schematic-bw-.jpg} 
+  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
+  \Description{} 
+  \label{fig:simple-schematic}
+\end{figure}
 
@@
 
-The optical sensors detect infrared light reflected from nearby surfaces, as per Figure \ref{fig:simple-schematic}.
-The gradient stickers affixed to each jack provided a consistent reflective surface for the sensors (Figure \ref{fig:jack-tags}), which were used to track jack displacement precisely.
+The optical sensors are wired in a voltage divider configuration \ref{fig:simple-schematic} and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
+The gradient stickers affixed to each jack provided a surface with varying reflectivity for the sensors, which were used to track jack displacement throughout the key dip.
 
@@
 These baffles, fabricated from dark-pigmented PLA, ensured that infrared reflections from neighbouring jacks did not interfere with sensor readings.
-Testing confirmed that this measure significantly improved data reliability.
+A manual visual check of plotted all sensors readings confirmed that interference from neighbouring jacks had been eliminated.
+
+\subsection{Controller Board}\label{controller-board}
 
@@
 
-The Arduino Nano was selected for its compact form factor and native USB support.
-The 33 BLE variationof the Nano had the additional benefit in its ability to be programmed as BLE MIDI device.
-Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/block-diagram.pdf}
+    \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
+    \label{fig:system-block-diagram}
+\end{figure*}
+
+The Arduino Nano small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets.
+The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as BLE MIDI device.
+%Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
 Custom cable looms were constructed using adapted crimping tools, ensuring robust connections without compromising accessibility.
 
-Non-volatile memory, implemented using Ferroelectric RAM (FRAM), provided a reliable means of storing calibration data.
+Non-volatile memory, implemented using a Fujitsu MB85RS64 Ferroelectric RAM (FRAM) chip, provided a reliable means of storing calibration data.
 This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in museums.
@@
 
-Each sensor's threshold was manually adjusted using RGB LEDs integrated into the system.
-These LEDs provided visual feedback during calibration, enabling the identification of malfunctioning sensors and simplifying the alignment process.
-The calibration workflow utilised the Arduino IDE serial plotter and open source MIDI monitoring software, as per Figure \ref{fig:serial_monitor}.
-As an expert harpsichord performer, finer calibration was carried out with the guidance of \anon{Catalina Vicens}.
+A rotary encoder was used as the interface to select a key, edit a threshold and save current thresholds to the FRAM.
+Thresholds were first set the midrange of possible values a a default.
+A rough calibration process was carried out where each sensor were selected individually and its readings were plotted against their current threshold value using the Arduino IDE serial plotter.
+Thresholds were adjusted with the rotary encoder until the key no longer passed the threshold line until the jack had plucked the string.
+During calibration RGB LEDs were used as a visual aid to quickly identify which key was currently being adjusted.
+These LEDs also provided visual feedback during calibration, and would change colour based on if readings were above below the threshold or if the were outside previously recorded maximum or minimum values.
+Such use of the LEDs meant for easy identification of malfunctioning sensors and simplified the alignment process.
+A finer calibration process was carried out with the guidance of expert harpsichord performer \anon{Catalina Vicens}.
 
-Power requirements for the system were estimated at 1.1 A at 5 V, with fluctuations during startup.
-While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors based on usage.
+Power requirements for the system were estimated at 1.1 A at 5 V, with some fluctuation when the system was first powered on.
+While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors based found in the McPherson piano \cite{McPherson2013}.
 
@@
 
-\subsection{Final Assembly and Integration}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{src/images/mac-mini.jpg}
+    \caption{Components and cabling hidden in the bottom section of the instrument case.}
+    \label{fig:mac-mini}
+\end{figure}
 
-The final 49-key harpsichord replica incorporated the sensor and controller boards into the rectangular frame.
+% \subsection{Final Assembly and Integration}
+
+% The final 49-key harpsichord replica incorporated the sensor and controller boards into the rectangular frame.
 Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
@@
 Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
-The sample library, purchased from Spitfire Audio and controlled via Native Instrument's Kontakt, was installed on a Mac Mini hosted inside the instrument's case.
-Special holes drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack.
-An iPad is used as the secondary monitor, displaying the library alone.
-Users interact with the library through it, while the main 12-inch display sits inside a custom-made drawer locked during usage.
+
+For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt.
+The software was installed on a Mac Mini hosted inside the instrument's case.
+Holes were drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack \ref{fig:mac-mini}.
+These components are accessible via locked drawer which is visible in \ref{fig:oratory}.
+An iPad is used as a monitor through which visitors can adjust parameters of playback such as tuning, voicing and which stops are engaged.
+The next iteration will use a library of samples from instruments in the collection, which is currently being carried out by the museum.
+%A main 12-inch display sits inside a custom-made drawer locked during usage.
+
+\begin{figure}[!hb]
+    \centering
+    \includegraphics[width=0.8\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \\
+    \includegraphics[width=0.8\linewidth]{src/images/baffles.png}
+    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
+\subsection{Controller Board}\label{controller-board}
 
@@
 
-In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
-The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of testers.
-The keyboard shows significant promise in enhancing the museum experience.
-Testers unfamiliar with harpsichords were particularly impressed by its responsiveness.
+% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
+The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
+Collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
+The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' as found in the McAlpine exhibition \cite{McAlpine2014} were absent in this initial feedback.
+
+A more formal surveying of museum visitors is to be carried out to determine to what degree the exhibition achieved it's core aims.
+% Testers unfamiliar with harpsichords were particularly impressed by its responsiveness.
 The keyboard is hosted in the \emph{Oratory} above the museum's main hall, an exceptional testimony to the Bolognese art, decorated by the finest students of the Carracci and presenting a series of frescoes covering most of the walls and ceiling.
@@
 Visitors often visit the room with the same caution and respect that is typical of worship spaces.
-Testers did not find that the exposed headphones and touchscreen affected such an experience negatively, though a more definitive answer will come as more feedback is collected after launch.
+Among the visitors interviewed, there were no comments made that the exposed headphones and touchscreen affected such an experience negatively, though a more definitive answer will come as more feedback is collected after launch.
 Pictures of the keyboard hosted in the \emph{Oratory} are shown in Figure \ref{fig:oratory}.
@@
 Additionally, the commercial sample library, while allowing the selection of registers such as 8', 4', and their combinations, restricts functionality to a single MIDI message per key, regardless of the number of string choirs controlled.
-Inspired by early Italian instruments, the keyboard's layout is designed to manage two 8' registers.
-However, one register was disengaged, and sensors were placed before a single set of jacks to accommodate these limitations.
-Technical improvements, such as refined calibration with hysteresis and developing a custom sample library, could address these constraints in future iterations.
+Inspired by early Italian instruments, the keyboard's layout is designed to manage two 8' registers with a sensor system fitted to each.
+Internally the instrument is therefore 2 individual MIDI devices however, one register needed to be disengaged as the software could only communicate with a single device.
+Multiple instances of the software could be used, but an arrangement of 2 sets GUI controls was deemed too confusing to visitors who would interact with the exhibition without orientation.
+Sensors are therefore active only on a single set of jacks to accommodate these limitations.
+Technical improvements, such as refined calibration with hysteresis and developing a custom sample library and interface, could address these constraints in future iterations.
+
+\begin{figure}
+\centering
+\includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
+\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}
+\label{fig:oratory}
+\end{figure}
 
@@
 This work supports an overarching narrative extending beyond its application in museum collections.
-The \anon{NEMUS project \cite{NEMUS}}, focusing on developing advanced physical models simulating the non-linear interaction between subcomponents, has commissioned a second keyboard to further explore the role of control interfaces in performance, exploiting the inherently bi-directional signal paths linking the mechanical world with the virtual, replicating the embodied relationships between performer and instrument and offering opportunities to modify or even disrupt these interactions.
+The \anon{NEMUS project \cite{NEMUS}}, focusing on developing advanced physical models simulating the non-linear interaction between subcomponents, has commissioned a second keyboard to further explore the role of control interfaces in performance, employing the sense of touch as link between the mechanical world with the digital, replicating the embodied relationships between performer and instrument and offering opportunities to modify or even disrupt these interactions.
 The upcoming Rem@ke project \cite{remake1} has also expressed an interest in engaging with the keyboard to explore meaningful embodied interactions between players and instruments.
 
-\section{Conclusion}\label{conclusion}
+% \section{Conclusion}\label{conclusion}
 
-This study has presented the design and implementation of an electronically augmented keyboard for the Tagliavini Collection at the San Colombano Museum.
+% This study has presented the design and implementation of an electronically augmented keyboard for the Tagliavini Collection at the San Colombano Museum.
 By leveraging optical sensors and MIDI-controlled sample libraries, the keyboard allows users to explore a harpsichord's sonic and mechanical essence without compromising the integrity of the museum's historical artefacts.

### 2025-01-31 src/article.tex
second draft

@@
 % (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?)
-, which differ significantly from a piano.
+% , which differ significantly from a piano.
 
@@
 
-Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected by a sensor surface.
-In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker \ref{fig:jack-tags} and applied to the side of each jack.
-Though the current configuration simply sends MIDI note on and off at passing a threshold, the displacement of all jacks is available continuously.
-The staggering of pluck point of the jack means that there are two very distinct tactile `anchors' along the key dip.
+Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected from a sensor surface.
+In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack.
+Though the current configuration simply sends MIDI note on and off messages when passing a threshold, the displacement of all jacks is available continuously.
+The staggering of the pluck point of the jacks between registers means that there are two very distinct tactile `anchors' along the key dip.
 A combination of string tension, quill voicing and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
@@
 Cost-effectiveness was also a central consideration.
-The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity.
+% The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity. 
 Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space.
@@
 
-\begin{figure}[!b]
-    \centering    
-    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
-    \label{fig:3key}
-\end{figure}
-
 \subsection{Materials and Construction}
@@
 The design was inspired by Venetian harpsichords, particularly the 1547 Alessandro Trasuntino harpsichord at the San Colombano Museum.
-An exception was made for the short octave --- the assigning of common keys in the first octave over a chromatic scale --- typically found in Italian harpsichords: this was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
+An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
 
 % % Option 2
-\begin{figure}
+\begin{figure}[!b]
     \centering
@@
 
-The rectangular poplar frame \ref{fig:rect-harp} also deviates from the traditional logarithmic \ref{fig:log-harp} form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
+The rectangular poplar frame also deviates from the traditional logarithmic \ref{fig:log-harp-comp} form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
 Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
@@
 The system evolved through iterative prototyping, beginning with simple threshold-based testing and ending in a fully functional multi-sensor interface capable of triggering MIDI events.
+
+\begin{figure}
+    \centering    
+    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
+    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
+    \label{fig:3key}
+\end{figure}
 
@@
     \item \emph{Reliability:} Sensor data should be dependable and consistent, with interference from anything external to the jack movement being absent or negligible.
-    \item \emph{Scalability.} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (figure \ref{fig:3key}) up to the 49 keys of the final design.
+    \item \emph{Scalability.} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (Figure \ref{fig:3key}) up to the 49 keys of the final design.
     % \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
@@
 The controller board, designed around the Arduino Nano format, was the central hub for processing sensor data and triggering MIDI messages.
-The in addition to solder terminals for sensor board channels, the controller board contained of:
+In addition to solder terminals for the sensor board channels, the controller board contained:
 
@@
 The Arduino Nano small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets.
-The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as BLE MIDI device.
+The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as a BLE MIDI device.
 %Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
@@
 Power requirements for the system were estimated at 1.1 A at 5 V, with some fluctuation when the system was first powered on.
-While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors based found in the McPherson piano \cite{McPherson2013}.
+While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors found in the McPherson piano \cite{McPherson2013}.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{src/images/mac-mini.jpg}
+    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
     \caption{Components and cabling hidden in the bottom section of the instrument case.}
@@
 The software was installed on a Mac Mini hosted inside the instrument's case.
-Holes were drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack \ref{fig:mac-mini}.
-These components are accessible via locked drawer which is visible in \ref{fig:oratory}.
+Holes were drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}).
+% These components are accessible via locked drawer which is visible in \ref{fig:oratory}. 
 An iPad is used as a monitor through which visitors can adjust parameters of playback such as tuning, voicing and which stops are engaged.
@@
 
-\begin{figure}[!hb]
+\begin{figure}[!b]
     \centering
-    \includegraphics[width=0.8\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
     \\
@@
 \end{figure}
-\subsection{Controller Board}\label{controller-board}
 

### 2025-01-31 src/article.tex

@@
 However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
-To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance and is the subject of this paper.
+To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance, which is the subject of this paper.
 The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
-The interface is currently connected to a software sampler loaded with the Spitfire sample library, though the ultimate goal is to use samples of instruments from the collection.
+The interface is currently connected to a commercial software sampler, though the ultimate goal is to use sample instruments from the collection.
 Visitors to the museum are invited to play the interface and listen through a pair of headphones.
-This work outlines the technological aspects of interface's construction and offers reflections on its role within the Tagliavini Collection and potential application within the wider musical instrument museum context.
+This work outlines the technological aspects of the interface's construction and offers reflections on its role within the Tagliavini Collection and potential application within the wider musical instrument museum context.
 % The paper outlines the design and motivations behind this replica keyboard interface.
@@
 
-These constraints demanded a non-invasive, robust design that was respectful of the historical authenticity of the instruments.
+These constraints demanded a non-invasive design that was respectful of the historical authenticity of the instruments.
 The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
-While actuators can generate considerable force, it is difficult to achieve the free-motion and resistance of a jack without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
-A tactile realism of the harpsichord can already be achieved by the instrument itself.
-It should first be confirmed if mechanical recreation is sufficient as part of a digital musical instrument before elaborate electromechanical emulation.
+While actuators can generate considerable force, it is difficult to achieve a jack's free motion and resistance without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
+The instrument itself can already achieve the tactile realism of the harpsichord.
+It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation. 
 % (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?)
-% , which differ significantly from a piano.
+% The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
+% The keyboard was consciously chosen to mimic the typical layout of 16th-century Italian harpsichords.
+This design decision was made to use humans' propensity to being influenced by visual components when musical making judgements \cite{Tsay2013} to our advantage.
+It is especially true concerning musical instruments, as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have shown.
+Rather than avoiding this effect, we have chosen to utilise it instead.
+This `musical instrument McGurk effect' is partly the reason behind hiding the electronics out of sight.
+% The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument.
+With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component may provide the necessary `push' of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
 
-The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
-The keyboard was consciously chosen to mimic that of the original XVI century Italian harpsichords.
-This design decision was made in order to use the propensity humans have to being influenced by visual components when musical making judgements \cite{Tsay2013} to our advantage.
-It is especially true with respect to musical instruments as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have clearly shown.
-Rather than avoiding this effect we have chosen instead to utilise it.
-This `musical instrument McGurk effect' is in part the reason behind hiding the electronics out of sight.
-The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument.
-With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component may provide the necessary `push' of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
+The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica.
+The keyboard layout was deliberately modelled after typical 16th-century Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
+This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}.
+Instead of attempting to mitigate this influence, the design embraces it.
+This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view.
+\textcolor{red}{Any good references on the McGurk effect?} The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument.
+Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.
 
@@
 In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack.
-Though the current configuration simply sends MIDI note on and off messages when passing a threshold, the displacement of all jacks is available continuously.
-The staggering of the pluck point of the jacks between registers means that there are two very distinct tactile `anchors' along the key dip.
-A combination of string tension, quill voicing and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
+Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
+The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip.
+A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
 Such tactile expectations provide fertile opportunities for extending the expressive capabilities of the interface.
@@
 \end{anonsuppress}
-% FILL OUT AS APPROPRIATE (and anonymise).
 
@@
 The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra.
-The design was inspired by Venetian harpsichords, particularly the 1547 Alessandro Trasuntino harpsichord at the San Colombano Museum.
+The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}.
 An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
@@
     \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from San Colombano, Bologna, 2023}
+    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023}
     \label{fig:3key}
@@
 \begin{itemize}
-    \item 1 Arduino Nano 33 IoT
+    \item 1 Arduino Nano 33 BLE
     \item 1 Fujitsu MB85RS64 SPI FRAM chip
@@
 
-For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt.
+For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt\footnote{\href{https://www.native-instruments.com/en/products/komplete/samplers/kontakt-8/?srsltid=AfmBOorKUf43SoIxGBS2-GnXmKHHkgcfcfWRskpweDhLSG3FiF0qrf2w}{Kontakt webpage} (accessed 31 Jan 2025)}.
 The software was installed on a Mac Mini hosted inside the instrument's case.
-Holes were drilled at the back of the instrument's frame, away from sight, allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}).
+Holes were drilled at the back of the instrument's frame, away from sight, to allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}).
 % These components are accessible via locked drawer which is visible in \ref{fig:oratory}. 
-An iPad is used as a monitor through which visitors can adjust parameters of playback such as tuning, voicing and which stops are engaged.
-The next iteration will use a library of samples from instruments in the collection, which is currently being carried out by the museum.
+An iPad is used as a monitor through which visitors can adjust playback parameters such as tuning, voicing and which stops are engaged.
+The next iteration will use a library of samples from instruments in the collection, which the museum is currently carrying out.
 %A main 12-inch display sits inside a custom-made drawer locked during usage.
@@
 The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
-Collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
-The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' as found in the McAlpine exhibition \cite{McAlpine2014} were absent in this initial feedback.
+The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
+The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.
+\textcolor{red}{This goes in the direction pointed by Andrew, but I wonder if it could be worded differently}
 
-A more formal surveying of museum visitors is to be carried out to determine to what degree the exhibition achieved it's core aims.
-% Testers unfamiliar with harpsichords were particularly impressed by its responsiveness.
+A more formal surveying of museum visitors is to be carried out to determine to what degree the exhibition achieved its core aims.
 The keyboard is hosted in the \emph{Oratory} above the museum's main hall, an exceptional testimony to the Bolognese art, decorated by the finest students of the Carracci and presenting a series of frescoes covering most of the walls and ceiling.
-The room hosts unique examples of the Italian Renaissance building tradition, including the 1547 harpsichord and the 1540 spinet by Alessandro Trasuntino.
-Visitors often visit the room with the same caution and respect that is typical of worship spaces.
-Among the visitors interviewed, there were no comments made that the exposed headphones and touchscreen affected such an experience negatively, though a more definitive answer will come as more feedback is collected after launch.
+The room hosts unique examples of the Italian Renaissance building tradition, including the 1547 harpsichord and the 1540 spinet by \anon{Alessandro Trasuntino}.
+Visitors often visit the room with the same caution and respect typical of worship spaces.
+Among the visitors interviewed, no comments were made that the exposed headphones and touchscreen affected such an experience negatively, though a more definitive answer will come as more feedback is collected after launch.
 Pictures of the keyboard hosted in the \emph{Oratory} are shown in Figure \ref{fig:oratory}.
 
-Expert feedback has highlighted certain limitations, including imprecise key calibration, which creates a temporal disconnect between the tactile plucking sensation and sound onset.
+Expert feedback has highlighted limitations, including imprecise key calibration, which creates a temporal disconnect between the tactile plucking sensation and sound onset.
 Additionally, the commercial sample library, while allowing the selection of registers such as 8', 4', and their combinations, restricts functionality to a single MIDI message per key, regardless of the number of string choirs controlled.
 Inspired by early Italian instruments, the keyboard's layout is designed to manage two 8' registers with a sensor system fitted to each.
-Internally the instrument is therefore 2 individual MIDI devices however, one register needed to be disengaged as the software could only communicate with a single device.
-Multiple instances of the software could be used, but an arrangement of 2 sets GUI controls was deemed too confusing to visitors who would interact with the exhibition without orientation.
-Sensors are therefore active only on a single set of jacks to accommodate these limitations.
-Technical improvements, such as refined calibration with hysteresis and developing a custom sample library and interface, could address these constraints in future iterations.
+Internally, the instrument functions as two separate MIDI devices.
+However, one register must be disengaged due to the software's limited communication with only a single device.
+While running multiple instances of the software is possible, having two sets of GUI controls was considered too confusing for visitors engaging with the exhibition without prior orientation.
+Consequently, sensors are only active on single jacks to accommodate these constraints.
+Future iterations could address these limitations through technical improvements, such as refined calibration with hysteresis and developing a custom sample library and interface.
 
@@
 \includegraphics[width = \linewidth]{src/images/keyboardMuseum.JPEG}
-\caption{The keyboard installed in the \emph{Oratory} at San Colombano, and a user wearing headphones playing it.}
+\caption{The keyboard installed in the \emph{Oratory} at \anon{San Colombano}, and a user wearing headphones playing it.}
 \label{fig:oratory}

### 2025-01-31 src/main.tex
Updates from Overleaf
second draft

@@
 \documentclass[sigconf]{nimeart}
+% \documentclass[sigconf,anonymous]{nimeart}
 % \documentclass[sigconf,anonymous,review]{nimeart}
-\usepackage{graphicx,subcaption}
-% \usepackage{tikz}
-% \usetikzlibrary{shapes,arrows}
+\usepackage{graphicx,subcaption,svg}
 
@@
 \affiliation{%
-  \institution{Master Harpsichord Maker}
+  \institution{Master Harpsichord Maker \\ Museo di San Colombano}
   \city{Pesaro}
@@
 \section{Ethical Standards}
-This work received funding from the European Commission's Horizon2020 programme, through grant \anon{ERC-StG 950084 NEMUS}
+This work is an empirical study of a keyboard system and
+involved no research participants.
+The authors obtained explicit consent to collect feedback.
+Seagull feathers, employed as a building material for the jacks, were ethically sourced and picked up along the Pesaro coastline.
 

### 2025-02-03 src/article.tex
Updates from Overleaf

@@
 
-The Tagliavini Collection in Bologna \cite{Tagliavini2007}, renowned for its historical keyboard instruments, exemplifies this dilemma.
+The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma.
 With more than seventy early keyboard instruments, including many unique Renaissance Italian harpsichords, spinets and virginals, the collection stands out as a valuable resource for musicologists, organologists and musicians alike.
-Preserving the instruments' authenticity was the cornerstone of Ferdinando Tagliavini’s vision.
+Preserving the instruments' authenticity was the cornerstone of \anon{Ferdinando Tagliavini}’s vision.
 This guiding principle led him to collect only playable instruments, or instruments which could be restored to playing condition after minimal intervention.
@@
 Visitors to the museum are invited to play the interface and listen through a pair of headphones.
-This work outlines the technological aspects of the interface's construction and offers reflections on its role within the Tagliavini Collection and potential application within the wider musical instrument museum context.
+This work outlines the technological aspects of the interface's construction and offers reflections on its role within the \anon{Tagliavini} Collection and potential application within the wider musical instrument museum context.
 % The paper outlines the design and motivations behind this replica keyboard interface.
@@
 
-McAlpine discusses a case similar to the Tagliavini collection in his examination of the Benton Fletcher Collection at National Trust Fenton House \cite{McAlpine2014}.
+McAlpine discusses a case similar to the \anon{Tagliavini} collection in his examination of the Benton Fletcher Collection at National Trust Fenton House \cite{McAlpine2014}.
 When these instruments were donated, Benton Fletcher stipulated that they remain playable and should continue to be maintained for tuition and public performance.
@@
 The keyboard presented here inherits some aspects from the Tromba Moderna project.
-The project aims to offer a tool to enhance the fruition of the Tagliavini collection while retaining a form of continuity with historical instrument-building traditions.
+The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions.
 However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms.
@@
 
-The design of the augmented replica keyboard for the Tagliavini Collection was guided by three principal constraints stipulated by the museum:
+The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:
 \begin{itemize}
@@
 Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles.
-Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software, reducing reliance on proprietary tools and simplifying the process for users.
+Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}, reducing reliance on proprietary tools and simplifying the process for users.
 
@@
     \end{subfigure}
-    \caption{Logarithmic shape of an original Trasuntino harpsichord contrast with rectangular shape of the replica}    
+    \caption{Logarithmic shape of an original \anon{Trasuntino} harpsichord contrast with the rectangular shape of the replica}    
     \Description{}
@@
     \item \emph{Non-invasiveness.} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
-    \item \emph{Low Latency.} The time required to process a single iteration of all sensor data should remain under 10 ms, with an upper limit of 25 ms deemed acceptable.
+    \item \emph{Low Latency.} The sampling period for reading and processing data from all sensors should remain under 10 ms to allow for latency introduced in other parts of the synthesis process.
 Empirical criteria found in previous studies \cite{Jack2016} was used as a guide.
@@
     \item 1 Texas Instruments CD4051BE multiplexer for signal aggregation.
+    \item 7 WS2812 RGB LEDs with integrated driver.
 \end{itemize}
@@
 A manual visual check of plotted all sensors readings confirmed that interference from neighbouring jacks had been eliminated.
+RGB LEDs are placed on the reverse side of the PCB, vertically in-line with the sensors.
 
@@
 Power requirements for the system were estimated at 1.1 A at 5 V, with some fluctuation when the system was first powered on.
-While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic activation of sensors found in the McPherson piano \cite{McPherson2013}.
+While the sensors were powered continuously in this iteration, future designs may incorporate power-saving measures, such as dynamic modulation of the optical emitters found in the McPherson piano \cite{McPherson2013}.
 

### 2025-02-03 src/article.tex

@@
   \centering
-  \includegraphics[width=0.9\linewidth]{src/images/tagging-jacks-3.jpg} 
+  \includegraphics[width=0.9\linewidth,trim={0 0 0 0.5cm},clip]{src/images/tagging-jacks-3.jpg} 
   \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
@@
 
-\begin{figure}[!b] 
+\begin{figure}[!h] 
   \centering
-  \includegraphics[width=0.7\linewidth]{src/images/simple-schematic-bw-.jpg} 
+  \includegraphics[width=0.7\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/simple-schematic-bw-.jpg} 
   \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
@@
 A manual visual check of plotted all sensors readings confirmed that interference from neighbouring jacks had been eliminated.
-RGB LEDs are placed on the reverse side of the PCB, vertically in-line with the sensors.
+RGB LEDs are placed on the reverse side of the PCB, vertically in-line with the sensors, and were added to provide a programmable means of providing visual feedback.
+The LEDs are controlled via a pulse width modulation (PWM) signal and are addressable individually.
+
+\begin{figure}[!b] 
+  \centering
+  \includegraphics[width=\linewidth,trim={0 10cm 0 7cm},clip]{src/images/sensor-board-reverse-side-anon.png} 
+  \caption{Reverse side of sensor board showing the RGB LEDs and terminal connection.}
+  \Description{} 
+  \label{fig:sensor-reverse}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{src/images/block-diagram.pdf}
+    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
     \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
@@
     \centering
-    \includegraphics[width=\linewidth]{src/images/sensor-board-w-baffles.jpeg}
+    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
     \\
-    \includegraphics[width=0.8\linewidth]{src/images/baffles.png}
+    \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
     \caption{Baffles designed to prevent cross-talk between adjacent sensors.}

### 2025-02-03 src/article.tex
Updates from Overleaf

@@
 The LEDs are controlled via a pulse width modulation (PWM) signal and are addressable individually.
+The output signal of the multiplexer is taken from each PCB and routed to a separate ADC channel of the Arduino.
+Multiplexer channel select and LED PWM signals are daisy-chained through each PCB (Figure \ref{fig:sensor-reverse}).
 
@@
   \centering
-  \includegraphics[width=\linewidth,trim={0 10cm 0 7cm},clip]{src/images/sensor-board-reverse-side-anon.png} 
-  \caption{Reverse side of sensor board showing the RGB LEDs and terminal connection.}
+  \includegraphics[width=\linewidth]{src/images/sensor-board-reverse-side-anon.png} 
+  \caption{Reverse side of the sensor board showing the RGB LEDs and terminal connections.
+The \texttt{SIG} terminal (Top Left) is routed to an ADC channel.
+Power \texttt{5V, 3V, GND}, multiplexer channels \texttt{A, B, C} and LED PWM input \texttt{DI} and output \texttt{DO}  connect from the previous sensor board (Left) and are routed to the next (Right).}
   \Description{} 

### 2025-02-03 src/article.tex
fix #39 #40

@@
 Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions.
-This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME \cite{Masu_NIME_2023}.
+This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
@@
 \includegraphics[width=0.8\linewidth]{src/images/details.jpg}
-\caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}\label{fig:details}
+\caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}
+\label{fig:details}
 \end{figure*}
@@
 The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:
+
 \begin{itemize}
@@
 The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
-\item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation, keeping artefacts to a minimum. 
+    \item \emph{Faithfulness}.
+The keyboard mechanism ensured fidelity to its authentic operation.
+The electronics system would not seek to `fix' or `improved' on limitations of the original 16th-century design.
 \end{itemize}
 
-These constraints demanded a non-invasive design that was respectful of the historical authenticity of the instruments.
+These constraints demanded an electronics system that was non-invasive in design and that was respectful of the historical aesthetics of the instruments.
 The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
 While actuators can generate considerable force, it is difficult to achieve a jack's free motion and resistance without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
-The instrument itself can already achieve the tactile realism of the harpsichord.
+A recreation of the mechanism can already achieve the tactile realism of the harpsichord.
 It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation.
-% (A REFERENCE WOULD BE GOOD HERE, ANY IDEAS?)
-% The final design, visible in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key two-register harpsichord keyboard replica.
-% The keyboard was consciously chosen to mimic the typical layout of 16th-century Italian harpsichords.
-This design decision was made to use humans' propensity to being influenced by visual components when musical making judgements \cite{Tsay2013} to our advantage.
-It is especially true concerning musical instruments, as the studies by Fritz \emph{et al}\cite{Fritz2012, Fritz2014, Fritz2017} have shown.
-Rather than avoiding this effect, we have chosen to utilise it instead.
-This `musical instrument McGurk effect' is partly the reason behind hiding the electronics out of sight.
-% The aesthetics of the interface make it more likely that the interface be accepted on the same terms as an `authentic' musical instrument.
-With all the work into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020} it is possible the visual component may provide the necessary `push' of acceptance similar to that which influences our judgment on performance \cite{Tsay2013}.
 
@@
 This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view.
-\textcolor{red}{Any good references on the McGurk effect?} The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument.
+The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument.
 Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.
@@
 Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
+The jacks for a single key typically have their pluck points offset from each other, a practice known as `staggering'.
+A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
 The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip.
-A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
-Such tactile expectations provide fertile opportunities for extending the expressive capabilities of the interface.
 The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.
@@
 
-\subsection{Project Deployment}
-
-A further objective, set forth by the authors, was to ensure accessibility and reproducibility by committing to an open source approach for all outputs of the project.
-The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
-Cost-effectiveness was also a central consideration.
-% The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity. 
-Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space.
-
-Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
-
-The Arduino Nano 33 BLE was chosen as the core microcontroller for its compatibility with open source tools and ability to support both USB and BLE MIDI.
-Ferroelectric RAM (FRAM) was employed for data storage since it offers an economical yet robust solution for preserving calibration settings across power cycles.
-Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}, reducing reliance on proprietary tools and simplifying the process for users.
-
-\begin{anonsuppress}
-Reference repositories for this project can be found here:
-
-    \begin{itemize}
-        \item 
-        \emph{Firmware}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-firmware}}
-        \item 
-        \emph{PCB CAD}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-cad}}
-        \item 
-        \emph{Interface Models}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-models}}
-    \end{itemize}
-\end{anonsuppress}
-
 \subsection{Materials and Construction}
-
 The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord.
@@
 
-The rectangular poplar frame also deviates from the traditional logarithmic \ref{fig:log-harp-comp} form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
+The rectangular poplar frame also deviates from the traditional logarithmic (Figure \ref{fig:log-harp-comp}) form to allow for the installation of the electronic sensors, as visible in Figure \ref{fig:49-key-bottom}, without compromising the visual or tactile authenticity.
 Two string choirs, crafted from yellow brass wire and anchored with wrought iron pins, were tensioned to replicate authentic plucking resistance.
@@
 
+\subsection{Project Deployment}
+
+A further objective, set forth by the authors, was to ensure reproducibility by committing to an open source approach for all outputs of the project.
+The commitment to open sourcing encompassed all aspects of the system, including hardware schematics, firmware, and calibration data.
+Cost-effectiveness was also a central consideration.
+% The system was initially developed for a 3-key prototype, shown in Figure \ref{fig:3key}, and successfully scaled to 49 keys without significant increases in cost or complexity. 
+Specifically, the system was designed to be easily assembled using resources typically available in a university-managed maker space.
+Components, such as QRE1113 optical sensors and CD4051BE multiplexers, are widely available from commercial resellers, while the modular PCB design ensures easy replication and maintenance.
+The Arduino Nano was chosen as the core microcontroller format for its compatibility with open source tools.
+Calibration workflows were optimised using the Arduino IDE’s serial plotter and open source MIDI Monitor software \footnote{\url{https://github.com/krevis/MIDIApps}}, reducing reliance on proprietary tools and simplifying the process for users.
+
+\begin{anonsuppress}
+Reference repositories for this project can be found here:
+
+    \begin{itemize}
+        \item 
+        \emph{Firmware}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-firmware}}
+        \item 
+        \emph{PCB CAD}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-cad}}
+        \item 
+        \emph{Interface Models}: \anon{\url{https://github.com/Nemus-Project/harpsichord-interface-models}}
+    \end{itemize}
+\end{anonsuppress}
+
 \section{Hardware Design}\label{hardware-design}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
+    \caption{Block diagram of PCB connections. A separate sensor signal is routed to the Arduino. LED and multiplexer (MUX) controls signal are daisy-chained through each sensor PCB.}
+    \label{fig:system-block-diagram}
+\end{figure*}
 
@@
     \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023}
+    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023.}
     \label{fig:3key}
@@
 \begin{itemize}
-    \item \emph{Non-invasiveness.} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
-    \item \emph{Low Latency.} The sampling period for reading and processing data from all sensors should remain under 10 ms to allow for latency introduced in other parts of the synthesis process.
+    \item \emph{Non-invasiveness:} No remarkable modifications were allowed on the harpsichord mechanics, particularly on all the visible parts.
+    \item \emph{Low Latency:} The sampling period for reading and processing data from all sensors should remain under 10 ms to allow for latency introduced in other parts of the synthesis process.
 Empirical criteria found in previous studies \cite{Jack2016} was used as a guide.
     \item \emph{Reliability:} Sensor data should be dependable and consistent, with interference from anything external to the jack movement being absent or negligible.
-    \item \emph{Scalability.} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (Figure \ref{fig:3key}) up to the 49 keys of the final design.
-    % \item \emph{Expandability.} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
+    \item \emph{Scalability:} The design needed to scale in both cost and time in order to be adaptable from a 3-key prototype (Figure \ref{fig:3key}) up to the 49 keys of the final design.
+    % \item \emph{Expandability:} The system should accommodate future functionality, such as additional MIDI parameters or data visualisation.
 \end{itemize}
 
-The hardware was designed to make assembly possible in standard maker spaces.
+The hardware was designed to make assembly possible in standard university maker spaces.
 
@@
   \centering
-  \includegraphics[width=0.7\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/simple-schematic-bw-.jpg} 
-  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer}
+  \includegraphics[width=0.7\linewidth,trim={0 0.5cm 0 0.5cm},clip]{src/images/simple-schematic-bw-.jpg} 
+  \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer.}
   \Description{} 
@@
 
-The optical sensors are wired in a voltage divider configuration \ref{fig:simple-schematic} and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
+The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
 The gradient stickers affixed to each jack provided a surface with varying reflectivity for the sensors, which were used to track jack displacement throughout the key dip.
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
+    \\
+    \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
+    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
+    \Description{}
+    \label{fig:baffles}
+\end{figure}
 
@@
 
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
+    \caption{Components and cabling hidden in the bottom section of the instrument case.}
+    \label{fig:mac-mini}
+\end{figure}
+
 \begin{itemize}
-    \item 1 Arduino Nano 33 BLE
-    \item 1 Fujitsu MB85RS64 SPI FRAM chip
-    \item 1 EC11 combined rotary encoder and tactile switch
+    \item 1 Arduino Nano 33 BLE.
+    \item 1 Fujitsu MB85RS64 SPI Ferroelectric RAM chip.
+    \item 1 EC11 combined rotary encoder and tactile switch.
     % \item 1 DFROBOT DFR0785 Tactile Switch
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
-    \caption{Block diagram of PCB connections. Sensor and LED are routed through each sensor PCB}
-    \label{fig:system-block-diagram}
-\end{figure*}
+The Arduino Nano's small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets.
+The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as both native USB MIDI and BLE MIDI devices.
+The 33 BLE was also able to achieve a sampling period of 1.7 milliseconds for all sensors.
+A 4-sample moving average filter was implemented to reduce noise and brought the effective sampling period to 5.1 milliseconds.
+% Samping period was measure by timing each iterations of the entire sensor read and processing cycle, polling the internal clock of the arduino a the start and end of each cycle and storing the difference in a buffer.
+After 10k cylces the buffer of timing was sent over serial and a mean average of 1.7 milliseconds was found with a jitter of ±0.2 milliseconds.
 
-The Arduino Nano small form factor made fitting the board inside the harpsichord easier and also allowed for testing of multiple chipsets.
-The 33 BLE variation of the Nano had the additional benefit in its ability to be programmed as a BLE MIDI device.
-%Initial iterations of the controller board used 2.54 mm pitch headers for cable connections, but issues with cable disconnections led to the adoption of JST-PH connectors in later designs.
-Custom cable looms were constructed using adapted crimping tools, ensuring robust connections without compromising accessibility.
-
-Non-volatile memory, implemented using a Fujitsu MB85RS64 Ferroelectric RAM (FRAM) chip, provided a reliable means of storing calibration data.
+Non-volatile memory, implemented using a Fujitsu MB85RS64 Ferroelectric RAM (FRAM) chip, provided a reliable means of storing and preserving calibration settings across power cycles.
 This ensured that sensor thresholds and other settings could be preserved across power cycles, enhancing the system's usability in museums.
-
-%\subsection{Calibration and Power Management}\label{calibration}
 
@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\linewidth]{src/images/serial_monitor.png}
-%     \caption{Arduino IDE serial plotter used for calibration of sensor thresholds.
-\texttt{M} and \texttt{m} represent the maximum an minimum value of digitised data.
-\texttt{K} and \texttt{T} are the values of the sensor and threshold being calibrated, followed by a two-digit number denoting the sensor index (in this case, sensor 32).}
-%     \label{fig:serial_monitor}
-% \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
-    \caption{Components and cabling hidden in the bottom section of the instrument case.}
-    \label{fig:mac-mini}
-\end{figure}
-
-% \subsection{Final Assembly and Integration}
-
-% The final 49-key harpsichord replica incorporated the sensor and controller boards into the rectangular frame.
-Ribbon cables connected the PCBs, while felt strips were interwoven between the strings to suppress vibrations and maintain the tactile response of the keys.
-The assembly process at the \anon{NEMUS} lab in \anon{Bologna} involved meticulous adjustments to key clearances and wiring layouts.
-Modifications included milling sections of the keys to ensure adequate clearance for the PCBs and varnishing the jacks to improve sticker adhesion.
-
 For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt\footnote{\href{https://www.native-instruments.com/en/products/komplete/samplers/kontakt-8/?srsltid=AfmBOorKUf43SoIxGBS2-GnXmKHHkgcfcfWRskpweDhLSG3FiF0qrf2w}{Kontakt webpage} (accessed 31 Jan 2025)}.
@@
 Holes were drilled at the back of the instrument's frame, away from sight, to allow the passage of cables, such as from power supplies, USB connectors and headphone jack (Figure \ref{fig:mac-mini}).
-% These components are accessible via locked drawer which is visible in \ref{fig:oratory}. 
 An iPad is used as a monitor through which visitors can adjust playback parameters such as tuning, voicing and which stops are engaged.
 The next iteration will use a library of samples from instruments in the collection, which the museum is currently carrying out.
-%A main 12-inch display sits inside a custom-made drawer locked during usage.
-
-\begin{figure}[!b]
-    \centering
-    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
-    \\
-    \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
-    \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
-    \Description{}
-    \label{fig:baffles}
-\end{figure}
 
@@
 The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.
-\textcolor{red}{This goes in the direction pointed by Andrew, but I wonder if it could be worded differently}
+
+%\textcolor{red}{This goes in the direction pointed by Andrew, but I wonder if it could be worded differently}
 

### 2025-02-04 src/article.tex
Start catalina comments

@@
 The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma.
-With more than seventy early keyboard instruments, including many unique Renaissance Italian harpsichords, spinets and virginals, the collection stands out as a valuable resource for musicologists, organologists and musicians alike.
+Housed at Museo San Colombano - Genus Bononiae, with more than seventy early keyboard instruments, primarily early plucked stringed keyboards of italian origin.
+The collection stands out as a valuable resource for musicologists, organologists and musicians alike.
 Preserving the instruments' authenticity was the cornerstone of \anon{Ferdinando Tagliavini}’s vision.
-This guiding principle led him to collect only playable instruments, or instruments which could be restored to playing condition after minimal intervention.
+This guiding principle led him to collect instruments which could be restored to playing condition after minimal intervention.
 
-However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator.
-To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance, which is the subject of this paper.
+However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—by experienced historical keyboard performers or by young musicians under strict supervision from the curator.
+To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian school, which is the subject of this paper.
 The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
-The interface is currently connected to a commercial software sampler, though the ultimate goal is to use sample instruments from the collection.
-Visitors to the museum are invited to play the interface and listen through a pair of headphones.
+The interface is currently connected to a commercial software sampler, though the ultimate goal is to render accessible the sound of instruments that can not be kept on a playing condition.
+Museum visitors are invited to play the interface and listen through a pair of headphones.
 This work outlines the technological aspects of the interface's construction and offers reflections on its role within the \anon{Tagliavini} Collection and potential application within the wider musical instrument museum context.
-% The paper outlines the design and motivations behind this replica keyboard interface.
+
+As part of the initiative to improve the accessibility of museums, the Museo San Colombano - Tagliavini Collection, part of the cultural itinerary of Fondazione Carisbo, has designed a series of projects to improve cognitive and sensory accessibility at different levels, in a project funded by NextGenerationEU.
+The aim of this project is to exploit the potential of the collection of musical instruments to create an immersive experience and interaction with the heritage on display.
+reinterpretation of the historical musical objects in their tactile-auditory dimension aims to open up the exhibition to blind and partially sighted visitors.
+The augmented keyboard fulfils many of the project's aims and is part of the initiative to allow visitors of different abilities to engage in a hands-on creative process within the museum.
 
@@
 The MIDI controller, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each.
-However, user tests identified a limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.
+However, user tests identified a limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical plucked keyboard instrument \cite{McAlpine2014}.
 
@@
 This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}.
+\textcolor{red}{ALT: This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in `The \emph{O} in NIME' \cite{Masu_NIME_2023}}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
@@
 
-Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \anon{\cite{NEMUS}}, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.
+Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \anon{\cite{NEMUS}}, aiming to virtually reproduce the sound of antique keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.
 
@@
 
-The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:
+The design of the augmented replica keyboard for the \anon{Museo San Colombano - Genus Bononiae} was guided by three principal constraints stipulated by the museum:
 
 \begin{itemize}
+    \item \emph{Faithfulness}.
+The keyboard mechanism ensured fidelity to its authentic operation.
+The electronics system would not seek to `fix' or `improved' on limitations of the original design.
+    \item \emph{Robustness and reliability}.
+The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
     \item \emph{No visible electronics}.
 To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
-    \item \emph{Robustness and reliability}.
-The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
-    \item \emph{Faithfulness}.
-The keyboard mechanism ensured fidelity to its authentic operation.
-The electronics system would not seek to `fix' or `improved' on limitations of the original 16th-century design.
+    \item \emph{Reduced use of space}. As museums often face space limitations, the instrument needed be compact enough not to compromise space for exhibition of the permanent display. 
 \end{itemize}
@@
 The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica.
-The keyboard layout was deliberately modelled after typical 16th-century Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
+The keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
 This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}.
@@
 The jacks for a single key typically have their pluck points offset from each other, a practice known as `staggering'.
-A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
+A combination of string tension, quill voicing, and jack staggering contribute to the overall resistance of the action, with each key differing from the low to the high octaves \cite{Veroli2012}.
 The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip.
@@
 \subsection{Materials and Construction}
-The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord.
+The keyboard was designed to replicate the tactile and aesthetic sensations of playing an historical Italian harpsichord.
 Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard.
@@
     \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023.}
+    \caption{
+    3-Key Model Harpsichord Mechanism    
+    %acquired from \anon{San Colombano, Bologna}, 2023.
+    }
     \label{fig:3key}
@@
 Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant internal measurement and layout discrepancies.
-Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was acquired from the museum and used as a foundation for prototyping.
+Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was used as a foundation for prototyping.
 This approach followed a methodology similar to that used in Timmermans \emph{et al.}'s Haptic Key project \cite{Timmermans2020}, since the 3-key model enabled iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances, before upscaling.
@@
 % In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
-The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
+The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 people -- consisting of staff, surveillance personnel and visitors who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
 The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.

### 2025-02-04 src/main.tex
Updates from Overleaf
Start catalina comments

@@
 %%
-\documentclass[sigconf]{nimeart}
-% \documentclass[sigconf,anonymous]{nimeart}
+% \documentclass[sigconf]{nimeart}
+\documentclass[sigconf,anonymous]{nimeart}
 % \documentclass[sigconf,anonymous,review]{nimeart}
@@
 \affiliation{%
-  \institution{Master Harpsichord Maker \\ Museo di San Colombano}
+  \institution{Master Harpsichord Maker \\ Museo San Colombano}
   \city{Pesaro}
@@
 \affiliation{%
-  \institution{Museo di San Colombano}
+  \institution{Museo San Colombano}
   \city{Bologna}
@@
 
-This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument.
+This paper describes the design and creation of an electronically augmented replica of a harpsichord keyboard with a typical early Italian layout to create a digital musical instrument. 
 The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
-The replica is designed to balance the competing demands of historical authenticity, public accessibility, and preservation. 
+The replica attempts to balance the competing demands of historical authenticity, public accessibility, and preservation. 
 It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques.
@@
 This work situates itself within broader discussions on the role of technology in museums. 
-A keyboard interface of this type offers an opportunity to enhance visitor interaction while safeguarding delicate artefacts. 
+A keyboard interface of this type offers an opportunity to enhance visitor interaction with cultural heritage while safeguarding delicate artefacts. 
 The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.
@@
 The authors obtained explicit consent to collect feedback.
-Seagull feathers, employed as a building material for the jacks, were ethically sourced and picked up along the Pesaro coastline.
+Seagull feathers, employed as a building material for the jacks, were ethically sourced and collected along the Pesaro coastline.
 

### 2025-02-04 src/article.tex

@@
 The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma.
-Housed at Museo San Colombano - Genus Bononiae, with more than seventy early keyboard instruments, primarily early plucked stringed keyboards of italian origin.
+Housed at \anon{Museo San Colombano - Genus Bononiae}, with more than seventy early keyboard instruments, primarily early plucked stringed keyboards of Italian origin.
 The collection stands out as a valuable resource for musicologists, organologists and musicians alike.
@@
 
+\textcolor{red}{
 As part of the initiative to improve the accessibility of museums, the Museo San Colombano - Tagliavini Collection, part of the cultural itinerary of Fondazione Carisbo, has designed a series of projects to improve cognitive and sensory accessibility at different levels, in a project funded by NextGenerationEU.
@@
 The augmented keyboard fulfils many of the project's aims and is part of the initiative to allow visitors of different abilities to engage in a hands-on creative process within the museum.
+}
 
@@
 Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions.
-This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}.
-\textcolor{red}{ALT: This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in `The \emph{O} in NIME' \cite{Masu_NIME_2023}}.
+This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in `The \emph{O} in NIME' \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
-Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.
+This project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.
 
@@
 
-% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors.
 The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 people -- consisting of staff, surveillance personnel and visitors who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
@@
 
-A more formal surveying of museum visitors is to be carried out to determine to what degree the exhibition achieved its core aims.
+A more formal surveying of museum visitors would be required to determine to what degree the exhibition achieved its core aims.
 The keyboard is hosted in the \emph{Oratory} above the museum's main hall, an exceptional testimony to the Bolognese art, decorated by the finest students of the Carracci and presenting a series of frescoes covering most of the walls and ceiling.

### 2025-02-04 src/article.tex
submission draft

@@
 
-The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma.
-Housed at \anon{Museo San Colombano - Genus Bononiae}, with more than seventy early keyboard instruments, primarily early plucked stringed keyboards of Italian origin.
-The collection stands out as a valuable resource for musicologists, organologists and musicians alike.
+The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, housed at \anon{Museo San Colombano - Genus Bononiae} and renowned for its historical keyboard instruments, exemplifies this dilemma.
+With over fifty early keyboard instruments, primarily early plucked stringed keyboards of Italian origin, the collection stands out as a valuable resource for musicologists, organologists and musicians alike.
 Preserving the instruments' authenticity was the cornerstone of \anon{Ferdinando Tagliavini}’s vision.
-This guiding principle led him to collect instruments which could be restored to playing condition after minimal intervention.
+This guiding principle led him to collect instruments that could be restored to their playing condition after minimal intervention.
 
-However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—by experienced historical keyboard performers or by young musicians under strict supervision from the curator.
+However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—by experienced historical keyboard performers or young musicians under strict supervision from the curator.
 To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian school, which is the subject of this paper.
 The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
-The interface is currently connected to a commercial software sampler, though the ultimate goal is to render accessible the sound of instruments that can not be kept on a playing condition.
+The interface is presently linked to a commercial software sampler; however, the ultimate aim is to make the sound of instruments in the collection that can no longer be maintained in playable condition accessible.
 Museum visitors are invited to play the interface and listen through a pair of headphones.
 This work outlines the technological aspects of the interface's construction and offers reflections on its role within the \anon{Tagliavini} Collection and potential application within the wider musical instrument museum context.
-
-\textcolor{red}{
-As part of the initiative to improve the accessibility of museums, the Museo San Colombano - Tagliavini Collection, part of the cultural itinerary of Fondazione Carisbo, has designed a series of projects to improve cognitive and sensory accessibility at different levels, in a project funded by NextGenerationEU.
-The aim of this project is to exploit the potential of the collection of musical instruments to create an immersive experience and interaction with the heritage on display.
-reinterpretation of the historical musical objects in their tactile-auditory dimension aims to open up the exhibition to blind and partially sighted visitors.
-The augmented keyboard fulfils many of the project's aims and is part of the initiative to allow visitors of different abilities to engage in a hands-on creative process within the museum.
-}
 
@@
 Karp \cite{Karp1979,Karp1985} advocates for a deeper understanding of musical instruments so that enough knowledge is generated to make them as ``copyable'' as possible.
+
+As part of a broader initiative to enhance accessibility in museums, \anon{the Museo San Colombano - Tagliavini Collection}, within the cultural itinerary of \anon{Fondazione Carisbo}, has developed a series of projects aimed at improving cognitive and sensory access, supported by funding from NextGenerationEU.
+The project seeks to leverage the museum’s collection of musical instruments to create an immersive and interactive engagement with the exhibited heritage.
+The reinterpretation of historical musical objects through their tactile and auditory dimensions is intended to make the exhibition more accessible to blind and partially sighted visitors.
+The augmented keyboard aligns with these objectives, serving as an important component of the initiative to provide visitors of varying abilities with a hands-on, creative experience within the museum.
 
@@
 The keyboard presented here inherits some aspects from the Tromba Moderna project.
-The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions.
+The project aims to offer a tool to enhance the visitor experience of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions.
 However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms.
+
 The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}.
-Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions.
-This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in `The \emph{O} in NIME' \cite{Masu_NIME_2023}.
+As such, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.
 The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.
-This project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.
 
@@
     \item \emph{Faithfulness}.
-The keyboard mechanism ensured fidelity to its authentic operation.
-The electronics system would not seek to `fix' or `improved' on limitations of the original design.
+The keyboard mechanism had to ensure fidelity to its authentic operation.
+The electronics system would not seek to `fix' or `improve' the limitations of the original design.
     \item \emph{Robustness and reliability}.
 The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
-    \item \emph{No visible electronics}.
+    \item \emph{No visible electronic components}.
 To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
@@
 
-These constraints demanded an electronics system that was non-invasive in design and that was respectful of the historical aesthetics of the instruments.
+These constraints demanded an electronics system that was non-invasive in design and respectful of the instruments' historical aesthetics.
 The requirement for faithfulness to the original mechanism immediately ruled out a mechanical design reliant on electromechanical actuators, as in previous works on piano haptics \cite{Timmermans2020,Gillespie1996}.
-While actuators can generate considerable force, it is difficult to achieve a jack's free motion and resistance without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
-A recreation of the mechanism can already achieve the tactile realism of the harpsichord.
-It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation.
+While actuators can generate considerable force, achieving a jack's free motion and resistance is difficult without extensive mechanical intervention such as that carried out by Gillespie \cite{Gillespie1996}.
+% Recreating the mechanism can already achieve the tactile realism of the harpsichord.
+% It should first be determined if mechanical recreation is sufficient as part of a digital musical instrument before pursuing elaborate electromechanical emulation.
 
 The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica.
-The keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
+Since the Tagliavini collection hosts primarily early plucked stringed keyboards of Italian origin, the keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
 This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}.
@@
   \centering
-  \includegraphics[width=0.9\linewidth,trim={0 0 0 0.5cm},clip]{src/images/tagging-jacks-3.jpg} 
+  \includegraphics[width=\linewidth,trim={0 0 0 0.5cm},clip]{src/images/tagging-jacks-3.jpg} 
   \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
@@
 A modular system of printed circuit boards (PCBs) was designed to manage the sensors and process their output.
-The system uses 49 optical sensors which are distributed across seven boards that contain 7 sensors each.
-The PCBs were secured to the underside of wrest plank and above the keys (Figure \ref{fig:49-key-bottom}), allowing them to be adjusted during installation.
+The system uses 49 optical sensors distributed across seven boards containing 7 sensors each.
+The PCBs were secured to the underside of the wrest plank and above the keys (Figure \ref{fig:49-key-bottom}), allowing them to be adjusted during installation.
 Ribbon cables connected the PCBs, providing flexibility during assembly while maintaining a compact form factor.
@@
 \subsection{Materials and Construction}
-The keyboard was designed to replicate the tactile and aesthetic sensations of playing an historical Italian harpsichord.
+The keyboard was designed to replicate the tactile and aesthetic sensations of playing a historical Italian harpsichord.
 Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard.
 The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra.
-The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}.
+The design was inspired by the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}, considered one of the first prototypes to have been conceived with two sets of 8-foot strings.
 An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.
@@
 
+\begin{figure}[!t]
+    \centering    
+    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
+    \caption{
+    3-Key Model Harpsichord Mechanism    
+    %acquired from \anon{San Colombano, Bologna}, 2023.
+    }
+    \label{fig:3key}
+\end{figure}
+
 \subsection{Project Deployment}
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
-    \caption{Block diagram of PCB connections. A separate sensor signal is routed to the Arduino. LED and multiplexer (MUX) controls signal are daisy-chained through each sensor PCB.}
-    \label{fig:system-block-diagram}
-\end{figure*}
-
 Figure \ref{fig:system-block-diagram} shows a block diagram of the finalised hardware setup. 
 The system evolved through iterative prototyping, beginning with simple threshold-based testing and ending in a fully functional multi-sensor interface capable of triggering MIDI events.
-
-\begin{figure}
-    \centering    
-    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
-    \caption{
-    3-Key Model Harpsichord Mechanism    
-    %acquired from \anon{San Colombano, Bologna}, 2023.
-    }
-    \label{fig:3key}
-\end{figure}
 
@@
 
-\begin{figure}[!h] 
+\begin{figure}[!b] 
   \centering
@@
 
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{src/images/block-diagram-2.pdf}
+    \caption{Block diagram of PCB connections. A separate sensor signal is routed to the Arduino. LED and multiplexer (MUX) controls signal are daisy-chained through each sensor PCB.}
+    \label{fig:system-block-diagram}
+\end{figure*}
+
 The optical sensors are wired in a voltage divider configuration (Figure \ref{fig:simple-schematic}) and the circuit outputs a voltage based on the infrared light reflected from nearby surfaces. 
@@
     \centering
-    \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
-    \\
     \includegraphics[width=0.8\linewidth,trim={0 2cm 0 2.5cm},clip]{src/images/baffles.png}
@@
 
-\begin{figure}[!b] 
-  \centering
-  \includegraphics[width=\linewidth]{src/images/sensor-board-reverse-side-anon.png} 
-  \caption{Reverse side of the sensor board showing the RGB LEDs and terminal connections.
-The \texttt{SIG} terminal (Top Left) is routed to an ADC channel.
-Power \texttt{5V, 3V, GND}, multiplexer channels \texttt{A, B, C} and LED PWM input \texttt{DI} and output \texttt{DO}  connect from the previous sensor board (Left) and are routed to the next (Right).}
-  \Description{} 
-  \label{fig:sensor-reverse}
-\end{figure}
-
 \subsection{Controller Board}\label{controller-board}
@@
 In addition to solder terminals for the sensor board channels, the controller board contained:
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
-    \caption{Components and cabling hidden in the bottom section of the instrument case.}
-    \label{fig:mac-mini}
-\end{figure}
 
@@
 
+\begin{figure}[!b] 
+  \centering
+  \includegraphics[width=\linewidth,trim={0 1.5cm 0 1.5cm},clip]{src/images/sensor-board-w-baffles.jpeg}
+    \\
+  \includegraphics[width=\linewidth]{src/images/sensor-board-reverse-side-anon.png} 
+  \caption{Sensor board showing the sensor side (top) and rear side with RGB LEDs and terminal connections (bottom).
+The \texttt{SIG} terminal (Top Left) is routed to an ADC channel.
+Power \texttt{5V, 3V, GND}, multiplexer channels \texttt{A, B, C} and LED PWM input \texttt{DI} and output \texttt{DO}  connect from the previous sensor board (Left) and are routed to the next (Right).}
+  \Description{} 
+  \label{fig:sensor-reverse}
+\end{figure}
+
 For the initial version of the exhibition, a harpsichord sample library was used and controlled via Native Instrument's Kontakt\footnote{\href{https://www.native-instruments.com/en/products/komplete/samplers/kontakt-8/?srsltid=AfmBOorKUf43SoIxGBS2-GnXmKHHkgcfcfWRskpweDhLSG3FiF0qrf2w}{Kontakt webpage} (accessed 31 Jan 2025)}.
@@
 An iPad is used as a monitor through which visitors can adjust playback parameters such as tuning, voicing and which stops are engaged.
-The next iteration will use a library of samples from instruments in the collection, which the museum is currently carrying out.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth,trim={0 5cm 0 0},clip]{src/images/mac-mini.jpg}
+    \caption{Components and cabling hidden in the bottom section of the instrument case.}
+    \label{fig:mac-mini}
+\end{figure}
 
@@
 The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 people -- consisting of staff, surveillance personnel and visitors who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
-The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
+Visitor feedback was collected through short informal interviews and primarily to identify technical problems.
 The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.
@@
 
+This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in the article `The \emph{O} in NIME' \cite{Masu_NIME_2023}.
+By integrating historical keyboard-building traditions with digital augmentation, it offers a possibility to extend the practical lifespan of musical heritage while maintaining the tactile qualities of the original instrument.
+Rather than prioritising technical novelty, the project demonstrates how digital interventions can support meaningful interaction with a shared musical heritage, ensuring its continued relevance in contemporary museum contexts.
+
 A longer discussion, but one going beyond the scope of this work, is whether the current setup or its future iterations may be effectively used to build legitimate replicas of historical musical instruments or even become a kind of \emph{new} musical instrument altogether.

### 2025-02-04 src/main.tex
submission draft

@@
 % \documentclass[sigconf]{nimeart}
-\documentclass[sigconf,anonymous]{nimeart}
+\documentclass[sigconf]{nimeart}
 % \documentclass[sigconf,anonymous,review]{nimeart}
@@
 \affiliation{%
-  \institution{Master Harpsichord Maker \\ Museo San Colombano}
+  \institution{Master Harpsichord Maker}
   \city{Pesaro}
@@
 
-This paper describes the design and creation of an electronically augmented replica of a harpsichord keyboard with a typical early Italian layout to create a digital musical instrument. 
+This paper describes the design and creation of an electronically augmented replica of a historical harpsichord keyboard with a typical 17th-century Italian layout to create a digital musical instrument.
 The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
-The replica attempts to balance the competing demands of historical authenticity, public accessibility, and preservation. 
-It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques.
+The replica balances the competing demands of historical authenticity, public accessibility, and preservation. 
+It replicates the original instrument's tactile feedback and mechanical resistance using historically informed construction techniques.
 Optical sensors integrated within the mechanism capture the jacks' motion data, enabling MIDI message generation.
 This work situates itself within broader discussions on the role of technology in museums. 
-A keyboard interface of this type offers an opportunity to enhance visitor interaction with cultural heritage while safeguarding delicate artefacts. 
+A keyboard interface of this type offers an opportunity to enhance visitor interaction with musical heritage while safeguarding delicate artefacts. 
 The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.
@@
 \begin{acks}
-This work received funding from the European Commission's Horizon2020 programme, through grant \anon{ERC-StG 950084 NEMUS}
+The construction of the keyboard received funding from NextGenerationEU within the framework of the accessibility project for the Removal of Physical and Cognitive Barriers in Museums, Libraries and Archives: Sound Heritage No Limits, in the Museum of San Colombano in Bologna - CUP: C38J23000040001 - in the context of PNRR, Mission 1.
+MH and MD's work was supported by the European Commission's Horizon2020 programme through grant \anon{ERC-StG 950084 NEMUS}.
 

### 2025-02-05 src/article.tex
fix #25

@@
 \caption{Details of the replica, including the key slots and sides, keys, and jacks with seagull plectra.}
+\Description{4 images contain details of the jack rows.
+A large image shows the end of the jack rows where an extrusion is put to allow jack rows to be disengaged.
+The smaller images show the keys and jack removed from the harpsichord exposing the materials from which they are constructed.
+Shown also is the shims for the jack which stop it from easily sliding.}
 \label{fig:details}
@@
   \caption{Gradient stickers applied to the side of the jack body. The coarse gradient scale was selected to maximise signal excursion while preserving signal readout stability.}
-  \Description{} 
+  \Description{Close of two jacks with gradient stickers applied to each.
+Both jack lie on top of the vinyl sheet that was used for printing.
+The gradient is a long black bar that terminates in 7 rectangles of grey over an 8mm span.
+The remainder of the sticker is white.
+The jacks have the numbers 46 and 49 written on them at different offset illustrating the numbering system to identify their position along the keyboard.}
   \label{fig:jack-tags}
@@
   \caption{Underside of the full model keyboard, showing two chambers: the front chamber (top) and the rear chamber (bottom).} 
-  \Description{} 
+  \Description{Underside view of the models with keys removed.
+A cross brace split the case into four quadrants the top two are the front chambe and the bottom are the rea chamber.
+Visible inside are the PCB which are connected to each other with multicoloured ribbon cable.
+From the right 2 long cables, one for sensor signals the other for LED and multiplexer signals.
+Both sets of cables terinate at the controller board.
+The PCBs are attached with white plastic supports to the underside of the wrest plank, a thick block of wood in which the tuning pegs are screwed.}
   \label{fig:49-key-bottom}
@@
     \caption{Logarithmic shape of an original \anon{Trasuntino} harpsichord contrast with the rectangular shape of the replica}    
-    \Description{}
+    \Description{Line drawing of two harpsichord shapes.
+The top line drawing shows the longer tapering form of a full scale instrument if it had the same key layout as the replica.
+The second line drawing is of the replica harpsichord.
+The line drawing share the same scale and it can be seen that the replica is roughly one quarter the depth of a full scale instrument.}
     \label{fig:log-harp-comp}
@@
     }
+    \Description{Side view of a small model mechanism of a harpsichord.
+The model is cut as though it were a cross section of a harpsichord mechanism.
+The Instrument is not to scale and only has a small sound board.
+Visible are the pivots through which the keys are mounted and how the two sets of jack rest on the end of the key.}
     \label{fig:3key}
@@
   \caption{Optical sensor in a simple voltage divider circuit. \texttt{V\_OUT} is routed to one of 8 channels on the CD4051BE multiplexer.}
-  \Description{} 
+  \Description{A circuit diagram of a voltage divider circuit.
+The Image shows a 100 ohm resistor in line with the LED and a 10 thousand ohm resistor connected to the emitter of the phototransistor and the collector connected to ground.
+The line to the bisects before the 10 thousand ohm resistor to a label titled V OUT.
+Arrows from the LED illustrate how light controls the base of the phototransistor.}
   \label{fig:simple-schematic}
@@
     \caption{Block diagram of PCB connections. A separate sensor signal is routed to the Arduino. LED and multiplexer (MUX) controls signal are daisy-chained through each sensor PCB.}
+    \Description{The block diagram shows seven blocks each with sensor board written on them.
+Each sensor block has a solid arrow which connects to a block labelled Arduino Nano 33 BLE.
+A dashed arrow is drawn from the Ardnuino Nano 33 BLE box and travels through each sensors board.
+A legend shows th soldi line represnts the flow of the sensor signal and the dashed arrow the flow of LED an multiplexer control signals.
+Angled arrows from the sensor board point towards and away from a hatched lined representing the surface of the jacks, which illustartes the reflection of light.
+Lines from boxes labelled FRAM and Rotary Encoder point to and from the box labelled Arduino Nano 33 BLE.}
     \label{fig:system-block-diagram}
@@
     \caption{Baffles designed to prevent cross-talk between adjacent sensors.}
-    \Description{}
+    \Description{A 3D model of the baffles used to reduce light leakage.
+Two slots are on the left and right sides.
+Two holes are visible beside each slot, through which the baffles can be attached to the PCB.
+A rectangular cutaway is visible for the multiplexer chip.
+The baffles contained14 angled extrusions 2 for each sensor.}
     \label{fig:baffles}
@@
 Power \texttt{5V, 3V, GND}, multiplexer channels \texttt{A, B, C} and LED PWM input \texttt{DI} and output \texttt{DO}  connect from the previous sensor board (Left) and are routed to the next (Right).}
-  \Description{} 
+  \Description{Two images one of the front and one of the rear of a single sensor board.
+The top image shows the baffles in place and how they enclose the field of view of each sensor.
+The bottom image shows the wires from neighbouring circuit boards terminating and the line of 7 RGB LEDs in the same spacing as the LEDs.}
   \label{fig:sensor-reverse}
@@
     \caption{Components and cabling hidden in the bottom section of the instrument case.}
+    \Description{The mac mini and its display fitted in a shallow draw with a channel drilled in the end through which cables can be connected.
+The draw has dividers in place to separate the mac mini from the display.}
     \label{fig:mac-mini}
@@
 \caption{The keyboard installed in the \emph{Oratory} at \anon{San Colombano}, and a user wearing headphones playing it.}
+\Description{2 images on showing the keyboard in situ in the museum and the second a user playing the keyboard.
+The harpsichord placed on top the drawer in which the mac mini and display reside.
+The whole harpsichord unit is positioned on a table.
+The right image shows a user standing playing the keyboard with a display for the exhibition visible in the background.}
 \label{fig:oratory}

### 2025-02-05 src/main.tex
fix #25

@@
 % \documentclass[sigconf]{nimeart}
-\documentclass[sigconf]{nimeart}
-% \documentclass[sigconf,anonymous,review]{nimeart}
+% \documentclass[sigconf]{nimeart}
+\documentclass[sigconf,anonymous,review]{nimeart}
 \usepackage{graphicx,subcaption,svg}
@@
   \caption{49-key Model Harpsichord Mechanism by \anon{Roberto Livi, Bologna 2024}}
-  \Description{49-key Model Harpsichord Mechanism by \anon{Roberto Livi, Bologna 2024}}
+  \Description{An overhead and front view of the replica harpsichord.
+The overhead view shows the exposed jacks rows, 49 per row and the slant of the bridge and change of sctring from the left to the right.
+The front view shows arch detailing on the edge of each key in the same style as is found on early Italian harpsichords.}
   \label{fig:teaser}

### 2025-04-09 src/article.tex
address part of #45

@@
 To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian school, which is the subject of this paper.
-The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages.
+% target audience: The target audience for the interface is visitors to the museum, who would interact with instrument if it were working order.
+% The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages. 
+The keyboard targets all visitors to the museum, regardless of playing ability, and can be used without special permission.
+Jacks that pluck two choirs of muted strings, across 49-keys, are used to generate MIDI messages that are sent to a connected computer for audio synthesis. 
 The interface is presently linked to a commercial software sampler; however, the ultimate aim is to make the sound of instruments in the collection that can no longer be maintained in playable condition accessible.
@@
 The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica.
-Since the Tagliavini collection hosts primarily early plucked stringed keyboards of Italian origin, the keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
+Since the \anon{Tagliavini collection} hosts primarily early plucked stringed keyboards of Italian origin, the keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}.
 This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}.
@@
 Both jack lie on top of the vinyl sheet that was used for printing.
-The gradient is a long black bar that terminates in 7 rectangles of grey over an 8mm span.
+The gradient is a long black bar that terminates in 7 rectangles of grey over an 8 millimetre span.
 The remainder of the sticker is white.
@@
   \Description{Underside view of the models with keys removed.
-A cross brace split the case into four quadrants the top two are the front chambe and the bottom are the rea chamber.
+A cross brace split the case into four quadrants the top two are the front chamber and the bottom are the rear chamber.
 Visible inside are the PCB which are connected to each other with multicoloured ribbon cable.
 From the right 2 long cables, one for sensor signals the other for LED and multiplexer signals.
-Both sets of cables terinate at the controller board.
+Both sets of cables terminate at the controller board.
 The PCBs are attached with white plastic supports to the underside of the wrest plank, a thick block of wood in which the tuning pegs are screwed.}
@@
 A dashed arrow is drawn from the Ardnuino Nano 33 BLE box and travels through each sensors board.
-A legend shows th soldi line represnts the flow of the sensor signal and the dashed arrow the flow of LED an multiplexer control signals.
-Angled arrows from the sensor board point towards and away from a hatched lined representing the surface of the jacks, which illustartes the reflection of light.
+A legend shows the solid line represents the flow of the sensor signal and the dashed arrow the flow of LED an multiplexer control signals.
+Angled arrows from the sensor board point towards and away from a hatched lined representing the surface of the jacks, which illustrates the reflection of light.
 Lines from boxes labelled FRAM and Rotary Encoder point to and from the box labelled Arduino Nano 33 BLE.}
@@
     \\
-  \includegraphics[width=\linewidth]{src/images/sensor-board-reverse-side-anon.png} 
+  \includegraphics[width=\linewidth,trim={4.45cm 4.5cm 5.25cm 4cm},clip]{src/images/sensor-board-reverse-side.png} 
   \caption{Sensor board showing the sensor side (top) and rear side with RGB LEDs and terminal connections (bottom).
@@
 By integrating historical keyboard-building traditions with digital augmentation, it offers a possibility to extend the practical lifespan of musical heritage while maintaining the tactile qualities of the original instrument.
-Rather than prioritising technical novelty, the project demonstrates how digital interventions can support meaningful interaction with a shared musical heritage, ensuring its continued relevance in contemporary museum contexts.
+Rather than prioritising technical novelty, the project demonstrates how digital interventions can support multi-modal interaction with a shared musical heritage, ensuring its continued relevance in contemporary museum contexts.
 
